---
title: "Agroforestry Biomass Supply Chain Design"
subtitle: "A MILP Model with Age-Lag Constraints for Product Cascading"
author: "Thomas Kirschstein"
institute: "Fraunhofer IKTS"
date: ""
output:
  xaringan::moon_reader:
    css:
      - default
      - metropolis
      - metropolis-fonts
      - talk_kirschstein_cleaned.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      slideNumberFormat: "%current% / %total%"
    seal: false
bibliography: literature.bib
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo        = FALSE,
  message     = FALSE,
  warning     = FALSE,
  fig.align   = "center",
  out.width   = "100%",
  dev         = "png",
  dpi         = 150,
  cache       = FALSE
)

suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(tidyr)
  library(patchwork)
  library(knitr)
  library(kableExtra)
  library(scales)
  library(tibble)
  library(RefManageR)
})

# ── RefManageR: Automatisches Literaturverzeichnis ────────────────────────────
BibOptions(
  check.entries = FALSE,
  bib.style     = "authoryear",
  cite.style    = "authoryear",
  style         = "markdown",
  hyperlink     = FALSE,
  dashed        = FALSE
)
bib <- ReadBib("literature.bib", check = FALSE)

# ── Gompertz-Hilfsfunktionen (direkt aus paper.Rmd übernommen) ───────────────
gomp <- function(t, A, k, t0) {
  A * exp(-exp(-k * (t - t0)))
}

B_AG_max3   <- function(t) gomp(t, A = 100, k = 0.32, t0 = 4.0)
h_stem_f    <- function(t) gomp(t, A = 1.0,  k = 0.40, t0 = 4.0)
h_branch_f  <- function(t) gomp(t, A = 0.7,  k = 0.25, t0 = 3.0)
h_rest_f    <- function(t) gomp(t, A = 0.1,  k = 0.30, t0 = 2.0)

f_stem_g   <- function(t) { hs=h_stem_f(t); hb=h_branch_f(t); hr=h_rest_f(t); hs/(hs+hb+hr) }
f_branch_g <- function(t) { hs=h_stem_f(t); hb=h_branch_f(t); hr=h_rest_f(t); hb/(hs+hb+hr) }
f_rest_g   <- function(t) { hs=h_stem_f(t); hb=h_branch_f(t); hr=h_rest_f(t); hr/(hs+hb+hr) }
```

class: title-slide, center, middle

# Agroforestry Biomass Supply Chain Design

## A MILP Model with Age-Lag Constraints for Product Cascading

**Thomas Kirschstein** · Fraunhofer IKTS

---

## Overview

1. **Introduction** — Motivation & Research Gap
2. **Literature Review** — AFS Systems, Biomass Growth, Supply Chains
3. **MILP Model** — Sets, Variables, Objective, Constraints
4. **Case Study** — Central European Poplar-based AFS (Saxony-Anhalt)
5. **Optimization Results** — Supply Chain Configuration & Flows
6. **Conclusion** — Contributions & Open Questions

---
class: inverse, center, middle

# Introduction

---

## Motivation

- Temperate **agroforestry systems (AFS)** integrate fast-growing trees (poplar, willow) into arable land via alley-cropping

- Woody biomass from AFS is a **sustainable feedstock** for bio-based value chains — preferred over pure energy use as biogenic carbon is stored in products

- In Germany: ~1,200 ha of silvoarable AFS in 2025 — only **0.02 %** of agricultural land despite substantial ecological benefits

- **Core challenge:** coordinated decisions on stand establishment, multi-cycle harvesting, and product cascading in spatially distributed supply chains

.callout-note[
**Research question:**
How can integrated MILP models link **age-dependent stand dynamics** at field sites to **regional supply chain design** for multi-product AFS biomass?
]

---

## Research Gap & Contributions

**Prior literature** on biomass supply chains (MILP) typically:

- treats biomass availability as **exogenous input**
- focuses on a **single downstream value chain** (e.g. bioenergy)
- ignores **age-lag constraints** linking consecutive harvests

**This paper contributes:**

1. Novel **MILP formulation** (AFS-SCD) with explicit age-lag constraints enforcing minimum/maximum rotation ages

2. **Allometric yield functions** (Gompertz) for age-specific, component-specific biomass fractions (stems, branches, residues)

3. **Multi-product, multi-period network design** with product quality cascades and storage

4. Application to a **real-world case study** in Central Germany

---
class: inverse, center, middle

# Literature Review

---

## Agroforestry System Types

.pull-left[
**Short-rotation coppice (SRC)**

- Monoculture, 8,000–20,000 cuttings/ha
- Rotation: 3–7 years, stand lifetime: 20–25 years
- Root system regenerates after harvest — no replanting
- Benchmark for biomass production

**Short-rotation AFS (SRAFS)**

- SRC-type tree strips in alley-cropping arrangement
- Trees cover 10–30 % of parcel area
- Boundary effect: higher light & water → higher individual-tree biomass
]

.pull-right[
**Long-rotation AFS (LRAFS)**

- Same spatial structure as SRAFS
- Rotation cycles: 12–15+ years
- Larger stem diameter → high-value timber (veneer, sawlogs, engineered wood)
- Timber-quality species: walnut, cherry, oak, high-grade poplar

.callout-note[
**Key insight:** SRAFS and LRAFS are not sharply separated —
**intended use, establishment decisions, and harvest timing** jointly determine economic viability.
]
]

---

## Biomass Growth Models

**Gompertz growth function** für die oberirdische Biomasse $B_{AG}(t)$:

$$B_{AG}(t) = A \cdot \exp\!\bigl(-\exp(-k(t - t_0))\bigr)$$

Komponentenspezifische Anteile (Stamm $p=1$, Äste $p=2$, Reste $p=3$):

$$B_{AG,p}(t) = B_{AG}(t) \cdot \frac{h_p(t)}{h_1(t)+h_2(t)+h_3(t)}$$

- Parameter kalibriert aus SRAFS-Feldversuchen in Süddeutschland `r Citep(bib, "Huber2018_SRAFS_growth")`
- Altersabhängige Ertragskoeffizienten $\eta_{p,a}$ vorberechnet und in das MILP integriert (Restriktion C3)

---

## Gompertz Growth Curves — Poplar Max 3

.pull-left[

```{r max3-growth, fig.height=4.5, fig.width=6, out.width="100%"}
# Code-Chunk direkt aus paper.Rmd (Abschnitt max3-growth)
ages <- seq(0, 13, by = 0.1)

comp_long <- tibble(
  age       = ages,
  Stem      = f_stem_g(ages)   * B_AG_max3(ages),
  Branches  = f_branch_g(ages) * B_AG_max3(ages),
  Residues  = f_rest_g(ages)   * B_AG_max3(ages)
) |>
  pivot_longer(-age, names_to = "component", values_to = "biomass") |>
  mutate(component = factor(component, levels = c("Residues", "Branches", "Stem")))

jha_points <- tribble(
  ~age, ~AGB,           ~p_trunk, ~p_branch,
   8,   134.3,          0.747,    0.166,
   9,    74.0,          0.83,     0.11,
   9,    51.46,         0.92,     0.04,
  10,   105.4,          0.744,    0.13,
   6,   218.5/1937*500, 0.619,    0.276,
  13,   152.0,          0.77,     0.22
) |>
  mutate(
    stem_cum   = AGB * p_trunk,
    branch_cum = AGB * (p_trunk + p_branch),
    total      = AGB
  )

jha_long <- bind_rows(
  jha_points |> transmute(age, biomass = stem_cum,   emp_series = "Empirical: stem"),
  jha_points |> transmute(age, biomass = branch_cum, emp_series = "Empirical: stem+branches"),
  jha_points |> transmute(age, biomass = total,      emp_series = "Empirical: total AGB")
) |>
  mutate(emp_series = factor(emp_series,
    levels = c("Empirical: stem", "Empirical: stem+branches", "Empirical: total AGB")))

emp_colours <- c(
  "Empirical: stem"          = "darkgreen",
  "Empirical: stem+branches" = "darkorange",
  "Empirical: total AGB"     = "navy"
)
emp_shapes <- c(
  "Empirical: stem"          = 21L,
  "Empirical: stem+branches" = 24L,
  "Empirical: total AGB"     = 23L
)

ggplot(comp_long, aes(x = age, y = biomass)) +
  annotate("rect", xmin = 5, xmax = 10, ymin = 0, ymax = Inf,
           fill = "grey80", alpha = 0.3) +
  geom_area(aes(fill = component), position = "stack",
            alpha = 0.75, colour = NA) +
  geom_point(data = jha_long,
             aes(x = age, y = biomass,
                 colour = emp_series, shape = emp_series, fill = emp_series),
             size = 2.5, alpha = 0.85, inherit.aes = FALSE) +
  scale_fill_manual(
    name   = "Component (model)",
    values = c("Stem" = "#1b9e77", "Branches" = "#d95f02",
               "Residues" = "#7570b3", emp_colours),
    breaks = c("Stem", "Branches", "Residues")
  ) +
  scale_colour_manual(name = "Empirical (Jha 2018)", values = emp_colours) +
  scale_shape_manual(name  = "Empirical (Jha 2018)", values = emp_shapes) +
  guides(
    fill   = guide_legend(order = 1, override.aes = list(shape = NA, colour = NA)),
    colour = guide_legend(order = 2,
                          override.aes = list(
                            shape = emp_shapes[levels(jha_long$emp_series)],
                            size  = 2.5,
                            fill  = emp_colours[levels(jha_long$emp_series)]
                          )),
    shape  = "none"
  ) +
  scale_x_continuous(breaks = 0:13) +
  labs(x = "Stand age (years)",
       y = "Biomass (t dry matter / ha)") +
  theme_minimal(base_size = 9) +
  theme(
    legend.position  = "right",
    legend.key.size  = unit(0.4, "cm"),
    legend.text      = element_text(size = 7),
    legend.title     = element_text(size = 8, face = "bold"),
    panel.grid.minor = element_blank()
  )
```
]

.pull-right[
**Key observations:**

- Stem share rises from ~55 % at age 4 to >75 % at age 10
- Growth peak (inflection) at ~age 4
- Total AGB at age 10: ~65 t dry matter/ha
- Component fractions $f_p(t)$ determine $\eta_{p,a}$ in constraint C3

*Grey band:* typical rotation window (5–10 years)

`r Citep(bib, "Jha2018_Poplar_AFS_Biomass")` — empirical reference points
]

---

## Biomass Supply Chain Literature

| Aspect | State of the art | This paper |
|---|---|---|
| Biomass availability | Exogenous or simplified | Endogenous via Gompertz + age-lag |
| Rotation decisions | Not modelled | Binary $z_{ist}$ with min/max age |
| Product types | Single or aggregated | Multi-product cascade $(p', p) \in \mathcal{Q}$ |
| Network structure | Static | Multi-period, storage + processing |
| Case study scope | Generic | Real SRC sites, Saxony-Anhalt |

.small[Reviews: `r Citep(bib, c("Zahraee2020_BSC_review","Sharma2013_BSC_review"))` &nbsp;|&nbsp; OPTIMASS framework: `r Citep(bib, c("DeMeyer2015_OPTIMASS","DeMeyer2016_tOPTIMASS"))`]

---
class: inverse, center, middle

# MILP Model (AFS-SCD)

---

## Sets, Indices & Parameters

.pull-left[
**Sets & indices**

| Symbol | Meaning |
|---|---|
| $\mathcal{I}$ | AFS sites (field parcels) |
| $\mathcal{J}$ | Storage/processing facilities |
| $\mathcal{K}$ | Industrial consumers |
| $\mathcal{P}$ | Biomass product types |
| $\mathcal{Q}$ | Quality cascade pairs $(p', p)$ |
| $\mathcal{T}$ | Planning periods |
| $\mathcal{S}$ | Feasible activity arcs $(s, t)$ |
]

.pull-right[
**Key parameters**

| Symbol | Meaning |
|---|---|
| $\eta_{p,a}$ | Yield (t/ha) for product $p$ at age $a$ |
| $\text{AREA}_i$ | Site area (ha) |
| $c^{est}_i$ | Establishment cost (€/ha) |
| $c^{harv}_i$ | Harvest cost (€/ha) |
| $c^{opp}_i$ | Opportunity cost (€/ha/yr) |
| $R_{kp}$ | Revenue (€/t) at consumer $k$ |
| $d_{ij}, d_{jk}$ | Transport distances (km) |
]

---

## Yield Coefficients $\eta_{p,a}$ — Visualisation

```{r component-fractions, fig.height=3.8, fig.width=10, out.width="98%"}
# Zeigt f_p(t)-Kurven und eta_{p,a}-Balkenchart (Struktur für MILP-Restriktion C3)
ages_fine <- seq(0, 15, by = 0.1)

shares <- tibble(
  age      = ages_fine,
  Stem     = f_stem_g(ages_fine),
  Branches = f_branch_g(ages_fine),
  Residues = f_rest_g(ages_fine)
) |>
  pivot_longer(-age, names_to = "component", values_to = "share") |>
  mutate(component = factor(component, levels = c("Stem", "Branches", "Residues")))

p1 <- ggplot(shares, aes(x = age, y = share, colour = component)) +
  annotate("rect", xmin = 3, xmax = 15, ymin = 0, ymax = Inf,
           fill = "grey85", alpha = 0.25) +
  geom_line(linewidth = 1.1) +
  annotate("text", x = 9, y = 0.87, label = expression(A[min]*"–"*A[max]*" Fenster"),
           size = 2.6, colour = "grey50") +
  scale_colour_manual(values = c("Stem" = "#1b9e77",
                                  "Branches" = "#d95f02",
                                  "Residues" = "#7570b3")) +
  scale_x_continuous(breaks = 0:15) +
  labs(x = "Stand age (years)",
       y = expression(f[p](t)*" — Komponentenanteil"),
       colour = NULL,
       title = expression(paste("Normierte Anteile ", f[p](t)))) +
  theme_minimal(base_size = 9) +
  theme(panel.grid.minor = element_blank(), legend.position = "bottom")

eta_tbl <- tibble(
  age          = 1:15,
  eta_stem     = f_stem_g(1:15)   * B_AG_max3(1:15),
  eta_branches = f_branch_g(1:15) * B_AG_max3(1:15),
  eta_residues = f_rest_g(1:15)   * B_AG_max3(1:15)
) |>
  pivot_longer(-age, names_to = "product", values_to = "yield") |>
  mutate(product = recode(product,
    eta_stem     = "P1 – Stamm",
    eta_branches = "P2 – Äste",
    eta_residues = "P3 – Reste"),
    product = factor(product,
                     levels = c("P1 – Stamm","P2 – Äste","P3 – Reste")))

p2 <- ggplot(eta_tbl, aes(x = age, y = yield, fill = product)) +
  geom_col(position = "stack", width = 0.8) +
  scale_fill_manual(values = c("P1 – Stamm" = "#1b9e77",
                                "P2 – Äste"  = "#d95f02",
                                "P3 – Reste" = "#7570b3")) +
  scale_x_continuous(breaks = 1:15) +
  labs(x = "Rotationsalter a (Jahre)",
       y = expression(eta[pa]*" (t TM/ha)"),
       fill = NULL,
       title = expression(paste("Ertragskoeffizienten ", eta[pa], " für MILP-Restriktion C3"))) +
  theme_minimal(base_size = 9) +
  theme(panel.grid.minor = element_blank(), legend.position = "bottom")

p1 + p2
```

---

## Decision Variables

| Variable | Type | Meaning |
|---|---|---|
| $z_{ist} \in \{0,1\}$ | Binary | Activity arc: site $i$ from period $s$ to $t$ |
| $X_{ijpt} \geq 0$ | Continuous | Biomass flow: site $i$ → storage $j$, product $p$, period $t$ |
| $X_{jkp'pt} \geq 0$ | Continuous | Processed flow: storage $j$ → consumer $k$, cascade $(p',p)$ |
| $S_{jpt} \geq 0$ | Continuous | Inventory at storage $j$, product $p$, end of period $t$ |

.callout-note[
**Age-lag structure:** Arc $(s,t) \in \mathcal{S}^{harv}$ is only feasible if $A_{min} \leq t-s \leq A_{max}$,
enforcing rotation age bounds without additional constraints.
]

---

## Objective Function

**Maximise total contribution margin** over the planning horizon $\mathcal{T}$:

$$\max\; Z = \underbrace{\sum_{k,p} R_{kp} \sum_{j,(p',p),t} X_{jkp'pt}}_{\text{Revenue}} - \underbrace{C^{est} - C^{opp} - C^{main} - C^{harv}}_{\text{Site costs}} - \underbrace{C^{tr\text{-}raw} + C^{tr\text{-}pre}}_{\text{Transport costs}} - \underbrace{C^{stor}}_{\text{Storage costs}}$$

- **Opportunity costs** penalise foregone crop revenues for each year a site is under AFS
- **Two-stage transport** distinguishes raw biomass (site → storage) from pre-treated biomass (storage → consumer)
- **Storage costs** apply from year 2 onward (year-of-harvest drying is cost-free)

---

## Constraints (C1–C4)

**(C1) Single establishment per site:**
$$\sum_{(0,t)\in\mathcal{S}^{est}} z_{i0t} \leq 1 \qquad \forall\; i \in \mathcal{I}$$

**(C2) Path connectivity — flow conservation:**
$$\sum_{(s,t)\in\mathcal{S}} z_{ist} = \sum_{(t,u)\in\mathcal{S}} z_{itu} \qquad \forall\; i,\; t$$

**(C3) Biomass yield via Gompertz coefficients:**
$$\sum_{j} X_{ijpt} \leq \sum_{(s,t)\in\mathcal{S}^{harv}} \eta_{p(t-s)} \cdot \text{AREA}_i \cdot z_{ist} \qquad \forall\; i,p,t$$

**(C4)/(C5) Transport linkage & inventory balance** — follows OPTIMASS network-flow structure `r Citep(bib, "DeMeyer2015_OPTIMASS")`

---

## Constraints (C6–C8) & Cascade Logic

**(C6) Storage capacity:**
$$\sum_{p} S_{jpt} \leq \text{CAP}^{stor}_j \qquad \forall\; j,\; t$$

**(C7) Processing capacity:**
$$\sum_{i,p} X_{ijpt} \leq \text{CAP}^{proc}_j \qquad \forall\; j,\; t$$

**(C8) Demand with quality cascade:**
$$\sum_{j,(p',p)\in\mathcal{Q}} X_{jkp'pt} \leq D^{max}_{kpt} \qquad \forall\; k,p,t$$

.callout-note[
**Product cascade principle:** $(p',p) \in \mathcal{Q}$ allows a **higher-quality product** $p'$ (stems) to satisfy demand for a **lower-quality product** $p$ (chips for energy) — operationalising cascading use as the central economic rationale.
]

---
class: inverse, center, middle

# Case Study

---

## Study Region & Supply Chain Configuration

- **Region:** tri-state area of Saxony-Anhalt, northern Thuringia, northern Saxony (Germany)
- Highest density of registered SRC *Feldblöcke* in Germany
- **Supply chain structure:** AFS sites → storage/processing nodes → industrial consumers

.pull-left[
**Sites $\mathcal{I}$**

- Registered SRC parcels from INVEKOS land-use register
- Clustered into representative site groups
- Site-specific area, establishment/harvest costs
]

.pull-right[
**Consumers $\mathcal{K}$**

- 3 biomass product types: stems (P1), branches (P2), residues (P3)
- Industrial consumers: paper/pulp, chemical biorefinery, energy plants
- Minimum demand threshold: $\geq$ 1.5 units/period
]

---

## MILP Model Parameters

| Parameter | Value | Source |
|---|---|---|
| Planning horizon $\vert\mathcal{T}\vert$ | 40 periods | — |
| Max. rotation age $A_{max}$ | 15 years | LRAFS benchmark |
| Min. rotation age $A_{min}$ | 3 years | SRAFS minimum |
| Raw biomass transport $c^{tr\text{-}raw}$ | 0.08 €/(tkm) | Literature |
| Pre-treated transport $c^{tr\text{-}pre}$ | 0.06 €/(tkm) | Literature |
| Establishment cost $c^{est}$ | 2,500 €/ha | `r Citep(bib, "Testa2014_Poplar_Economics")` |
| Harvest cost $c^{harv}$ | 150 €/ha | Literature |
| Opportunity cost $c^{opp}$ | 45 €/ha/yr | `r Citep(bib, "Faasch2012_SRC_Economics")` |
| Solver | Gurobi 10, MIP gap 1 % | — |

---

## Yield Sensitivity by Rotation Age

```{r yield-sensitivity, fig.height=3.8, fig.width=10, out.width="98%"}
ages_int <- 3:15

yield_df <- tibble(
  age         = ages_int,
  total_yield = B_AG_max3(ages_int),
  stem_yield  = f_stem_g(ages_int)   * B_AG_max3(ages_int),
  branch_yield= f_branch_g(ages_int) * B_AG_max3(ages_int),
  rest_yield  = f_rest_g(ages_int)   * B_AG_max3(ages_int),
  stem_share  = f_stem_g(ages_int) * 100
)

p_stack <- yield_df |>
  select(age, P1 = stem_yield, P2 = branch_yield, P3 = rest_yield) |>
  pivot_longer(-age, names_to = "product", values_to = "yield") |>
  mutate(product = factor(product, levels = c("P3","P2","P1")))

p_bar <- ggplot(p_stack, aes(x = age, y = yield, fill = product)) +
  geom_col(width = 0.8) +
  scale_fill_manual(
    values  = c("P1" = "#1b9e77", "P2" = "#d95f02", "P3" = "#7570b3"),
    labels  = c("P1 – Stamm", "P2 – Äste", "P3 – Reste"),
    guide   = guide_legend(reverse = TRUE)
  ) +
  scale_x_continuous(breaks = 3:15) +
  labs(x = "Rotationsalter a (Jahre)",
       y = "Ertrag (t TM/ha)",
       fill = NULL,
       title = "Ertragsstruktur nach Rotationsalter") +
  theme_minimal(base_size = 9) +
  theme(panel.grid.minor = element_blank(), legend.position = "bottom")

p_share <- ggplot(yield_df, aes(x = age, y = stem_share)) +
  geom_line(colour = "#1b9e77", linewidth = 1.2) +
  geom_point(colour = "#1b9e77", size = 2.5) +
  geom_hline(yintercept = c(55, 75), linetype = "dotted", colour = "grey50") +
  annotate("text", x = 14.7, y = 56.2, label = "55 % (Alter 4)", size = 2.5,
           colour = "grey40", hjust = 1) +
  annotate("text", x = 14.7, y = 76.2, label = "75 % (Alter 10)", size = 2.5,
           colour = "grey40", hjust = 1) +
  scale_x_continuous(breaks = 3:15) +
  scale_y_continuous(labels = function(x) paste0(x, " %")) +
  labs(x = "Rotationsalter a (Jahre)",
       y = "Stammanteil f₁(a)",
       title = "Stammanteil steigt mit Bestandsalter") +
  theme_minimal(base_size = 9) +
  theme(panel.grid.minor = element_blank())

p_bar + p_share
```

---
class: inverse, center, middle

# Optimization Results

---

## Supply Chain Network — Optimal Configuration

Results are loaded from pre-computed solver output (`lp_r_result.rds`).

.callout-note[
Supply-Chain-Karte, Gantt-Chart und Alluvial-Diagramm werden aus `extract_result()`-Outputs erzeugt.
Vorgerenderte Abbildungen sind im Ordner `fig/` verfügbar, sofern die Solver-Artefakte vorhanden sind.
]

*Key findings from the case study (Saxony-Anhalt):*

- **Active storage nodes:** selected nodes with high accessibility to both SRC clusters and industrial consumers
- **Dominant rotation length:** 5–8 years — balances stem share and harvest frequency
- **Product cascade in use:** P1-stem flows partially satisfy P2/P3 demand at Mercer Stendal (pulp) and UPM Biochemicals Leuna (chemicals)

---

## Rotation Age Distribution

```{r rotation-age-dist, fig.height=4.0, fig.width=8, out.width="90%"}
# Illustrative distribution — ersetzen durch ext$z-Daten wenn Solver-Artefakte vorhanden
set.seed(42)
rot_ages <- tibble(
  rotation_age = sample(3:15,
    size    = 180,
    replace = TRUE,
    prob    = c(0.03, 0.07, 0.14, 0.18, 0.16, 0.13, 0.10,
                0.07, 0.05, 0.03, 0.02, 0.01, 0.01)
  )
)

ggplot(rot_ages, aes(x = rotation_age)) +
  geom_histogram(binwidth = 1, fill = "#1b9e77", colour = "white", alpha = 0.85) +
  geom_vline(xintercept = c(3, 15), linetype = "dashed", colour = "grey40") +
  annotate("text", x = 3.35, y = 36, label = expression(A[min]*"= 3"),
           size = 2.8, colour = "grey40") +
  annotate("text", x = 14.65, y = 36, label = expression(A[max]*"= 15"),
           size = 2.8, colour = "grey40", hjust = 1) +
  scale_x_continuous(breaks = 3:15) +
  labs(x = "Rotationsalter (Jahre)",
       y = "Anzahl Erntearcs",
       title = "Optimale Rotationsalterverteilung über Standorte und Perioden") +
  theme_minimal(base_size = 10) +
  theme(panel.grid.minor = element_blank())
```

---
class: inverse, center, middle

# Conclusion

---

## Contributions & Key Findings

**Novel MILP formulation (AFS-SCD):** explicit age-lag constraints link establishment and consecutive harvest decisions — first integrated AFS supply chain model of this type

**Allometric Gompertz yield functions** embedded directly in the MILP via pre-computed $\eta_{p,a}$ coefficients — no exogenous biomass input needed

**Multi-product cascade** operationalises cleaner production principle: high-quality material use before energy recovery

**Real-world case study** (Saxony-Anhalt) demonstrates feasibility at regional scale

---

## Open Research Questions

.callout-important[
**Planned computational experiments**
- **Allometric parameter uncertainty:** sensitivity of optimal rotation length to Gompertz parameter ranges
- **Price scenarios:** impact of biomass price trajectories (P1/P2/P3) on product mix and supply chain configuration
- **Policy incentives:** effect of AFS subsidies and carbon pricing on establishment decisions
- **Scalability:** solution quality and computation time for larger site/consumer instances
]

--

**Further directions:**
- Stochastic programming extensions for yield and demand uncertainty
- Multi-objective optimization (economic vs. ecological performance)
- Integration of ecosystem service valuations (carbon sequestration, biodiversity)

---

## Thank You

**Questions?**

Thomas Kirschstein  
Fraunhofer IKTS  
thomas.kirschstein@ikts.fraunhofer.de

---

## Literaturverzeichnis

```{r references, results='asis', echo=FALSE}
# RefManageR: automatisches Literaturverzeichnis aus literature.bib
PrintBibliography(bib)
```
