---
title: "Agroforestry SmartAgroforst: Poplar biomass modelling"
author:
  - name: "Thomas Kirschstein"
    affiliation: 
      - "Fraunhofer IKTS, Dresden, Germany"
    email: "thomas.kirschstein@ikts.fraunhofer.de"
    correspondingauthor: true
  
journal: "Journal Name"
# elsarticle options: preprint preview style, 1p or 3p for layout
classoption: "preprint,review,1p"
date: "`r Sys.Date()`"
linenumbers: true
numbersections: true
always_allow_html: true


keywords: 
  - Agroforestry 
  - Short-rotation coppice 
  - Biomass supply chain
  - Product cascading

abstract: |
  Short-rotation agroforestry systems are increasingly promoted as a low-carbon feedstock option for bio-based value chains, but their economic viability depends on coordinated decisions on stand establishment, multi-cycle harvesting, and product cascading in spatially distributed supply chains. This paper develops a mixed-integer linear programming (MILP) model for the integrated design and operation of an agroforestry biomass supply chain with explicit age-lag constraints that link consecutive harvests at each site. The formulation couples allometric yield functions for different biomass types with a multi-product, multi-period network design model including storage, processing, and product quality cascades. The model is applied to a stylised case study of Central European poplar-based alley-cropping systems to explore trade-offs between harvest rotation length, product portfolios, and infrastructure configuration. We outline a computational experimentation plan to test the impact of allometric parameter uncertainty, price scenarios, and policy incentives on the optimal deployment of agroforestry systems in cleaner bioeconomy pathways.

output:
  rticles::elsevier_article:
    keep_tex: true
    latex_engine: xelatex

header-includes:
  # Use elsarticle’s review layout and set 1.5 line spacing
#  - \usepackage{setspace}
#  - \onehalfspacing        % or \doublespacing for 1.5 lines
    # Line numbers
#  - \usepackage[left]{lineno}
#  - \linenumbers

bibliography: references.bib
csl: elsevier-harvard.csl
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.path = "figures/"
)

library(magick)
library(dplyr)
library(plotly)
library(ggplot2)
library(tidyr)
library(patchwork)
#library(webshot2)

# Chapman–Richards biomass function
B_cr <- function(t, Bmax, k, m) {
  Bmax * (1 - exp(-k * t))^m
}

# gompertz growth function
gomp <- function(t, A, k, t0) {
  A * exp(-exp(-k * (t - t0)))
}


```

# Introduction

Agroforestry systems that integrate trees with agricultural crops (silvoarable systems) are increasingly recognised as a promising land-use option to reconcile climate mitigation, biodiversity conservation, and rural income generation on the same hectare of land [@Toensmeier2017_Carbon_Agroforestry]. Temperate agroforestry systems (AFS) based on fast-growing species such as poplar, willow, alder, and black locust can deliver substantial biomass yields while improving soil quality and providing multiple ecosystem services compared to conventional monocultures [@Huber2018_SRAFS_growth]. Biomasses produced in AFS is considered as a sustainable feedstock and can be used either energetically, chemically or materially (e.g. in the building sector). The latter two options are generally preferred over the energetic usage from a cleaner production perspective as biogenic carbon is stored in products during their life span which contributes to climate change mitigation. 

Despite substantial research on ecological advantages and pioneering initiatives to establish AFS in practice, the share of AFS in agricultural production systems in still small. E.g. in Germany about 1200 hectars of silvoarable AFS are established in 2025, which accounts for about 0.02\% of total agricultural land use [DEFAF]. One of the core challenges in establishing AFS systems in large scale in practice, is a lack of opportunities to generate substantial revenues from woody biomasses of AFS. On the other side, the rise of bioeconomy creating substantial demands of woody biomass requires a reliable and cost-efficient feedstock supply potentials. As woody biomasses are bulky feedstocks with low energy density, the design of cost-efficient supply chains is crucial to ensure the economic viability of AFS-based value chains. This requires coordinated decisions on stand establishment, multi-cycle harvesting, and product cascading in spatially distributed supply chains, which can be supported by integrated optimisation models that link stand-level growth dynamics to regional supply chain design.

In the last two decades, a large body of research has analysed biomass supply chains for bioenergy and biorefineries using mathematical programming models, often in the form of mixed-integer linear programming (MILP) [@Zahraee2020_BSC_review; @Sharma2013_BSC_review]. These models typically optimise network configuration, feedstock sourcing, and logistics under cost or profit objectives. They often treat biomass availability as an exogenous input or rely on simplified agricultural production decisions. Moreover, downstream processing of biomasses is concentrated on one particular value chain e.g. bioenergy plants consuming all produced biomasses. Recent work has started to integrate operational details such as various biomass types, biomass growth models and regeneration processes in multi-period supply chain models. 

The present paper contributes to this literature in four ways. First, we propose a novel MILP formulation for AFS-based biomass supply chain design that explicitly links establishment decisions and consecutive harvest decisions, ensuring that harvesting decisions at each site are consistent with minimum and maximum rotation ages. Second, we embed allometric yield functions for different biomass types generating age-specific yields for different woody biomass types (like stems, branches, or leaves and twigs). Third, these different woody biomass types can be processed in specific downstream industries, like chemical or paper industry as well as energy sector. The model aims to provide strategic decision support for establishing regiona .AFS-based bioeconomy ecosystems connecting multiple industries and product markets. 

The remainder of the paper is structured as follows. Section \@ref(sec:litreview) reviews the relevant literature on agroforestry biomass systems, allometric biomass modelling, and biomass supply chain optimisation. Section \@ref(sec:allometrics) introduces the allometric models adopted to represent biomass growth of different tree species and biomass types. Section \@ref(sec:milp) summarises the MILP formulation for the agroforestry supply chain with age-lag constraints. Section \@ref(sec:experiments) outlines the planned computational experiments for the case study and discusses expected insights for cleaner production strategies.

# Literature review {#seclitreview}

Agroforestry biomass systems, tree and stand biomass models, and biomass supply chain optimisation have all been studied extensively, but they are rarely combined in an integrated modelling framework that links age‑dependent stand dynamics to regional product‑cascading value chains [@Zahraee2020_BSC_review; @Sharma2013_BSC_review; @Espinoza2017_Biorefinery_SC_review]. This section reviews (i) agroforestry and short‑rotation systems with emphasis on yields and economics, (ii) allometric and age‑dependent biomass yield models, and (iii) biomass supply chain optimisation and cascading use of biomass, and concludes by identifying the modelling gap addressed in this paper.

## Biomass supply chain optimisation

Biomass supply chain design and planning have been widely studied using MILP models at strategic, tactical, and operational levels @Zahraee2020_BSC_review; @Sharma2013_BSC_review; @Espinoza2017_Biorefinery_SC_review. Comprehensive reviews classify models according to decision scope (location, capacity, routing, inventory), time horizon, uncertainty treatment, and sustainability metrics, and show a strong emphasis on bioenergy and biorefinery systems @Lamers2015_Feedstock_Design; @Espinoza2017_Biorefinery_SC_review. Typical models optimise the number, size, and locations of processing facilities, biomass sourcing patterns, multimodal transport flows, and inventory policies under cost‑ or profit‑oriented objectives, sometimes complemented by greenhouse gas emissions or other environmental indicators @Eksioglu2009_Biorefinery_SC; @Arabi2019_Microalgae_Biobutanol_SC.

A key methodological feature of many biomass supply chain models is the integration of long‑term location and capacity decisions with multi‑period logistics decisions such as harvesting, transport, storage, and processing @Eksioglu2009_Biorefinery_SC; @Zahraee2020_BSC_review. For example, mixed‑integer formulations have been used to simultaneously determine biorefinery locations, their capacities, biomass collection and storage sites, and time‑dependent flows of biomass and biofuel, explicitly accounting for seasonality, deterioration, and inventory trade‑offs @Eksioglu2009_Biorefinery_SC. These models demonstrate that a significant share of biofuel production costs is attributable to logistics, and that integrated optimisation of network design and operations can substantially reduce costs compared to decoupled decision making @Eksioglu2009_Biorefinery_SC; @Arabi2019_Microalgae_Biobutanol_SC.

However, in virtually all of these models, biomass availability over time is treated as an exogenous input, either as annual supply potentials per region or as simple time profiles based on harvest windows and assumed yields @Zahraee2020_BSC_review; @DeMeyer2016_tOPTIMASS. Stand‑level growth dynamics, age‑class structure, and rotation decisions are not represented explicitly; instead, land‑use or yield scenarios enter as fixed parameters @DeMeyer2016_tOPTIMASS; @Lamers2015_Feedstock_Design. As a result, these models cannot explore trade‑offs between rotation length, stand management, and supply chain configuration in a unified optimisation framework, especially not for agroforestry systems with repeated coppice cycles and strong interactions between tree growth and harvesting decisions.


## Agroforestry and short‑rotation systems

Agroforestry deliberately integrates woody perennials with crops and/or livestock on the same management unit, with the aim of enhancing productivity, profitability, and environmental performance @Nair1993_IntroAgroforestry; @Toensmeier2017_Carbon_Agroforestry. In temperate regions, short‑rotation agroforestry systems (SRAFS) often take the form of alley‑cropping, where double or multiple tree rows alternate with crop strips, enabling simultaneous production of agricultural and woody biomass @Huber2018_SRAFS_growth. Empirical trials in Southern Germany and Central Europe show that SRAFS with poplar, willow, black alder, and black locust can achieve mean annual increments (MAI) of roughly 7–10 t per ha and year in aboveground woody biomass during the first rotation, with pronounced species‑specific differences in growth trajectories, stand structure, and mortality (see @Huber2018_SRAFS_growth; @Trnka2008_Poplar_SRF).

For short‑rotation coppice (SRC) and very short‑rotation coppice (vSRC) poplar plantations in Italy and Central Europe, economic analyses indicate that rotations of about 5–7 years at moderate planting densities can be competitive with arable crops, but profitability is highly sensitive to yield assumptions, establishment costs, and logistics [@Testa2014_Poplar_Economics]. This sensitivity implies to couple stand‑level growth and  models with supply chain optimisation models, rather than treating biomass availability as an exogenous, time‑invariant parameter [@Zahraee2020_BSC_review]. Beyond yields and economics, agroforestry and SRC systems have been widely analysed for carbon sequestration, nutrient cycling, and biodiversity, reinforcing their role as candidate feedstock systems for regional bioeconomy strategies that emphasise material‑before‑energy use of biomass (see e.g. @Toensmeier2017_Carbon_Agroforestry).

Thus, a fundamental cornerpiece of AFS modeling is the representation of biomass growth dynamics, which can be achieved through allometric models that link tree biometric variables to biomass, and through age‑dependent stand biomass functions that capture growth trajectories over rotation cycles. These models provide the necessary link between stand structure and biomass availability, which can then be integrated into supply chain optimisation models to explore trade‑offs between rotation length, product portfolios, and infrastructure configuration in AFS supply chains.

<!-- Agroforestry value‑chain studies in other regions, notably India, highlight how institutional arrangements, contract tree farming, and dedicated market support structures can transform fragmented wood supply into organised industrial value chains @Singh2023_Agroforestry_ValueChains. These contributions mainly analyse governance, market access, and stakeholder coordination, whereas the present work focuses on techno‑economic optimisation of spatially explicit biomass flows and therefore treats prices, demands, and policy incentives parametrically. -->


## Tree‑level allometric models

Allometric tree biomass models relate easily measured variables, such as stem base diameter (SBD) or diameter at breast height, to aboveground dry biomass using power‑law functions of the form \(M = b_0\,\mathrm{SBD}^{b_1}\), with species‑specific coefficients \(b_0\) and exponents \(b_1\) [@Huber2016_Allometric_SRAFS]. These models allow non‑destructive estimation of individual‑tree and stand biomass from diameter measurements and have been widely used in forestry and agroforestry for both carbon accounting and management planning [@Huber2016_Allometric_SRAFS; @Toensmeier2017_Carbon_Agroforestry].

For SRAFS in Southern Germany, species‑specific allometric functions based on SBD measured 10 cm above ground were developed for black alder, black locust, poplar (Max 3), and willow (Inger) @Huber2016_Allometric_SRAFS. The reported models share a common exponent \(b_1 = 2.603\) across species, while the scale parameter \(b_0\) ranges from about 0.025 to 0.041, reflecting differences in wood density and branching pattern @Huber2016_Allometric_SRAFS. Applying these functions to all shoots per tree and aggregating over observed tree densities yields annual aboveground leafless dry biomass at plot and stand level, from which current annual increment (CAI) and MAI curves can be derived for each species and management system @Huber2018_SRAFS_growth.

These SRAFS studies demonstrate that stand structure, diameter distributions, and mortality strongly affect aggregate yields, with black locust showing high wood density and strong size inequality and willow showing high shoot densities but low wood density and diameter growth @Huber2018_SRAFS_growth. From a modelling perspective, the key point is that tree‑level allometrics provide a robust mapping from biometric state variables to biomass, which can be exploited either to build detailed stand simulators or to construct age‑class yield tables for optimisation models @Huber2016_Allometric_SRAFS.


## Age-dependent biomass yield functions

Existing allometric models link tree biomass to diameter or height but do not yet provide **stand-level** biomass as an explicit function of stand age. Age-dependent biomass functions are useful for yield projections in short-rotation forestry and agroforestry, where management is organized in discrete rotations (e.g. 5–15 years). [@Niemczyk2021_Poplar_Rotation; @Testa2014_Poplar_Economics]

A common approach is to derive age functions from rotation trials that report mean annual increment (MAI) over time and identify the age at which MAI plateaus. For poplar SRWC in northern Poland, Niemczyk et al. (2021) showed that MAI in aboveground dry biomass increases up to about age 10; beyond this, the MAI curve flattens, suggesting a biologically optimal rotation age around 10 years because the slope of MAI is near zero. [@Niemczyk2021_Poplar_Rotation] Reported MAI values for single-stem poplar plantations in a 10-year rotation ranged approximately from 1–15 Mg DM ha\(^{-1}\) yr\(^{-1}\) depending on cultivar. [@Niemczyk2021_Poplar_Rotation] Similar magnitudes and rotation lengths (about 5–7 years for SRC, up to 10–15 years for SRC/SRF) are reported for poplar SRC systems in Italy, where whole-rotation averages of roughly 15 Mg DM ha\(^{-1}\) yr\(^{-1}\) are considered economically viable. [@Testa2014_Poplar_Economics]

A simple and widely used parametric form for age-dependent stand biomass \(B(t)\) is the Chapman–Richards (or Mitscherlich/monomolecular) function:
\[
B(t) = B_{\max}\left(1 - e^{-k \cdot t}\right)^m
\]
with stand age \(t\) in years, asymptotic biomass \(B_{\max}\) (Mg DM ha\(^{-1}\)), rate parameter \(k\) (year\(^{-1}\)), and shape parameter \(m\). This formulation captures the rapid juvenile growth, a phase of approximately linear biomass accumulation, and a saturating phase where growth slows. [@Niemczyk2021_Poplar_Rotation] The corresponding current annual increment (CAI) and mean annual increment (MAI) are
\[
\text{CAI}(t) = \frac{dB}{dt}, \qquad \text{MAI}(t) = \frac{B(t)}{t}.
\]
The age maximizing MAI (optimal rotation under a volume/biomass criterion) can be derived analytically or determined numerically. Fitting such functions to observed age–biomass or age–volume series from rotation trials allows cultivar-specific yield curves that are directly usable in agroforestry simulation models. [@Niemczyk2021_Poplar_Rotation; @Testa2014_Poplar_Economics]

For illustration, we parameterize three stylized poplar curves representing low, medium, and high productivity on temperate mineral soils. Parameter values are loosely informed by the ranges reported by Niemczyk et al. (2021) and Testa et al. (2014) and should be recalibrated when site-specific data are available. [@Niemczyk2021_Poplar_Rotation; @Testa2014_Poplar_Economics]

- Low productivity: \(B_{\max} = 80\) Mg DM ha\(^{-1}\), \(k = 0.18\), \(m = 1.1\)
- Medium productivity: \(B_{\max} = 120\) Mg DM ha\(^{-1}\), \(k = 0.20\), \(m = 1.2\)
- High productivity: \(B_{\max} = 160\) Mg DM ha\(^{-1}\), \(k = 0.22\), \(m = 1.3\)

These values yield MAI maxima between about 8 and 12 years with peak MAI values of roughly 8–16 Mg DM ha\(^{-1}\) yr\(^{-1}\), consistent with the empirical range where 10-year rotations produced MAI of about 10–15 Mg DM ha\(^{-1}\) yr\(^{-1}\) for the best poplar cultivars in northern Poland and Italy. [@Niemczyk2021_Poplar_Rotation; @Testa2014_Poplar_Economics]

In the agroforestry context, such age functions can be combined with tree density and allometric partitioning to above- and below-ground biomass to obtain tree-row biomass over time for different designs. Calibration against local inventory data or growth measurements (e.g. basal area increment and height growth) is recommended before using the curves for scenario analysis.

```{r max3-growth, echo=F, message=FALSE, warning=FALSE, fig.cap='Modeled stand-level biomass components for poplar clone Max 3 over age, represented as stacked Gompertz curves and compared with empirical reference values from agroforestry trials.'}
library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)

# ------------------------------------------------------------
# 1. SBD(t) for Max 3 (Gompertz)
# ------------------------------------------------------------

SBD_max3_gomp <- function(t){
  # Asymptotic SBD ~ 10 cm, inflection around year 3–4
  gomp(t, A = 10, k = 0.45, t0 = 3.5)
}

# ------------------------------------------------------------
# 2. Total aboveground biomass B_AG(t) (Gompertz, Mg/ha)
# ------------------------------------------------------------

B_AG_max3 <- function(t) {
  # Asymptotic AG biomass ~ 100 Mg/ha, inflection around year 4
  gomp(t, A = 100, k = 0.32, t0 = 4.0)
}

# ------------------------------------------------------------
# 3. Component shares via Gompertz helper functions h_i(t)
# ------------------------------------------------------------

h_stem   <- function(t) gomp(t, A = 1.0, k = 0.40, t0 = 4.0)
h_branch <- function(t) gomp(t, A = 0.7, k = 0.25, t0 = 3.0)
h_rest   <- function(t) gomp(t, A = 0.1, k = 0.30, t0 = 2.0)

f_stem_gomp <- function(t) {
  hs <- h_stem(t); hb <- h_branch(t); hr <- h_rest(t)
  hs / (hs + hb + hr)
}

f_branch_gomp <- function(t) {
  hs <- h_stem(t); hb <- h_branch(t); hr <- h_rest(t)
  hb / (hs + hb + hr)
}

f_rest_gomp <- function(t) {
  hs <- h_stem(t); hb <- h_branch(t); hr <- h_rest(t)
  hr / (hs + hb + hr)
}

# ------------------------------------------------------------
# 4. Component biomass functions (Mg/ha dry matter)
# ------------------------------------------------------------

B_stem_max3   <- function(t) f_stem_gomp(t)   * B_AG_max3(t)
B_branch_max3 <- function(t) f_branch_gomp(t) * B_AG_max3(t)
B_rest_max3   <- function(t) f_rest_gomp(t)   * B_AG_max3(t)

# ------------------------------------------------------------
# 5. Trajectories for ages 0–12 years
# ------------------------------------------------------------

ages <- seq(0, 13, by = 0.1)

comp_df <- tibble(
  age              = ages,
  SBD_cm           = SBD_max3_gomp(ages),
  B_AG_Mg_ha       = B_AG_max3(ages),
  f_stem           = f_stem_gomp(ages),
  f_branch         = f_branch_gomp(ages),
  f_rest           = f_rest_gomp(ages),
  B_stem_Mg_ha     = B_stem_max3(ages),
  B_branch_Mg_ha   = B_branch_max3(ages),
  B_rest_Mg_ha     = B_rest_max3(ages)
)

comp_wide <- comp_df |>
  select(age, B_stem_Mg_ha, B_branch_Mg_ha, B_rest_Mg_ha)

# ------------------------------------------------------------
# 6. Empirical reference values (Huber & Jha, approximate)
# ------------------------------------------------------------

# Huber 2017/2018: Max 3, SRAFS, 4-year stand biomass ~ 35.8 Mg/ha
B_huber4_total <- 35.8  # Mg/ha

# Jha Table S1: selected rows (values taken from Jha2018_MOESM1_ESM Table S1)
# Age 4, D-121, agroforestry/forestry, 666 ha^-1, AGB 82 Mg/ha
# 55.1% trunk, 16.9% branch, 10.0% leaf --> does no sum up to 1
# Age 8, D-121, agroforestry/forestry, 400 ha^-1, AGB 134.3 Mg/ha
# 74.7% trunk, 16.6% branch, 8.6% leaf --> Lodhiyal and Lodhiyal (1997)
# Age 9, G-3, agroforestry, 500 ha^-1, AGB 74 Mg/ha
# 83% trunk, 11% branch, 6% leaf --> Das and Chaturvedi (2005)
# Age 9, G-48, agroforestry (wheat), 500 ha^-1, AGB 51.46 Mg/ha
# 92% trunk, 4% branch, 4% leaf --> Yadava (2010)
# Age 12, hybrid poplar SRC (Canada), 1111 ha^-1, AGB range 110–162 Mg/ha --> Zabek and Prescott (2006)
# Age 10, Populus deltoides, 500 ha^-1, AGB 105.4 Mg/ha --> 74.4% trunk, 13% branch, 12.6% leaf --> Singh (1989)
# Age 6 hybrid poplar SRC (Canada), 1937 ha^-1, AGB 218.5/1937*500 Mg/ha --> 61.9% trunk, 27.6% branch, 10.3% leaf --> Fortier et al. (2010a)
# Age 13 I-214 poplar SRC (France), 135 ha^-1, AGB 152 Mg/ha --> 77% trunk, 22% branch, 1% leaf --> Jha (2018)


jha_points <- tribble(
  ~age, ~system,                  ~AGB,   ~p_trunk, ~p_branch, ~p_leaf,
   #4,   "D-121 AF/forestry",      82.0,   0.551,    0.169+.551,     0.100+ 0.169+.551,
   8,   "D-121 AF/forestry",      134.3,  0.747,    0.166+.747,     0.086+ 0.166+.747,
   9,   "G-3 AF",                 74.0,   0.83,     0.11+.83,      0.06+0.11+.83,
   9,   "G-48 AF wheat",          51.46,  0.92,     0.04+.92,      0.04 + 0.04+.92,
  10 ,  "Poplus delt.",          105.4,  0.744,     0.744+.13,      0.744 + 0.13+.126,
  6 , "hybrid, canada",          218.5/1937*500,  0.619,     0.619+.276,      0.619 + 0.276+.103,
  13, "I-214, france",  152.0,  0.77,     0.77+.22,      0.77 + 0.22+.01
)

jha_components <- jha_points |>
  mutate(
    stem_Mg_ha   = AGB * p_trunk,
    branch_Mg_ha = AGB * p_branch,
    rest_Mg_ha   = AGB * p_leaf
  ) |>
  select(age, system, stem_Mg_ha, branch_Mg_ha, rest_Mg_ha)

# 12-year total biomass band from Zabek & Prescott (2006), 110–162 Mg/ha
band12 <- tibble(
  age   = 12,
  ymin  = 110/1111*500,
  ymax  = 162/1111*500
)

comp_long_biomass <- comp_df |>
  select(age, B_stem_Mg_ha, B_branch_Mg_ha, B_rest_Mg_ha) |>
  tidyr::pivot_longer(-age, names_to = "component", values_to = "biomass") |>
  mutate(component = dplyr::recode(
           component,
           B_stem_Mg_ha   = "Stem (model)",
           B_branch_Mg_ha = "Branches (model)",
           B_rest_Mg_ha   = "Rest (model)"
         ),
         component = factor(component,
                            levels = c("Rest (model)",
                                       "Branches (model)",
                                       "Stem (model)")))

jha_long <- jha_components |>
  tidyr::pivot_longer(cols = c(stem_Mg_ha, branch_Mg_ha, rest_Mg_ha),
                      names_to = "comp",
                      values_to = "biomass") |>
  mutate(category = dplyr::recode(
    comp,
    stem_Mg_ha   = "Stem (emp.)",
    branch_Mg_ha = "Branches (emp.)",
    rest_Mg_ha   = "Rest (emp.)"
  ))

# gemeinsame Kategorien (optional, nur zur Reihenfolge)
all_categories <- c("Stem (model)", "Branches (model)", "Rest (model)",
                    "Stem (emp.)",  "Branches (emp.)",  "Rest (emp.)")

fill_vals <- c(
  "Stem (model)"     = "#1b9e77",
  "Branches (model)" = "#d95f02",
  "Rest (model)"     = "#7570b3",
  "Stem (emp.) "      = NA,
  "Branches (emp.) "  = NA,
  "Rest (emp.) "      = NA          
)

colour_vals <- c(
  "Stem (model)"     = NA,
  "Branches (model)" = NA,
  "Rest (model)"     = NA,
  "Stem (emp.)"      = "darkgreen",
  "Branches (emp.)"  = "orange",
  "Rest (emp.)"      = "blue"
)

shape_vals <- c(
  "Stem (model)"     = NA,
  "Branches (model)" = NA,
  "Rest (model)"     = NA,
  "Stem (emp.)"      = 21,
  "Branches (emp.)"  = 24,
  "Rest (emp.)"      = 23
)

ggplot(comp_long_biomass,
       aes(x = age, y = biomass)) +
  geom_area(aes(fill = component),
            position = "stack", alpha = 0.8, colour = NA) +
  #geom_point(aes(x = 4, y = B_huber4_total),
  #           inherit.aes = FALSE,
  #           colour = "black", shape = 4, size = 3,
  #           show.legend = FALSE) +
  geom_point(
    data = jha_long,
    aes(x = age, y = biomass,
        colour = category, fill = category, shape = category),
    inherit.aes = FALSE,
    size = 3
  ) +
  scale_fill_manual(
    name   = "",
    values = fill_vals,
    breaks = all_categories
  ) +
  scale_colour_manual(
    name   = "",
    values = colour_vals,
    breaks = all_categories
  ) +
  scale_shape_manual(
    name   = "",
    values = shape_vals,
    breaks = all_categories
  ) +
  labs(
    x = "Age (years)",
    y = "Component biomass (Mg/ha dry matter)"
  ) +
  theme_minimal()
```


## Cascading use of biomass and product quality

From a cleaner production perspective, cascading use of biomass—prioritising high‑value material uses before energy recovery—is widely advocated to maximise value creation and reduce environmental burdens @Lamers2015_Feedstock_Design; @Espinoza2017_Biorefinery_SC_review. In wood‑based value chains, this typically implies allocating higher‑quality stem fractions to products such as sawn timber, veneer, panels, or pulp, while directing lower‑quality fractions and residues to energy applications @DeMeyer2015_OPTIMASS. Multi‑product MILP models such as OPTIMASS implement this concept by defining a hierarchy of product quality classes and allowing higher‑quality products to satisfy lower‑quality demands via substitution constraints along the cascade @DeMeyer2015_OPTIMASS; @DeMeyer2016_tOPTIMASS.

For agroforestry biomass such as poplar and willow, this logic translates into a set of product streams (e.g. larger‑diameter logs for material uses, medium‑diameter fractions for pulp and panels, small‑diameter wood and residues for energy), whose relative shares depend on stand age and growth conditions @Huber2018_SRAFS_growth; @Trnka2008_Poplar_SRF. In the present work, the age‑class yield construction described above is extended by associating each age class with product‑specific yield fractions, thereby capturing age‑dependent quality in a simplified but optimisation‑compatible form. The MILP formulation then enforces a product hierarchy in which higher‑quality products can satisfy lower‑quality demands, subject to the available yields and infrastructure capacities, thus embedding a simplified cascading logic into the regional supply chain design @DeMeyer2015_OPTIMASS.

## Synthesis and modelling gap

The literature summarised above provides three critical ingredients for modelling agroforestry biomass supply chains. First, empirical SRAFS and SRC studies offer species‑ and site‑specific information on growth, MAI, stand structure, and economic performance, including allometric models that link diameter distributions to biomass @Huber2016_Allometric_SRAFS; @Huber2018_SRAFS_growth; @Trnka2008_Poplar_SRF; @Testa2014_Poplar_Economics. Second, age‑dependent stand biomass functions such as Chapman–Richards provide a convenient parametric representation of growth over rotation, from which discrete age‑class yields can be derived @Niemczyk2021_Poplar_Rotation. Third, biomass supply chain MILP models and multi‑product cascading formulations offer a rich toolkit for optimising infrastructure, logistics, and product allocation in regional bioeconomy systems @Zahraee2020_BSC_review; @DeMeyer2015_OPTIMASS; @DeMeyer2016_tOPTIMASS; @Espinoza2017_Biorefinery_SC_review.

At the same time, existing models fall short of explicitly coupling agroforestry stand dynamics with supply chain design in several respects. Most biomass supply chain formulations assume exogenous biomass availability profiles and do not represent stand age or rotation decisions, while forestry‑oriented growth and yield models are rarely embedded in regional optimisation models that include multiple products and cascading uses @DeMeyer2016_tOPTIMASS; @Lamers2015_Feedstock_Design. In particular, there is a lack of models that (i) link age‑dependent stand biomass functions to age‑class yield tables for different product types, (ii) enforce multi‑cycle age‑lag constraints that ensure consistency of harvesting decisions at each agroforestry site across successive rotations, and (iii) allocate biomass along a product quality cascade within a spatially explicit, multi‑period supply chain optimisation framework @DeMeyer2015_OPTIMASS; @DeMeyer2016_tOPTIMASS; @Eksioglu2009_Biorefinery_SC.

The MILP formulation developed in this paper addresses this gap by integrating age‑dependent biomass yield modelling, stand‑level age‑lag constraints, and a product‑cascading hierarchy into a single optimisation model for the design and operation of agroforestry biomass supply chains.


# Biomass supply chain modelling 

Biomass supply chain design and planning has been widely studied using MILP models at strategic, tactical, and operational levels [@Zahraee2020_BSC_review; @Sharma2013_BSC_review]. Comprehensive reviews classify models according to decision scope (location, capacity, routing, inventory), time horizon, uncertainty treatment, and sustainability metrics [@Zahraee2020_BSC_review; @Espinoza2017_Biorefinery_SC_review]. Many models focus on bioenergy supply chains for electricity, heat, or biofuels, considering feedstocks such as agricultural residues, forestry residues, dedicated energy crops, and municipal wastes [@Sharma2013_BSC_review]. Environmental aspects, including greenhouse gas emissions and land use, are increasingly integrated, often using life-cycle assessment (LCA) data or emission factors [@Espinoza2017_Biorefinery_SC_review].

Among generic models, OPTIMASS is a notable example of a multi-product MILP that optimises biomass supply chains with changing biomass characteristics and by-product re-injection [@DeMeyer2015_OPTIMASS]. It models the flow of multiple biomass types and products through supply chain stages, capturing conversion yields, storage, and cascading uses. The t-OPTIMASS extension further incorporates biomass growth and regeneration over multiple periods, enabling the representation of time-dependent availability for biomass supply chains [@DeMeyer2016_tOPTIMASS]. However, these formulations generally assume exogenous biomass availability profiles per period, and do not explicitly model stand-level age dynamics or rotation constraints.

Several scenario-based approaches have been developed for forest biorefinery supply chains, linking product portfolios, process choices, and supply chain design under market volatility [@Mansoornejad2013_Forest_Biorefinery_SC]. @Lamers2015_Feedstock_Design discuss strategic feedstock supply system design for biorefineries, emphasising the importance of economies of scale, conversion yields, and logistics in achieving cost-efficient and robust systems. Reviews highlight that only a limited subset of models explicitly consider biomass growth, spatially explicit land-use change, or agroforestry systems [@Zahraee2020_BSC_review; @Sharma2013_BSC_review]. There is therefore a clear opportunity to develop models that couple agroforestry stand dynamics with network design and cleaner production objectives.

## Cascading use of biomass

From a cleaner production perspective, the cascading use of biomass—prioritising high-value material uses before energy recovery—is widely advocated to maximise resource efficiency and mitigate environmental impacts [@Lamers2015_Feedstock_Design]. In wood-based value chains, this implies that high-quality timber or chemical feedstock fractions should first be allocated to material products such as engineered wood products, pulp and paper, or biochemicals, while lower-quality fractions and residues are used for energy [@DeMeyer2015_OPTIMASS; @Lamers2015_Feedstock_Design]. MILP models can represent such cascades by defining product quality classes and allowing substitution of higher-quality products for lower-quality demands, subject to hierarchy constraints.

The OPTIMASS framework explicitly models such product cascades by defining a product quality order and allowing flows from higher-quality products to lower-quality demand nodes [@DeMeyer2015_OPTIMASS]. Similar ideas have been applied in forest-based biorefinery supply chains, where product portfolios include platform chemicals, fuels, and electricity [@Mansoornejad2013_Forest_Biorefinery_SC]. For agroforestry biomass, poplar and willow can supply different product streams, including sawlogs, peeled veneers, pulpwood, and energy chips, with quality primarily defined by diameter, straightness, and defect rates [@Testa2014_Poplar_Economics; @Spinelli2009_SRC_Harvesting]. Economic analyses in Italy and Central Europe suggest that the economic viability of SRC systems depends on access to higher-value markets such as panel products and pulp, in addition to energy markets [@Testa2014_Poplar_Economics].

Cleaner production strategies that integrate agroforestry biomass into regional bioeconomies therefore require models that can allocate biomass flows along such cascades while respecting stand growth dynamics, rotation constraints, and logistical capacities. Explicit representation of age-dependent yield quality (e.g., diameter distributions, wood density) remains challenging in large-scale MILP models, but can be approximated through age- and species-specific product yield fractions.

## Stand-level yield functions and age classes

To integrate allometric growth into a MILP formulation, we abstract from individual trees and construct age-class yield functions for each biomass type \(p\) and age \(a\). Following common practice in forest and SRC modelling, we define discrete annual age classes and compute the stand biomass yield \(Y_{pa}\) (t ha\(^{-1}\)) associated with harvesting at age \(a\) by combining allometric functions with empirically observed diameter distributions and stand densities [@Huber2018_SRAFS_growth; @Trnka2008_Poplar_SRF]. Where detailed stand structure data are available, this can be done by simulating SBD distributions over time (e.g., via mixed-effects models) and applying tree-level allometrics to each diameter class [@Huber2018_SRAFS_growth]. In the absence of such detailed data, age-dependent MAI and CAI curves reported in the literature can be used to approximate \(Y_{pa}\) [@Trnka2008_Poplar_SRF; @Huber2018_SRAFS_growth].

In the context of the agroforestry supply chain design (SCD) model, we assume that each agroforestry site \(i\) is characterised by a species or species mix, a planting density, and management regime, which jointly determine a site-specific yield profile \(\eta_{pa}\) (t ha\(^{-1}\)) for each product type \(p\) at age \(a\). These yield profiles can be interpreted as the maximum harvestable biomass per hectare for product \(p\) if the stand is harvested at age \(a\), conditional on the stand having been established at an earlier time. In the current MILP formulation, these profiles are captured through the parameter \(\eta_{pa}\), termed base yield, and are linked to decision variables via the constraint
\[
Y_{ipt} = \sum_{s=1}^{t-1} \eta_{p(t-s)} \cdot \text{AREA}_i \cdot z_{ist} \quad \forall i,p,t \in \mathcal{T}^{\text{harv}}.
\]
Here, \(Y_{ipt}\) is the maximum harvest quantity of product \(p\) at site \(i\) in period \(t\), \(\text{AREA}_i\) is the site area (ha), and \(z_{ist}\) is a binary variable indicating that harvests at site \(i\) in periods \(s\) and \(t\) form consecutive harvests with an age lag \(t-s\) within allowed bounds. The term \(\eta_{p(t-s)}\) thus plays the role of an allometrically derived yield coefficient that depends only on stand age \(a = t-s\), not on calendar time.

## Biomass types and product quality classes

In the model, three product types are distinguished: \(p = 1\) (chemical-grade biomass), \(p = 2\) (pulp-grade biomass), and \(p = 3\) (energy-grade biomass). This tripartite classification reflects a product quality hierarchy, with product 1 having the highest quality and price, and product 3 being the lowest-quality fraction typically used for energy [@DeMeyer2015_OPTIMASS; @Lamers2015_Feedstock_Design]. From an allometric perspective, the differentiation among these products can be linked to diameter classes, stem straightness, and defect rates, as these determine the suitability of biomass fractions for different processing routes [@Testa2014_Poplar_Economics; @Spinelli2009_SRC_Harvesting].

In practice, allometric models provide estimates of total aboveground woody biomass, which can be split into product classes using empirical conversion factors or allocation rules derived from product recovery studies (e.g., share of biomass suitable for veneer vs. pulp vs. chips as a function of diameter and log length) [@Spinelli2009_SRC_Harvesting]. For the purposes of the SCD model, we can represent this by specifying product-specific yield coefficients \(\eta_{pa}\) that sum to the total biomass yield at age \(a\) and reflect the expected proportion of each product class under a given management regime.

The cascading hierarchy in the MILP formulation is implemented through demand satisfaction constraints that allow higher-quality products to satisfy lower-quality demands. Specifically, demand \(D_{kpt}^{\max}\) at consumer \(k\) for product class \(p\) in period \(t\) can be met by flows of product \(p\) and all higher-quality classes \(p' \leq p\) from storage hubs, as captured by
\[
\sum_{j \in \mathcal{J}, p' \in \mathcal{P}: p' \leq p } X_{jkp'pt} \leq D_{kpt}^{\max} \quad \forall k,p,t.
\]
This structure supports cleaner production by favouring material uses when possible and relegating residual fractions to energy uses, depending on revenue parameters \(R_{k,p}\) and system constraints [@DeMeyer2015_OPTIMASS; @Lamers2015_Feedstock_Design].


# MILP model for agroforestry supply chain design {#sec:milp}

## Sets, indices, and time structure

The model considers a set of agroforestry sites \(\mathcal{I}\), storage/processing facilities \(\mathcal{J}\), consumer sites \(\mathcal{K}\), and product types \(\mathcal{P} = \{1,2,3\}\). Time is discretised into annual periods \(t \in \mathcal{T} = \{1,\dots,T_{\max}\}\). To represent stand establishment and harvest timing, extended time sets \(\mathcal{T}^{+} = \{0,\dots,T_{\max}+1\}\) and harvest periods \(\mathcal{T}^{\text{harv}} = \{A^{\min}+1,\dots,T_{\max}\}\) are defined, where \(A^{\min}\) and \(A^{\max}\) are minimum and maximum stand ages (years) at harvest. The set \(\mathcal{S}\) comprises all ordered pairs \((s,t)\) of consecutive harvest periods (including establishment at \(s=0\)) satisfying the age-lag bounds:
\[
\mathcal{S} = \{(s,t) \mid s,t \in \{0,\dots,T_{\max}+1\},\; t-s \in \{A^{\min},\dots,A^{\max}\} \}.
\]

## Decision variables

Binary variables \(z_{ist}\) indicate whether, at site \(i\), periods \(s\) and \(t\) form consecutive harvests (or establishment and first harvest if \(s=0\)), with \((s,t) \in \mathcal{S}\). These variables define a path of consecutive harvest cycles at each site over the planning horizon. Continuous variables \(Y_{ipt} \ge 0\) represent the maximum harvest quantity (t) of product \(p\) at site \(i\) in harvest period \(t\). Flow variables \(X_{ijpt}\) and \(X_{jkpp't}\) represent transport of products from sites to storage and from storage to consumers, respectively, and \(S_{jpt}\) represent inventories at storage locations.

## Objective function

The model maximises net present value over the planning horizon, expressed as total revenues minus establishment, opportunity, harvest, transport, and storage costs:
\[
\begin{aligned}
\max Z =\;&
\sum_{k,p,t} R_{k,p} \cdot \sum_{j \in \mathcal{J},\, p' \in \mathcal{P}: p' \leq p} X_{jkp'pt} \\
& - \sum_{i} C^{\text{est}}_{i} \cdot \text{AREA}_i \cdot \sum_{t \in \mathcal{T}} z_{i0t} \\
& - \sum_{i} C^{\text{opp}}_{i} \cdot \text{AREA}_i \cdot \sum_{t \in \mathcal{T}} (T_{\max}-t) \cdot z_{i0t} \\
& - \sum_{i,(s,t)\in \mathcal{S}: s>0} C^{\text{harv}}_{i} \cdot \text{AREA}_i \cdot z_{ist} \\
& - c^{\text{tr-raw}} \sum_{i,j,p,t} d_{ij} \cdot X_{ijpt}
- c^{\text{tr-pre}} \sum_{j,k,p,t} d_{jk} \cdot X_{jkpt}
- \sum_{j,p,t} c^{\text{stor}}_{j} \cdot S_{jpt}.
\end{aligned}
\]

Here, \(R_{k,p}\) are product- and consumer-specific revenues (€/t), \(C^{\text{est}}_{i}\) are establishment costs per hectare, \(C^{\text{opp}}_{i}\) are annual opportunity costs per hectare, and \(C^{\text{harv}}_{i}\) are harvest and refitting costs per hectare [@Testa2014_Poplar_Economics]. Transport costs are proportional to flow and distance with rates \(c^{\text{tr-raw}}\) and \(c^{\text{tr-pre}}\) for raw and pre-treated biomass, respectively, and storage incurs linear holding costs \(c^{\text{stor}}_j\) per tonne.

## Key constraints

The path connectivity constraint ensures that each harvest period at a site has exactly one predecessor and one successor in the age-lag graph, effectively forming a path over \(\mathcal{T}^{+}\):
\[
\sum_{s=0: (s,t)\in \mathcal{S}}^{t} z_{ist} = \sum_{u = t+1: (t,u) \in \mathcal{S}}^{T_{\max}+1} z_{itu}
\quad \forall i \in \mathcal{I}, t \in \mathcal{T}.
\]

Establishment is restricted to at most one time per site:
\[
\sum_{t \in \mathcal{T}^{+}} z_{i0t} \le 1 \quad \forall i \in \mathcal{I}.
\]

Biomass yield is linked to age-lag decisions through the allometric yield coefficients:
\[
Y_{ipt} = \sum_{s=1}^{t-1} \eta_{p(t-s)} \cdot \text{AREA}_i \cdot z_{ist}
\quad \forall i \in \mathcal{I}, p \in \mathcal{P}, t \in \mathcal{T}^{\text{harv}}.
\]

Harvest quantities bound outbound flows from each site:
\[
\sum_{j \in \mathcal{J}} X_{ijpt} \le Y_{ipt}
\quad \forall i,p,t \in \mathcal{T}^{\text{harv}}.
\]

At storage hubs, inventories obey standard balance equations with initial inventory zero:
\[
S_{jpt} = S_{jp,t-1} + \sum_{i} X_{ijpt} - \sum_{k, p' \ge p} X_{jkpp't}
\quad \forall j,p,t \in \mathcal{T}^{\text{harv}}.
\]

Storage and processing capacities are enforced via:
\[
\sum_{p} S_{jpt} \le \text{CAP}^{\text{stor}}_j \quad \forall j,t, \qquad
\sum_{i,p} X_{ijpt} \le \text{CAP}^{\text{proc}}_j \quad \forall j,t.
\]

Demand satisfaction with cascades is modelled as described above, where higher-quality products can satisfy lower-quality demands up to maximum demand levels \(D^{\max}_{kpt}\).


# Case study and computational experiment plan {#sec:experiments}

The proposed MILP model is intended to support the design of agroforestry biomass supply chains for cleaner production in specific regional contexts. In a forthcoming case study, we focus on poplar-based alley-cropping systems in a temperate European setting, informed by empirical data from SRAFS trials in Southern Germany and SRC plantations in Central Europe [@Huber2018_SRAFS_growth; @Trnka2008_Poplar_SRF; @Testa2014_Poplar_Economics].

We envisage the following steps and experimental scenarios for computational analysis:

1. **Parameterisation of allometric yield profiles:** Use species- and site-specific allometric models and observed growth trajectories to construct age-dependent yield coefficients \(\eta_{pa}\) for chemical, pulp, and energy product classes at representative agroforestry sites. Where necessary, scenario ranges will be defined to capture uncertainty in MAI and CAI due to climate variability and management differences [@Huber2016_Allometric_SRAFS; @Huber2018_SRAFS_growth; @Trnka2008_Poplar_SRF].

2. **Baseline optimisation:** Solve the MILP model for a baseline configuration with fixed prices, capacities, and policy settings, obtaining optimal establishment timing, harvest schedules, and supply chain structure (locations, flows, inventories).

3. **Rotation length scenarios:** Vary the age-lag bounds \(A^{\min}\) and \(A^{\max}\) to represent alternative rotation strategies (e.g., 3–5 years vs. 8–10 years) and analyse their impacts on profitability, product mix, and land use. This will shed light on trade-offs between shorter rotations favouring energy chips and longer rotations enabling higher shares of material products [@Trnka2008_Poplar_SRF; @Testa2014_Poplar_Economics].

4. **Product price and cascading scenarios:** Explore scenarios with different relative prices for chemical, pulp, and energy products, reflecting changes in market demand and policy incentives (e.g., bioenergy subsidies, green chemicals markets). Assess how the product cascade hierarchy and yield quality structure influence the marginal value of species and rotation decisions [@DeMeyer2015_OPTIMASS; @Lamers2015_Feedstock_Design].

5. **Infrastructure and logistics scenarios:** Evaluate the effect of alternative storage and processing network configurations (e.g., decentralised vs. centralised hubs, different capacity levels, transport cost parameters) on the optimal spatial deployment of agroforestry sites and biomass flows [@DeMeyer2015_OPTIMASS; @DeMeyer2016_tOPTIMASS; @Lamers2015_Feedstock_Design].

6. **Sensitivity to allometric and growth uncertainty:** Conduct parametric sensitivity and, where computationally feasible, stochastic analyses to quantify the impact of uncertainty in allometric parameters and growth responses (e.g., under drought or management differences) on optimal decisions and system robustness [@Huber2018_SRAFS_growth; @Trnka2008_Poplar_SRF].

7. **Cleaner production indicators:** For selected solutions, compute derived indicators such as biomass cascading index (share of biomass used for material products before energy), utilisation rates of higher-quality fractions, and, where coupled with LCA data, greenhouse gas emission savings relative to fossil benchmarks [@Espinoza2017_Biorefinery_SC_review; @Zahraee2020_BSC_review].


# Conclusions

This paper outlines a research framework that couples allometric biomass growth models with a MILP-based agroforestry biomass supply chain design model featuring age-lag constraints and product cascading. Building on empirical evidence from temperate SRAFS and SRC systems and established biomass supply chain models, the proposed approach aims to capture key interactions between stand-level growth dynamics and network-level design decisions that are central to the cleaner production potential of agroforestry [@Huber2018_SRAFS_growth; @DeMeyer2015_OPTIMASS; @DeMeyer2016_tOPTIMASS]. Future work will implement and calibrate the model for a regional case study, conduct the computational experiments sketched above, and extend the framework towards multi-objective optimisation including environmental indicators such as greenhouse gas emissions and biodiversity proxies.
