---
title: "Agroforestry SmartAgroforst: Poplar biomass modelling"
author:
  - name: "Thomas Kirschstein"
    affiliation: 
      - "Fraunhofer IKTS, Dresden, Germany"
    email: "thomas.kirschstein@ikts.fraunhofer.de"
    correspondingauthor: true
  
journal: "Journal Name"
# elsarticle options: preprint preview style, 1p or 3p for layout
classoption: "preprint,review,1p,authoryear"
date: "`r Sys.Date()`"
linenumbers: true
numbersections: true
always_allow_html: true


keywords: 
  - Agroforestry 
  - Short-rotation coppice 
  - Biomass supply chain
  - Product cascading

abstract: |
  Short-rotation agroforestry systems are increasingly promoted as a low-carbon feedstock option for bio-based value chains, but their economic viability depends on coordinated decisions on stand establishment, multi-cycle harvesting, and product cascading in spatially distributed supply chains. This paper develops a mixed-integer linear programming (MILP) model for the integrated design and operation of an agroforestry biomass supply chain with explicit age-lag constraints that link consecutive harvests at each site. The formulation couples allometric yield functions for different biomass types with a multi-product, multi-period network design model including storage, processing, and product quality cascades. The model is applied to a stylised case study of Central European poplar-based alley-cropping systems to explore trade-offs between harvest rotation length, product portfolios, and infrastructure configuration. We outline a computational experimentation plan to test the impact of allometric parameter uncertainty, price scenarios, and policy incentives on the optimal deployment of agroforestry systems in cleaner bioeconomy pathways.

output:
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
    keep_tex: true
    latex_engine: xelatex
    number_sections: true
    

header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{float}
  - \usepackage{xcolor}

bibliography: references.bib
csl: elsevier-harvard.csl
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.path = "figures/"
)

library(magick)
library(dplyr)
library(plotly)
library(ggplot2)
library(tidyr)
library(patchwork)

# Chapman-Richards biomass function
B_cr <- function(t, Bmax, k, m) {
  Bmax * (1 - exp(-k * t))^m
}

# Gompertz growth function
gomp <- function(t, A, k, t0) {
  A * exp(-exp(-k * (t - t0)))
}
```

```{r load-model-data, echo=FALSE, message=FALSE, warning=FALSE}
library(sf)
library(leaflet)
library(osrm)
library(ompr)
library(ompr.roi)
library(ROI.plugin.glpk)
library(gurobi)
library(ROI.plugin.gurobi)
library(ggrepel)
library(ggnewscale)
library(rnaturalearth)
library(rnaturalearthdata)

source("../Modell/!helper_func.r")
source("../Modell/!helper_instance_builder_v8a.R")
source("../Modell/build_agroforestry_lp_v10.R")

load("../Modell/afs_workspace.RData")

sites_sf   <- afs_workspace$sites_sf
sites      <- afs_workspace$sites
storages   <- afs_workspace$storages
consumers  <- afs_workspace$consumers

# Yield table from Chapman-Richards (3 products)
yields_by_age <- expand_grid(product = 1:3, age = 1:15) %>%
  mutate(
    B_total = B_cr(age,
                   Bmax = c(80, 120, 160)[product],
                   k    = c(0.18, 0.20, 0.22)[product],
                   m    = c(1.1, 1.2, 1.3)[product]),
    yield_ha = B_total * c(0.20, 0.35, 0.45)[product]
  )
```

# Introduction

Agroforestry systems that integrate trees with agricultural crops (silvoarable systems) are increasingly recognised as a promising land-use option to reconcile climate mitigation, biodiversity conservation, and rural income generation on the same hectare of land [@Toensmeier2017_Carbon_Agroforestry]. Temperate agroforestry systems (AFS) based on fast-growing species such as poplar, willow, alder, and black locust can deliver substantial biomass yields while improving soil quality and providing multiple ecosystem services compared to conventional monocultures [@Huber2018_SRAFS_growth]. Biomasses produced in AFS is considered as a sustainable feedstock and can be used either energetically, chemically or materially (e.g. in the building sector). The latter two options are generally preferred over the energetic usage from a cleaner production perspective as biogenic carbon is stored in products during their life span which contributes to climate change mitigation.

Despite substantial research on ecological advantages and pioneering initiatives to establish AFS in practice, the share of AFS in agricultural production systems is still small. E.g. in Germany about 1200 hectares of silvoarable AFS are established in 2025, which accounts for about 0.02\% of total agricultural land use [DEFAF]. One of the core challenges in establishing AFS systems at large scale in practice is a lack of opportunities to generate substantial revenues from woody biomasses of AFS. On the other side, the rise of bioeconomy creates substantial demands for woody biomass and requires a reliable and cost-efficient feedstock supply. As woody biomasses are bulky feedstocks with low energy density, the design of cost-efficient supply chains is crucial to ensure the economic viability of AFS-based value chains. This requires coordinated decisions on stand establishment, multi-cycle harvesting, and product cascading in spatially distributed supply chains, which can be supported by integrated optimisation models that link stand-level growth dynamics to regional supply chain design.

In the last two decades, a large body of research has analysed biomass supply chains for bioenergy and biorefineries using mathematical programming models, often in the form of mixed-integer linear programming (MILP) [@Zahraee2020_BSC_review; @Sharma2013_BSC_review]. These models typically optimise network configuration, feedstock sourcing, and logistics under cost or profit objectives. They often treat biomass availability as an exogenous input or rely on simplified agricultural production decisions. Moreover, downstream processing of biomasses is concentrated on one particular value chain e.g. bioenergy plants consuming all produced biomasses. Recent work has started to integrate operational details such as various biomass types, biomass growth models and regeneration processes in multi-period supply chain models.

The present paper contributes to this literature in four ways. First, we propose a novel MILP formulation for AFS-based biomass supply chain design that explicitly links establishment decisions and consecutive harvest decisions, ensuring that harvesting decisions at each site are consistent with minimum and maximum rotation ages. Second, we embed allometric yield functions for different biomass types generating age-specific yields for different woody biomass types (like stems, branches, or leaves and twigs). Third, these different woody biomass types can be processed in specific downstream industries, like chemical or paper industry as well as energy sector. The model aims to provide strategic decision support for establishing regional AFS-based bioeconomy ecosystems connecting multiple industries and product markets.

The remainder of the paper is structured as follows. Section \@ref(sec:litreview) reviews the relevant literature on agroforestry biomass systems, allometric biomass modelling, and biomass supply chain optimisation. Section \@ref(sec:milp) summarises the MILP formulation for the agroforestry supply chain with age-lag constraints. Section \@ref(sec:experiments) outlines the planned computational experiments for the case study and discusses expected insights for cleaner production strategies.

```{r fig-supply-chain-network, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="AFS supply chain network in Saxony-Anhalt. Grey lines: German federal state boundaries. Blue lines: major rivers. Green polygons: potential KUP sites. Orange symbols: pre-treatment facilities (circle~=~sawmill, triangle~=~logistics hub, size proportional to processing capacity). Coloured squares: consumer sites (violet: P1 chemical/pulp, blue: P2 pulp/paper, red: P3 energy/biogas, size proportional to total demand).", fig.width=6.5, fig.height=5.5, out.width="\\textwidth"}

COL_SITE_F <- "#74c476"; COL_SITE_B <- "#2d6a2d"
COL_SAW    <- "#e65c00"; COL_HUB    <- "#c05000"
COL_P1     <- "#6a0dad"; COL_P2     <- "#08519c"; COL_P3     <- "#a50026"

de_states <- ne_states(country = "Germany", returnclass = "sf") %>%
  st_transform(4326)

rivers_eu <- ne_download(
  scale = 10, type = "rivers_lake_centerlines",
  category = "physical", returnclass = "sf"
) %>%
  st_transform(4326) %>%
  st_crop(xmin = 10.6, xmax = 13.2, ymin = 50.9, ymax = 52.8)

consumers_plt <- consumers %>%
  mutate(
    total_demand = demand_P1 + demand_P2 + demand_P3,
    kategorie = case_when(
      demand_P1 >= demand_P2 & demand_P1 >= demand_P3 & demand_P1 > 0 ~ "Chemical / Pulp (P1)",
      demand_P2 >= demand_P1 & demand_P2 >= demand_P3 & demand_P2 > 0 ~ "Pulp / Paper (P2)",
      demand_P3 > 0 ~ "Energy / Biogas (P3)",
      TRUE          ~ "Other"
    ),
    pt_size = pmin(4.5, pmax(1.5, 1.5 + log1p(total_demand) * 0.55))
  )

storages_plt <- storages %>%
  mutate(
    subtype = if_else(type == "Hub", "Logistics Hub", "Sawmill"),
    pt_size = pmin(3.5, pmax(1.2, 1.2 + log1p(CAP_proc) * 0.38))
  )

stor_sf <- st_as_sf(storages_plt, coords = c("lng", "lat"), crs = 4326)
cons_sf <- st_as_sf(consumers_plt, coords = c("lng", "lat"), crs = 4326)

ggplot() +
  geom_sf(data = de_states,
          fill = "grey96", colour = "grey55", linewidth = 0.35) +
  geom_sf(data = rivers_eu,
          colour = "#4292c6", linewidth = 0.45, alpha = 0.7) +
  geom_sf(data = sites_sf,
          fill = COL_SITE_F, colour = COL_SITE_B,
          linewidth = 0.12, alpha = 0.45) +
  geom_sf(data = stor_sf,
          aes(colour = subtype, shape = subtype, size = pt_size),
          stroke = 1.2, alpha = 0.92, show.legend = TRUE) +
  scale_colour_manual(
    name   = "Pre-treatment",
    values = c("Sawmill" = COL_SAW, "Logistics Hub" = COL_HUB)
  ) +
  scale_shape_manual(
    name   = "Pre-treatment",
    values = c("Sawmill" = 16, "Logistics Hub" = 17)
  ) +
  ggnewscale::new_scale_fill() +
  geom_sf(data  = cons_sf,
          aes(fill = kategorie, size = pt_size),
          shape = 22, colour = "white", stroke = 0.45,
          alpha = 0.88, show.legend = TRUE) +
  scale_fill_manual(
    name   = "Consumer type",
    values = c(
      "Chemical / Pulp (P1)" = COL_P1,
      "Pulp / Paper (P2)"    = COL_P2,
      "Energy / Biogas (P3)" = COL_P3,
      "Other"                = "grey60"
    )
  ) +
  scale_size_continuous(range = c(1.5, 5), guide = "none") +
  guides(
    colour = guide_legend(order = 1, override.aes = list(size = 4, stroke = 1.5)),
    shape  = guide_legend(order = 1, override.aes = list(size = 4, stroke = 1.5)),
    fill   = guide_legend(order = 2, override.aes = list(size = 4, shape = 22, colour = "white")),
    size   = "none"
  ) +
  coord_sf(xlim = c(10.6, 13.2), ylim = c(50.9, 52.8), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude") +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid      = element_line(colour = "grey88", linewidth = 0.25),
    legend.position = "right",
    legend.key.size = unit(0.45, "cm"),
    legend.text     = element_text(size = 7),
    legend.title    = element_text(size = 8, face = "bold"),
    axis.text       = element_text(size = 7)
  )
```

# Literature review {#sec:litreview}

Agroforestry biomass systems, stand-level growth models, and biomass supply chain optimisation have each been studied extensively, but they are rarely combined in an integrated framework that links age-dependent stand dynamics to regional product-cascading value chains [@Zahraee2020_BSC_review; @Sharma2013_BSC_review; @Espinoza2017_Biorefinery_SC_review].
This section reviews (i) agroforestry systems with emphasis on system types, economics, and the cascading use of biomass as the primary revenue driver; (ii) biomass growth models at tree and stand level with explicit attention to the allometric--diameter--stand-biomass pathway; and (iii) biomass supply chain design using mathematical programming.
The section concludes by identifying the modelling gap addressed in this paper.


## Agroforestry systems {#sec:lit-afs}

Agroforestry deliberately integrates woody perennials with crops and/or livestock on the same management unit, with the aim of enhancing productivity, profitability, and environmental performance [@Nair1993_IntroAgroforestry; @Toensmeier2017_Carbon_Agroforestry]. In the temperate zone, three distinct system types can be distinguished based on the spatial arrangement of trees and the rotation length at which woody biomass is harvested.

**Short-rotation coppice (SRC)** is defined as a monoculture plantation of fast-growing woody species covering the entire management unit, with planting densities of 8\,000--20\,000 cuttings~ha\(^{-1}\). After each harvest the root system regenerates without replanting. Rotation cycles are typically 3--7~years and total stand lifetime is 20--25~years [@Trnka2008_Poplar_SRF; @Faasch2012_SRC_Economics]. SRCs are no AFSs in the narrow sense, but it is helpful to benchmark biomass production.

**Short-rotation agroforestry systems (SRAFS)** integrate SRC-type tree strips (typically 1--4 rows of the same fast-growing species ike poplar or willow) into an otherwise arable field in an alley-cropping arrangement. Crop production continues in the inter-row strips [@Huber2018_SRAFS_growth; @Graves2007_Agroforestry_Economics]. Thus, tree strips typically cover 10--30\% of the total parcel area. Rotation lengths and total stand lifetime in SRAFS are comparable to SRCs. In SRAFS all trees are effectively edge trees, exposed to higher light and water availability than interior SRC trees. This boundary effect raises individual-tree biomass accumulation but also intensifies competition with adjacent crop rows [@Huber2018_SRAFS_growth].

**Long-rotation agroforestry systems (LRAFS)** use the same alley-cropping spatial structure as SRAFS but with rotation cycles of 12--15~years or longer [@Graves2007_Agroforestry_Economics; @Nair1993_IntroAgroforestry].
At this rotation length, individual trees develop substantial stem diameters and the proportion of high-value  timber increases markedly, enabling premium material uses (veneer, sawlogs, engineered wood products) that are inaccessible in short-rotation systems. Therefore, often timber-quality tree species (such as walnut, cherry, oak, or high-grade poplar clones) are planted. 

SRAFS and LRAFS are not clearly differentiated in practice. First, a clear cut-off age is hard to define and, mor importantly, the intended use of AFS biomass is not always clear and can also change over time. Nonetheless, when establishing AFS, the structure of AFS and its intended use after harvest drives planting decisions and, subsequently, economic viability. Thus, in the present paper SRAFS and LRAFS are not considered as separate AFS systems, but the interplay of intended use, establishment, and harvesting decisions is analysed.

### Economic rationale of landowners

The decision of a landowner to establish an AFS involves weighing substantial up-front costs and long-term opportunity costs against a diversified revenue stream. On the **cost side**, establishment requires site preparation, planting material, and planting labour, typically amounting to 1\,500--3\,500 \euro/ha for SRC or SRAFS [@Faasch2012_SRC_Economics; @Testa2014_Poplar_Economics]. Recurrent harvest and refitting costs (machine hire, chipping, transport to roadside) add approximately 200--400 \euro/ha per rotation depending on stand density and equipment availability [@Testa2014_Poplar_Economics; @Spinelli2009_SRC_Harvesting]. During the rotation, maintenance cost of AFS arise on a yearly basis. Next to expenses for establishment, maintenance and harvesting, opportunity cost arise as no revenues from crops are generate on tree strips. In regions with fertile soils and high cereal prices, yearly crop revenues may sum up to 400--600 \euro/ha [@Faasch2012_SRC_Economics; @Graves2007_Agroforestry_Economics]. Thus, opportunity cost are often the dominant cost driver in SRAFS profitability analyses .

On the **benefit side**, the primary revenue of AFS is the sale of harvested woody biomass, whose magnitude depends critically on rotation length, species, and site productivity. For poplar-based SRAFS in Central Europe, typical yields range from 10 to 25 t of dry matter biomass per ha for typical rotation lengths [@Huber2018_SRAFS_growth; @Niemczyk2021_Poplar_Rotation]. Additionally, positive crop-yield effects in the inter-row strips of SRAFS and LRAFS may result. Tree rows act as windbreaks, reducing soil erosion and evapotranspiration; they provide partial shading that moderates heat stress during drought years; and their root systems may improve water retention in the topsoil [@Graves2007_Agroforestry_Economics; @Grunewald2007_Robinia_AFS]. Empirical measurements in Central European alley-cropping trials indicate that these microclimate effects can increase cereal yields in adjacent strips by 5--15\% relative to open-field controls, partially or fully compensating the loss of arable area occupied by the tree strips [@Graves2007_Agroforestry_Economics]. Beyond direct yield effects, AFS deliver broader ecosystem services  (like soil organic matter accumulation, carbon sequestration, biodiversity enhancement) that may lead to  agri-environment subsidies and carbon credits [@Toensmeier2017_Carbon_Agroforestry; @Faasch2012_SRC_Economics].

Nonetheless, economic attractiveness of AFS depends primarily on revenues of AFS biomass. Thereby, revenues are different for the types of biomass and in which value chain it is used. Often AFS biomasses are used energetically, i.e. for heat and power generation, or anaerobic digestion, which typically creates low prices per tonne of biomass. Alternatvely, woody parts of the biomass (i.e., the stem and large branches) can also be used as pulp and paper grade material as well as feedstock for chemical industry or as timber achieving higher prices [@Lamers2015_Feedstock_Design; @DeMeyer2015_OPTIMASS]. Thus, fostering AFS establishment by maximizing economic returns from AFS biomass requires a strategic approach to product cascading that allocates different biomass fractions to the highest-value applications, while ensuring that the rotation length is aligned with the growth dynamics of the stand and the market demand for different biomass types [@Lamers2015_Feedstock_Design].

Thus, for making educated decisions on AFS establishment and harvesting, the expected biomass quantities at harvest  for stem, branches, and residues need  to be estimated. Thereby, the shares of the biomass types change with stand age. E.g.~the share of stem weight rises from roughly 55\% at age 4 to over 75\% at age 10 [@Huber2018_SRAFS_growth; @Testa2014_Poplar_Economics]. Therefore, in Section \@ref(sec:lit-growth) growth modelling approaches are reviewed that capture the age-dependent growth dynamics of different biomass types in AFS.

## Biomass growth models {#sec:lit-growth}

```{r tab-notation-growth, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
options(kableExtra.latex.load_packages = FALSE)
options(knitr.table.format = "latex")

notation <- tribble(
  ~Symbol, ~Description,
  # --- GROWTH MODEL PARAMETERS ---
  "$M^{sh}$",     "Dry biomass (kg) per shoot",
  "$M^{sh}(t)$",     "Mean shoot biomass (kg) at stand age $t$",
  "$M^{sl}$",     "aboveground dry biomass per ha (t) ",
  "$M^{sl}(t)$",    "aboveground dry biomass per ha (t)  at stand age $t$",
  "$SBD$",     "Stem base diameter (cm) measured 10\\,cm above ground",
  "$b^0$, $b^1$",     "Allometric coefficients for shoot biomass estimation",
  "$SBD(t)$",     "Mean stem base diameter (cm) at stand age $t$",
  "$a^{\\infty}$",             "Asymptotic stand-level biomass (t\\,dry\\,matter/ha)",
  "$b^{\\infty}$",    "Asymptotic mean shoot biomass ($= d^{\\infty} \\cdot b^0$)",
  "$d^{\\infty}$",             "Asymptotic mean stem base diameter (cm)",
  "$r$",                       "Yearly diameter growth rate ",
  "$b'^1$",     "Effective shoot-biomass growth rate ($= r \\cdot b^1$)",
  "$\\tau$",                   "Age at inflection point of diameter growth curve (yr)",
  "$N(t)$",                    "Stand density (trees/ha) at age $t$",
  "$\\bar{n}_{sh}(t)$",     "Mean number of shoots per tree at age $t$",
  "$h_p(t)$",     "Gompertz helper function for biomass component $p$",
  "$f_p(t)$",     "Normalised share of component $p$ in total AGB at age $t$",
  "$a^p$",                     "Asymptotic biomass of component $p$",
  "$b^p$",                     "Growth rate of biomass component $p$",
  "$\\tau_{p}$", "Inflection age of component $p$ (yr)"
)

tbl <- kable(notation,
      format    = "latex",
      booktabs  = TRUE,
      escape    = FALSE,
      col.names = c("Symbol", "Description"),
      caption   = "Notation for the biomass growth models.",
      label     = "tab-notation") |>
  kable_styling(
    font_size     = 8,
    latex_options = c("hold_position", "scale_down"),
    full_width    = FALSE
  ) |>
  column_spec(1, width = "3cm") |>
  column_spec(2, width = "11.0cm")


tbl
```

Total biomass on tree level is commonly modeled using allometric functions that relate tree characteristics like stem diameter or height to total biomass weight. There are numerous allometric models in the literature, often specific to species, site conditions, and stand age. For fast-growing species used in SRAFS (like poplar and willow), allometric functions of the power-law form are used [Reference]. I.e., dry biomass \(M^{sh}\) (kg) of a shoot is approximated by
\begin{equation}
  M^{sh} = b^0 \cdot SBD^{b^1}
  \label{eq:allometry}
\end{equation}
where $SBD$ is the shoot's stem base diameter (in cm) measured 10 cm above ground. Parameters \(b^0\) and  \(b^1\) are species-specific [@Huber2016_Allometric_SRAFS]. 
In the following it is assumed that each tree is pruned to one dominant shoot. For SRAFS species in Southern Germany, @Huber2018_SRAFS_growth reports that most trees produce not more than 2 shoots and @Huber2016_Allometric_SRAFS report a common exponent \(b_1 = 2.603\) for black alder, black locust, poplar (Max~3), and willow (Inger), while \(b^0\) ranges from 0.025 to 0.041, reflecting differences in wood density and branching structure. 

Allometric rules like \eqref{eq:allometry} generate estimates of biomass quantities for a given shoot, i.e. they  can be used to support operational harvest decisions. For planning on a strategic level, it is necessary to estimate biomass development over time, i.e. an SBD growth model. The mean stem base diameter \(SBD(t)\) of coppice shoots increases rapidly in the first growing seasons as shoots exploit stored root carbohydrates, then decelerates as intra-stand competition for light and water intensifies.
Such trajectories can be described by sigmoidal growth functions. Empirical evidence from poplar SRC and SRAFS trials indicates that the SBD growth curve exhibits maximal growth rate (i.e., inflection point) at a relatively early stand age (2--4 years) [@BarrioAnta2008_Poplar_Growth; @Niemczyk2021_Poplar_Rotation].Therefore, a Gompertz model for \(\mathrm{SBD}(t)\) can be formulated as

\begin{equation}
  SBD(t) = d^{\infty} \cdot e^{-e^{-r \cdot (t - \tau)}},
  \label{eq:sbd-gompertz}
\end{equation}
where \(d^{\infty}\) (cm) is the asymptotic mean diameter, \(r\) controls the rate of yearly diameter increment, and \(\tau\) (in years) is the age at the inflection point of the diameter curve. Given \(SBD(t)\) from \eqref{eq:sbd-gompertz}, the mean shoot biomass growth follows directly from the allometric equation \eqref{eq:allometry}:
\begin{equation}
  M^{sh}(t) = b^0 \cdot \mathrm{SBD}(t)^{b^1} = d^{\infty} \cdot b^0 \cdot e^{-e^{-r \cdot b^1 \cdot (t - \tau)}}.
  \label{eq:shoot-biomass}
\end{equation}
Thereby, $b^\infty = d^{\infty} \cdot b^0$ denotes the asymptotic mean shoot biomass, and the effective growth rate of shoot biomass is $b'^1 = r \cdot b^1$, which is amplified relative to the biomass growth rate \(r\) by the allometric exponent \(b^1\).

Total aboveground biomass quantities on stand level \(M^{sl}(t)\) (in t per ha) is then obtained by aggregating individual shoot biomasses over all trees and multiplying by stand density \(N(t)\) (in trees per ha):
\begin{equation}
  M^{sl}(t)
  = N(t) \cdot \bar{n}^{sh}(t) \cdot M^{sh}(t),
  \label{eq:stand-biomass}
\end{equation}
where \(\bar{n}^{shoot}\) is the mean number of shoots per tree [@Huber2018_SRAFS_growth]. Both \(\bar{n}^{sh}(t)\) and \(N(t)\) may change over time. The number of shoots is often controlled in tree care processes and reduced to 1 or 2 shoots per tree [@Huber2018_SRAFS_growth]. Stand level density $N(t)$ is initially defined when the AFS is established and often declines somewhat over time, but usually saturates quickly [REFERENCE]. For the product \(M^{sl}(t)\) in \eqref{eq:stand-biomass} a right-skewed sigmoidal shape with an inflection occurring at a relatively low fraction of the asymptote can be assumed which can be captured in a Gompertz model as well [@Niemczyk2021_Poplar_Rotation; @Testa2014_Poplar_Economics]. Assuming constant tree density  and an average shoot number of 1, \(M^{sl}(t)\) can be simplified as 
\begin{equation}
  M^{sl}(t) = a^{\infty} \cdot e^{-e^{-b'^1 \cdot (t - \tau)}}.
  \label{eq:total-stand-biomass}
\end{equation}
where $a^\infty = N \cdot \bar{n}^{sh} \cdot d^{\infty} \cdot b^0$ is the asymptotic stand-level biomass. For poplar clone Max~3 under SRAFS conditions in Central Europe, representative parameter values are \(d^{\infty} = 9.5\) cm, \(r = 0.55\), and \(\tau = 2.5\) years, yielding a mean SBD of approximately 4--5 cm at age 3 and 7--9 cm at age 7 which is consistent with the diameter ranges reported by @Huber2018_SRAFS_growth and @BarrioAnta2008_Poplar_Growth. For total aboveground biomass of poplar clone Max 3 under SRAFS management, the Gompertz model \eqref{eq:total-stand-biomass} is parameterised as \(a^{\infty} = 100\) t of dry biomass per ha, \(b'^1 = 0.32\) per year, and \(\tau = 4.0\) years, yielding a growth peak at approximately age 4 and a total aboveground biomass of roughly 65 t of dry biomass per ha at age 10 which is consistent with the empirical reference range of 50--135 t per ha reported for comparable poplar systems in the literature [@Huber2018_SRAFS_growth; @Niemczyk2021_Poplar_Rotation].

Yet, for the purpose of modelling AFS supply chains, the three biomass components merchantable stem, structural branches, and fine residues need to be distinguished. Thus, the total biomass growth model \eqref{eq:total-stand-biomass} needs to be adjusted those types. Thereby, growth characteristics are different each type and, thus, require paramterization. Thereby, the differentiation between the merchantable stem and branch depends on the requirements defined by downstream processes. E.g. for chemical production, the merchantable stem is defined as the portion of the tree with a certain minimum diameter (e.g. 10 cm), while branches and residues are defined as the portion of the tree that cannot be chipped into such large chips and thus require different processing or are used for lower-value applications. The share of these components in total aboveground biomass changes with stand age, which has important implications for product cascading decisions in AFS-based supply chains.

Empirical studies show that the stem fraction rises from roughly 55\% of AGB at age~4 to over 75\% at age~10, while branch and residue shares decline as stands mature [@Huber2018_SRAFS_growth; @Testa2014_Poplar_Economics,@Jha2018_Poplar_AFS_Biomass]. This age-dependence is biologically driven by the increasing dominance of secondary stem growth (radial increment) relative to primary shoot elongation and branching in later stand development stages. To capture these dynamics analytically, each component share \(f_p(t)\) is represented as a normalised Gompertz function:
\begin{align}
  h_p(t) = & a_p \cdot e^{-e^{-b_p \cdot (t - \tau_{p})}}  \label{eq:biomass-partial}\\
  f_p(t) =& \frac{h_p(t)}{\sum_{p'} h_{p'}(t)}.
  \label{eq:component-share}
\end{align}
For Max~3 the stem helper is parameterised with a high asymptote (\(a_\mathrm{stem} = 1.0\)) and a relatively late inflection (\(\tau_{\mathrm{stem}} = 4.0\) years), whereas the branch uses a lower asymptote (\(a_\mathrm{branch} = 0.7\)) and an earlier inflection (\(\tau_{\mathrm{branch}} = 3.0\) years), and the residue helper saturates earliest (\(a_\mathrm{rest} = 0.1\), \(\tau_{\mathrm{rest}} = 2.0\) years). This parameterisation reproduces the observed shift towards stem dominance with age and is consistent with the empirical component fractions compiled across poplar agroforestry and SRC systems spanning ages 6--13 years, as illustrated in Figure \@ref(fig:max3-growth).

```{r max3-growth, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Modelled stand-level biomass components for poplar (Max~3) over stand age, represented as stacked Gompertz curves (shaded areas: stem green, branches orange, residues purple) and compared with empirical aboveground biomass observations from poplar agroforestry and SRC systems. Symbols denote cumulative component totals: stem (filled circles), stem+branches (triangles), total AGB (diamonds). Empirical data compiled from Jha (2018) and cited primary sources. Grey shading: typical rotation-age window (5--10 years).", fig.width=6.5, fig.height=5.0, out.width="\\textwidth"}


# Component Gompertz functions (parameters as defined above)
B_AG_max3     <- function(t) gomp(t, A = 100, k = 0.32, t0 = 4.0)
h_stem_f      <- function(t) gomp(t, A = 1.0,  k = 0.40, t0 = 4.0)
h_branch_f    <- function(t) gomp(t, A = 0.7,  k = 0.25, t0 = 3.0)
h_rest_f      <- function(t) gomp(t, A = 0.1,  k = 0.30, t0 = 2.0)

f_stem_g   <- function(t) { hs=h_stem_f(t); hb=h_branch_f(t); hr=h_rest_f(t); hs/(hs+hb+hr) }
f_branch_g <- function(t) { hs=h_stem_f(t); hb=h_branch_f(t); hr=h_rest_f(t); hb/(hs+hb+hr) }
f_rest_g   <- function(t) { hs=h_stem_f(t); hb=h_branch_f(t); hr=h_rest_f(t); hr/(hs+hb+hr) }

ages <- seq(0, 13, by = 0.1)

comp_long <- tibble(
  age       = ages,
  Stem      = f_stem_g(ages)   * B_AG_max3(ages),
  Branches  = f_branch_g(ages) * B_AG_max3(ages),
  Residues  = f_rest_g(ages)   * B_AG_max3(ages)
) |>
  pivot_longer(-age, names_to = "component", values_to = "biomass") |>
  mutate(component = factor(component, levels = c("Residues", "Branches", "Stem")))

# Empirical reference points (Jha 2018, Table S1 and primary sources)
jha_points <- tribble(
  ~age, ~AGB,            ~p_trunk, ~p_branch,
   8,   134.3,           0.747,    0.166,
   9,    74.0,           0.83,     0.11,
   9,    51.46,          0.92,     0.04,
  10,   105.4,           0.744,    0.13,
   6,   218.5/1937*500,  0.619,    0.276,
  13,   152.0,           0.77,     0.22
) |>
  mutate(
    stem_cum   = AGB * p_trunk,
    branch_cum = AGB * (p_trunk + p_branch),
    total      = AGB
  )

# Reshape to long format → enables unified ggplot2 legend
jha_long <- bind_rows(
  jha_points |> transmute(age, biomass = stem_cum,   emp_series = "Empirical: stem"),
  jha_points |> transmute(age, biomass = branch_cum, emp_series = "Empirical: stem+branches"),
  jha_points |> transmute(age, biomass = total,      emp_series = "Empirical: total AGB")
) |>
  mutate(emp_series = factor(emp_series,
    levels = c("Empirical: stem", "Empirical: stem+branches", "Empirical: total AGB")))

emp_colours <- c(
  "Empirical: stem"          = "darkgreen",
  "Empirical: stem+branches" = "darkorange",
  "Empirical: total AGB"     = "navy"
)
emp_shapes <- c(
  "Empirical: stem"          = 21L,
  "Empirical: stem+branches" = 24L,
  "Empirical: total AGB"     = 23L
)

ggplot(comp_long, aes(x = age, y = biomass)) +
  annotate("rect", xmin = 5, xmax = 10, ymin = 0, ymax = Inf,
           fill = "grey80", alpha = 0.3) +
  geom_area(aes(fill = component), position = "stack",
            alpha = 0.75, colour = NA) +
  geom_point(data = jha_long,
             aes(x = age, y = biomass,
                 colour = emp_series, shape = emp_series, fill = emp_series),
             size = 2.5, alpha = 0.85, inherit.aes = FALSE) +
  scale_fill_manual(
    name   = "Component (model)",
    values = c("Stem" = "#1b9e77", "Branches" = "#d95f02", "Residues" = "#7570b3",
               emp_colours),
    breaks = c("Stem", "Branches", "Residues")
  ) +
  scale_colour_manual(
    name   = "Empirical data\n(Jha 2018)",
    values = emp_colours
  ) +
  scale_shape_manual(
    name   = "Empirical data\n(Jha 2018)",
    values = emp_shapes
  ) +
  guides(
    fill   = guide_legend(order = 1,
                          override.aes = list(shape = NA, colour = NA)),
    colour = guide_legend(order = 2,
                          override.aes = list(
                            shape = emp_shapes[levels(jha_long$emp_series)],
                            size  = 2.5,
                            fill  = emp_colours[levels(jha_long$emp_series)]
                          )),
    shape  = "none"
  ) +
  scale_x_continuous(breaks = 0:13) +
  labs(
    x = "Stand age (years)",
    y = expression("Biomass (t dry biomass per ha)")
  ) +
  theme_minimal(base_size = 9) +
  theme(
    legend.position  = "right",
    legend.key.size  = unit(0.4, "cm"),
    legend.text      = element_text(size = 7),
    legend.title     = element_text(size = 8, face = "bold"),
    panel.grid.minor = element_blank()
  )
```


## Biomass supply chain design {#sec:lit-sc}

Biomass supply chain design problems have been extensively studied in literature [@Zahraee2020_BSC_review; @Sharma2013_BSC_review; @Espinoza2017_Biorefinery_SC_review]. Most models jointly optimise facility locations and capacities, biomass sourcing patterns, multimodal transport flows, and inventory policies under cost or profit objectives, sometimes augmented by greenhouse gas emission indicators [@Eksioglu2009_Biorefinery_SC; @Arabi2019_Microalgae_Biobutanol_SC].
A key methodological feature of the more advanced formulations is the integration of long-term network design with multi-period logistics, enabling simultaneous optimisation of infrastructure and operational decisions across seasonal or annual time steps [@Eksioglu2009_Biorefinery_SC; @Zahraee2020_BSC_review].

Among generic multi-product frameworks, OPTIMASS explicitly models product quality cascades by defining a hierarchy of biomass classes and allowing higher-quality products to satisfy lower-quality demands, while the t-OPTIMASS extension incorporates biomass growth and regeneration over multiple periods [@DeMeyer2015_OPTIMASS; @DeMeyer2016_tOPTIMASS].
However, in virtually all existing formulations biomass availability is treated as an exogenous time-series input: stand age structure, rotation decisions, and coppice cycles are not represented endogenously, so that trade-offs between rotation length, stand management, and supply chain configuration cannot be explored in a unified model [@DeMeyer2016_tOPTIMASS; @Lamers2015_Feedstock_Design].
This gap is particularly pronounced for agroforestry systems with repeated coppice cycles, where establishment timing, consecutive harvest decisions, and age-dependent biomass quality jointly determine both the feasible supply profile and the achievable product cascade.

The MILP formulation developed in this paper addresses this gap by endogenising stand age dynamics through binary age-lag variables, coupling the component-specific yield coefficients \(\eta_{pa}\) derived from the Gompertz growth models above with supply chain flow variables, and embedding a product quality cascade that allocates stem, branch, and residue fractions to chemical, pulp, and energy markets within a single optimisation model.

## Biomass supply chain design 

The design and planning of biomass supply chains has attracted substantial research attention over the past two decades, driven by the need to ensure cost-efficient and sustainable feedstock provision for bioenergy and
biorefinery applications. Several systematic reviews provide comprehensive overviews of this literature.
@Cambero2014_ForestBSC_Review survey economic, social, and environmental assessment and optimisation studies for forest biomass supply chains, identifying MILP as the dominant modelling approach and noting a growing
trend towards multi-objective formulations integrating life cycle assessment.

@Yue2014_BSC_Optimization_Overview review biomass-to-bioenergy and biofuel supply chain models with a focus on multi-scale modelling, uncertainty treatment, and sustainability metrics, and conclude that the coupling of
upstream feedstock dynamics with downstream network design remains an underexplored area.
@Malladi2018_BSC_Logistics survey logistics-specific features of biomass supply chain models—seasonal availability, moisture content degradation, preprocessing steps—and highlight that most models treat biomass as a
homogeneous commodity, neglecting quality differentiation. @Zahraee2020_BSC_review provide a recent meta-review of 140+ studies, confirming that cost minimisation is still the dominant objective, and that
multi-product cascading formulations are rare.

The structure of biomass supply chain models typically encompasses three decision layers: (i) strategic network design (facility location, capacity sizing, technology selection), (ii) tactical planning (harvest scheduling,
seasonal flows, inventory), and (iii) operational routing. The following paragraphs discuss key individual contributions and their specific methodological advances.

**Strategic multi-product network design.**
@Eksioglu2009_Biorefinery_SC develop one of the earliest large-scale MILP models for biomass-to-biorefinery supply chain design, integrating facility location, capacity sizing, and multimodal transport for lignocellulosic
feedstocks. Their model minimises total system cost and introduces binary facility-opening variables that have since become standard in the field. A key limitation, however, is that biomass supply is treated as an exogenous input, so feedstock availability is not endogenised as a function of land-use or growth decisions. @Mansoornejad2013_Forest_Biorefinery_SC extend the strategic design perspective to forest biorefineries by integrating product portfolio design with supply chain configuration in a single MILP. Their model explicitly accounts for product conversion efficiencies along different processing pathways, enabling joint optimisation of process technology selection and logistics.

**Multi-period and tactical formulations.**
@DeMeyer2015_OPTIMASS introduce the OPTIMASS framework, a generic MILP for biomass-based supply chain optimisation that explicitly models product quality cascades: higher-quality biomass classes can substitute for lower-quality demands downstream, whereas the reverse is infeasible. This cascade hierarchy, formalised through ordered product sets, is particularly relevant for woody biomass where stem, branch, and residue fractions command different market prices. The companion t-OPTIMASS extension [@DeMeyer2016_tOPTIMASS] incorporates temporal dynamics by explicitly representing biomass growth and regeneration over multiple periods, enabling simultaneous optimisation of harvest timing and network flows. However, in both OPTIMASS formulations the growth dynamics are represented as exogenous time-series rather than as age-structured stand models, so rotation-length decisions and age-dependent biomass quality are not endogenised. @Arabi2019_Microalgae_Biobutanol_SC demonstrate multi-period MILP extensions for bio-based supply chains with production dynamics, but focus on microalgae rather than woody biomass and do not address product quality cascades.

**Uncertainty and stochastic formulations.**
@Sharma2013_BSC_review identify supply-side uncertainty (yield variability, seasonal availability, moisture content) as the most critical source of risk in biomass supply chains. In response, a strand of literature has developed stochastic and robust optimisation extensions. @Zahraee2020_BSC_review document that two-stage stochastic MILP and robust optimisation approaches now account for roughly 30\% of the modelling literature, with biomass yield uncertainty and price volatility as the dominant stochastic parameters.

**Agroforestry and short-rotation coppice logistics.**
Biomass supply chain research focusing specifically on agroforestry or SRC contexts is comparatively sparse.
@Espinoza2017_Biorefinery_SC_review note that most supply chain models aassume continuous feedstock availability and do not represent the periodic harvest cycles characteristic of coppice systems. @Lamers2015_Feedstock_Design analyse feedstock design principles for lignocellulosic supply chains and argue that the allocation of different biomass fractions to valorisation pathways—product cascading—is a key lever for economic viability, yet is rarely optimised jointly with network design decisions. The model developed in Section~\@ref(sec:milp) of this paper directly
addresses this gap by endogenising stand age dynamics through binary age-lag variables and coupling component-specific yield coefficients with a product quality cascade.

**Modelling features overview.**
Table~\@ref(tab:lit-sc-features) summarises the key modelling features of
the reviewed contributions.

```{r lit-sc-features, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
options(kableExtra.latex.load_packages = FALSE)
options(knitr.table.format = "latex")
knitr::opts_chunk$set(dev = "pdf")

lit_tab <- tribble(
  ~Study,                            ~Objective,          ~`Multi-product`,  ~`Quality cascade`, ~`Age dynamics`,    ~`Uncertainty`,    ~`Level`,
  "Eksioglu et al. (2009)",          "Cost min.",         "No",              "No",               "No",               "No",              "Strategic",
  "Sharma \\& Taaffe (2013)",        "Cost min.",         "Partial",         "No",               "No",               "No",              "Strategic/Tactical",
  "Mansoornejad et al. (2013)",      "Profit max.",       "Yes",             "No",               "No",               "No",              "Strategic",
  "Cambero \\& Sowlati (2014)",      "Cost/GHG/jobs",     "No",              "No",               "No",               "Stochastic",      "Strategic",
  "Yue et al. (2014)",               "Cost/GHG",          "Yes",             "No",               "No",               "No",              "Multi-scale",
  "De Meyer et al. (2015)",          "Cost min.",         "Yes",             "Yes",              "Exogenous",        "No",              "Strategic/Tactical",
  "De Meyer et al. (2016)",          "Cost min.",         "Yes",             "Yes",              "Exogenous TS",     "No",              "Multi-period",
  "Lamers et al. (2015)",            "Cost/quality",      "Yes",             "Partial",          "No",               "Scenario",        "Strategic",
  "Espinoza et al. (2017)",          "Sustainability",    "Yes",             "No",               "No",               "Scenario",        "Strategic",
  "Malladi \\& Sowlati (2018)",      "Cost min.",         "Partial",         "No",               "No",               "No",              "Tactical/Oper.",
  "Arabi et al. (2019)",             "Profit max.",       "Yes",             "No",               "Exogenous",        "No",              "Multi-period",
  "Zahraee et al. (2020)",           "Multi-obj.",        "Yes",             "No",               "No",               "Stochastic",      "Strategic/Tactical",
  "\\textbf{This paper}",            "NPV max.",          "Yes",             "Yes",              "\\textbf{Endogenous}", "Scenario",   "Strategic/Multi-period"
) |>
  mutate(across(everything(), ~ gsub("Yes", "\\\\checkmark", .))) |>
  mutate(across(everything(), ~ gsub("No",  "---",            .)))

kable(lit_tab,
      format    = "latex",
      booktabs  = TRUE,
      escape    = FALSE,
      caption   = "Key modelling features of selected biomass supply chain design studies.",
      label     = "lit-sc-features") |>
  kable_styling(
  font_size     = 7,
  latex_options = c("hold_position", "scale_down"),
  full_width    = FALSE
) |>
  row_spec(nrow(lit_tab), bold = TRUE) |>
  column_spec(1, width = "3.2cm") |>
  column_spec(2, width = "1.6cm") |>
  column_spec(3:7, width = "1.4cm")
```

# MILP model for agroforestry supply chain design {#sec:milp}

## Sets, indices, and time structure

The model considers a set of agroforestry sites \(\mathcal{I}\), storage/processing facilities \(\mathcal{J}\), consumer sites \(\mathcal{K}\), and product types \(\mathcal{P} = \{1,2,3\}\). Time is discretised into annual periods \(t \in \mathcal{T} = \{1,\dots,T_{\max}\}\). To represent stand establishment and harvest timing, extended time sets \(\mathcal{T}^{+} = \{0,\dots,T_{\max}+1\}\) and harvest periods \(\mathcal{T}^{\text{harv}} = \{A^{\min}+1,\dots,T_{\max}\}\) are defined, where \(A^{\min}\) and \(A^{\max}\) are minimum and maximum stand ages (years) at harvest. The set \(\mathcal{S}\) comprises all ordered pairs \((s,t)\) of consecutive harvest periods (including establishment at \(s=0\)) satisfying the age-lag bounds:
\begin{equation}
\mathcal{S} = \{(s,t) \mid s,t \in \{0,\dots,T_{\max}+1\},\; t-s \in \{A^{\min},\dots,A^{\max}\} \}.
\label{eq:age-lag-set}
\end{equation}


```{r tab-notation-milp, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
options(kableExtra.latex.load_packages = FALSE)
options(knitr.table.format = "latex")

# Hilfsfunktion: bricht einen langen Math-Ausdruck an einer definierten
# Stelle um — teilt bei | in zwei Zeilen mit \shortstack
math_break <- function(line1, line2) {
  paste0("\\shortstack[l]{$", line1, "$\\\\$", line2, "$}")
}

notation <- tribble(
  ~Symbol, ~Description,

  # --- SETS & INDICES ---
  "$i \\in \\mathcal{I}$",   "Agroforestry sites",
  "$j \\in \\mathcal{J}$",   "Storage / pre-treatment facilities",
  "$k \\in \\mathcal{K}$",   "Consumer sites",
  "$p \\in \\mathcal{P}$",   "Product types (stem, branches, residues)",
  "$(p,p') \\in \\mathcal{Q}$",   "Product usage (e.g. $p=stem$, it holds $p' \in \\{stem, branches, residues \\}$) ",
  "$\\mathcal{T}=\\{1,\\dots,T_{\\max}\\}$",
    "Planning periods (years)",
  "$\\mathcal{T}^{+}=\\{0,\\dots,T_{\\max}+1\\}$",
    "Extended time set incl.\\ dummy nodes",
    "$\\mathcal{T}^{\\text{harv}}=\\{A^{\\min}+1,\\dots,T_{\\max}\\}$",
    "Feasible harvest periods",
    "$\\mathcal{S}^{\\text{est}}=\\{(0,t)\\mid t\\in\\{1,\\dots,T_{\\max}-A^{\\min}\\}\\}$",
    "Establishment arcs: dummy source to period $t$",
  math_break(
    "\\mathcal{S}^{\\text{harv}}=\\{(s,t)\\mid s,t\\in \\mathcal{T},",
    "t-s\\in\\{A^{\\min},\\dots,A^{\\max}\\}\\}"),
    "Harvest cycles within age-lag bounds",
  math_break(
    "\\mathcal{S}^{\\text{term}}=\\{(s,T_{\\max}+1)\\mid",
    "s\\in\\{A^{\\min}+1,\\dots,T_{\\max}\\}\\}"),
    "Termination arcs to dummy sink $T_{\\max}+1$",
  
    "$\\mathcal{S} = \\mathcal{S}^{\\text{est}} \\cup \\mathcal{S}^{\\text{harv}} \\cup\\; \\mathcal{S}^{\\text{term}}$",
    "Age-lag arc set (establishment $\\cup$ harvest $\\cup$ termination)",

  # --- DECISION VARIABLES ---
  "$z_{ist}\\in\\{0,1\\}$",
    "1 if $s$ and $t$ are consecutive harvest cycles at site $i$",
  "$Y_{ipt}\\ge 0$",
    "Harvest quantity (t) of product $p$ at site $i$ in period $t$",
  "$X_{ijpt}\\ge 0$",
    "Transport flow (t) of $p$ from site $i$ to facility $j$ in $t$",
  "$X_{jkpp't}\\ge 0$",
    "Flow of $p$ from facility $j$ to consumer $k$ satisfying $p'$ in $t$",
  "$S_{jpt}\\ge 0$",
    "Inventory (t) of product $p$ at facility $j$ end of period $t$",

  # --- PARAMETERS ---
  "$T_{\\max}$",              "Length of planning horizon (years)",
  "$A^{\\min}$, $A^{\\max}$", "Minimum / maximum stand age at harvest (years)",
  "$\\text{AREA}_i$",         "Area of site $i$ (ha)",
  "$\\eta_{pa}$",             "Yield coefficient (t/ha) of product $p$ at age $a$",
  "$R_{kp}$",                 "Revenue (\\euro/t) for product $p$ at consumer $k$",
  "$C^{\\text{est}}_i$",      "Establishment cost (\\euro/ha) at site $i$",
  "$C^{\\text{opp}}_i$",      "Annual opportunity cost (\\euro/ha) at site $i$",
  "$C^{\\text{harv}}_i$",     "Harvest \\& refitting cost (\\euro/ha) at site $i$",
  "$c^{\\text{tr-raw}}$",     "Transport cost rate (\\euro/t\\,km) for raw biomass",
  "$c^{\\text{tr-pre}}$",     "Transport cost rate (\\euro/t\\,km) pre-treated biomass",
  "$c^{\\text{stor}}_j$",     "Storage holding cost (\\euro/t) at facility $j$",
  "$d_{ij}$, $d_{jk}$",       "Distance (km) between nodes",
  "$\\text{CAP}^{\\text{stor}}_j$", "Storage capacity (t) at facility $j$",
  "$\\text{CAP}^{\\text{proc}}_j$", "Processing capacity (t/period) at facility $j$",
  "$D^{\\max}_{kpt}$",        "Maximum demand (t) of $p$ at consumer $k$ in period $t$"
)

tbl <- kable(notation,
      format    = "latex",
      booktabs  = TRUE,
      escape    = FALSE,
      col.names = NULL,
      caption   = "Notation for the MILP model.",
      label     = "tab-notation-milp") |>
  kable_styling(
    font_size     = 8,
    latex_options = c("hold_position", "scale_down","striped"),
    full_width    = FALSE
  ) |>
  column_spec(1, width = "5cm") |>
  column_spec(2, width = "9cm") |>
  pack_rows("Sets \\& Indices",    1,  11,
            bold = TRUE, italic = FALSE, hline_before = TRUE, escape = FALSE) |>
  pack_rows("Decision Variables", 12,  16,
            bold = TRUE, italic = FALSE, hline_before = TRUE) |>
  pack_rows("Parameters",         17,  31,
            bold = TRUE, italic = FALSE, hline_before = TRUE)

# Spaltentyp p{} → >{\raggedright\arraybackslash}p{} direkt im LaTeX-String ersetzen
tbl <- gsub(
  "p\\{(5\\.8cm|8\\.5cm)\\}",
  ">\\{\\\\raggedright\\\\arraybackslash\\}p\\{\\1\\}",
  tbl
)

tbl
```


```{tikz fig-set-S-tikz, echo=FALSE,     fig.cap="Network structure of set $\\mathcal{S}$ for $T=8$, $A_{\\min}=3$,    $A_{\\max}=5$. Dashed blue arcs: establishment arcs $(0,t)$,     $t\\in\\{1,\\ldots,T-1\\}$. Solid green arcs above: harvest arcs $(s,t)$     with odd $s$; below: harvest arcs with even $s$. Solid red arcs:     termination arcs $(s,T+1)$, $s\\in\\{A_{\\min}+1,\\ldots,T\\}$.",    fig.align="center", out.width="\\textwidth",     engine.opts=list(extra.preamble=c(      "\\usetikzlibrary{arrows.meta,bending,positioning}"    ))}

\begin{tikzpicture}[
  x=1.55cm, y=1cm,
  node/.style={circle, draw, fill=#1, text=white,
               font=\small\bfseries, inner sep=2pt, minimum size=18pt},
  est/.style  ={-{Stealth[length=4pt]}, blue!70!black,  dashed, semithick},
  harva/.style={-{Stealth[length=4pt]}, green!40!black, semithick},
  harvb/.style={-{Stealth[length=4pt]}, green!40!black, semithick},
  term/.style ={-{Stealth[length=4pt]}, red!70!black,   semithick},
]

%% ---- Nodes ----
\node[node=blue!70!black]  (n0) at (0,0) {$0$};
\foreach \i in {1,...,8} {
  \node[node=gray!50!black] (n\i) at (\i,0) {$\i$};
}
\node[node=red!70!black]   (n9) at (9,0) {$T\!+\!1$};

%\node[below=10pt of n0, font=\scriptsize\itshape, text=blue!70!black]
%  {Estab.\,(dummy)};
%\node[below=10pt of n9, font=\scriptsize\itshape, text=red!70!black]
%  {Term.\,(dummy)};

%% ---- Establishment arcs ----
%% (0,1) straight
\draw[est] (n0) -- (n1);
\foreach \t/\l in {2/1.3, 3/1.0, 4/0.80, 5/0.65, 6/0.55, 7/0.48}{
  \draw[est] (n0) to[out=55, in=125, looseness=\l] (n\t);
}

%% ---- Harvest arcs (odd s above, even s below) ----
% span 3
\draw[harva] (n1) to[out=55,  in=125,  looseness=0.55] (n4);
\draw[harvb] (n2) to[out=-55, in=-125, looseness=0.55] (n5);
\draw[harva] (n3) to[out=55,  in=125,  looseness=0.55] (n6);
\draw[harvb] (n4) to[out=-55, in=-125, looseness=0.55] (n7);
\draw[harva] (n5) to[out=55,  in=125,  looseness=0.55] (n8);
% span 4
\draw[harva] (n1) to[out=60,  in=120,  looseness=0.42] (n5);
\draw[harvb] (n2) to[out=-60, in=-120, looseness=0.42] (n6);
\draw[harva] (n3) to[out=60,  in=120,  looseness=0.42] (n7);
\draw[harvb] (n4) to[out=-60, in=-120, looseness=0.42] (n8);
% span 5
\draw[harva] (n1) to[out=65,  in=115,  looseness=0.35] (n6);
\draw[harvb] (n2) to[out=-65, in=-115, looseness=0.35] (n7);
\draw[harva] (n3) to[out=65,  in=115,  looseness=0.35] (n8);

%% ---- Termination arcs ----
%% (8,T+1) straight
\foreach \s/\l in {4/0.80, 5/0.65, 6/0.55, 7/0.48}{
  \draw[term] (n\s) to[out=55, in=125, looseness=\l] (n9);
}
\draw[term] (n8) -- (n9);

%% ---- Legend (no frame, side by side) ----
\node[anchor=north west] at (0.0,-0.85) {%
  \begin{tikzpicture}[x=0.6cm,y=0.35cm,baseline=0pt]
    \draw[est]   (0,0)--(1,0);
    \node[font=\small, anchor=west] at (1.1,0)
      {Establishment arc $(0,t)$};
  \end{tikzpicture}};
\node[anchor=north west] at (4.2,-0.85) {%
  \begin{tikzpicture}[x=0.6cm,y=0.35cm,baseline=0pt]
    \draw[harva] (0,0)--(1,0);
    \node[font=\small, anchor=west] at (1.1,0)
      {Harvest arc $(s,t)$};
  \end{tikzpicture}};
\node[anchor=north west] at (7.6,-0.85) {%
  \begin{tikzpicture}[x=0.6cm,y=0.35cm,baseline=0pt]
    \draw[term]  (0,0)--(1,0);
    \node[font=\small, anchor=west] at (1.1,0)
      {Termination arc $(s,T{+}1)$};
  \end{tikzpicture}};

\end{tikzpicture}
```


```{tikz fig-set-S-path, echo=FALSE,     fig.cap="Example of a feasible path through the network for $T=8$,     $A_{\\min}=3$, $A_{\\max}=5$: establishment in period~1, harvests in periods~5 and~8, termination at $T+1$.",     fig.align="center", out.width="\\textwidth",     engine.opts=list(extra.preamble=c(      "\\usetikzlibrary{arrows.meta,bending,positioning}"    ))}

\begin{tikzpicture}[
  x=1.55cm, y=1cm,
  node/.style    ={circle, draw, fill=#1, text=white,
                   font=\small\bfseries, inner sep=2pt, minimum size=18pt},
  inactive/.style={node=gray!25, text=gray!50},
  pathnode/.style={node=green!40!black},
  n0style/.style ={node=blue!70!black},
  n9style/.style ={node=red!70!black},
  patharc/.style ={-{Stealth[length=5pt]}, green!40!black, very thick},
  est/.style     ={patharc, blue!70!black, dashed},
  term/.style    ={patharc, red!70!black},
]

%% ---- Nodes ----
\node[n0style]   (n0) at (0,0) {$0$};
\foreach \i in {2,3,4,6,7} {
  \node[inactive] (n\i) at (\i,0) {$\i$};
}
\foreach \i in {1,5,8} {
  \node[pathnode]  (n\i) at (\i,0) {$\i$};
}
\node[n9style]   (n9) at (9,0) {$T\!+\!1$};

%\node[below=10pt of n0, font=\scriptsize\itshape, text=blue!70!black]
%  {Estab.\,(dummy)};
%\node[below=10pt of n9, font=\scriptsize\itshape, text=red!70!black]
%  {Term.\,(dummy)};

%% ---- Path arcs ----
\draw[est]     (n0) -- (n1);
\draw[patharc] (n1) to[out=60, in=120, looseness=0.42] (n5);
\draw[patharc] (n5) to[out=55, in=125, looseness=0.55] (n8);
\draw[term]    (n8) -- (n9);

%% ---- Arc labels ----
\node[font=\footnotesize, text=blue!70!black]   at (0.5, 0.28) {est.};
\node[font=\footnotesize, text=green!40!black]  at (3.0, 0.6) {$t-s=4$};
\node[font=\footnotesize, text=green!40!black]  at (6.5, 0.55) {$t-s=3$};
\node[font=\footnotesize, text=red!70!black]    at (8.5, 0.28) {term.};

%% ---- Legend (no frame, side by side) ----
\node[anchor=north west] at (0.0,-0.78) {%
  \begin{tikzpicture}[x=0.6cm,y=0.35cm,baseline=0pt]
    \draw[est]     (0,0)--(1,0);
    \node[font=\small, anchor=west] at (1.1,0)
      {Establishment arc $(0,t)$};
  \end{tikzpicture}};
\node[anchor=north west] at (4.2,-0.78) {%
  \begin{tikzpicture}[x=0.6cm,y=0.35cm,baseline=0pt]
    \draw[patharc] (0,0)--(1,0);
    \node[font=\small, anchor=west] at (1.1,0)
      {Harvest arc $(s,t)$};
  \end{tikzpicture}};
\node[anchor=north west] at (7.6,-0.78) {%
  \begin{tikzpicture}[x=0.6cm,y=0.35cm,baseline=0pt]
    \draw[term]    (0,0)--(1,0);
    \node[font=\small, anchor=west] at (1.1,0)
      {Termination arc $(s,T{+}1)$};
  \end{tikzpicture}};

\end{tikzpicture}
```

## Decision variables

Binary variables \(z_{ist}\) indicate whether, at site \(i\), periods \(s\) and \(t\) form consecutive harvests (or establishment and first harvest if \(s=0\)), with \((s,t) \in \mathcal{S}\) as defined in \eqref{eq:age-lag-set}. These variables define a path of consecutive harvest cycles at each site over the planning horizon. Continuous variables \(Y_{ipt} \ge 0\) represent the maximum harvest quantity (t) of product \(p\) at site \(i\) in harvest period \(t\). Flow variables \(X_{ijpt}\) and \(X_{jkpp't}\) represent transport of products from sites to storage and from storage to consumers, respectively, and \(S_{jpt}\) represent inventories at storage locations.

## Objective function

The model maximises net present value over the planning horizon, expressed as total revenues minus establishment, opportunity, harvest, transport, and storage costs:
\begin{equation}
\begin{aligned}
\max Z =\;&
\sum_{k,p,t} R_{k,p} \cdot \sum_{j \in \mathcal{J},\, p' \in \mathcal{P}: p' \leq p} X_{jkp'pt} \\
& - \sum_{i} C^{\text{est}}_{i} \cdot \text{AREA}_i \cdot \sum_{t \in \mathcal{T}} z_{i0t} \\
& - \sum_{i} C^{\text{opp}}_{i} \cdot \text{AREA}_i \cdot \sum_{t \in \mathcal{T}} (T_{\max}-t) \cdot z_{i0t} \\
& - \sum_{i,(s,t)\in \mathcal{S}: s>0} C^{\text{harv}}_{i} \cdot \text{AREA}_i \cdot z_{ist} \\
& - c^{\text{tr-raw}} \sum_{i,j,p,t} d_{ij} \cdot X_{ijpt}
- c^{\text{tr-pre}} \sum_{j,k,p,t} d_{jk} \cdot X_{jkpt}
- \sum_{j,p,t} c^{\text{stor}}_{j} \cdot S_{jpt}.
\end{aligned}
\label{eq:objective}
\end{equation}

Here, \(R_{k,p}\) are product- and consumer-specific revenues (€/t), \(C^{\text{est}}_{i}\) are establishment costs per hectare, \(C^{\text{opp}}_{i}\) are annual opportunity costs per hectare, and \(C^{\text{harv}}_{i}\) are harvest and refitting costs per hectare [@Testa2014_Poplar_Economics]. Transport costs are proportional to flow and distance with rates \(c^{\text{tr-raw}}\) and \(c^{\text{tr-pre}}\) for raw and pre-treated biomass, respectively, and storage incurs linear holding costs \(c^{\text{stor}}_j\) per tonne.

## Key constraints

The path connectivity constraint ensures that each harvest period at a site has exactly one predecessor and one successor in the age-lag graph, effectively forming a path over \(\mathcal{T}^{+}\):
\begin{equation}
\sum_{(s,t)\in \mathcal{S}} z_{ist} = \sum_{(t,u) \in \mathcal{S}} z_{itu}
\quad \forall i \in \mathcal{I}, t \in \mathcal{T}.
\label{eq:path-connect}
\end{equation}

Establishment is restricted to at most one time per site:
\begin{equation}
\sum_{(0,t) \in \mathcal{S}^{est}} z_{i0t} \le 1 \quad \forall i \in \mathcal{I}.
\label{eq:one-establishment}
\end{equation}

Biomass yield is linked to age-lag decisions through the allometric yield coefficients:
\begin{equation}
Y_{ipt} = \sum_{s=1}^{t-1} \eta_{p(t-s)} \cdot \text{AREA}_i \cdot z_{ist}
\quad \forall i \in \mathcal{I}, p \in \mathcal{P}, t \in \mathcal{T}^{\text{harv}}.
\label{eq:yield-link}
\end{equation}

Harvest quantities bound outbound flows from each site:
\begin{equation}
\sum_{j \in \mathcal{J}} X_{ijpt} \le Y_{ipt}
\quad \forall i,p,t \in \mathcal{T}^{\text{harv}}.
\label{eq:harvest-flow}
\end{equation}

At storage hubs, inventories obey standard balance equations with initial inventory zero:
\begin{equation}
S_{jpt} = S_{jp,t-1} + \sum_{i} X_{ijpt} - \sum_{k, p' \ge p} X_{jkpp't}
\quad \forall j,p,t \in \mathcal{T}^{\text{harv}}.
\label{eq:inventory-balance}
\end{equation}

Storage and processing capacities are enforced via:
\begin{equation}
\sum_{p} S_{jpt} \le \text{CAP}^{\text{stor}}_j \quad \forall j,t, \qquad
\sum_{i,p} X_{ijpt} \le \text{CAP}^{\text{proc}}_j \quad \forall j,t.
\label{eq:capacity}
\end{equation}

Demand satisfaction with cascades is modelled as described above, where higher-quality products can satisfy lower-quality demands up to maximum demand levels \(D^{\max}_{kpt}\).


# Case study and computational experiment plan {#sec:experiments}

The proposed MILP model is intended to support the design of agroforestry biomass supply chains for cleaner production in specific regional contexts. In a forthcoming case study, we focus on poplar-based alley-cropping systems in a temperate European setting, informed by empirical data from SRAFS trials in Southern Germany and SRC plantations in Central Europe [@Huber2018_SRAFS_growth; @Trnka2008_Poplar_SRF; @Testa2014_Poplar_Economics].

We envisage the following steps and experimental scenarios for computational analysis:

1. **Parameterisation of allometric yield profiles:** Use species- and site-specific allometric models and observed growth trajectories to construct age-dependent yield coefficients \(\eta_{pa}\) for chemical, pulp, and energy product classes at representative agroforestry sites. Where necessary, scenario ranges will be defined to capture uncertainty in MAI and CAI due to climate variability and management differences [@Huber2016_Allometric_SRAFS; @Huber2018_SRAFS_growth; @Trnka2008_Poplar_SRF].

2. **Baseline optimisation:** Solve the MILP model \eqref{eq:objective}--\eqref{eq:capacity} for a baseline configuration with fixed prices, capacities, and policy settings, obtaining optimal establishment timing, harvest schedules, and supply chain structure (locations, flows, inventories).

3. **Rotation length scenarios:** Vary the age-lag bounds \(A^{\min}\) and \(A^{\max}\) in \eqref{eq:age-lag-set} to represent alternative rotation strategies (e.g., 3--5 years vs. 8--10 years) and analyse their impacts on profitability, product mix, and land use. This will shed light on trade-offs between shorter rotations favouring energy chips and longer rotations enabling higher shares of material products [@Trnka2008_Poplar_SRF; @Testa2014_Poplar_Economics].

4. **Product price and cascading scenarios:** Explore scenarios with different relative prices for chemical, pulp, and energy products, reflecting changes in market demand and policy incentives (e.g., bioenergy subsidies, green chemicals markets). Assess how the product cascade hierarchy and yield quality structure influence the marginal value of species and rotation decisions [@DeMeyer2015_OPTIMASS; @Lamers2015_Feedstock_Design].

5. **Infrastructure and logistics scenarios:** Evaluate the effect of alternative storage and processing network configurations (e.g., decentralised vs. centralised hubs, different capacity levels, transport cost parameters) on the optimal spatial deployment of agroforestry sites and biomass flows [@DeMeyer2015_OPTIMASS; @DeMeyer2016_tOPTIMASS; @Lamers2015_Feedstock_Design].

6. **Sensitivity to allometric and growth uncertainty:** Conduct parametric sensitivity and, where computationally feasible, stochastic analyses to quantify the impact of uncertainty in allometric parameters and growth responses (e.g., under drought or management differences) on optimal decisions and system robustness [@Huber2018_SRAFS_growth; @Trnka2008_Poplar_SRF].

7. **Cleaner production indicators:** For selected solutions, compute derived indicators such as biomass cascading index (share of biomass used for material products before energy), utilisation rates of higher-quality fractions, and, where coupled with LCA data, greenhouse gas emission savings relative to fossil benchmarks [@Espinoza2017_Biorefinery_SC_review; @Zahraee2020_BSC_review].


# Conclusions

This paper outlines a research framework that couples allometric biomass growth models \eqref{eq:allometry}--\eqref{eq:component-share} with a MILP-based agroforestry biomass supply chain design model \eqref{eq:objective}--\eqref{eq:capacity} featuring age-lag constraints and product cascading. Building on empirical evidence from temperate SRAFS and SRC systems and established biomass supply chain models, the proposed approach aims to capture key interactions between stand-level growth dynamics and network-level design decisions that are central to the cleaner production potential of agroforestry [@Huber2018_SRAFS_growth; @DeMeyer2015_OPTIMASS; @DeMeyer2016_tOPTIMASS]. Future work will implement and calibrate the model for a regional case study, conduct the computational experiments sketched above, and extend the framework towards multi-objective optimisation including environmental indicators such as greenhouse gas emissions and biodiversity proxies.
