---
title: "Strategic Planning of Agroforestry Supply Chains"

format:
  wiley-njd-pdf:
    latex-tinytex: false      # use MiKTeX
    pdf-engine: xelatex
    classoption: [APA, TimesLARGE]
    journal: "Preprint submitted to Journal of Industrial Ecology"
    article-type: "Research Article"
    author-mark: "T. Kirschstein"
    keep-tex: true

execute:
  echo: false

author:
  - name: Thomas Kirschstein
    corresponding: true
    email: thomas.kirschstein@ikts.fraunhofer.de
    affiliations:
      - name: Fraunhofer IKTS Center for Economics and Management of Materials
        address: "Leipziger Str. 70/71"
        city: Halle
        postal-code: "06108"
        country: Germany
      - name: "Hochschule RheinMain – University of Applied Sciences and Arts, Wiesbaden Business School"
        address: "Bleichstr. 44"
        city: Wiesbaden
        postal-code: "65183"
        country: Germany

keywords:
  - Industrial symbiosis
  - Agroforestry
  - Biomass cascading
  - Bioeconomy network design
  - Supply chain optimisation
  - Biobased production

abstract: |
  Agroforestry systems integrate fast-growing woody perennials into agricultural land forming a  biological-industrial symbiosis. They simultaneously deliver multiple biomass fractions of differing quality to food, chemical, pulp-and-paper, as well as energy industries through a shared regional supply network. The woody biomass feedstock produced in agroforestry systems can be used to substitute fossil feedstocks particularly in chemical and energy industry. Additionally, biogenic carbon can be stored in long-lived products and agroforestry increases climate resiliency and sustainability of agricultural production. In contrast to the traditional forestry-based supply of woody biomass, agroforestry-based woody biomass shows substantial feedstock potentials as most land is used agriculture in central Europe. Despite these environmental and economic potentials, agroforestry is only applied in marginal scale in central European agriculture. A core challenge in fostering agroforestry is to establish economically viable regional bioeconomy networks which connect partners from agriculture as well as downstream industries. Such a agroforestry-based supply chain needs to be structured and organized in a way that allows all partners to benefit economically in the long run. This paper presents a supply chain design model that helps to support design decisions for agroforestry-based supply chains. Therefore, it models agroforestry establishment and harvest scheduling on farm-level in a large-scaled linear multi-echelon network design model. It relies on biomass growth models which describe the production of multiple biomass types in agroforestry systems. These biomass fractions are pretreated and transported to consumers from  chemical, paper-and-pulp, and energy industry depending on quality and demand restrictions. The model is applied in a case study focussing on central Germany where a heterogenuous industrial cluster is located within a fertile and large agricultural production area. The optimization results indicate that agroforestry-based supply chains can be configured in an economically viable way. Agroforestry systems can serve as feedstock suppliers for industrial consumers in the same region at a considerable scale. Profitability of agroforestry-based agricultural practice seem to be competitive with traditional agricultural production and at the same time allow producing woody biomass at a competitive cost, provided transports are not too long. 

bibliography: references.bib
csl: apa.csl

knitr:
  opts_chunk:
    fig.align: "center"
    fig.pos: "htbp"
    out.width: "0.99\\textwidth"
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo        = FALSE,
  message     = FALSE,
  warning     = FALSE,
  fig.align   = "center",
  fig.path    = "figures/",
  # --- additions ---
  cache       = FALSE,          # set TRUE per-chunk for heavy computations
  out.width   = "\\textwidth",  # consistent figure width for elsevier 1p
  dev         = "pdf"           # vector output for LaTeX/XeLaTeX pipeline
)

# Suppress noisy package startup messages globally
suppressPackageStartupMessages({
  library(magick)
  library(dplyr)
  library(plotly)
  library(ggplot2)
  library(tidyr)
  library(patchwork)
  library(knitr)
  library(kableExtra)
  library(sf)
  library(tibble)
  library(stringr)
  library(leaflet)
  library(osrm)
  library(ROI)
  library(ROI.plugin.glpk)
  library(gurobi)
  library(ROI.plugin.gurobi)
  library(ggrepel)
  library(ggnewscale)
  library(rnaturalearth)
  library(rnaturalearthdata)
  library(Rcpp)
  library(RcppArmadillo)
  library(scatterpie)
  library(networkD3)
  library(htmlwidgets)
  library(webshot2)
  library(ggalluvial)
  library(scales)
  library(purrr)
  library(broom)
  library(lme4)        # mixed-effects models
  library(lmerTest)    # Satterthwaite p-values for lmer
  library(mgcv)        # GAM + GAMM
  library(gratia)      # tidy GAM diagnostics & draw()
  library(MuMIn)       # R² for mixed models (r.squaredGLMM)
  library(AICcmodavg)  # AICc comparison table
  library(performance) # model_performance(), compare_performance(
  library(broom.mixed) # tidy() for lmer objects
  library(kableExtra)
})
options(kableExtra.latex.load_packages = FALSE)
options(knitr.table.format = "latex")

```

```{r load-model-data, echo=FALSE, message=FALSE, warning=FALSE, cache=T, cache.extra = tools::md5sum(c(     "../Modell/!helper_func.r", "../Modell/!helper_instance_builder_v8a.R","../Modell/!build_AFS_milp.r",      "../Modell/build_lp_rcpp_wrapper.R","../Modell/gurobi_adapter.R","../Modell/gurobi_warmstart.R",      "../Modell/gurobi_feascheck.R","../Modell/extract_result.R","../Modell/build_lp_rcpp.cpp","afs_workspace_red.RData",      "lp_r.RDS","lp_r_result.rds","lp_instance_data.rds" ))}
source("../Modell/!afs_biomass_setup.r")
source("../Modell/!helper_func.r")
source("../Modell/!helper_instance_builder_v8a.R")
source("../Modell/!build_AFS_milp.r")
source("../Modell/build_lp_rcpp_wrapper.R")
source("../Modell/gurobi_adapter.R")
source("../Modell/gurobi_warmstart.R")
source("../Modell/gurobi_feascheck.R")
source("../Modell/extract_result.R")
source("!helper_extract_site_profit_paper.R")
#Rcpp::sourceCpp("../Modell/build_lp_rcpp.cpp")
#Rcpp::sourceCpp("../Modell/build_and_solve_afs_milp.cpp")
Rcpp::sourceCpp("../Modell/build_and_solve_afs_lp_v11.cpp")

#load("../Modell/afs_workspace.RData")
load("afs_workspace_red.RData")

sites_sf   <- afs_workspace$sites_sf
sites      <- afs_workspace$sites
sites_clustered      <- afs_workspace$sites_clustered
storages   <- afs_workspace$storages
consumers  <- afs_workspace$consumers

# scale down industrial demand quantities by dividing by 5
consumers <- consumers %>%
  mutate(demand_P1 = demand_P1 / 5)

# scale down prices to 40-70 €/t chemical, 30-40 €/t pulp + 15-20 €/t
consumers <- consumers %>% 
  mutate(P1 = P1/2, P2 = P2/2, P3 = P3*.6 )

# drop small Biogas plants --> compute total demand and drop all consumer with D<2
consumers <- consumers %>%
  mutate(total_demand = demand_P1 + demand_P2 + demand_P3) %>%
  filter(total_demand >= 1.5) %>%
  select(-total_demand) 


dist_jk <- afs_workspace$dist_jk[,consumers$consumer_id]

yields_by_age <- build_scenario_ts(ages = 1:20 , N = 250, C_site = 6.500, k = 0.194, t0 = 9.7, label = "Conservative (250 trees/ha)")


yields_by_age <- yields_by_age %>% 
  # drop column scenario from yields_by_age
  select(-scenario) %>% 
  # make long --> bring columns Merch. stem Merch. branch Residue to column "product" and values to "yield_ha"
  pivot_longer(-1 , names_to = "product", values_to = "yield_ha") %>% 
  # rename entries in "product" to 1:3
  mutate(product = as.integer(as.character(factor(product, labels = c(2,1,3) ))))

# scale yields to fresh biomass
yields_by_age$yield_ha <- yields_by_age$yield_ha / (1-0.5)


set.seed(123578)
# generate random vector of opportunity cost for clustered sites in range [0,600]
c.opp.vec <- rnorm(nrow(sites_clustered), 1, .3)


sites_clustered <- sites_clustered %>%
   mutate(
     C_est  = 2500,   # €/ha  establishment (Section 2.1 mid-range)
     C_harv =  150,    # €/ha  harvest cost  (Section 2.1 mid-range)
     C_main = 10 , #  €/ha  maintenance cost  (Section 2.1 mid-range)
     C_opp = 500*c.opp.vec, #  €/ha  opportunity cost  (Section 2.1 mid-range)
   ) #%>%   # change column name area_ha to area if necessary  rename(area_ha = area)


# # change storage costs
# storages$c_stor <- storages$c_stor
# # change capacities
 storages$CAP_stor <- storages$CAP_stor * 100000
 storages$CAP_proc <- storages$CAP_proc * 100000
# 
data_obj <- list(
   sites    = sites_clustered,
  #sites    = sites_filt,
   storages = storages,
   consumers = consumers,
   dist_ij  = afs_workspace$dist_ij_clust,
   #dist_ij  = dist_ij_filt$distance_matrix_km,
   dist_jk  = dist_jk,
   yields_by_age = yields_by_age
 )
# rescale demands in consumers to tons
data_obj$consumers[,c("demand_P1","demand_P2","demand_P3")]  <- data_obj$consumers[,c("demand_P1","demand_P2","demand_P3")] * 1000

# 
# # ── 3. PARAMS OBJECT — all scalar parameters ──────────────────────────────────
params_obj <- list(
   n_periods  = 40,
   max_age    = 20L,    # A_max: LRAFS upper rotation bound
   min_age    =  3L,    # A_min: SRAFS lower rotation bound
   c_tr_raw   = 0.08,   # €/(t·km)  raw biomass  site → storage
   c_tr_pre   = 0.06   # €/(t·km)  chips        storage → consumer
 )
# 
# ── 4. generate instance ─────────────────────────────────────
milp_instance <- build_optimization_instance(
    data         = data_obj,
    params       = params_obj
  ) 

#result <- build_and_solve_afs_lp_v11(milp_instance, 
#                            gurobi_params = list(TimeLimit = 300),
#                            verbose = TRUE)
 
#saveRDS(result, "result_lp_v11.rds")

result <- readRDS("result_lp_v11.rds")

ext <- result$solution_vector

#  renaming

ext$Xjk <- ext$Xjk %>% 
  rename(src_product = p, del_product = q, period = t, hub_id = j, consumer_id = k)

ext$Xij <- ext$Xij %>% 
  rename(period = t, site_id = i, hub_id = j)

ext$z <- ext$z %>% 
  rename(arc_type = type, site_id = i )

ext$S <- ext$S %>% 
  rename(hub_id = j, period = t)

# sensitivity analysis ##########################################
#
# scen.mat <- expand.grid(opp = c(150, 300, 600), cost.log = c(.5, 1.5, 2), cost.est = c(.5, 1.5, 2), revenue = c(.5, .75, 1.25)  )
# nb.rep <- 10
# 
# res.sens.list <- vector("list", nrow(scen.mat)*nb.rep)
# 
# for(i in seq_len(nrow(scen.mat))){
#   tmp.setting <- list(
#     opp = scen.mat$opp[i],
#     cost.log = scen.mat$cost.log[i],
#     cost.est = scen.mat$cost.est[i],
#     revenue = scen.mat$revenue[i]
#     )
# 
#   # storage cost
#   tmp.storages <- storages
#   tmp.storages$c_stor <- tmp.storages$c_stor * tmp.setting$cost.log
#   # consumers
#   tmp.consumers <- consumers
#   tmp.consumers$P1 <- tmp.consumers$P1 * tmp.setting$revenue
#   tmp.consumers$P2 <- tmp.consumers$P2 * tmp.setting$revenue
#   tmp.consumers$P3 <- tmp.consumers$P3 * tmp.setting$revenue
# 
#   # # ── 3. PARAMS OBJECT — all scalar parameters ──────────────────────────────────
#   tmp.params_obj <- list(
#        n_periods  = 40,
#        max_age    = 20L,    # A_max: LRAFS upper rotation bound
#        min_age    =  3L,    # A_min: SRAFS lower rotation bound
#        c_tr_raw   = 0.08 * tmp.setting$cost.log,   # €/(t·km)  raw biomass  site → storage
#        c_tr_pre   = 0.06 * tmp.setting$cost.log   # €/(t·km)  chips        storage → consumer
#      )
#     #
# 
#   for(j in seq_len(nb.rep)){
#     tmp.c.opp.vec <- rnorm(nrow(sites_clustered), 1, .3)
# 
#     # opportunity cost
#     tmp.c.opp.vec <- tmp.c.opp.vec * tmp.setting$opp
# 
#     # site details
#     tmp.sites_clustered <- sites_clustered %>%
#        mutate(
#          C_est  = 2500 * tmp.setting$cost.est,   # €/ha  establishment
#          C_harv = 150 * tmp.setting$cost.log,    # €/ha  harvest cost
#          C_main = 10 * tmp.setting$opp/150, #  €/ha  maintenance cost
#          C_opp  = tmp.c.opp.vec #  €/ha  opportunity cost
#        )
#     #
#     tmp.data_obj <- list(
#        sites    = tmp.sites_clustered,
#        storages = tmp.storages,
#        consumers = tmp.consumers,
#        dist_ij  = afs_workspace$dist_ij_clust,
#        dist_jk  = dist_jk,
#        yields_by_age = yields_by_age
#      )
# 
#     # rescale demands in consumers to tons
#     tmp.data_obj$consumers[,c("demand_P1","demand_P2","demand_P3")]  <- tmp.data_obj$consumers[,c("demand_P1","demand_P2","demand_P3")] * 1000
# 
#     # ── 4. generate instance ─────────────────────────────────────
#     tmp.milp_instance <- build_optimization_instance(
#         data         = tmp.data_obj,
#         params       = tmp.params_obj
#       )
#     #
#     tmp.result <- build_and_solve_afs_lp_v11(
#        tmp.milp_instance,
#        gurobi_params = list(
#          TimeLimit = 600),
#        verbose = TRUE)
#     #
# 
#     tmp.ext <- tmp.result$solution_vector
#     #  renaming
# 
#     tmp.ext$Xjk <- tmp.ext$Xjk %>%
#       rename(src_product = p, del_product = q, period = t, hub_id = j, consumer_id = k)
# 
#     tmp.ext$Xij <- tmp.ext$Xij %>%
#       rename(period = t, site_id = i, hub_id = j)
# 
#     tmp.ext$z <- tmp.ext$z %>%
#       rename(arc_type = type, site_id = i )
# 
#     tmp.ext$S <- tmp.ext$S %>%
#       rename(hub_id = j, period = t)
# 
#     res.sens.list[[(i-1)*nb.rep + j]] <- list(setting = tmp.setting, milp_instance = tmp.milp_instance, ext = tmp.ext)
#   }
# 
# }
# 
# saveRDS(res.sens.list, "res_sens_list_lp_v11.rds")

res.sens.list <- readRDS("res_sens_list_lp_v11.rds")

# ── 2. Benchmark + alle Sensitivitätsszenarien zusammenführen ────────────────
res_bench_full <- list(
  setting       = list(opp = 500, cost.log = 1.0, cost.est = 1.0, revenue = 1.0),
  milp_instance = milp_instance,
  ext           = ext,
  result        = result
)

df_bench <- extract_site_profit(res_bench_full, "Bench.")

df_bench <- df_bench %>% mutate(opp_cost = opp_cost/500)

df_sens <- purrr::imap_dfr(
  res.sens.list,
  ~ extract_site_profit(.x, scenario_name = paste0("S", .y))
)

df_sens <- df_sens %>% mutate(opp_cost = opp_cost/500)

```

# Introduction
 
The industrial bio-economy depends on connecting primary biological production systems to multiple industrial sectors through networks that efficiently allocate biomass fractions of varying quality to their highest-value application [@Hildebrandt2019]. Agroforestry systems (AFS) that deliberately integrate fast-growing woody perennials into agricultural land represent realisation of this principle. By co-producing heterogeneous biomass fractions like stems, branches, twigs, leaves etc. which are more or less suitable for chemical, paper-and-pulp industry, as well as energy generation, AFS create diverse feedstocks that can serve multiple industrial consumer sectors through a shared regional supply network [@Toensmeier2017_Carbon_Agroforestry,@trujillo2026agribusiness]. This biological-industrial symbiosis substitutes fossil feedstocks with renewable bio-based alternatives and stores biogenic carbon in long-lived material products [@Millard2011,@montagnini2004carbon].

Despite their ecological and industrial potentials, temperate AFS occupy only a marginal share of agricultural land in practice. In Germany, approximately 1,200 hectares of silvoarable AFS were established in 2025, accounting for barely 0.02% of total agricultural land @DEFAF. This gap between supposed potentials and observed practices indicates implementation challenges which often root in economic, organizational, and regulatory obstacles. Establishing AFS at regional scale requires coordinated decisions on stand establishment, harvest scheduling, and biomass allocation across spatially distributed industrial consumers @Low2023mixed. E.g. farmers will only invest in AFS when they are convinced to generate revenues from selling wood to industrial customers at the time of harvest which is, however, 5-15 years after establishing an AFS. This long time span imposes a severe investment risk as investment cost are clear and immediate but revenue is uncertain and far away @Millard2011. On the other side, industrial consumers particularly in process industries like chemical and paper-and-pulp industry prefer reliable and high-volume biomass supply contracts which imposes a challenge for farmers as AFS growth depends on weather as do harvest decisions [REF]. Next to harvest scheduling, a lot of other business decisions need to be made. Early harvests promise earlier revenues, but harvested biomass is dominated by small-scaled branches, twigs, and residues with limited usability in downstream industries and lower value [REFS]. At longer rotations, biomass quantity is higher and mostly composed of stems whose diameters surpass quality thresholds such tha the share of chemical- and paper-grade material rises markedly [@Huber2018_SRAFS_growth; @Niemczyk2021_Poplar_Rotation]. The biomass feedstock quantity and quality in an AFS-based supply chain therefore depend on the rotation strategies and the industrial demand structures of multiple farms and industrial consumers. Current biomass supply chain models, reviewed in \autoref{sec:lit-sc}, often do not consider this aspect as biomass availability is usually given at exogenous time series or optimise a single downstream supply chain rather [@Zahraee2020_BSC_review; @DeMeyer2016_tOPTIMASS].

This paper closes this gap by formulating an AFS supply chain design (\problem) problem. The linear programm integrates design and operation decisions of biomass production in AFS with multiple downstream industries via logistical processes. This way, interdependencies between biomass production decisions in AFS, logistical handling decisions, and biomass allocation decisions in down-stream processes are modeled explicitly. Biomass production decisions are based on growth models which estimate the quantity of biomass types in AFS. Thereby, biomass types are defined based on its suitability for downstream processing. Three different grades are distinguished: chemical/biorefinery grade, pulp and paper grade, and
energy grade. Thereby, a quality cascade is assumed. I.e., biorefinery-grade biomass can also be used in  pulp-and-paper as well as energy applications, but the reverse doesn't hold.  Depending on stand age,  quantities of these grades are available in AFS which guide harvest decisions. Once harvests are scheduled, biomass flows to pre-treatment and storage hubs are planned where biomass grades are allocated and transshipments to consumers are prepared depending on demand quantity and the consumers' willingness-to-pay. This way, the proposed \problemm configures a supply chain for a regional AFS-based industrial ecosystem. It helps to study effects of ecological, regulatory, and economic parameters on the profitability of AFS-based supply chains. Thereby, it integrates relevant decisions of different actors from various industries all relevant in AFS-based supply chains including farmers and land-owners, wood logistics, as well as industrial down-stream sectors (like chemicals, paper or energy).    

The remainder of the paper is structured as follows. Section \ref{sec:litreview} reviews the relevant literature on AFS, biomass growth modelling, and biomass supply chain design. Section \ref{sec:milp} presents the mathematical formulation of the AFS-SCD. Section \ref{sec-casestudy} describes a case study illustrating the application of the \problemm to a real-world problem. Section \ref{sec-conclusion} summarizes the findings of the paper and discusses directions for future research.

# Literature review {#sec:litreview}

Agroforestry biomass systems, stand-level growth models, and biomass supply chain optimisation have each been studied extensively, but they are rarely combined in an integrated framework that links age-dependent stand dynamics to regional product-cascading value chains. This section reviews literature on (i) agroforestry systems and economics, (ii) biomass growth models for AFS considering multiple biomass grades, and (iii) biomass supply chain design problems. 

## Agroforestry systems {#sec:lit-afs}

Agroforestry deliberately integrates woody perennials with crops and/or livestock on the same management unit, with the aim of enhancing productivity, profitability, and environmental performance [@Nair1993_IntroAgroforestry; @Toensmeier2017_Carbon_Agroforestry]. In temperate climate, three distinct system types can be distinguished based on the spatial arrangement of trees and the rotation length at which woody biomass is harvested.

**Short-rotation coppice (SRC)** is defined as a monoculture plantation of fast-growing woody species covering the entire management unit, with planting densities of 8,000--20,000 trees per hectare. After each harvest, the root system regenerates without replanting. Rotation cycles
are typically 3--7 years and total stand lifetime is 20--25 years [@Trnka2008_Poplar_SRF; @Faasch2012_SRC_Economics]. SRCs are no AFS in the narrow sense, but it is helpful to benchmark biomass production.

**Short-rotation agroforestry systems (SRAFS)** integrate SRC-type tree strips (typically 1--4 rows of the same fast-growing species like poplar or willow) into an otherwise arable field in an alley-cropping arrangement. Crop production continues in the inter-row strips [@Huber2018_SRAFS_growth]. Thus, tree strips typically cover 10--30% of the total parcel area [@Graves2007_Agroforestry_Economics]. Rotation lengths and total stand lifetime in SRAFS are comparable to SRCs. In SRAFS most trees are effectively edge trees, exposed to higher light and water availability than interior SRC trees. This boundary effect raises individual-tree biomass accumulation but also intensifies competition with adjacent crop rows [@Huber2018_SRAFS_growth].

**Long-rotation agroforestry systems (LRAFS)** use the same alley-cropping spatial structure as SRAFS but with rotation cycles of 12--15 years or longer [@Graves2007_Agroforestry_Economics;
@Nair1993_IntroAgroforestry]. At this rotation length, individual trees develop substantial stem diameters and the proportion of high-value timber increases markedly, enabling premium material uses (veneer, sawlogs, engineered wood products). Therefore, also timber-quality tree species (such as walnut, cherry, oak, or high-grade poplar clones) are planted [REFS].

SRAFS and LRAFS are not clearly differentiated in practice. First, a clear cut-off age is hard to define and, more importantly, the intended use of AFS biomass is not always clear when establishing the AFS and can also change over time. Nonetheless, when establishing AFS, the structure of AFS and its intended use after harvest affect economic viability substantially. Thus, in the present paper SRAFS and LRAFS are not considered as separate AFS systems, but the interplay of intended use, establishment, and harvesting decisions is analysed.

The decision of a landowner to establish an AFS involves weighing substantial up-front costs and long-term opportunity costs against a diversified revenue stream. On the cost side, establishment requires site preparation, planting material, and planting labour, typically
amounting to 1,500--3,500 \euro{}/ha for SRC or SRAFS [@Faasch2012_SRC_Economics; @Testa2014_Poplar_Economics]. Recurrent harvest and refitting costs (machine hire, chipping, transport to roadside) add approximately 200--400 \euro{}/ha per rotation depending on stand density and equipment availability [@Testa2014_Poplar_Economics;@Spinelli2009_SRC_Harvesting]. During the rotation, maintenance costs of AFS arise on a yearly basis. Next to expenses for establishment, maintenance and harvesting, opportunity costs arise as no revenues from crops are generated on tree strips. In regions with fertile soils and high cereal prices, yearly crop revenues may sum up to 400--600 \euro{}/ha [@Faasch2012_SRC_Economics; @Graves2007_Agroforestry_Economics]. Thus, opportunity costs are often the dominant cost driver in SRAFS profitability analyses.

On the benefit side, the primary revenue of AFS is the sale of harvested woody biomass, whose magnitude and frequency depends on rotation length, species, and site productivity. For poplar-based SRAFS in Central Europe, typical yields range from 10 to 25 t of dry matter biomass per ha for typical rotation lengths [@Huber2018_SRAFS_growth; @Niemczyk2021_Poplar_Rotation]. Additionally, positive crop-yield effects in the inter-row strips of SRAFS and LRAFS may result. Tree rows act as windbreaks, reducing soil erosion and evapotranspiration. They provide partial shading that moderates heat stress during drought years and their root systems may improve water retention in top-level soil [@Graves2007_Agroforestry_Economics; @Grunewald2007_Robinia_AFS]. Empirical measurements in Central European alley-cropping trials indicate that these microclimate effects can increase cereal yields in adjacent strips by 5--15% relative to open-field controls, partially or fully compensating the loss of arable area occupied by the tree strips [@Graves2007_Agroforestry_Economics]. Beyond direct yield effects, AFS deliver broader ecosystem services (like soil organic matter accumulation, carbon sequestration, biodiversity enhancement) that may lead to agri-environment subsidies and carbon credits which have to be considered as additional financial benefits [@Toensmeier2017_Carbon_Agroforestry; @Faasch2012_SRC_Economics].

Nonetheless, economic attractiveness of AFS depends primarily on revenues of AFS biomass. Thereby, revenues are different for the types of biomass and in which value chain it is used. Often AFS biomasses are used energetically, i.e. for heat and power generation, or anaerobic
digestion, which typically creates low prices per tonne of biomass [REF]. Alternatively, woody parts of the biomass (i.e., the stem and large branches) can also be used as pulp and paper grade material as well as feedstock for chemical industry or as timber achieving higher prices
[@Lamers2015_Feedstock_Design] [EXTEND REFS + Explain Forestry Product Classification]. Thus, fostering AFS establishment by maximizing economic returns from AFS biomass requires a strategy for allocating biomass fractions to the highest-value applications, while ensuring that the rotation length is aligned with the growth dynamics of the stand and the market demand for different biomass types [@Lamers2015_Feedstock_Design].

Thus, for making educated decisions on AFS establishment and harvesting, the expected biomass quantities at harvest for stem, branches, and residues need to be estimated. Thereby, the shares of the biomass types change with stand age. E.g.\~the share of stem weight rises from roughly 55% at age 4 to over 75% at age 10 [@Huber2018_SRAFS_growth; @Testa2014_Poplar_Economics]. Therefore, in \autoref{sec:lit-growth} growth modelling approaches are reviewed that capture the age-dependent growth dynamics of different biomass types in AFS.

## Biomass growth models {#sec:lit-growth}

Reliable estimation of stand-level biomass quantities and their age-dependent composition is a necessary prerequisite for strategic analysis of AFS-based supply chains. Individual-tree biomass is commonly approximated by species-specific allometric power laws relating stem base diameter to tree biomass. Such relationships have been calibrated e.g. for Central European poplar clones under SRC and SRAFS conditions [@Huber2016_Allometric_SRAFS; @Huber2018_SRAFS_growth]. For supply chain planning, however, the relevant quantity is stand-level biomass per hectare over a multi-decade rotation horizon, which requires a temporal growth model. In the following, temporal biomass growth models for different biomass fractions are derived. Therefore, a set of variables is used which are summarized in \autoref{tab:tab-notation-growth} in the appendix.

Stand-level biomass per hectare is denoted by $M(t)$. Thereby, stand-level biomass growth can be described by different growth functions like the logistic, Gompertz, von Bertallanfy, and Richards models which are also applied for a multitude of other growth processes in biology [@tjorve2010unified,@vanclay1994modelling,@garcia1993stand]. All these functions show a right-skewed sigmoidal pattern and their shape is controlled by various parameters [@tjorve2010unified]. Thereby more parameters usually imply a more flexible growth model, but more parameters require more calibration efforts which have to be counterweighted with modelling accuracy [@dieguez2006dynamic,@Fekedulegn1999_GrowthModels]. Empirical studies indicate that most growth models can be calibrated to empirical observations quite accurately [@BarrioAnta2008_Poplar_Growth,@Paula2013]. Therefore, in the following, a standard 3-parameter Gompertz model [@Tjorve2017] is chosen for modelling biomass quantity, which is defined as
\begin{equation}
  M(t) \;=\; A(N) \cdot \exp\!\bigl(-\exp(-k \cdot (t - t_0))\bigr),
  \label{eq:gomp-stand}
\end{equation} 
where $t_0$ is the age of maximum instantaneous biomass increment and $k$ determines how quickly the stand approaches its site-specific asymptote $A(N)$, i.e. the maximum stand-level biomass quantity for $t \rightarrow \infty$. For Central European poplar under medium-rotation AFS management (10--20 yr), analysing empirical data from @Huber2018_SRAFS_growth,@Niemczyk2021_Poplar_Rotation, and @Jha2018_Poplar_AFS_Biomass yields parameter intervals of $k \in [0.17,0.2]$ and $t_0 \in [9,10]$ years, depending on local specifications of the AFS under study. Parameter $t_0$ is empirically estimated by analyzing current annual increments (CAI) and indicates an orientation for the optimal rotation age as growth slows for higher ages. 

The asymptote $A(N)$ biomass quantity depends on multiple factors, particularly on planting density and soil quality. It can be modelled by a competition-density power law
[@Shinozaki1961_CD_Effect]: 
\begin{equation}
  A(N) \;=\; C_{\mathrm{site}} \cdot N^{1-\beta},
  \label{eq:A-density}
\end{equation} 
where $C_{\mathrm{site}}$ encodes site quality indicator, $N$ indiactes tree density (e.g. trees per hectar), and $\beta$ is the competition exponent. Equivalently, the per-tree asymptote $A_{\mathrm{tree}}(N) = A(N)/N = C_{\mathrm{site}}\cdot N^{-\beta}$ decreases with density as individual trees have access to less growing space and more nutrition competition. Parameters $C_{\mathrm{site}}$ and $\beta$ are highly site- and specicies-specific. For Poplar clones under European conditions,  $\beta \in [0.3,0.5]$ and $C_{\mathrm{site}} \in [4,6.5]$ [@Niemczyk2021_Poplar_Rotation]. \autoref{fig:growth-asymptote-functions} in the appendix shows both asymptote functions according for advantageous and disadvantageous site conditions.

When assessing potential downstream application of AFS biomass, biomass fractions are to be defined meeting the requirements of downstream consumers [@Petras2008Assortment,@Spinelli2011WoodChips,@Civitarese2019_Poplar_Harvest]. E.g. for timber and biochemical applications, stem diameters of AFS trees need to exceed a minimum threshold in order to qualify as a suitable feedstock [@Zhu2011WoodSizeReduction]. In the following, three biomass fractions are distinguished based on common forestry standards: stems, branch wood, and residue. The stem fraction consists of stems with a diameter of at least 15 cm, while branch wood requires a diameter of at least 7 cm. The residue fraction comprises are smaler parts of the tree. To subdivide the total AFS biomass modelled by \eqref{eq:gomp-stand} into the three fractions depending on tree age, a logistic growth model [@Zeide1993_AnalysisGrowth] is used: 
\begin{equation}
  q_p(t) \;=\; \frac{f_p}{1 + e^{-r_p \cdot (t - t_{50,p})}},
  \label{eq:logistic-merch}
\end{equation} 
where $f_p$ denotes the asymptotic share of fraction $p$ for mature trees. Parameter $t_{50,p}$ indicates the infliction point and conincides with the age at which 50% of the asymptotic share of fraction $p$ is reached. Using empirical data from @Civitarese2019_Poplar_Harvest, @Jha2018_Poplar_AFS_Biomass, and @Niemczyk2021_Poplar_Rotation, $r_{stem} = 0.467$,yr$^{-1}$, $t_{50,stem} = 8.19$ years, and $f_{stem} = 0.75$ is estimated for the stem fraction. For the branch fraction, parameters $r_{bran} = 0.698$ yr$^{-1}$, $t_{50,bran} = 7.55$,yr, and $f_{bran} = 0.15$ can be assumed. \autoref{fig:afs-model-functions} illustrates the development of the fraction shares over stand age. The share of the residue fraction is implicitly defined as $q_{resid}(t) = 1 - q_{stem}(t) - q_{branch}(t)$. 

Combining equations \eqref{eq:gomp-stand}--\eqref{eq:logistic-merch}, the dry biomass yield of fraction $p$ per hectare at harvest age $t$ is: 
\begin{equation}
  \eta_p(t) \;=\; q_p(t)\cdot M(t).
  \label{eq:yield}
\end{equation} 
Fresh-weight yields at felling are obtained from dry matter yields via $\eta_p^{\mathrm{fw}}(t) = \eta_p(t)/(1-\omega)$, where the moisture content at felling $\omega = 0.55$ [@Civitarese2019_Poplar_Harvest]. \autoref{fig:afs-growth-curves} shows the fractions' stacked biomass quantities $\eta_p(t)$  depending on stand age compared with empirical observations from @Jha2018_Poplar_AFS_Biomass and @Huber2018_SRAFS_growth.

```{r afs-growth-curves, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Stand-level aboveground biomass components for poplar under AFS management with $N=250$ trees\\, ha$^{.1}$.  Grey band: typical rotation times (5-15 years).\\label{fig:afs-growth-curves}", fig.pos="htp",  cache=T,fig.width=7, fig.height=3, dev="pdf", out.width="\\textwidth"}

#Empirical AGB observations compiled from \\citet{Jha2018_Poplar_AFS_Biomass}, \\citet{Civitarese2019_Poplar_Harvest}, and \\citet{Niemczyk2021_Poplar_Rotation}.

age.vec <- seq(0, 20, by = 0.1)

cons <- build_scenario_ts(ages = age.vec , N = 300, C_site = 6.500, k = 0.194, t0 = 9.7, label = "Conservative (250 trees/ha)")

comp_long <- bind_rows(cons) |>
  rename(stem = "Merch. stem", branch = "Merch. branch", residue = Residue) %>% 
  pivot_longer(c("stem", "branch", "residue"),
               names_to = "component", values_to = "biomass") |>
  mutate(
    component = factor(component,
      levels = c("residue", "branch", "stem")),
    scenario  = factor(scenario,
      levels = c("Conservative (113 trees/ha)", "Good site (139 trees/ha)"))
  ) %>% 
  select(-scenario)

# totals_long <- comp_long |>
#   group_by(age, scenario) |>
#   summarise(total = sum(biomass), .groups = "drop")

emp_pts <- tribble(
  ~age, ~AGB,   ~p_stem, ~p_branch,
   8,   134.3,  0.747,   0.166, # N = 400
   9,    74.0,  0.830,   0.110, # N = 500
   9,    51.5,  0.920,   0.040, # N = 500
  10,   119,    0.76,   0.210,# N = 139
  13,   152.0,  0.770,   0.220 # N = 139
) |>
  mutate(stem_cum   = AGB * p_stem,
         branch_cum = AGB * (p_stem + p_branch),
         total      = AGB)

emp_long <- bind_rows(
  emp_pts |> transmute(age, biomass = stem_cum,   series = "Empirical: stem"),
  emp_pts |> transmute(age, biomass = branch_cum, series = "Empirical: stem+branches"),
  emp_pts |> transmute(age, biomass = total,      series = "Empirical: total biomass")
) |> mutate(series = factor(series, levels = c(
  "Empirical: stem", "Empirical: stem+branches", "Empirical: total biomass")))

emp_col   <- c("Empirical: stem"          = "darkgreen",
               "Empirical: stem+branches" = "darkorange",
               "Empirical: total biomass"     = "navy")
emp_shape <- c("Empirical: stem"          = 21L,
               "Empirical: stem+branches" = 24L,
               "Empirical: total biomass"     = 23L)

ggplot(comp_long, aes(x = age, y = biomass)) +
  annotate("rect", xmin = 5, xmax = 15, ymin = 0, ymax = Inf,
           fill = "grey80", alpha = 0.25) +
  geom_area(aes(fill = component),  #, linetype = scenario
            position = "stack", alpha = 0.70, colour = "white", linewidth = 0.2) +
  # geom_line(data    = totals_long,
  #           aes(x = age, y = total, linetype = scenario),
  #           colour  = "grey30", linewidth = 0.45, inherit.aes = FALSE) +
  geom_point(data = emp_long,
             aes(x = age, y = biomass,
                 colour = series, shape = series, fill = series),
             size = 2.5, alpha = 0.90, inherit.aes = FALSE) +
  scale_fill_manual(
    name   = "Biomass fraction",
    values = c("stem"   = "#1b9e77",
               "branch" = "#d95f02",
               "residue"       = "#7570b3"),
    breaks = c("stem", "branch", "residue")) +
  scale_colour_manual(name = "Empirical data", values = emp_col) +
  scale_shape_manual (name = "Empirical data", values = emp_shape) +
  # scale_linetype_manual(
  #   name   = "Scenario",
  #   values = c("Conservative (113 trees/ha)" = "solid",
  #              "Good site (139 trees/ha)"    = "dashed")) +
  guides(
    fill     = guide_legend(order = 1,
                 override.aes = list(shape = NA, colour = NA, linetype = 0)),
    colour   = guide_legend(order = 3,
                 override.aes = list(
                   shape  = emp_shape[levels(emp_long$series)],
                   size   = 2.5,
                   fill   = emp_col[levels(emp_long$series)])),
    linetype = guide_legend(order = 2,
                 override.aes = list(fill = NA, colour = "grey30")),
    shape    = "none") +
  scale_x_continuous(breaks = seq(0, 20, 2)) +
  labs(x = "Stand age (years)",
       y = expression("biomass (t DM "*ha^{-1}*")")) +
  theme_minimal(base_size = 9) +
  theme(legend.position  = "right",
        legend.key.size  = unit(0.4, "cm"),
        legend.text      = element_text(size = 7),
        legend.title     = element_text(size = 8, face = "bold"),
        panel.grid.minor = element_blank())
```

## Biomass supply chain design {#sec:lit-sc}

The design and planning of supply chains based on industrial biomass feedstocks has attracted substantial research attention over the past decades, driven by the need to ensure cost-efficient and sustainable feedstock provision for bioenergy and biorefinery applications (see e.g. extensive reviews from @Gital2024BiomassSupplyChain, @Zahraee2020_BSC_review, and @DeMeyer2014BiomassSupplyChains). Despite its name, the term biomass supply chain (BSC) is often associated with exclusive energetic uses of biomass and often equivalent to biorefinery supply chain (see e.g. @Espinoza2017_Biorefinery_SC_review). The considered biomass feedstocks range from agricultural energy crops, over urban biomass residues to forestry wood (see e.g. @Sharma2013_BSC_review for an overview). Thereby, literature on forestry supply chains (FSC) is the oldest sub-branch in this literature stream @Cambero2014_ForestBSC_Review. Here, the focus is on one particular biomass feedstock, but usually multiple fields of applications are considered (see e.g. @DAmours2008ORForestSupplyChain for an overview of fields of applications). In general, the structure of BSC and FSC typically encompasses three process types: i) biomass production, ii) harvesting and logistics, and iii) biomass processing. These processes are modeled at various levels of details depending on the scope of decisions to be supported. The decision scopes can be distinguished in: (i) strategic network design (e.g. facility location, tree selection), (ii) tactical planning (e.g. harvest scheduling, inventory capacities), and (iii) operations (e.g. routing of vehicles, inventory control) @DeMeyer2014BiomassSupplyChains. This manuscript is dedicated to agroforestry supply chain design. I.e. the focus is put on a specific type of biomass and multiple fields of application. In the following relevant literature from FSC and BSC literature is reviewed which contributes to this type of problem. Relevant aspects from BSC literature mainly involve biomass production economics and strategic design modelling while relevant aspects from FSC literature is concentrated on wood logistics  and industrial applications. Note that there is no existing literature dedicated on agroforestry supply chain design to the best of the author's knowledge.

One of the earliest FSC models is proposed by @Gunnarsson2004. Here, a cost-optimal distribution and processing network for supplying heating plants with wood chips is to be found. Therefore, a multi-period MILP model is formulated satisfying feedstock demand at heating plants by selecting harvesting areas and deciding in each period (i.e., each year) about chipping, storing, and transporting forest biomass (like logs). @Eksioglu2009_Biorefinery_SC develop one of the earliest MILP models for biomass-to-biorefinery supply chain design. The focus is on biomass conversion processes by integrating facility location, capacity sizing, and transport mode selection  for lignocellulosic feedstocks. The model minimises total system cost assuming, that biomass supply is an exogenous input, i.e. biomass production is not modelled explicitely. @DeMeyer2016_tOPTIMASS proposes a similar profit-maximising MILP for strategic and tactical decision support. In this model, the focus is on pre-treatment and storage processes of different types of grassland biomass types (grass, hay, straw, etc.) supplying anaerobic digesters for energy generation. @walther2012design describe wood-based biomass supply chain for biofuel production where the biomass is converted either directly into biofuels or indirectly via an intermediate material (slurry). The objective is here maximize the net present value (NPV) over the planning horizon discounting the yearly cash flows of the supply chain. The revenue is created by supplying refineries with the supply chain's biofuels. Similar to @Gunnarson2004, the refineries' biofuel demand can be satisfied by imports, such that a possible solution is that no regional biomass is used to produce biofuel. The aforementioned models consider only biomass-to-energy supply chains.

@Mansoornejad2013_Forest_Biorefinery_SC extend this perspective by integrating product and process portfolios. Their model considers multiple biomass conversion processes as well as sources of lignocellulosic biomass. It accounts for product conversion efficiencies along different processing pathways and maximises profit over a multi-period time horizon. Likewise, @cambero2015strategic model energetic and chemical conversion pathways of lignocellulosic biomass in a strategic supply chain design MILP for maximizing NPV over a 20-year period. @GarciaVelasquez2022 propose a bi-objective MILP for designing a supply chain for producing Polyethylene Terephthalate (PET) based on wheat and sugar beets. Thereby, both biomass types are required to produce intermediate chemicals (mono ethylene glycol and purified terephthalic acid) which are merged into PET such that a 4-stage supply chain is modeled. For each processing step locations have to be selected among a st of candidate locations. The objectives pursued involve total supply chain  costs and GHG emissions. The trade-off between cost and GHG emissions is studied by incorporating GHG cost and varying GHG prices. A similar 4-stage bio-chemical network design problem for biobutanol production based on microalgae is studied in @Arabi2019_Microalgae_Biobutanol_SC. Recently, @Roudneshin2025 presents a novel procedure for a bioenergy supply chain design problem by combining mathematical programming with an Analytic Hierarchy Process (AHP). The AHP pre-processes potential candidate locations for biorefineries where biomass wastes are converted into biogas which is subsequently fed into the local gas grid. 

All BSC and FSC literature reviewed above does not consider biomass production processes explicitly, but model biomass as an exogenous resource. This is because most approaches concentrate on application where waste biomass is considered as feedstock. Also in FSC applications production decisions are usually not part of supply chain design decisions as growing cycles in forest management are even longer than the time horizons in supply chain design problems. This manuscript contributes to literature by proposing a novel agroforestry suppply chain design problem (\problem) which explicitly models biomass production is an integral part of the decision support. In agroforestry applications, biomass production decisions depend on the existence of downstream markets for the produced  wood. Moreover, multiple types of biomass and multiple application industries are considered. In the follwoing the \problemm is is outlined in detail. 

<!-- \autoref{tab:lit-sc-features} summarises the key modelling features -->
<!-- of the reviewed contributions. -->

<!-- ```{r lit-sc-features, echo=FALSE, message=FALSE, warning=FALSE} -->
<!-- library(knitr) -->
<!-- library(kableExtra) -->
<!-- options(kableExtra.latex.load_packages = FALSE) -->
<!-- options(knitr.table.format = "latex") -->
<!-- knitr::opts_chunk$set(dev = "pdf") -->

<!-- lit_tab <- tribble( -->
<!--   ~Study,                            ~Objective,          ~`Multi-product`,  ~`Quality cascade`, ~`Age dynamics`,    ~`Uncertainty`,    ~`Level`, -->
<!--   "Eksioglu et al. (2009)",          "Cost min.",         "No",              "No",               "No",               "No",              "Strategic", -->
<!--   "Sharma \\& Taaffe (2013)",        "Cost min.",         "Partial",         "No",               "No",               "No",              "Strategic/Tactical", -->
<!--   "Mansoornejad et al. (2013)",      "Profit max.",       "Yes",             "No",               "No",               "No",              "Strategic", -->
<!--   "Cambero \\& Sowlati (2014)",      "Cost/GHG/jobs",     "No",              "No",               "No",               "Stochastic",      "Strategic", -->
<!--   "Yue et al. (2014)",               "Cost/GHG",          "Yes",             "No",               "No",               "No",              "Multi-scale", -->
<!--   "De Meyer et al. (2015)",          "Cost min.",         "Yes",             "Yes",              "Exogenous",        "No",              "Strategic/Tactical", -->
<!--   "De Meyer et al. (2016)",          "Cost min.",         "Yes",             "Yes",              "Exogenous",     "No",              "Strategic", -->
<!--   "Lamers et al. (2015)",            "Cost/quality",      "Yes",             "Partial",          "No",               "Scenario",        "Strategic", -->
<!--   "Espinoza et al. (2017)",          "Sustainability",    "Yes",             "No",               "No",               "Scenario",        "Strategic", -->
<!--   "Malladi \\& Sowlati (2018)",      "Cost min.",         "Partial",         "No",               "No",               "No",              "Tactical/Oper.", -->
<!--   "Arabi et al. (2019)",             "Profit max.",       "Yes",             "No",               "Exogenous",        "No",              "Multi-period", -->
<!--   "Zahraee et al. (2020)",           "Multi-obj.",        "Yes",             "No",               "No",               "Stochastic",      "Strategic/Tactical", -->
<!--   "\\textbf{This paper}",            "Profit max.",          "Yes",             "Yes",              "\\textbf{Endogenous}", "No",   "Strategic" -->
<!-- ) |> -->
<!--   mutate(across(everything(), ~ gsub("Yes", "\\\\checkmark", .))) |> -->
<!--   mutate(across(everything(), ~ gsub("No",  "---",            .))) -->

<!-- kable(lit_tab, -->
<!--       format    = "latex", -->
<!--       booktabs  = TRUE, -->
<!--       escape    = FALSE, -->
<!--       caption   = "Key modelling features of selected biomass supply chain design studies.", -->
<!--       label     = "lit-sc-features") |> -->
<!--   kable_styling( -->
<!--   font_size     = 18, -->
<!--   latex_options = c("hold_position", "striped"), -->
<!--   full_width    = FALSE -->
<!-- ) |> -->
<!--   row_spec(nrow(lit_tab), bold = TRUE) |> -->
<!--   row_spec(0, bold = TRUE) #|> -->
<!--   #column_spec(1, width = "3.2cm") |> -->
<!--   #column_spec(2, width = "1.6cm") |> -->
<!--   #column_spec(3:7, width = "1.6cm") -->
<!-- ``` -->


# MILP Model for Agroforestry Supply Chain Design {#sec:milp}

Building on the biomass growth models introduced in
\autoref{sec:lit-growth} and the supply chain design literature reviewed in
\autoref{sec:lit-sc}, this section describes the AFS supply chain
design problem (\problem), a mixed-integer linear programming (MILP)
model for the integrated design and operation of an agroforestry biomass
supply chain with multiple AFS biomass types (i.e., stem, branch,
and residue fractions) used as feedstocks for potential consumers in chemical, pulp and
paper as well as energy industry.

## Sets, Indices, and Time Structure

The model considers a set of agroforestry sites $\mathcal{I}$, storage/pre-treatment facilities $\mathcal{J}$, consumer sites $\mathcal{K}$, and biomass types $\mathcal{P} = \{stem, branches, residues\}$, whose growth models are outlined above. Time is discretised into annual periods $t \in \Tc = \{1,\dots,T_{\max}\}$. To reduce complexity, reflect legal definitions of SRAFS and LRAFS as well as discard implausible options (e.g. extremly long rotation cycles), $A^{\min}$ and $A^{\max}$ define minimum and maximum stand ages. Thus, harvest are possible in periods $\mathcal{T}^{\text{harv}} = \{A^{\min}+1,\dots,T_{\max}\}$. To consistently model establishment and consecutive harvest decisions, the following arc sets are defined: $\Sc^{est}$ defines potential establishments of AFS. If an AFS is established in period $t \in \Tc$, arc $(0,t)$ is selected. Harvest set $\Sc^{harv}$ summarizes all potential consecutive harvest periods within the age interval $[A^{\min},A^{\max}]$. I.e., if an AFS is established or harvested in period $s$ and next harvest is in period $t$, arc $(s,t)$ is selected. Finally, if there is no subsequent harvest after period $t$, the AFS is terminated by selecting arc $(t,T^{\max}+1)$. Termination arcs are summarized in set $\Sc^{term}$. All arc sets are formally defined as
follows: \begin{align}
  \Sc^{{est}}  &= \{(0,t) \mid t \in \{1,\dots,T^{\max}-A^{\min}\}\}, \notag\\
  \Sc^{{harv}} &= \{(s,t) \mid s,t \in \Tc \wedge t-s \in \{A^{\min},\dots,A^{\max}\}\},  \notag\\
  \Sc^{{term}} &= \{(s,T^{\max}+1) \mid s \in \{A^{\min}+1,\dots,T_{\max}\}\}, \notag\\
  \Sc               &= \Sc^{{est}} \cup \Sc^{{harv}} \cup \Sc^{{term}}. \notag
\end{align}

\autoref{fig:fig-set-S-tikz} illustrates this network structure of all arc sets for $T=8$, $A^{\min}=3$, $A^{\max}=5$.

```{=latex}
\begin{figure}[htbp]
  \centering

  % ---- Panel (a): Arc/timeline diagram ----
  \begin{subfigure}[b]{\linewidth}
    \centering
    \begin{tikzpicture}[
  x=1.55cm, y=1cm,
  node/.style={circle, draw, fill=#1, text=white,
               font=\bfseries, inner sep=2pt, minimum size=18pt},
  est/.style  ={-{Stealth[length=4pt]}, blue!70!black,  dashed, semithick},
  harva/.style={-{Stealth[length=4pt]}, green!40!black, semithick},
  harvb/.style={-{Stealth[length=4pt]}, green!40!black, semithick},
  term/.style ={-{Stealth[length=4pt]}, red!70!black,   semithick},
]

\node[node=blue!70!black]  (n0) at (0,0) {$0$};
\foreach \i in {1,...,8} {
  \node[node=gray!50!black] (n\i) at (\i,0) {$\i$};
}
\node[node=red!70!black]   (n9) at (9,0) {$9$};

\draw[est] (n0) -- (n1);
\foreach \t/\l in {2/1.3, 3/1.0, 4/0.80, 5/0.65}{
  \draw[est] (n0) to[out=55, in=125, looseness=\l] (n\t);
}

\draw[harva] (n1) to[out=55,  in=125,  looseness=0.55] (n4);
\draw[harvb] (n2) to[out=-55, in=-125, looseness=0.55] (n5);
\draw[harva] (n3) to[out=55,  in=125,  looseness=0.55] (n6);
\draw[harvb] (n4) to[out=-55, in=-125, looseness=0.55] (n7);
\draw[harva] (n5) to[out=55,  in=125,  looseness=0.55] (n8);
\draw[harva] (n1) to[out=60,  in=120,  looseness=0.42] (n5);
\draw[harvb] (n2) to[out=-60, in=-120, looseness=0.42] (n6);
\draw[harva] (n3) to[out=60,  in=120,  looseness=0.42] (n7);
\draw[harvb] (n4) to[out=-60, in=-120, looseness=0.42] (n8);
\draw[harva] (n1) to[out=65,  in=115,  looseness=0.35] (n6);
\draw[harvb] (n2) to[out=-65, in=-115, looseness=0.35] (n7);
\draw[harva] (n3) to[out=65,  in=115,  looseness=0.35] (n8);

\foreach \s/\l in {4/0.80, 5/0.65, 6/0.55, 7/0.48}{
  \draw[term] (n\s) to[out=55, in=125, looseness=\l] (n9);
}
\draw[term] (n8) -- (n9);

\node[anchor=north west] at (0.0,2.5) {%
  \begin{tikzpicture}[x=0.6cm,y=0.35cm,baseline=0pt]
    \draw[est]   (0,0)--(1,0);
    \node[font=\small, anchor=west] at (1.1,0) {Establishment arc $(0,t)$};
  \end{tikzpicture}};
\node[anchor=north west] at (3.5,2.5) {%
  \begin{tikzpicture}[x=0.6cm,y=0.35cm,baseline=0pt]
    \draw[harva] (0,0)--(1,0);
    \node[font=\small, anchor=west] at (1.1,0) {Harvest arc $(s,t)$};
  \end{tikzpicture}};
\node[anchor=north west] at (6.3,2.5) {%
  \begin{tikzpicture}[x=0.6cm,y=0.35cm,baseline=0pt]
    \draw[term]  (0,0)--(1,0);
    \node[font=\small, anchor=west] at (1.1,0) {Termination arc $(s,9)$};
  \end{tikzpicture}};
%\fontsize{10}{12}\selectfont 
\end{tikzpicture}
    \caption{Complete sets of arcs $\Sc^{{est}}$, $\Sc^{{harv}}$, and $\Sc^{{term}}$.}
    \label{fig:fig-set-S-tikz}
  \end{subfigure}

  \vspace{1em}   % vertical gap between panels

  % ---- Panel (b): Supply chain network diagram ----
  \begin{subfigure}[b]{\linewidth}
    \centering
    
    \begin{tikzpicture}[
  x=1.55cm, y=1cm,
  node/.style    ={circle, draw, fill=#1, text=white,
                   font=\bfseries, inner sep=2pt, minimum size=18pt},
  inactive/.style={node=gray!25, text=gray!50},
  pathnode/.style={node=green!40!black},
  n0style/.style ={node=blue!70!black},
  n9style/.style ={node=red!70!black},
  patharc/.style ={-{Stealth[length=5pt]}, green!40!black, very thick},
  est/.style     ={patharc, blue!70!black, dashed},
  term/.style    ={patharc, red!70!black},
]

\node[n0style]   (n0) at (0,0) {$0$};
\foreach \i in {2,3,4,6,7} {
  \node[inactive] (n\i) at (\i,0) {$\i$};
}
\foreach \i in {1,5,8} {
  \node[pathnode]  (n\i) at (\i,0) {$\i$};
}
\node[n9style]   (n9) at (9,0) {$9$};

\draw[est]     (n0) -- (n1);
\draw[patharc] (n1) to[out=60, in=120, looseness=0.42] (n5);
\draw[term] (n5) to[out=-50, in=-130, looseness=0.42] (n9);
\draw[patharc] (n5) to[out=55, in=125, looseness=0.55] (n8);
\draw[term]    (n8) -- (n9);

\node[font=\footnotesize, text=blue!70!black]   at (0.5, 0.28) {est.};
\node[font=\footnotesize, text=blue!70!black]   at (0.5, -0.3) {$100$};
\node[font=\footnotesize, text=green!40!black]  at (3.0, 1.1)  {$100$};
\node[font=\footnotesize, text=green!40!black]  at (3.0, 0.6)  {$t-s=4$};
\node[font=\footnotesize, text=green!40!black]  at (6.5, 1) {$70$};
\node[font=\footnotesize, text=red!70!black]  at (7, -1.0) {$30$};
\node[font=\footnotesize, text=red!70!black]  at (8.5, -0.3) {$70$};
\node[font=\footnotesize, text=green!40!black]  at (6.5, 0.55) {$t-s=3$};
\node[font=\footnotesize, text=red!70!black]    at (8.5, 0.28) {term.};

\end{tikzpicture}
    \caption{Example of a feasible path through the network with establishment in period~1, harvests in periods~5 and~8, termination in period~9.}
    \label{fig:fig-set-S-path}
  \end{subfigure}

  \caption{Network structure of set $\mathcal{S}$ for $T=8$, $A_{\min}=3$, $A_{\max}=5$. Dashed blue arcs: establishment arcs $\Sc^{{est}}$, solid green arcs: harvest arcs $\Sc^{{harv}}$, solid red arcs: termination arcs $\Sc^{{term}}$.}
  \label{fig:fig-set-S}
\end{figure}
```


## Decision Variables

Temporal flow Variables $z_{ist}$ indicate on which quantity of the area of site $i$ an an AFS is operated between $(s,t) \in \Sc$. Depending on the arc type, this implies establishment, harvest, or termination. Selected arcs buiild a network flow over the complete planning horizon from $t=0$ to $T^{max}+1$. \autoref{fig:fig-set-S-path} illustrates a simple network flow with establishment of an AFS in period 1 and harvests in periods 5 and 8. 

Spatial flow variables $X_{ijpt}$ and $X_{jkpp't}$ represent transport of products from AFS sites to storage/pre-treatment facilities and from there to consumers, respectively, while $S_{jpt}$ captures end-of-period inventory at facility $j$. \autoref{tab:tab-notation-milp} in the Appendix summarises all sets, decision variables, and parameters used in the MILP formulation, while $\eta_{pa}$ refers to \eqref{eq:yield}.

## Objective Function

As discussed in Section \autoref{sec:lit-afs}, the economic attractiveness of AFS supply chains is modelled from a global perspective taking into account all processes, activities, and participants involved in finally serving customer markets with AFS-based feedstocks. Therefore, revenues for the different biomass fractions and are weighed against costs occurring at all value-adding stages of the supply chains. Revenues created by selling AFS-based feedstocks to different industrial consumers are counterweighted by establishment, opportunity, harvest, transport, and storage costs of landowners, agricultural firms as well as biomass logistics companies. The model therefore maximizes total profit of all participants in the AFS supply chain over the planning horizon as follows: 
\begin{align}
\max\; Z \;=\;&
  \sum_{k \in \Kc}\sum_{p \in \Pc} R_{kp} \cdot \sum_{j \in \Jc, }\sum_{(p',p) \in \Qc} \sum_{t \in \Tc}  X_{jkp'pt}  \label{eq:obj-revenue} \\
&- \sum_{i \in \Ic} c^{{est}}_i \cdot \sum_{(0,t) \in \Sc^{{est}}} z_{i0t} \notag \\
% &- \sum_{i \in \Ic} c^{{opp}}_i \cdot \left(\sum_{(s,T+1) \in \Sc^{term}} (s+1) \cdot z_{is(T+1)}  -  \sum_{(0,t) \in \Sc^{{est}}}  t \cdot z_{i0t} \right) \notag \\ 
&- \sum_{i \in \Ic} (c^{main}_{i} + c^{{opp}}_i) \cdot (t-s) \cdot \sum_{(s,t) \in \Sc^{{harv}}} z_{ist} \notag \\
&- \sum_{i \in \Ic} c^{{harv}}_i  \cdot \sum_{(s,t) \in \Sc^{{harv}}} z_{ist} \notag \\
&- \sum_{i \in \Ic}\sum_{j \in \Jc}\sum_{p \in \Pc}\sum_{t \in \Tc} c^{\text{tr-raw}}_p \cdot d_{ij} \cdot X_{ijpt} \notag \\
&- \sum_{j \in \Jc}\sum_{k \in \Kc}\sum_{(p,p') \in \Qc} \sum_{t \in \Tc} c^{\text{tr-pre}}_{p'} \cdot d_{jk} \cdot X_{jkpp't} \notag \\
&- \sum_{j \in \Jc}\sum_{p \in \Pc}\sum_{t \in \Tc}  c^{{stor}}_j \cdot S_{jpt}  \notag
\end{align}

Here, $R_{kp}$ are product- and consumer-specific revenues for each ton of biomass fraction (\euro{}{}/t). The opportunity cost term penalises the land use when AFS are established. Thereby, $c^{{opp}}_i$ can be negative or positive depending on the site-specific conditions. It comprises effect of lost field crop revenues due to AFS as well as higher yields of remaining field crop areas e.g. due to wind and water erosion protection by AFS. While the assessment of lost field crop revenues is usually well predictable, the economic evaluation of positive ecological effects of AFS (e.g. on neighboring crop yields) is more difficult. Thus, $c^{{opp}}_i$ can also contain regulatory subsidies to support establishing and maintaining AFS. Transport cost terms mirror the two-stage flow structure of the AFS supply chain with differentiated rates for raw and pre-treated biomass depending on the biomass type. Cost rates differ because as e.g. the density and water content of the shipped biomass types differ or because specialized equipment (e.g. for log transports) is used [REFERENCES]. Moreover, site-specific establishment cost for planting the AFS trees is considered, next to yearly maintenance and harvesting costs. All these costs scale with the size of a site ($AREA_i$) measured in hectar. Maintenance cost comprise e.g. costs for trimming, replanting, and pruning unnecessary shoots. Storage costs are considered if biomass types are stored for more than a year. Storage cost rate $c^{{stor}}_j$ measures the costs for storing a ton of biomass for one year and comprises direct storage cost (e.g. for storage site maintenance) as well as opportunity cost due to potential material losses and capital commitment. Biomass types harvested and consumed in the same year incur no storage cost, as usually AFS are harvested at the beginning of a year. Harvested biomass has to dry during spring and summer such that the biomass can be used in autumn or winter the same year. As under-year drying is necessary for some industrial applications and beneficial for transport economics, drying storage in the year of harvest is not associated with additional cost in this study.

## Constraints

Constraints \eqref{eq:C1} limits the area dedicated to AFS operations to the maximum size of site $i$. Constraints \eqref{eq:C2} enforce flow conservation in each period. If AFS systems of size $z_{ist}$ are established or harvested in period $t \in \Tc$, there needs to be a succeeding activity (harvest or termination) of this area. This way, \eqref{eq:C2} guarantee valid (possibly empty) AFS activity  over the complete time horizon for each site. Constraints \eqref{eq:C3} link the harvest-cycle variables $z_{ist}$ to physical yield quantities $\eta^{fw}_{p(t-s)}$ for age $t-s$, which are pre-computed using \eqref{eq:gomp-stand}--\eqref{eq:yield} scaled by moisture content $\omega$. Thus, constant yields per hectar are assumed for each site $i$. Note that this simplification can be dropped and site-specific yields $\eta^{fw}_{ip(t-s)}$ could be introduced without increasing numerical complexity of the \problem. Moreover, harvest quantities can be smaller than available AFS biomass implying that unharvested biomass quantities are lost. Constraints \eqref{eq:C5} link harvest quantities to transport flows and ensure inventory balance at each facility. Thereby, biomass losses due to drying ($\Delta \leq 1$) and biomass conversion processes ($\delta_{pp'} \geq 1$) are considered. Constraints \eqref{eq:C6}--\eqref{eq:C7} enforce storage and processing capacity limits at each facility $j$. Constraints \eqref{eq:C8} implement product usage at multiple consumption sites $k$. Thereby, $(p',p) \in \Qc$ defines a product mapping allowing high-quality product $p'$ (e.g. stems) satisfying demand of a lower-quality products $p$ up to maximum demand $D^{\max}_{kpt}$.

\begin{align}
  \sum_{(0,t) \in \Sc^{{est}}} z_{i0t} &\leq AREA_i && \forall\; i \in \Ic \label{eq:C1} \\[1pt]
  %
  \sum_{(s,t) \in \Sc} z_{ist} &= \sum_{(t,u) \in \Sc} z_{itu} && \forall\; i \in \Ic,\; t \in \Tc \label{eq:C2} \\[1pt]
  %
  \sum_{j \in \Jc} X_{ijpt}    &\leq \sum_{(s,t) \in \Sc^{{harv}}} \eta^{fw}_{p(t-s)} \cdot z_{ist} && \forall\; i \in \Ic,\; p \in \Pc,\; t \in \Tc^{{harv}}     \label{eq:C3} \\[1pt]
  %
  S_{jpt}    = \Delta \cdot S_{jp,t-1} + &\sum_{i \in \Ic} X_{ijpt} - \sum_{k \in \Kc,\,(p,p') \in \Qc} \delta_{pp'} \cdot  X_{jkpp't}   && \forall\; j \in \Jc,\; p \in \Pc,\; t \in \Tc^{harv}  \label{eq:C5} \\[1pt]
  %
  \sum_{p \in \Pc} S_{jpt}    &\leq \text{CAP}^{{stor}}_j    && \forall\; j \in \Jc,\; t \in \Tc     \label{eq:C6} \\[1pt]
  %
  \sum_{i \in \Ic,\, p \in \Pc} X_{ijpt}  &\leq \text{CAP}^{{proc}}_j && \forall\; j \in \Jc,\; t \in \Tc^{harv} \label{eq:C7} \\[1pt]
  %
  \sum_{j \in \Jc,\,(p',p) \in \Qc} X_{jkp'pt} &\leq D^{\max}_{kpt} && \forall\; k \in \Kc,\; p \in \Pc,\; t \in \Tc^{harv}     \label{eq:C8} \\[1pt]
  %
  z_{ist},\;X_{ijpt},\; &X_{jkp'pt},\; S_{jpt} \geq 0 && \forall\; i,j,k,p,p',t     \label{eq:D2}
  %
\end{align}

In summary, \problemm constitutes are large-scaled LP which allows to decide on which fraction of a site, AFS are established and how the AFS operations are scheduled in a multi-year planning horizon. The \problemm takes a supply chain perspective and maximizes total profit of all actors involved in these processes. Thus, it does not explain how the profits are split e.g. between farmers, AFS service providers and logistics service providers. The continuous modelling of AFS areas in $z_{ist}$ leads to a comparatively easy to solve problem. On the other side, it is unclear which part of a site's total area is chosen for AFS operations if $z_{ist} < AREA_i$, which holds for establishment as well as harvesting and termination.

# Case Study {#sec-casestudy}

The \problemm is intended to support the design of regional agroforestry biomass supply chains fostering regional bioeconomy clusters thereby connecting multiple industrial sectors. The proposed \problemm model is instantiated on a real-world biomass supply chain case study located in the tri-state region of Saxony-Anhalt, northern Thuringia, and northern Saxony, Germany. This region is characterized by a high density of agricultural land usage as well as cluster of paper and wood as well as chemical industry. The regional chemical industry relies on fossil ressources so fa, but is under substantial transformative pressure which led to investments in sustainable production processes requiring sustainable regional bio-based feedstocks. \autoref{fig:fig-supply-chain-network} in the Appendix shows agricultural production sites along with the set of potential wood-processing hubs, and the regional users of AFS biomasses from chemical, paper and wood processing industry as well as the energy sector. 

The supply chain comprises `r nrow(afs_workspace$sites)` agricultural production parcels which are grouped into  $|\mathcal{I}|=$ `r nrow(afs_workspace$sites_clustered)` potential AFS production sites. Clustering is based on geographical distances between parcel centroid using a  k-means approach in order to reduce problem size. For distance calculations, for each site a centroid is calculated. Pre-treatment and intermediate storage facilities $\mathcal{J}$ are located on existing wood processing facilities as well as potential new log or wood chip yards. All harvested biomass quantities at AFS sites are transported to logistics hubs first. At the hubs sorting and transshipment processes take place before biomass types are forwarded to the consumer facilities. Storing and transshipment capacities at the hibs are assumed to be constant in time and universally set to `r round(mean(storages$CAP_stor)/1000,0)` kt per year.  Finally, $|\mathcal{K}|=$ `r nrow(consumers)` consumer sites are included. Each consumer site belongs to an existing facility from chemical, paper/wood processing industry, or the energy sector. Distances between the facilities $d_{ij}$ and $d_{jk}$ are calculated as road distances between the facilities' coordinates. For each consumer facility, an yearly upper demand capacity for each of the three distinguished biomass product types are given along with consumer-specific prices. The following biomass types are distinguished:

- **P1: stem wood – chemical/paper grade**: long-fibre wood chips from stems suitable for conversion in a local biorefinery and paper factory. In total, a yearly demand of `r round(sum(consumers$demand_P1), 0)` kt and an average prices of `r round(mean(milp_instance$consumers$P1[milp_instance$consumers$P1>0]),0)` \euro{} is assumed. 

- **P2: branch wood – sawmill/pulp grade**: medium/short-fibre wood chips from larger branches suitable for mechanical processing in sawmills and lower-grade pulping. Demand is distributed across `r sum(consumers$demand_P2 > 0)` smaller facilities and accumulates to `r round(sum(consumers$demand_P2), 0)` kt/year. Gate prices range from `r round(min(milp_instance$consumers$P2[milp_instance$consumers$P2>0]),0)`-`r round(max(milp_instance$consumers$P2[milp_instance$consumers$P2>0]),0)` \euro/t.

- **P3: residue – energy grade**: residual wood fractions (chips, bark, off-cuts) used for heat and electricity generation or lower-grade pellet production. Demand is accumulates among `r sum(consumers$demand_P3 > 0)` consumers to `r round(sum(consumers$demand_P3), 0)` kt/year. Gate prices range from `r round(min(milp_instance$consumers$P3[milp_instance$consumers$P3>0]),0)`-`r round(max(milp_instance$consumers$P3[milp_instance$consumers$P3>0]),0)` \euro/t.

\autoref{tab:tab-consumers} in the appendix summarizes the consumer-specific demand data. It is assumed that the biomass product types show a quality cascade which means that the demand of pulp grade (P2) can also be satisfied by chemical/paper grade biomass (P1). Likewise, energy grade demand (P3) can be satisfied by all biomass types. Note that constant demands over the planning horizon are assumed, although an increasing biomass demand can be expected due to the green transition in those industries. From this perspective, a rather conservative market demand scenario is assumed.

Growth model introduced in  \autoref{sec:lit-growth} is used to model biomass quantities depending on age and the area dedicated to AFS. The growth model is parameterized for poplar clones at central European conditions as introduced above with a planting density of $N=400$ trees per hectar. If AFS are established, they have to be harvested within $A^{\\min}=$ `r params_obj$min_age` to $A^{\\max}=$ `r params_obj$max_age` years allowing a wide range of potential harvesting strategies between short-rotation and long-rotation AFS. As the total planning horizon is $|\mathcal{T}|=$ `r params_obj$n_periods` years multiple harvesting cycles can be established. The sites do not differ with respect to growth conditions as the climatic and soil quality variation within the comparatively small region is rather limited. However, the sites differ with respect to size (given in hectar) and opportunity cost rates (given in \euro{} per hectar and year). Site-specific opportunity cost rates $c^{opp}_i$ is  normally distributed with mean `r round(mean(sites_clustered$C_opp),0)` and standard deviation `r round(0.75*mean(sites_clustered$C_opp),0)`. For most sites, substantial opportunity costs are assumed reflecting the loss in field-crop revenue, risk premiums of landowners as well as higher costs for field crop maintenance. The `r length(sites_clustered$C_opp<0)` sites with negative opportunity cost rates reflect sites where AFS lead to an increase in field crop productivity on the remaining field area. The variation in opportunity cost rates to study the effect of opportunity cost on site selection for AFS and gives valuable insights particularly for communication with farmers and politics. Subsidies for establishing AFS are one of the core instruments to foster AFS. The height and effectiveness of those subsidies, however, is controversially disputed. Remaining cost parameters are universally set for all sites and comprise maintenance cost rates $c^{main}_i$ (\euro{} per year and hectar), establishment cost $c^{est}_i$ (\euro{} per hectar) and harvesting cost $c^{harv}_i$ (\euro{} per hectar). Likewise, transport cost rates for harvested biomasses $c^{tr-raw}_p$ and pre-treated/sorted biomasses $c^{tr-raw}_{p'}$ are set universally. All scalar parameters are shown in \autoref{tab:tab-milp-params} in the Appendix.

## Optimization results

The AFS-SCD model is solved for the instance data described above with Gurobi. The model selects a subset of sites for AFS establishment and identifies the optimal harvest schedules, intermediate facility utilisation, and product allocation across the planning horizon.

\autoref{fig:fig-results-sc-map} shows the spatial configuration of the profit-optimal supply chain. 

```{r fig-results-sc-map, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Site selection and biomass production in optimal supply chain configuration. Green polygons: active sites with biomass harvest; grey polygons: non-selected sites. Pie charts at active sites show the product mix of harvested biomass, scaled by total harvested volume.\\label{fig:fig-results-sc-map}", fig.width=8, fig.height=7, dev="pdf", out.width=".9\\textwidth", cache=T, cache.extra = c("ext","sites_clustered", "storages" , "customers", "milp_instance")}


crs_map <- 25833

# ── Colour palette constants ─────────────────────────────────────────────────
COL_SAW       <- "#e65c00"
COL_HUB       <- "grey40"
COL_P1        <- "#6a0dad"
COL_P2        <- "#08519c"
COL_P3        <- "#a50026"
COL_SITE_NONE <- "#bdbdbd"

pal_products <- c(
  p1 = "#1b9e77",
  p2 = "#d95f02",
  p3 = "#7570b3"
)

# ── Map extent ────────────────────────────────────────────────────────────────
bbox_wgs84 <- st_as_sfc(
  st_bbox(c(xmin = 10.6, xmax = 13.2, ymin = 50.9, ymax = 52.8), crs = st_crs(4326))
)
bbox_map <- st_transform(bbox_wgs84, crs_map)
bbox_num <- st_bbox(bbox_map)

# ── Base-map layers ───────────────────────────────────────────────────────────
de_states <- ne_states(country = "Germany", returnclass = "sf") %>%
  st_transform(crs_map) %>%
  st_intersection(bbox_map)

rivers_eu <- ne_download(
  scale = 10,
  type = "rivers_lake_centerlines",
  category = "physical",
  returnclass = "sf"
) %>%
  st_transform(crs_map) %>%
  st_intersection(bbox_map)

# ── Aggregate harvest volumes per site over all periods ──────────────────────
site_harvest <- ext$Xij %>%
  group_by(site_id, p) %>%
  summarise(vol = sum(value, na.rm = TRUE), .groups = "drop") %>%
  tidyr::pivot_wider(
    names_from   = p,
    values_from  = vol,
    values_fill  = 0,
    names_prefix = "p"
  )


for (col in c("p1", "p2", "p3")) {
  if (!col %in% names(site_harvest)) site_harvest[[col]] <- 0
}

site_harvest <- site_harvest %>%
  mutate(total_vol = p1 + p2 + p3) %>%
  filter(total_vol > 0)

active_site_ids <- unique(site_harvest$site_id)

# ── Site polygons ─────────────────────────────────────────────────────────────
sites_sf_map <- sites_sf %>%
  mutate(site_id = seq_len(nrow(sites_sf))) %>%
  inner_join(afs_workspace$site_cluster_assig, "site_id") %>%
  st_transform(crs_map) %>%
  mutate(active = hac_cluster %in% active_site_ids)

tmp.opp <- sites_clustered %>%
  select(site_id, C_opp) %>%
  rename(hac_cluster = site_id)

sites_sf_map <- sites_sf_map %>%
  left_join(tmp.opp, by = "hac_cluster")

# Split into active and inactive subsets for separate fill scales
# CHANGED: split polygons so inactive uses fixed grey fill and
#          active uses a continuous C_opp colour scale
sites_inactive <- sites_sf_map %>% filter(!active)
sites_active   <- sites_sf_map %>% filter(active)

# ── Pie-chart data ────────────────────────────────────────────────────────────
sites_pie_sf <- milp_instance$sites %>%
  inner_join(site_harvest, by = "site_id") %>%
  st_as_sf(coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(crs_map)

coords_sites <- st_coordinates(sites_pie_sf)

sites_pie <- sites_pie_sf %>%
  st_drop_geometry() %>%
  mutate(
    x = coords_sites[, 1],
    y = coords_sites[, 2]
  )

vol_med <- median(sites_pie$total_vol[sites_pie$total_vol > 0], na.rm = TRUE)
if (is.na(vol_med) || vol_med <= 0) vol_med <- 1

sites_pie <- sites_pie %>%
  mutate(r = 3500 * sqrt(total_vol / vol_med))

# ── Consumers & storages ──────────────────────────────────────────────────────
consumers_plt <- consumers %>%
  mutate(
    total_demand = demand_P1 + demand_P2 + demand_P3,
    kategorie = case_when(
      demand_P1 >= demand_P2 & demand_P1 >= demand_P3 & demand_P1 > 0 ~ "P1: stem/chemical",
      demand_P2 >= demand_P1 & demand_P2 >= demand_P3 & demand_P2 > 0 ~ "P2: branch/pulp",
      demand_P3 > 0 ~ "P3: residue/energy",
      TRUE ~ "Other"
    ),
    pt_size = pmin(4.5, pmax(1.5, 1.5 + log1p(total_demand) * 0.55))
  )

storages_plt <- storages %>%
  mutate(
    subtype = if_else(type == "Hub", "Hub", "Hub"),
    pt_size = 1.5
  )

stor_sf <- st_as_sf(storages_plt, coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(crs_map)

cons_sf <- st_as_sf(consumers_plt, coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(crs_map)

# Gemeinsamen Bereich aus ALLEN Sites berechnen (nicht nur sitesactive!)
copp_range <- range(sites_sf_map$C_opp, na.rm = TRUE)
copp_lim   <- c(min(copp_range[1], -1), max(copp_range[2], 1))

# ── Plot ──────────────────────────────────────────────────────────────────────
ggplot() +
  # Base map
  geom_sf(data = de_states,
          fill = "grey96", colour = "grey55", linewidth = 0.35) +
  geom_sf(data = rivers_eu,
          colour = "#4292c6", linewidth = 0.35, alpha = 0.7) +

  # CHANGED: Inactive site polygons — fixed grey fill, no legend scale needed
  geom_sf(data = sites_inactive,
          fill = COL_SITE_NONE,
          colour = NA, alpha = 0.55, show.legend = FALSE) +

  # CHANGED: Active site polygons — continuous fill mapped to C_opp
  geom_sf(data = sites_active,
          aes(fill = C_opp),
          colour = NA, alpha = 0.75, show.legend = TRUE) +
 scale_fill_gradient2(
  name     = "Opportunity cost (€/ha)",
  low      = "#1a9641",   # green  → low / negative C_opp
  mid      = "#ffffbf",   # pale yellow → midpoint
  high     = "#d7191c",   # red    → high / positive C_opp
  midpoint = 0,           # centre the diverging scale on zero
  limits   = copp_lim,        # <-- NEU: expliziter, identischer Bereich
  oob      = scales::squish,  # <-- NEU: Werte außerhalb → Endfarbe statt NA
  na.value = COL_SITE_NONE,
  guide    = guide_colourbar(
    barwidth       = 0.5,
    barheight      = 4,
    title.position = "top"
  ) 
  ) +

  # Pre-treatment facilities
  ggnewscale::new_scale_colour() +
  ggnewscale::new_scale("shape") +
  geom_sf(data = stor_sf,
          aes(colour = subtype, size = pt_size),
          stroke = 1.0, alpha = 0.92, show.legend = TRUE) +
  scale_colour_manual(
    name   = "Logistics hubs",
    values = c("Hub" = COL_HUB)
  ) +
  scale_shape_manual(
    name   = "Logistics hubs",
    values = c("Hub" = 16)
  ) +

  # Consumer sites
  ggnewscale::new_scale_fill() +
  geom_sf(data = cons_sf,
          aes(fill = kategorie, size = pt_size),
          shape = 22, colour = "white", stroke = 0.45,
          alpha = 0.9, show.legend = TRUE) +
  scale_fill_manual(
    name   = "Consumers",
    values = c(
     "P1: stem/chemical" = COL_P1,
      "P2: branch/pulp"    = COL_P2,
      "P3: residue/energy" = COL_P3,
      "Other"                = "grey60"
    ),
  guide = guide_legend(
    order        = 2,
    override.aes = list(
      shape  = 22,          # gefülltes Quadrat
      size   = 4,
      stroke = 0.4,
      colour = "white"      # Rand wie im geom_sf
    )
  )
  ) +

  # Pie charts at active sites
  ggnewscale::new_scale_fill() +
  scatterpie::geom_scatterpie(
    data         = sites_pie,
    aes(x = x, y = y, r = r),
    cols         = c("p1", "p2", "p3"),
    colour       = "white",
    linewidth    = 0.15,
    donut_radius = .5,
    alpha        = 0.7,
    show.legend  = TRUE
  ) +
  scale_fill_manual(
    name   = "Harvested biomass",
    values = pal_products,
    labels = c(p1 = "Stem", p2 = "Branches", p3 = "Residues"),
  guide  = guide_legend(
    order        = 3,
    override.aes = list(shape = 21, size = 4, stroke = 0.3)
  )
  ) +

  scale_size_continuous(range = c(1.5, 5), guide = "none") +
  guides(
  colour = guide_legend(
    order        = 1,
    override.aes = list(shape = 16, size = 4, stroke = 0, colour = COL_HUB)
  ),
  shape  = "none"   # shape-Skala unterdrücken (colour-Skala reicht für Hubs)
) +

  coord_sf(
    xlim   = c(bbox_num["xmin"], bbox_num["xmax"]),
    ylim   = c(bbox_num["ymin"], bbox_num["ymax"]),
    expand = FALSE
  ) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid      = element_line(colour = "grey88", linewidth = 0.25),
    legend.position = "right",
    legend.key.size = unit(0.40, "cm"),
    legend.text     = element_text(size = 6),
    legend.title    = element_text(size = 7.5, face = "bold"),
    axis.text       = element_text(size = 7)
  )
```

In total, on `r length(active_site_ids)` sites  AFS are established which corresponds to about `r round(length(active_site_ids)/milp_instance$n_sites*100,0)` % of all sites and `r round(sum(ext$z[ext$z$arc_type=="establishment","value"])/sum(milp_instance$sites$area_ha)*100,0)` % of the total farm area. Over the planning horizon, `r  round(sum(site_harvest$total_vol)/1000)` kt of AFS biomasses are produced on these sites which suffices to satisfy `r round(sum(site_harvest$total_vol)/sum(milp_instance$demand$D_max)*100,0)` % of total demand. \autoref{fig:fig-results-product-mix} shows the harvested biomasses over time along with total demand of all consumer sites. 

```{r fig-results-product-mix, echo=FALSE, message=FALSE, warning=FALSE,     fig.cap="Total biomass delivered to consumers by product type over the planning horizon. Stacked areas show period-level volumes of stem (P1), branches (P2), and residues (P3) flowing from storage hubs to all consumers. Dashed line: total harvested raw biomass shipped from sites to hubs.\\label{fig:fig-results-product-mix}", fig.width=6.5, fig.height=3, dev="pdf", out.width="\\textwidth", cache=F, cache.extra = c("ext","sites_clustered", "storages" , "customers", "milp_instance")}

prod_labels <- c("1" = "Stem (P1)", "2" = "Branches (P2)", "3" = "Residues (P3)")
prod_colors <- c("1" = "#1b9e77", "2" = "#d95f02", "3" = "#7570b3")


# Deliveries aggregated by period × del_product
delivery_ts <- ext$Xjk %>%
  group_by(period, product = factor(del_product)) %>%
  summarise(volume = sum(value), .groups = "drop")

# Raw biomass shipped (Xij) aggregated by period — shown as dashed reference
shipped_ts <- ext$Xij %>%
  group_by(period) %>%
  summarise(shipped = sum(value), .groups = "drop")

# Maximum total consumer demand per product type
# Assumes milp_instance$demand is a data frame with columns: customer, product, period, volume
# Adjust column names to match your actual milp_instance structure
demand_max <- milp_instance$demand %>%
  group_by(product = factor(product, levels = c("3", "2", "1"))) %>%
  summarise(max_demand = sum(D_max)/milp_instance$n_periods, .groups = "drop") %>% 
  mutate(max_demand = cumsum(max_demand))

ggplot(delivery_ts, aes(x = period, y = volume, fill = product)) +
  geom_area(position = "stack", alpha = 0.80, colour = "white", linewidth = 0.2) +
  # Raw shipped biomass reference (dashed grey)
  geom_line(data = shipped_ts,
            aes(x = period, y = shipped),
            inherit.aes = FALSE,
            colour = "#636363", linetype = "dashed", linewidth = 0.7) +
  # Max consumer demand per product (dotted, colour-matched to product)
  geom_hline(data = demand_max,
             aes(yintercept = max_demand, colour = product),
             linetype = "dashed", linewidth = 1.0, inherit.aes = FALSE) +
  scale_fill_manual(
    name   = "Product",
    values = prod_colors,
    labels = prod_labels
  ) +
  scale_colour_manual(
    name   = "Max. demand",
    values = prod_colors,
    labels = prod_labels
  ) +
  scale_x_continuous(name = "Planning period (year)",
                     breaks = seq(1, milp_instance$n_periods, by = 2)) +
  scale_y_continuous(name = "Biomass quantity (t fresh biomass)",
                     labels = scales::comma) +
  annotate("text", x = max(shipped_ts$period) - 1,
           y = max(shipped_ts$shipped) * 0.95,
           label = "Raw shipped\n(site-hub)", size = 2.5,
           colour = "#636363", hjust = 1) +
  theme_minimal(base_size = 9) +
  theme(legend.position = "right",
        panel.grid.minor = element_blank())
```

During the first 10 years no AFS biomasses are harvested as trees of established AFS in first years grow up to a certain size before first harvest. Thus, about 25\% of consumers' total biomass demand cannot be satisfied due to this grow-up period. Afterwards, a cyclical pattern appears with a harvesting peak around year 20 and a breakdown in year 28 before plateau is reached from year 35-40. Thereby, the harvesting quantities of stems and branches show a similar pattern while quantities of harvested energy grade biomass are rather stable and close to the demand limit over the planning horizon. Thus, rather long-rotation AFS are established as \autoref{fig:fig-results-harvest-cycle-dist} confirms showing the histogram of harvesting ages across all sites.

```{r fig-results-harvest-cycle-dist, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of harvesting cycle lengths at active agroforestry sites. Each bar represents the number of individual harvest events (over all periods and active sites) whose cycle length (stand age at harvest, i.e. $t - s$) falls in that year bin.\\label{fig:fig-results-harvest-cycle-dist}", fig.width=7, fig.height=3, dev="pdf", out.width="\\textwidth", cache=T, cache.extra = c("ext", "milp_instance")}

# --- Extract harvest-arc cycle lengths from binary z variables ---------------
# ext$z contains one row per selected arc; columns: site_id, s, t, arc_type
# Harvest arcs have arc_type == "harvest" and cycle_length = t - s

cycle_df <- ext$z %>%
  filter(arc_type == "harvest") %>%
  mutate(cycle_length = t - s)

# Restrict to active sites (at least one harvest arc selected)
active_sites <- unique(cycle_df$site_id)

# Summary statistics for annotation
mean_cl  <- mean(cycle_df$cycle_length)
med_cl   <- median(cycle_df$cycle_length)
n_events <- nrow(cycle_df)
n_sites  <- length(active_sites)


# --- Plot --------------------------------------------------------------------
ggplot(cycle_df, aes(x = cycle_length)) +
  geom_bar(fill = "#1b9e77", colour = "white", linewidth = 0.25, width = 0.75) +
  geom_vline(xintercept = mean_cl,
             linetype = "dashed", colour = "grey30", linewidth = 0.6) +
  annotate("text",
           x     = mean_cl + 0.25,
           y     = Inf,
           label = paste0("Mean = ", round(mean_cl, 1), " yr"),
           hjust = 0, vjust = 1.5, size = 2.8, colour = "grey20") +
  scale_x_continuous(
    name   = "Harvesting cycle length (years)",
    breaks = seq(params_obj$min_age, params_obj$max_age, by = 1),
    limits = c(params_obj$min_age - 0.5, params_obj$max_age + 0.5)
  ) +
  scale_y_continuous(
    name   = "Number of harvest events",
    expand = expansion(mult = c(0, 0.12))
  ) +
  labs(
    caption = paste0(
      n_events, " harvest events across ", n_sites,
      " active sites; median = ", med_cl, " yr"
    )
  ) +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    plot.caption       = element_text(size = 7, colour = "grey40", hjust = 0)
  )
```

The histogram illustrates that majority of harvest events occur at stand ages between 14--18 years, indicating that long-rotation AFS seem to be more profitable in this setting. Short-rotation cycles (<10 years) occur only in marginal frequencies. This optimal rotation length might be caused either due to the limited demand of energy-grade biomass or the price gaps between biomass types or a combination of effects. The quantity of energy-grade biomass created in long-rotation AFS suffices to cover total energy-grade biomass demand almost completely. Rotation length is focused on production of chemical-grade biomass where highest prices are assumed. Focusing on chemical industry supply requires a production of trees with a comparatively large stems. This strategy maximizes revenue, but in the same time increases opportunity costs as wood harvests appear in larger time intervals. \autoref{fig:fig-results-cost-profit-breakdown} presents the aggregate cost--profit waterfall over all periods and actors. 

```{r fig-results-cost-profit-breakdown, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Total cost–profit breakdown across all actors and years. The left panel (waterfall) decomposes the objective value into revenue and individual cost categories; the right panel shows the period-level trajectory of revenues and stacked costs.\\label{fig:fig-results-cost-profit-breakdown}", fig.width = 7, fig.height = 3, dev = "pdf", out.width = "\\textwidth", cache = T, cache.extra = c("ext", "milp_instance", "sites_clustered", "storages", "params_obj")}

# ── 0. Colour palette (consistent with paper) ──────────────────────────────
COL_REV   <- "#1b9e77"   # revenue: green
COL_EST   <- "#d95f02"   # establishment
COL_HARV  <- "#7570b3"   # harvesting
COL_MAIN  <- "#e7298a"   # maintenance
COL_OPP   <- "#66a61e"   # opportunity
COL_TR_R  <- "#e6ab02"   # transport raw
COL_TR_P  <- "#a6761d"   # transport pre-processed
COL_STOR  <- "#666666"   # storage
COL_P1_REV <- "#1b9e77"   # P1 Stem
COL_P2_REV <- "#d95f02"   # P2 Branches
COL_P3_REV <- "#7570b3"   # P3 Residue

prod_labels <- c("1" = "P1 Stem", "2" = "P2 Branches", "3" = "P3 Residue")
prod_colors_rev <- c("1" = COL_P1_REV, "2" = COL_P2_REV, "3" = COL_P3_REV)

# ── Colour maps — revenues and costs clearly separated ───────────────────────
prod_colors_rev <- c(
  "Revenue P1 (Stem)"     = "#1b9e77",   # teal-green
  "Revenue P2 (Branches)" = "#d95f02",   # orange
  "Revenue P3 (Residue)"  = "#7570b3"    # purple
)

cost_colors <- c(
  "Establishment"         = "#b2182b",   # dark red
  "Harvesting"            = "#ef8aa2",   # salmon
  "Maintenance"           = "#fddbc7",   # pale peach
  "Opportunity"           = "#67001f",   # deep burgundy
  "Transport (raw)"       = "#2166ac",   # medium blue
  "Transport (pre-proc.)" = "#92c5de",   # light blue
  "Storage"               = "#4d4d4d"    # dark grey
)

col_map <- c(prod_colors_rev, cost_colors)


# ── 1. Site meta: area_ha and cost parameters for active sites only ─────────
active_site_ids <- ext$z %>%
  filter(arc_type == "harvest") %>%
  pull(site_id) %>% unique()

site_meta <- milp_instance$sites %>%
  mutate(site_id = row_number()) %>%
  filter(site_id %in% active_site_ids)

# ── 2. Cultivation costs (derived from arc decisions in ext$z) ──────────────
cost_est <- ext$z %>%
  filter(arc_type == "establishment") %>%
  #select(-area_ha) %>%
  left_join(site_meta, by = "site_id") %>%
  mutate(cost = C_est * value, period = t) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Establishment")

cost_harv <- ext$z %>%
  filter(arc_type == "harvest") %>%
  #select(-area_ha) %>%
  left_join(site_meta, by = "site_id") %>%
  mutate(cost = C_harv * value, period = t) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Harvesting")

cost_main <- ext$z %>%
  filter(arc_type == "harvest") %>%
  #select(-area_ha) %>%
  mutate(age_len = t-s) %>% 
  left_join(site_meta, by = "site_id") %>%
  mutate(cost = C_main * value * age_len, period = t) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Maintenance")


cost_opp <- ext$z %>%
  filter(arc_type == "harvest") %>%
  #select(-area_ha) %>%
  left_join(site_meta, by = "site_id") %>%
  mutate(age_len = t-s) %>% 
  mutate(cost = C_opp * value * age_len, period = t) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Opportunity")


# ── 3. Transport costs ───────────────────────────────────────────────────────
dist_ij_m <- as.matrix(milp_instance$d_ij)

cost_tr_raw <- ext$Xij %>%
  mutate(
    dist_km = mapply(function(i, j) dist_ij_m[i, j], site_id, hub_id),
    cost    = params_obj$c_tr_raw * dist_km * value
  ) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Transport (raw)")

dist_jk_m <- as.matrix(milp_instance$d_jk)

cost_tr_pre <- ext$Xjk %>%
  mutate(
    dist_km = mapply(function(j, k) dist_jk_m[j, k], hub_id, consumer_id),
    cost    = params_obj$c_tr_pre * dist_km * value
  ) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Transport (pre-proc.)")

# ── 4. Storage costs ─────────────────────────────────────────────────────────
cost_stor <- ext$S %>%
  left_join(
    milp_instance$storages %>%
      mutate(hub_id = row_number()) %>%
      select(hub_id, c_stor),
    by = "hub_id"
  ) %>%
  mutate(cost = c_stor * value) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Storage")

# ── 5. Revenue per period ─────────────────────────────────────────────────────
price_df <- as.data.frame(milp_instance$consumer_prices) %>%
  setNames(c("consumer_id", "del_product", "price")) %>%
  mutate(consumer_id = as.integer(consumer_id),
         del_product  = as.integer(del_product))

revenue_by_product_ts <- ext$Xjk %>%
  left_join(price_df, by = c("consumer_id", "del_product")) %>%
  mutate(rev = value * coalesce(price, 0)) %>%
  group_by(period, del_product) %>%
  summarise(rev = sum(rev, na.rm = TRUE), .groups = "drop") %>%
  mutate(del_product = as.character(del_product))

# Total revenue per period (for cumulative cashflow calc)
revenue_ts <- revenue_by_product_ts %>%
  group_by(period) %>%
  summarise(rev = sum(rev), .groups = "drop")

# ── 6. Assemble long-format time series (unchanged) ──────────────────────────
all_periods <- tibble(period = seq_len(milp_instance$n_periods))

cost_ts <- bind_rows(
  cost_est, cost_harv, cost_main, cost_opp,
  cost_tr_raw, cost_tr_pre, cost_stor
) %>%
  right_join(crossing(period    = seq_len(milp_instance$n_periods),
                      component = unique(.$component)),
             by = c("period", "component")) %>%
  replace_na(list(cost = 0)) %>%
  mutate(component = factor(component, levels = c(
    "Establishment", "Harvesting", "Maintenance", "Opportunity",
    "Transport (raw)", "Transport (pre-proc.)", "Storage"
  )))

rev_ts <- all_periods %>%
  left_join(revenue_ts, by = "period") %>%
  replace_na(list(rev = 0))

# ── 7. Waterfall totals — Revenue split into 3 product bars ─────────────────
rev_totals_by_product <- revenue_by_product_ts %>%
  group_by(del_product) %>%
  summarise(value = sum(rev), .groups = "drop") %>%
  mutate(
    component = recode(del_product,
      "1" = "Revenue P1 (Stem)",
      "2" = "Revenue P2 (Branches)",
      "3" = "Revenue P3 (Residue)"
    )
  ) %>%
  select(component, value)

cost_totals <- cost_ts %>%
  group_by(component) %>%
  summarise(value = -sum(cost), .groups = "drop")

totals <- bind_rows(rev_totals_by_product, cost_totals) %>%
  mutate(component = factor(component, levels = c(
    "Revenue P1 (Stem)", "Revenue P2 (Branches)", "Revenue P3 (Residue)",
    "Establishment", "Harvesting", "Maintenance", "Opportunity",
    "Transport (raw)", "Transport (pre-proc.)", "Storage"
  ))) %>%
  arrange(component) %>%
  mutate(
    end      = cumsum(value),
    start    = lag(end, default = 0),
    bar_low  = pmin(start, end),
    bar_high = pmax(start, end)
  )

col_map <- c(
  "Revenue P1 (Stem)"      = COL_P1_REV,
  "Revenue P2 (Branches)"  = COL_P2_REV,
  "Revenue P3 (Residue)"   = COL_P3_REV,
  "Establishment"          = COL_EST,
  "Harvesting"             = COL_HARV,
  "Maintenance"            = COL_MAIN,
  "Opportunity"            = COL_OPP,
  "Transport (raw)"        = COL_TR_R,
  "Transport (pre-proc.)"  = COL_TR_P,
  "Storage"                = COL_STOR
)

# Revenue rows: one row per product, all starting from 0 (stacked bar)
rev_waterfall <- rev_totals_by_product %>%
  arrange(component) %>%                     # consistent stacking order
  mutate(
    x_pos    = 1L,                             # all revenue in column 1
    bar_low  = cumsum(lag(value, default = 0)),
    bar_high = cumsum(value),
    component = factor(component, levels = c(
      "Revenue P1 (Stem)", "Revenue P2 (Branches)", "Revenue P3 (Residue)"
    ))
  )

# Cost rows: floating waterfall segments starting where previous bar ended
# Revenue total = anchor for cost cascade
rev_total <- sum(rev_totals_by_product$value)

cost_waterfall <- cost_totals %>%             # negative values
  arrange(component) %>%
  mutate(
    x_pos    = row_number() + 1L,             # columns 2, 3, 4 …
    end      = rev_total + cumsum(value),
    start    = lag(end, default = rev_total),
    bar_low  = pmin(start, end),
    bar_high = pmax(start, end)
  )

# X-axis labels: "Revenue" for col 1, then cost component names
x_labels <- c("Revenue\n(stacked)", as.character(cost_waterfall$component))

# ── Add a grouping column so we can split the legend ─────────────────────────
rev_waterfall <- rev_waterfall %>%
  mutate(legend_group = "Revenue by product")

cost_waterfall <- cost_waterfall %>%
  mutate(legend_group = "Costs")

# ── 8. Right panel: Cumulative cash flow — stacked revenue by product ────────
rev_product_wide <- revenue_by_product_ts %>%
  right_join(crossing(period = seq_len(milp_instance$n_periods),
                      del_product = c("1", "2", "3")),
             by = c("period", "del_product")) %>%
  replace_na(list(rev = 0)) %>%
  mutate(del_product = factor(del_product, levels = c("3", "2", "1"),
                               labels = c("P3 Residue", "P2 Branches", "P1 Stem")))

total_cost_ts <- cost_ts %>%
  group_by(period) %>%
  summarise(total_cost = sum(cost, na.rm = TRUE), .groups = "drop")

cashflow_ts <- all_periods %>%
  left_join(rev_ts, by = "period") %>%
  left_join(total_cost_ts, by = "period") %>%
  replace_na(list(rev = 0, total_cost = 0)) %>%
  mutate(
    net_period = rev - total_cost,
    cum_cost   = cumsum(total_cost),
    cum_net    = cumsum(net_period)
  )

breakeven_period <- cashflow_ts %>%
  filter(cum_net >= 0) %>%
  slice_min(period, n = 1) %>%
  pull(period)


# ── Compute shared y range ────────────────────────────────────────────────────
y_min <- min(
  min(cost_waterfall$bar_low  / 1e6, na.rm = TRUE),
  min(cashflow_ts$cum_net     / 1e6, na.rm = TRUE),
  0
)
y_max <- max(
  max(rev_waterfall$bar_high  / 1e6, na.rm = TRUE),
  max(cashflow_ts$cum_rev_prod/ 1e6, na.rm = TRUE),   # top of stacked revenue area
  0
)

# add 5% padding
y_range <- c(y_min * 1.05, y_max * 1.05)


p_waterfall <- ggplot() +

  # — stacked revenue bar —
  geom_rect(
    data = rev_waterfall,
    aes(xmin = x_pos - 0.38, xmax = x_pos + 0.38,
        ymin = bar_low  / 1e6,
        ymax = bar_high / 1e6,
        fill = component),
    colour = "white", linewidth = 0.25
  ) +

  # — floating cost bars —
  geom_rect(
    data = cost_waterfall,
    aes(xmin = x_pos - 0.38, xmax = x_pos + 0.38,
        ymin = bar_low  / 1e6,
        ymax = bar_high / 1e6,
        fill = component),
    colour = "white", linewidth = 0.25
  ) +

  # — connector lines —
  geom_segment(
    data = cost_waterfall,
    aes(x    = x_pos - 0.5,   xend = x_pos - 0.38,
        y    = bar_low / 1e6, yend = bar_low / 1e6),
    colour = "grey60", linewidth = 0.3, linetype = "dotted",
    inherit.aes = FALSE
  ) +

  geom_hline(yintercept = 0, colour = "grey40", linewidth = 0.35) +

  # — split legend via two scale calls using ggnewscale —
  scale_fill_manual(
    name   = "Revenue by product",
    values = prod_colors_rev,
    breaks = names(prod_colors_rev),
    guide  = guide_legend(
      order     = 1,
      ncol      = 1,
      keywidth  = unit(0.35, "cm"),
      keyheight = unit(0.35, "cm")
    )
  ) +

  ggnewscale::new_scale_fill() +

  geom_rect(                                   # dummy layer to register cost fill
    data = cost_waterfall,
    aes(xmin = x_pos - 0.38, xmax = x_pos + 0.38,
        ymin = bar_low  / 1e6,
        ymax = bar_high / 1e6,
        fill = component),
    colour = "white", linewidth = 0.25
  ) +

  scale_fill_manual(
    name   = "Costs",
    values = cost_colors,
    breaks = names(cost_colors),
    guide  = guide_legend(
      order     = 2,
      ncol      = 1,
      keywidth  = unit(0.35, "cm"),
      keyheight = unit(0.35, "cm")
    )
  ) +

  scale_x_continuous(
    breaks = seq_along(x_labels),
    labels = x_labels,
    expand = expansion(add = 0.6)
  ) +

  scale_y_continuous(
    name   = "Cumulative value (M€)",
    labels = scales::label_number(suffix = " M"),
    limits = y_range
  ) +

  labs(x = NULL, title = "Waterfall") +

  theme_minimal(base_size = 8) +
  theme(
    panel.grid.major.x  = element_blank(),
    panel.grid.minor    = element_blank(),
    axis.text.x         = element_blank(),# element_text(size = 6.5, angle = 30, hjust = 1),
    plot.title          = element_text(size = 8, face = "bold", hjust = 0.5),
    legend.position     = "right",
    legend.text         = element_text(size = 6),
    legend.title        = element_text(size = 6.5, face = "bold"),
    legend.spacing.y    = unit(0.4, "cm")   # gap between the two legend groups
  )

# 2nd plot -----------------------------------------------------------------------

p_cashflow <- ggplot() +
  # Stacked area of cumulative revenue by product type
  geom_area(
    data = rev_product_wide %>%
      arrange(period, del_product) %>%
      group_by(del_product) %>%
      mutate(cum_rev_prod = cumsum(rev)) %>%
      ungroup(),
    aes(x = period, y = cum_rev_prod / 1e6, fill = del_product),
    position = "stack", alpha = 0.75, colour = "white", linewidth = 0.15
  ) +
  scale_fill_manual(
    name   = "Revenue by product",
    values = c("P1 Stem"     = COL_P1_REV,
               "P2 Branches" = COL_P2_REV,
               "P3 Residue"  = COL_P3_REV)
  ) +
  # Cumulative cost line
  geom_line(data = cashflow_ts,
            aes(x = period, y = cum_cost / 1e6),
            colour = "grey30", linewidth = 0.9, linetype = "dashed",
            inherit.aes = FALSE) +
  # Cumulative net cash flow
  geom_line(data = cashflow_ts,
            aes(x = period, y = cum_net / 1e6),
            colour = "grey10", linewidth = 0.65, linetype = "dotted",
            inherit.aes = FALSE) +
  geom_hline(yintercept = 0, colour = "grey50", linewidth = 0.35) +
  {if (length(breakeven_period) == 1)
    list(
      geom_vline(xintercept = breakeven_period,
                 colour = "grey40", linewidth = 0.45, linetype = "dotdash"),
      annotate("text",
               x = breakeven_period + 0.4, y = 0,
               label = paste0("BEP: t=", breakeven_period),
               size = 2.4, hjust = 0, vjust = -0.4,
               colour = "grey30", fontface = "italic")
    )
  } +
  annotate("text",
           x = milp_instance$n_periods - 0.3,
           y = last(cashflow_ts$cum_cost) / 1e6,
           label = "Cum. cost", size = 2.4, hjust = 1, vjust = 1.3,
           colour = "grey30", fontface = "italic") +
  annotate("text",
           x = milp_instance$n_periods - 0.3,
           y = last(cashflow_ts$cum_net) / 1e6,
           label = "Cum. net CF", size = 2.4, hjust = 1, vjust = -0.4,
           colour = "grey10", fontface = "italic") +
  scale_x_continuous(name   = "Planning period (year)",
                     breaks = seq(1, milp_instance$n_periods, by = 2)) +
  scale_y_continuous(name   = "Cumulative value (M€)",
                     labels = scales::label_number(suffix = " M"),
                     limits = y_range) +
  labs(title = "Cumulative cash flow by product") +
  theme_minimal(base_size = 8) +
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(size = 8, face = "bold", hjust = 0.5),
        legend.position = "none",
        #legend.key.size = unit(0.35, "cm"),
        #legend.text = element_text(size = 6.5)
        )

(p_waterfall + p_cashflow +
  plot_layout(
    widths  = c(1, 1.6),            # waterfall narrower, cashflow wider
    guides  = "collect"                # keep separate legends per panel
  ) #+
  #plot_annotation(
    #title   = "Cost–profit breakdown",
    #caption = "Left: cumulative waterfall; Right: period-by-period cash flow by product type.",
    #theme   = theme(
    #  plot.title   = element_text(size = 9, face = "bold", hjust = 0.5),
    #  plot.caption = element_text(size = 6, colour = "grey50", hjust = 0)
    #)
  #)
)

```

It appears that revenue is dominated by selling chemical-grade biomass and dominating cost component are opportunity cost although active sites are those sites with a comparatively small opportunity cost rate. Next to opportunity cost, establishment and transport costs are the most relevant cost drivers. Analyzing total cashflow over time shows that break-even is reached after about 20 years, i.e. after the first large-scale harvest occurred. Afterwards, cashflow is continuously increasing indicating that establishing AFS is a strategic change in agricultural practice and pays only if this strategy is implemented and maintained over a comparatively long time spans. This underpins the importance of strategic buying contracts with industrial partners in order to reduce revenue risks of farmers when producing AFS biomass. 

Although profitability of the total AFS suply chain is given for the case study data, profitability of AFS operations on site-level are more heterogeneous. \autoref{fig:fig-results-site-breakdown} shows revenues and costs split by component and normalized to AFS size and active years of AFS.

```{r fig-results-site-breakdown, echo=FALSE, message=FALSE, warning=FALSE,     fig.cap="Revenues, costs and net profit for sites with AFS operations split by type and expressed in € per hectare per year of AFS operation.\\label{fig:fig-results-site-breakdown}",  fig.width = 7, fig.height = 3, dev = "pdf", out.width = "\\textwidth",   cache = T, cache.extra = c("ext", "milp_instance", "sites_clustered", "params_obj")}

# ── 0. Colour palette — identical to fig-results-cost-profit-breakdown ───────
prod_colors_rev <- c(
  "Revenue P1 (Stem)"     = "#1b9e77",
  "Revenue P2 (Branches)" = "#d95f02",
  "Revenue P3 (Residue)"  = "#7570b3"
)
cost_colors <- c(
  "Establishment"         = "#b2182b",
  "Harvesting"            = "#ef8a62",
  "Maintenance"           = "#fddbc7",
  "Opportunity"           = "#67001f",
  "Transport (raw)"       = "#2166ac",
  "Transport (pre-proc.)" = "#92c5de",
  "Storage"               = "#4d4d4d"
)

# ── 1. Site meta ─────────────────────────────────────────────────────────────
active_site_ids_sb <- ext$z %>%
  filter(arc_type == "harvest") %>%
  pull(site_id) %>% unique()

site_meta_sb <- milp_instance$sites %>%
  mutate(site_id = row_number()) %>%
  filter(site_id %in% active_site_ids_sb)

# ── 2. Hub-site share ────────────────────────────────────────────────────────
hub_site_share_sb <- ext$Xij %>%
  group_by(hub_id, period) %>%
  mutate(hub_total = sum(value)) %>%
  ungroup() %>%
  mutate(share = ifelse(hub_total > 0, value / hub_total, 0)) %>%
  select(site_id, hub_id, period, share)

# ── 3. Price table ───────────────────────────────────────────────────────────
price_df_sb <- as.data.frame(milp_instance$consumer_prices) %>%
  setNames(c("consumer_id", "del_product", "price")) %>%
  mutate(consumer_id = as.integer(consumer_id),
         del_product  = as.integer(del_product))

# ── 4. Normalisation denominator: area_afs × active_years (ha·yr) ────────────
active_p_sb <- ext$z %>%
  filter(arc_type == "establishment") %>%
  group_by(site_id) %>%
  summarise(tot_value = sum(value), .groups = "drop")

est_p_sb <- ext$z %>%
  filter(arc_type == "establishment") %>%
  group_by(site_id) %>%
  summarise(t_est = sum(t * value) / sum(value), .groups = "drop")

term_p_sb <- ext$z %>%
  filter(arc_type == "termination") %>%
  group_by(site_id) %>%
  summarise(t_term = sum(s * value) / sum(value), .groups = "drop")

site_denom_sb <- site_meta_sb %>%
  left_join(active_p_sb, by = "site_id") %>%
  left_join(est_p_sb,    by = "site_id") %>%
  left_join(term_p_sb,   by = "site_id") %>%
  mutate(
    t_term       = ifelse(is.na(t_term), milp_instance$n_periods, t_term),
    active_years = pmax(t_term - t_est, 1L),
    denom        = tot_value * active_years   # ha·yr
  ) %>%
  select(site_id, denom)

# ── 5. Revenue per site × product type, normalised to €/ha/yr ────────────────
site_rev_by_product <- ext$Xjk %>%
  left_join(price_df_sb, by = c("consumer_id", "del_product")) %>%
  mutate(rev_jkt = value * coalesce(price, 0)) %>%
  left_join(hub_site_share_sb, by = c("hub_id", "period"),
            relationship = "many-to-many") %>%
  mutate(site_rev = rev_jkt * share) %>%
  filter(!is.na(site_id)) %>%
  group_by(site_id, del_product) %>%
  summarise(value_raw = sum(site_rev, na.rm = TRUE), .groups = "drop") %>%
  left_join(site_denom_sb, by = "site_id") %>%
  mutate(
    value     = value_raw / denom,
    component = recode(as.character(del_product),
      "1" = "Revenue P1 (Stem)",
      "2" = "Revenue P2 (Branches)",
      "3" = "Revenue P3 (Residue)"
    )
  ) %>%
  select(site_id, component, value)

# ── 6. Costs per site × component, normalised to €/ha/yr (negative) ──────────
dist_ij_m_sb <- as.matrix(milp_instance$d_ij)
dist_jk_m_sb <- as.matrix(milp_instance$d_jk)

normalise_cost <- function(df, denom_df) {
  df %>%
    left_join(denom_df, by = "site_id") %>%
    mutate(value = value / denom) %>%
    select(site_id, component, value)
}

sc_est_sb <- ext$z %>% filter(arc_type == "establishment") %>%
  left_join(site_meta_sb, by = "site_id") %>%
  mutate(cost = C_est * value) %>%
  group_by(site_id) %>%
  summarise(value = -sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Establishment") %>%
  normalise_cost(site_denom_sb)

sc_harv_sb <- ext$z %>% filter(arc_type == "harvest") %>%
  left_join(site_meta_sb, by = "site_id") %>%
  mutate(cost = C_harv * value) %>%
  group_by(site_id) %>%
  summarise(value = -sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Harvesting") %>%
  normalise_cost(site_denom_sb)

sc_main_sb <- ext$z %>% filter(arc_type == "harvest") %>%
  mutate(age_len = t - s) %>%
  left_join(site_meta_sb, by = "site_id") %>%
  mutate(cost = C_main * value * age_len) %>%
  group_by(site_id) %>%
  summarise(value = -sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Maintenance") %>%
  normalise_cost(site_denom_sb)

sc_opp_sb <- ext$z %>% filter(arc_type == "harvest") %>%
  mutate(age_len = t - s) %>%
  left_join(site_meta_sb, by = "site_id") %>%
  mutate(cost = C_opp * value * age_len) %>%
  group_by(site_id) %>%
  summarise(value = -sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Opportunity") %>%
  normalise_cost(site_denom_sb)

sc_tr_raw_sb <- ext$Xij %>%
  mutate(dist_km = mapply(function(i, j) dist_ij_m_sb[i, j], site_id, hub_id),
         cost    = milp_instance$c_tr_raw * dist_km * value) %>%
  group_by(site_id) %>%
  summarise(value = -sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Transport (raw)") %>%
  normalise_cost(site_denom_sb)

sc_tr_pre_sb <- ext$Xjk %>%
  mutate(dist_km  = mapply(function(j, k) dist_jk_m_sb[j, k], hub_id, consumer_id),
         cost_jkt = milp_instance$c_tr_pre * dist_km * value) %>%
  left_join(hub_site_share_sb, by = c("hub_id", "period"),
            relationship = "many-to-many") %>%
  mutate(site_cost = cost_jkt * share) %>%
  group_by(site_id) %>%
  summarise(value = -sum(site_cost, na.rm = TRUE), .groups = "drop") %>%
  filter(!is.na(site_id)) %>%
  mutate(component = "Transport (pre-proc.)") %>%
  normalise_cost(site_denom_sb)

sc_stor_sb <- ext$S %>%
  left_join(milp_instance$storages %>%
              mutate(hub_id = row_number()) %>%
              select(hub_id, c_stor),
            by = "hub_id") %>%
  mutate(stor_cost = c_stor * value) %>%
  select(hub_id, period, stor_cost) %>%
  left_join(hub_site_share_sb, by = c("hub_id", "period"),
            relationship = "many-to-many") %>%
  mutate(site_cost = stor_cost * share) %>%
  group_by(site_id) %>%
  summarise(value = -sum(site_cost, na.rm = TRUE), .groups = "drop") %>%
  filter(!is.na(site_id)) %>%
  mutate(component = "Storage") %>%
  normalise_cost(site_denom_sb)

# ── 7. Combine & order sites by descending net profit/ha/yr ──────────────────
site_breakdown_df <- bind_rows(
  site_rev_by_product,
  sc_est_sb, sc_harv_sb, sc_main_sb, sc_opp_sb,
  sc_tr_raw_sb, sc_tr_pre_sb, sc_stor_sb
)

site_order <- site_breakdown_df %>%
  group_by(site_id) %>%
  summarise(profit_ha_yr = sum(value, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(profit_ha_yr)) %>%
  pull(site_id)

# ── 7b. Net profit per site (for point overlay) ───────────────────────────────
site_profit_df <- site_breakdown_df %>%
  group_by(site_id) %>%
  summarise(net = sum(value, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    site_id  = factor(site_id, levels = site_order),
    label    = "Net profit"
  )

site_profit_df_real <- site_breakdown_df  %>% 
  group_by(site_id) %>% 
  filter(component %in% c("Revenue P1 (Stem)","Revenue P2 (Branches)", "Revenue P3 (Residue)","Opportunity") ) %>% 
  summarize(net = sum(value))

# ── 8. Factor levels ─────────────────────────────────────────────────────────
component_levels <- c(
  "Revenue P1 (Stem)", "Revenue P2 (Branches)", "Revenue P3 (Residue)",
  "Establishment", "Harvesting", "Maintenance", "Opportunity",
  "Transport (raw)", "Transport (pre-proc.)", "Storage"
)

site_breakdown_df <- site_breakdown_df %>%
  mutate(
    site_id   = factor(site_id, levels = site_order),
    component = factor(component, levels = component_levels)
  )

# ── 9. Plot ───────────────────────────────────────────────────────────────────
ggplot() +

  # Revenue stacked bars (positive, €/ha/yr)
  geom_col(
    data = site_breakdown_df %>%
      filter(component %in% names(prod_colors_rev)),
    aes(x = site_id, y = value, fill = component),
    position = "stack", width = 0.7,
    colour = "white", linewidth = 0.15
  ) +
  scale_fill_manual(
    name   = "Revenue by product",
    values = prod_colors_rev,
    breaks = names(prod_colors_rev),
    guide  = guide_legend(
      order     = 1,
      ncol      = 1,
      keywidth  = unit(0.35, "cm"),
      keyheight = unit(0.35, "cm")
    )
  ) +

  ggnewscale::new_scale_fill() +

  # Cost stacked bars (negative, €/ha/yr)
  geom_col(
    data = site_breakdown_df %>%
      filter(component %in% names(cost_colors)),
    aes(x = site_id, y = value, fill = component),
    position = "stack", width = 0.7,
    colour = "white", linewidth = 0.15
  ) +
  scale_fill_manual(
    name   = "Costs",
    values = cost_colors,
    breaks = names(cost_colors),
    guide  = guide_legend(
      order     = 2,
      ncol      = 1,
      keywidth  = unit(0.35, "cm"),
      keyheight = unit(0.35, "cm")
    )
  ) +

  geom_hline(yintercept = 0, colour = "grey40", linewidth = 0.35) +

  # ── Net profit points with legend entry via colour aesthetic ─────────────
  ggnewscale::new_scale_colour() +

  geom_point(
    data = site_profit_df,
    aes(x = site_id, y = net, colour = label),
    shape = 18, size = 1.8
  ) +
  scale_colour_manual(
    name   = NULL,
    values = c("Net profit" = "grey15"),
    guide  = guide_legend(
      order     = 3,
      ncol      = 1,
      keywidth  = unit(0.35, "cm"),
      keyheight = unit(0.35, "cm"),
      override.aes = list(shape = 18, size = 1.8)
    )
  ) +

  scale_y_continuous(
    name   = "€ per hectare and year",
    labels = scales::label_number(big.mark = ",")
  ) +
  labs(x = "Site ID") +
  theme_minimal(base_size = 8) +
  theme(
    axis.text.x        = element_text(size = 5.5, angle = 60, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    legend.position    = "right",
    legend.key.size    = unit(0.35, "cm"),
    legend.text        = element_text(size = 6),
    legend.title       = element_text(size = 6.5, face = "bold"),
    legend.spacing.y   = unit(0.4, "cm")
  )
```

It appears that net profitability of AFS ranges between `r round(max(site_profit_df$net),0)` and `r round(min(site_profit_df$net),0)` \euro{} per hectare and active year. Note that this value constitutes additional profit after subtracting operational cost as well as imputed opportunity cost of lost profit from field crop. I.e., for calculating the realized net margin, opportunity can be added such that realized yearly net profit ranges between `r round(max(site_profit_df_real$net),0)` and `r round(min(site_profit_df_real$net),0)` \euro{} per hectare. Still, relative net profit is comparatively low compared to ordinary agricultural practices with profitable field crops (like winter wheat). The revenue splits of active sites indicates that at some sites medium-grade and low-grade biomass types contribute substantially or even solely to total revenue (see e.g. site 12, 135, 145, and 112). This indicates that site-specific AFS strategies might differ from the overall supply chain strategy. In other words, there seem to be strategic niches w.r.t. the consumer segments to be supplied. This might also be reflected in the rotation length and, although not discussed here, the selection of trees and clones for the AFS. Moreover, profitability of AFS sites seems to depend on various factors like local opportunity costs as well as closeness to consumers. \autoref{fig:fig-results-site-scatter} shows net margins and opportunity cost rates of sites with active AFS. 

```{r fig-results-site-scatter, echo=FALSE, message=FALSE, warning=FALSE,     fig.cap="Scatterplot of opportunity cost rate and net margin (both in € per ha and year) for each site with active AFS. Size of AFS reflected in point size.\\label{fig:fig-results-site-scatter}", fig.width = 7, fig.height = 3, dev = "pdf", out.width = "\\textwidth", cache = T, cache.extra = c("ext", "milp_instance", "params_obj")}


ggplot(df_bench, aes(x = opp_cost_site, y = profit_ha_yr)) +
  geom_point(aes(size = area_afs),
             colour = "grey30", fill = "steelblue", alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, colour = "firebrick", linewidth = 0.7) +
  scale_shape_manual(values = c(`TRUE` = 21, `FALSE` = 1),
                     guide = "none") +
  scale_size_area(
     name     = "AFS size (ha)",
     max_size = 8,
     breaks   = scales::breaks_extended(n = 8),
     labels   = scales::label_number(big.mark = ",")
   ) +
  #scale_size_continuous(name = "AFS size (ha)", range = c(1, 8)) +
  labs(
    x = "Opportunity cost (€ per hectare and year)",
    y = "| Net margin |\n(€ per hectare and year)"
  ) +
  theme_minimal(base_size = 9) +
  theme(legend.position  = "right",
        panel.grid.minor = element_blank())
```

As expected net margins of AFS depend heavily on a site's opportunity cost rate showing a clear negative trend. Moreover, it appears that AFS sizes vary quite heavily between `r round(min(df_bench$area_afs))` and `r round(max(df_bench$area_afs))` hectare. Probably very small sites won't operate efficiently because cost for establishment, harvesting, and maintenance are assumed to scale with size, but are rather fixed below a certain critical threshold.

# Conclusion {#sec-conclusion}

This paper has developed and applied the AFS-SCD model, a MILP
formulation for the integrated design and operation of agroforestry
biomass supply chains. The model endogenises stand age dynamics through
binary arc-based age-lag variables, couples component-specific Gompertz
yield coefficients for stem, branch, and residue fractions with
multi-echelon logistics decisions, and embeds a product quality cascade
that allocates biomass grades to chemical, pulp, and energy markets
within a unified optimisation framework. The formulation is, to the best
of our knowledge, the first supply chain model to simultaneously
optimise AFS establishment timing, consecutive harvest scheduling under
rotation-age constraints, and product-grade allocation across spatially
heterogeneous consumer markets.

The case study for the Saxony-Anhalt tri-state region demonstrates the
practical relevance of the approach. The optimal supply chain
concentrates AFS establishments in the Elbe and Saale floodplains,
where high biomass productivity and proximity to large pulp mills create
the most favourable economics. The sensitivity analysis confirms that
profitability depends primarily on achievable gate prices for
high-quality (P1) biomass grades: scenarios combining reduced revenue
with high opportunity costs yield negative net profits, whereas
scenarios with full or elevated prices remain viable under cost pressure.
Opportunity cost is the dominant site-level discriminator, with parcels
exceeding approximately 300 \euro{}\ ha$^{-1}$ yr$^{-1}$ rarely selected
in the optimal solution. These findings underscore the importance of
product cascading as a strategic lever for AFS viability: maximising
the share of stem biomass directed to high-value chemical and pulp
applications substantially improves profitability compared to
energy-only offtake strategies.

Methodologically, the AFS-SCD model bridges the gap between stand-level
growth modelling and regional supply chain optimisation identified in
the literature [@Zahraee2020_BSC_review; @Lamers2015_Feedstock_Design].
The age-lag constraint structure generalises to other perennial biomass
crops with periodic coppice cycles, such as willow or black locust, and
to long-rotation silvoarable systems in which rotation length is a
strategic variable. The model is extensible to stochastic settings by
replacing deterministic yield coefficients with scenario-weighted
counterparts or by incorporating robust optimisation over yield
uncertainty intervals [@Sharma2013_BSC_review].

Several limitations point to directions for future research. First,
consumer demand is treated as exogenous and time-invariant; endogenising
demand responses to biomass price changes would require a
partial-equilibrium extension. Second, the case study uses a single
allometric scenario calibrated to Central European poplar; propagating
Gompertz parameter uncertainty through to supply chain outcomes would
strengthen the robustness analysis. Third, environmental co-benefits of
AFS --- including carbon sequestration, nitrate leaching reduction, and
biodiversity enhancement --- are not monetised in the current objective.
Integrating life cycle assessment indicators as additional objectives
or constraints would support a more comprehensive sustainability
evaluation, consistent with the cleaner production framing of this work.


# Author contributions {.bm .unnumbered}

The author is responsible for all contents of this text.

# Acknowledgments {.bm .unnumbered}

The author thanks all partners in the aforementioned project for their help in compiling and reviewing this manuscript in its early stages. Special thanks are due to Nico Tauchnitz, David Anthony, Jan Lukas Wenzel und Christopher Conrad.

# Financial disclosure {.bm .unnumbered}

This research was funded by the German Federal Ministry of Education and Research (BMBF) within the programme “Modellregion der Bioökonomie (DiP)” under the project “DiP: SMART-Agroforst – Digitalisierung zur Förderung der Etablierung von Agroforstsystemen auf der Landschaftsebene als Beitrag zur Klimaresilienz Süd-Sachsen-Anhalts und Dekarbonisierung seiner chemischen Industrie – Teilvorhaben J” (funding code 031B1457J). The funder had no role in the study design, data collection and analysis, decision to publish, or preparation of the manuscript.

# Conflict of interest {.bm .unnumbered}

The author declares no potential conflict of interests.

\bibliography{ {{< meta bibliography >}} }

# Supporting information {.bm .unnumbered}

Additional supporting information may be found in the online version of the article at the
publisher's website.

\appendix


```{r afs-model-functions, echo=FALSE, message=FALSE, warning=FALSE,     fig.cap="Logistic diameter-class merchantability functions $q_p(t)$ for stem  ($d > 15$\\,cm, green) and branches ($d > 7$\\,cm, orange); the 50\\% threshold ages $t_{50}$ are indicated by dashed vertical lines.\\label{fig:afs-model-functions}", fig.width=6.5, fig.height=3, out.width="\\textwidth", cache=FALSE}

# ── Panel (b): logistic merchantability q_p(t) ─────────────────────────────
ages_c <- seq(1, 20, by = 0.1)
df_c <- tibble(
  age                    = ages_c,
  `Stem (d > 15 cm)`    = f_stem_merch(ages_c),
  `Branch (d > 7 cm)`   = f_branch_merch(ages_c)
) |>
  pivot_longer(-age, names_to = "compartment", values_to = "q") |>
  mutate(compartment = factor(compartment,
           levels = c("Stem (d > 15 cm)", "Branch (d > 7 cm)")))

t50_vals <- tibble(
  compartment = factor(c("Stem (d > 15 cm)", "Branch (d > 7 cm)"),
                       levels = c("Stem (d > 15 cm)", "Branch (d > 7 cm)")),
  t50 = c(8.19, 7.55)
)

merch_cols <- c("Stem (d > 15 cm)"  = "#1b9e77",
                "Branch (d > 7 cm)" = "#d95f02")

p_c <- ggplot(df_c, aes(x = age, y = q, colour = compartment)) +
  geom_vline(data = t50_vals,
             aes(xintercept = t50, colour = compartment),
             linetype = "dashed", linewidth = 0.45, show.legend = FALSE) +
  geom_hline(yintercept = 0.5,
             linetype = "dotted", colour = "grey60", linewidth = 0.4) +
  geom_line(linewidth = 0.8) +
  annotate("text", x = 8.19 + 0.3, y = 0.08,
           label = expression(italic(t)[50*","*s]*" = 8.2 yr"),
           size = 2.4, hjust = 0, colour = "#1b9e77") +
  annotate("text", x = 7.55 - 0.3, y = 0.20,
           label = expression(italic(t)[50*","*b]*" = 7.6 yr"),
           size = 2.4, hjust = 1, colour = "#d95f02") +
  scale_colour_manual(name = "Compartment", values = merch_cols) +
  scale_x_continuous(breaks = seq(0, 20, 2)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, max(df_c$q)*1.1)) +
  labs(x = "Stand age (yr)",
       y = expression(italic(q[p])*"(t)  [merchantable fraction]")) +
  theme_minimal(base_size = 8) +
  theme(legend.position  = c(0.15, 0.75),
        legend.key.size  = unit(0.4, "cm"),
        legend.text      = element_text(size = 7),
        legend.title     = element_text(size = 7, face = "bold"),
        legend.background = element_rect(fill = "white", colour = NA),
        panel.grid.minor = element_blank())

# ── Assemble with patchwork ─────────────────────────────────────────────────
p_c
```


```{r growth-asymptote-functions, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Tree-level and stand-level Gompertz asymptote functions $A_{tree}(N)$ and $A(N)$ for advanteguous ($N=4$, $\\beta=-0.3$) and disadvantageous ($N=6.5$, $\\beta=-0.5$) site conditions.\\label{fig:growth-asymptote-functions}",  fig.width=6.5, fig.height=4, out.width="\\textwidth", cache=FALSE}

# ── Panel (a): A_stand vs N ─────────────────────────────────────────────────
N_seq <- seq(20, 400, by = 1)
df_a  <- bind_rows(
  tibble(N = N_seq,
         A = A_stand(N_seq, C_site = 4.000, beta=-.3),
         scenario = "advanteguous"),
  tibble(N = N_seq,
         A = A_stand(N_seq, C_site = 6.500, beta = -.5),
         scenario = "disadvantageous")
)

df_tree  <- bind_rows(
  tibble(N = N_seq,
         A = A_tree(N_seq, C_site = 4.000, beta=-.3),
         scenario = "advanteguous"),
  tibble(N = N_seq,
         A = A_tree(N_seq, C_site = 6.500, beta = -.5),
         scenario = "disadvantageous")
)


p_a <- ggplot(df_a, aes(x = N, y = A, linetype = scenario)) +
  geom_line(colour = "grey20", linewidth = 0.7) +
  scale_linetype_manual(
    name   = "Scenario",
    values = c("advanteguous" = "solid", "disadvantageous" = "dashed")) +
  labs(x = expression("Density "*italic(N)*" (trees ha"^{-1}*")"),
       y = expression(italic(A(N))*" (t DM ha"^{-1}*")")) +
  theme_minimal(base_size = 8) +
  theme(legend.position  = c(0.72, 0.82),
        legend.key.size  = unit(0.4, "cm"),
        legend.text      = element_text(size = 7),
        legend.title     = element_text(size = 7, face = "bold"),
        legend.background = element_rect(fill = "white", colour = NA),
        panel.grid.minor = element_blank())

p_b <- ggplot(df_tree, aes(x = N, y = A, linetype = scenario)) +
  geom_line(colour = "grey20", linewidth = 0.7) +
  scale_linetype_manual(
    name   = "Scenario",
    values = c("advanteguous" = "solid", "disadvantageous" = "dashed")) +
  labs(x = expression("Density "*italic(N)*" (trees ha"^{-1}*")"),
       y = expression(italic(A[tree](N))*" (t DM per tree)")) +
  theme_minimal(base_size = 8) +
  theme(legend.position  = c(0.72, 0.82),
        legend.key.size  = unit(0.4, "cm"),
        legend.text      = element_text(size = 7),
        legend.title     = element_text(size = 7, face = "bold"),
        legend.background = element_rect(fill = "white", colour = NA),
        panel.grid.minor = element_blank())
# ── Assemble with patchwork ─────────────────────────────────────────────────
p_b / p_a
```



```{r tab-notation-growth, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr); library(kableExtra)
options(kableExtra.latex.load_packages = FALSE)
options(knitr.table.format = "latex")

notation <- tribble(
  ~Symbol, ~Description,
  # --- stand-level Gompertz ---
  "$M(t)$",       "Total above- and belowground dry biomass per ha of AFS (t\\,DM\\,ha$^{-1}$) at stand age $t$",
  "$A(N)$",            "Stand-level Gompertz asymptote (t\\,DM\\,ha$^{-1}$); density- and site-dependent",
  "$k$",               "Gompertz growth-rate coefficient",
  "$t_0$",             "Inflection age of stand biomass accumulation (years)",
  # --- density--asymptote power law ---
  "$N$",               "Stand density (trees\\,ha$^{-1}$)",
  "$C_{\\mathrm{site}}$", "Site-quality constant (kg\\,ha$^{\\beta-1}$); conservative: 4\\,475, good site: 6\\,471",
  "$\\beta$",          "Competition-density exponent; fitted $\\hat{\\beta}=0.380$",
  # --- AGB/BGB split ---
#  "$\\phi_{\\mathrm{ag}}$", "Aboveground fraction of total stand biomass; $\\phi_{\\mathrm{ag}}=0.89$",
#  "$M^{ag}(t)$",       "Aboveground dry biomass per ha (t\\,DM\\,ha$^{-1}$) at age $t$",
  # --- compartment fractions ---
  "$f_p(t)$",          "asymtotic fraction of AGB attributed to compartment $p \\in \\{s,b,r\\}$ at age $t$",
#  "$\\alpha_p,\\,\\delta_p$", "Intercept and slope of the linear fraction model for compartment $p$",
  # --- logistic merchantability ---
  "$q_p(t)$",          "Merchantable fraction of compartment $p$ biomass at age $t$ (logistic)",
  "$r_p$",             "Logistic growth rate for merchantability of compartment $p$",
  "$t_{50,p}$",        "Age at which 50\\% of compartment $p$ biomass is merchantable (yr)",
  # --- harvest yields ---
  "$\\eta_p(t)$",          "Merchantable dry biomass of compartment $p$ per ha at harvest age $t$ (t\\,DM\\,ha$^{-1}$)",
  "$\\omega$",         "Moisture content at felling (mass fraction); $\\omega = 0.55$"
)

kable(notation,
      format    = "latex", booktabs = TRUE, escape = FALSE,
      col.names = c("Symbol", "Description"),
      caption   = "Notation for the biomass growth and harvest yield models.",
      label     = "tab-notation-growth") |>
  kable_styling(font_size = 8,
                latex_options = c("hold_position", "scale_down", "striped"),
                full_width = FALSE) |>
  column_spec(1, width = "3.2cm") |>
  column_spec(2, width = "10.8cm")
```



```{r tab-notation-milp, echo=FALSE, message=FALSE, warning=FALSE, cache=T}
library(knitr)
library(kableExtra)
options(kableExtra.latex.load_packages = FALSE)
options(knitr.table.format = "latex")

math_break <- function(line1, line2) {
  paste0("\\shortstack[l]{$", line1, "$\\\\$", line2, "$}")
}

notation <- tribble(
  ~Symbol, ~Description,
  # --- SETS & INDICES ---
  "$i \\in \\mathcal{I}$",   "Agroforestry sites",
  "$j \\in \\mathcal{J}$",   "Storage / pre-treatment facilities",
  "$k \\in \\mathcal{K}$",   "Consumer sites",
  "$p \\in \\mathcal{P}$",   "Product types (stem, branches, residues)",
  "$(p,p') \\in \\mathcal{Q}$",   "Admissible product-substitution pairs (quality cascade)",
  "$\\mathcal{T}=\\{1,\\dots,T\\}$",
    "Planning periods (years)",
  "$\\mathcal{T}^{\\text{harv}}=\\{A^{\\min}+1,\\dots,T\\}$",
    "Feasible harvest periods",
  "$\\mathcal{S}^{\\text{est}}=\\{(0,t)\\mid t\\in\\{1,\\dots,T-A^{\\min}\\}\\}$",
    "Establishment arcs: dummy source to period $t$",
  math_break(
    "\\mathcal{S}^{\\text{harv}}=\\{(s,t)\\mid s,t\\in \\mathcal{T},",
    "t-s\\in\\{A^{\\min},\\dots,A^{\\max}\\}\\}"),
    "Harvest cycles within age-lag bounds",
  math_break(
    "\\mathcal{S}^{\\text{term}}=\\{(s,T+1)\\mid",
    "s\\in\\{A^{\\min}+1,\\dots,T\\}\\}"),
    "Termination arcs to dummy sink $T+1$",
  "$\\mathcal{S} = \\mathcal{S}^{\\text{est}} \\cup \\mathcal{S}^{\\text{harv}} \\cup\\; \\mathcal{S}^{\\text{term}}$",
    "Age-lag arc set (establishment $\\cup$ harvest $\\cup$ termination)",
  # --- DECISION VARIABLES ---
  "$z_{ist} \\geq 0$",
    "Area of site $i$ with AFS operation between $s$ and $t$",
  "$X_{ijpt}\\ge 0$",
    "Transport flow (t) of $p$ from site $i$ to facility $j$ in $t$",
  "$X_{jkpp't}\\ge 0$",
    "Flow of $p$ from facility $j$ to consumer $k$ satisfying demand $p'$ in $t$",
  "$S_{jpt}\\ge 0$",
    "Inventory (t) of product $p$ at facility $j$ end of period $t$",
  # --- PARAMETERS ---
  "$T$",              "Length of planning horizon (years)",
  "$A^{\\min}$, $A^{\\max}$", "Minimum / maximum stand age at harvest (years)",
  "$\\text{AREA}_i$",         "Area of site $i$ (ha)",
  "$\\eta_{pa}$",
    "Yield coefficient (t/ha) of fraction $p$ at stand age $a$, derived from \\eqref{eq:yield}",
  "$R_{kp}$",                 "Revenue (€/t) for product $p$ at consumer $k$",
  "$C^{\\text{est}}_i$",      "Establishment cost (€/ha) at site $i$",
  "$C^{\\text{opp}}_i$",      "Annual opportunity cost (€/ha) at site $i$",
  "$C^{\\text{harv}}_i$",     "Harvest \\& refitting cost (€/ha) at site $i$",
  "$c^{\\text{tr-raw}}$",     "Transport cost rate (€/t\\,km) for raw biomass",
  "$c^{\\text{tr-pre}}$",     "Transport cost rate (€/t\\,km) for pre-treated biomass",
  "$c^{\\text{stor}}_j$",     "Storage holding cost (€/t) at facility $j$",
  "$d_{ij}$, $d_{jk}$",       "Distance (km) between nodes",
  "$\\text{CAP}^{\\text{stor}}_j/\\text{CAP}^{\\text{proc}}_j$", "Storage and processing capacity (t) at facility $j$",
  "$D^{\\max}_{kpt}$",        "Maximum demand (t) of fraction $p$ at consumer $k$ in period $t$"
)

tbl <- kable(notation,
      format    = "latex",
      booktabs  = TRUE,
      escape    = FALSE,
      col.names = NULL,
      caption   = "Notation for the MILP model.",
      label     = "tab-notation-milp") |>
  kable_styling(
    font_size     = 12,
    latex_options = c("hold_position", "striped"),
    full_width    = FALSE
  ) |>
  column_spec(1, width = "5cm") |>
   column_spec(2, width = "9cm") |>
   pack_rows("Sets \\& Indices",    1,  11,
             bold = TRUE, italic = FALSE, hline_before = TRUE, escape = FALSE) |>
   pack_rows("Decision Variables", 12,  15,
             bold = TRUE, italic = FALSE, hline_before = TRUE) |>
   pack_rows("Parameters",         16,  29,
             bold = TRUE, italic = FALSE, hline_before = TRUE)

tbl <- gsub(
  "p\\{(5\\.8cm|8\\.5cm)\\}",
  ">\\{\\\\raggedright\\\\arraybackslash\\}p\\{\\1\\}",
  tbl
)

tbl
```


```{r tab-consumers, echo=FALSE, message=FALSE, warning=FALSE, cache=F}

consumers %>%
  mutate(name_pseudo = paste("consumer ", seq_len(nrow(consumers))) ) %>% 
  select(name_pseudo, demand_P1, demand_P2, demand_P3, P1, P2, P3) %>%
  mutate(across(c(demand_P1, demand_P2, demand_P3), ~ ifelse(.x == 0, "–", format(.x, big.mark = ",")))) %>%
  mutate(across(c(P1, P2, P3), ~ ifelse(.x == 0, "–", paste0(.x)))) %>%
  kbl(
    col.names = c("Consumer", "P1 (kt/yr)", "P2 (kt/yr)", "P3 (kt/yr)",
                  "P1 (€/t)", "P2 (€/t)", "P3 (€/t)"),
    caption   = "Consumer demand quantities and gate prices by product grade. P1: chemical/pulp; P2: sawlog/pulpwood; P3: energy/pellets. Demands in kt wood input per year; prices in €/t wood input at the factory gate.",
    booktabs  = TRUE,
    align     = c("l", rep("r", 6)),
    label     = "tab-consumers"
  ) %>%
  kable_styling(latex_options = c("striped", "hold_position", "scale_down"),
                font_size = 8) %>%
  add_header_above(c(" " = 1, "Annual demand" = 3, "Gate price" = 3))
```


```{r fig-supply-chain-network, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="AFS supply chain network in Saxony-Anhalt. Grey lines: German federal state boundaries. Blue lines: major rivers. Coloured polygons: potential AFS parcels grouped by site (colours encode cluster membership; no legend). Orange symbols: pre-treatment facilities (circle = sawmill, triangle = logistics hub, size proportional to processing capacity). Coloured squares: consumer sites (violet: P1 chemical/pulp, blue: P2 pulp/paper, red: P3 energy/biogas, size proportional to total demand).\\label{fig:fig-supply-chain-network}", fig.width=6.5, fig.height=5.5, out.width="\\textwidth", cache=T}

crs_map <- 25833

# ── Colour palette constants (identical to fig-results-sc-map) ────────────────
COL_SAW <- "#e65c00"
COL_HUB <- "grey40"
COL_P1  <- "#6a0dad"
COL_P2  <- "#08519c"
COL_P3  <- "#a50026"

# ── Map extent: defined in WGS84, transformed to projected CRS ───────────────
bbox_wgs84 <- st_as_sfc(
  st_bbox(c(xmin = 10.6, xmax = 13.2, ymin = 50.9, ymax = 52.8),
          crs = st_crs(4326))
)
bbox_map <- st_transform(bbox_wgs84, crs_map)
bbox_num <- st_bbox(bbox_map)

# ── Base-map layers ────────────────────────────────────────────────────────────
de_states <- ne_states(country = "Germany", returnclass = "sf") %>%
  st_transform(crs_map) %>%
  st_intersection(bbox_map)

rivers_eu <- ne_download(
  scale    = 10,
  type     = "rivers_lake_centerlines",
  category = "physical",
  returnclass = "sf"
) %>%
  st_transform(crs_map) %>%
  st_intersection(bbox_map)

# ── Cluster colour palette (no legend printed) ────────────────────────────────
cluster_assig <- afs_workspace$site_cluster_assig %>%
  select(site_id, hac_cluster) %>%
  mutate(hac_cluster = as.factor(hac_cluster))

n_clusters <- length(unique(cluster_assig$hac_cluster))

cluster_pal <- setNames(
  if (n_clusters <= 12) {
    colorRampPalette(RColorBrewer::brewer.pal(min(n_clusters, 12), "Set3"))(n_clusters)
  } else {
    scales::viridis_pal(option = "turbo")(n_clusters)
  },
  levels(cluster_assig$hac_cluster)
)

# ── Site polygons: join cluster assignment, project ───────────────────────────
sites_sf_clust <- sites_sf %>%
  mutate(site_id = seq_len(nrow(sites_sf))) %>%
  left_join(cluster_assig, by = "site_id") %>%
  st_transform(crs_map)

# merge opportunity cost of 
tmp.opp <- sites_clustered %>% 
  select(site_id, C_opp) %>% 
  rename(hac_cluster = site_id) %>% 
  mutate(hac_cluster = as.factor(hac_cluster))

sites_sf_clust <- sites_sf_clust %>% 
  left_join(tmp.opp,by = "hac_cluster")

# ── Consumers & storages ───────────────────────────────────────────────────────
consumers_plt <- consumers %>%
  mutate(
    total_demand = demand_P1 + demand_P2 + demand_P3,
    kategorie = case_when(
      demand_P1 >= demand_P2 & demand_P1 >= demand_P3 & demand_P1 > 0 ~ "P1: stem/chemical",
      demand_P2 >= demand_P1 & demand_P2 >= demand_P3 & demand_P2 > 0 ~ "P2: branch/pulp",
      demand_P3 > 0                                                     ~ "P3: residue/energy",
      TRUE                                                              ~ "Other"
    ),
    pt_size = pmin(4.5, pmax(1.5, 1.5 + log1p(total_demand) * 0.55))
  )

storages_plt <- storages %>%
  mutate(
    #subtype = if_else(type == "Hub", "Logistics Hub", "Sawmill"),
    subtype = if_else(type == "Hub", "Hub", "Hub"),
    pt_size = 1.5
  )

stor_sf <- st_as_sf(storages_plt, coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(crs_map)

cons_sf <- st_as_sf(consumers_plt, coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(crs_map)

# color range C_opp
copp_range <- range(sites_sf_clust$C_opp, na.rm = TRUE)
copp_lim   <- c(min(copp_range[1], -1), max(copp_range[2], 1))

# ── Plot ───────────────────────────────────────────────────────────────────────
ggplot() +

  # Base map
  geom_sf(data = de_states,
          fill = "grey96", colour = "grey55", linewidth = 0.35) +
  geom_sf(data = rivers_eu,
          colour = "#4292c6", linewidth = 0.45, alpha = 0.7) +

  # KUP site polygons coloured by cluster — legend suppressed via guide = "none"
  geom_sf(data      = sites_sf_clust,
          aes(fill  = C_opp),
          colour    = "grey30",
          linewidth = 0.05,
          alpha     = 0.65,
          show.legend = FALSE) +
  scale_fill_gradient2(
  name     = "Opportunity cost (€/ha)",
  low      = "#1a9641",   # green  → low / negative C_opp
  mid      = "#ffffbf",   # pale yellow → midpoint
  high     = "#d7191c",   # red    → high / positive C_opp
  midpoint = 0,           # centre the diverging scale on zero
  na.value = "grey",
  limits   = copp_lim,        
  oob      = scales::squish,  
  guide    = guide_colourbar(
    barwidth       = 0.5,
    barheight      = 4,
    title.position = "top"
  ) 
  ) +
  # scale_fill_manual(
  #   values = cluster_pal,
  #   guide  = "none"          # colours visible on map; no legend entry printed
  # ) +

  # Pre-treatment facilities — new colour + shape scale after cluster fill
  ggnewscale::new_scale_colour() +
  ggnewscale::new_scale("shape") +
  ggnewscale::new_scale_fill() +
  geom_sf(data = stor_sf,
          aes(size = pt_size, colour = subtype),
          stroke = 1.2, alpha = 0.92, show.legend = T) +
   scale_colour_manual(
    name   = "Logistics hubs",
      values = c("Hub" = COL_HUB)
    ) +
  # scale_shape_manual(
  #   name   = "Pre-treatment",
  #   values = c("Hub" = 17)
  # ) +

  # Consumer sites — fill scale registered after new_scale_fill() above
  geom_sf(data  = cons_sf,
          aes(fill = kategorie, size = pt_size),
          shape = 22, colour = "white", stroke = 0.45,
          alpha = 0.88, show.legend = TRUE) +
  scale_fill_manual(
    name   = "Consumers",
    values = c(
      "P1: stem/chemical" = COL_P1,
      "P2: branch/pulp"    = COL_P2,
      "P3: residue/energy" = COL_P3,
      "Other"                = "grey60"
    )
  ) +

  scale_size_continuous(range = c(1.5, 5), guide = "none") +
  guides(
    fill   = guide_legend(order = 1,
                          override.aes = list(size = 3.5, stroke = 0.4,
                                              colour = "white")),
  colour = guide_legend(order = 2,
                        override.aes = list(shape  = 16,      
                                            size   = 4,
                                            stroke = 0,       
                                            colour = COL_HUB)),
    shape  = guide_legend(order = 2,
                          override.aes = list(size = 4, stroke = 1.5))
  ) +

  coord_sf(
    xlim   = c(bbox_num["xmin"], bbox_num["xmax"]),
    ylim   = c(bbox_num["ymin"], bbox_num["ymax"]),
    expand = FALSE
  ) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid      = element_line(colour = "grey88", linewidth = 0.25),
    legend.position = "right",
    legend.key.size = unit(0.40, "cm"),
    legend.text     = element_text(size = 6),
    legend.title    = element_text(size = 7.5, face = "bold"),
    axis.text       = element_text(size = 7)
  )
```


```{r tab-milp-params, echo=FALSE, message=FALSE, warning=FALSE, cache=T}
params_df <- data.frame(
  Symbol          = c("$|\\mathcal{T}|$", "$|\\mathcal{I}|$", "$|\\mathcal{J}|$",
                      "$|\\mathcal{K}|$", "$|\\mathcal{P}|$",
                      "$A^{\\min}$", "$A^{\\max}$",
                      "$c^{tr\\text{-}raw}$", "$c^{tr\\text{-}pre}$",
                      "$c^{est}_i$",
                      "$c^{harv}_i$",
                      "$c^{main}_i$",
                      "$c_i^{opp}$",
                      "$R_{P1}$", "$R_{P2}$", "$R_{P3}$"),
  Description     = c("Planning periods", "Sites",
                      "Pre-treatment / storage nodes", "Consumers", "Product grades",
                      "Minimum harvest age (years)", "Maximum stand age (years)",
                      "Raw transport cost", "Pre-processed transport cost",
                      "Establishment cost rate",
                      "harvest cost rate",
                      "Maintenance cost rate",
                      "aver. opportunity cost",
                      "Mean gate price P1", "Mean gate price P2", "Mean gate price P3"),
  Value           = c(milp_instance$n_periods, milp_instance$n_sites, nrow(afs_workspace$storages),
                      milp_instance$n_consumers, "3",
                      milp_instance$min_age, milp_instance$max_age,
                      milp_instance$c_tr_raw, milp_instance$c_tr_pre,
                      round(mean(milp_instance$sites$C_est)),
                      round(mean(milp_instance$sites$C_harv)),
                      round(mean(milp_instance$sites$C_main)),
                      round(mean(milp_instance$sites$C_opp),0),
                      round(mean(milp_instance$consumers$P1[milp_instance$consumers$P1>0]),0),
                      round(mean(milp_instance$consumers$P2[milp_instance$consumers$P2>0]),0),
                      round(mean(milp_instance$consumers$P3[milp_instance$consumers$P3>0]),0)),
  Unit            = c("yr", "–", "–", "–", "–",
                      "yr", "yr",
                      "€ t$^{-1}$ km$^{-1}$", "€ t$^{-1}$ km$^{-1}$",
                      "€ ha$^{-1}$ yr$^{-1}$",
                      "€ ha$^{-1}$ yr$^{-1}$",
                      "€ ha$^{-1}$ yr$^{-1}$",
                      "€ ha$^{-1}$ yr$^{-1}$",
                      "€ t$^{-1}$", "€ t$^{-1}$", "€ t$^{-1}$"),
  stringsAsFactors = FALSE
)

params_df %>%
  kbl(
    col.names = c("Parameter", "Description", "Base-case value", "Unit"),
    caption   = "Parameter values for the \\problem case study. Transport costs refer to road distance",
    booktabs  = TRUE,
    escape    = FALSE,
    align     = c("l", "l", "r", "l"),
    label     = "tab-milp-params"
  ) %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 16) %>%
  pack_rows("Index sets", 1, 5) %>%
  pack_rows("Biomass dynamics", 6, 7) %>%
  pack_rows("Cost parameters", 8, 13) %>%
  pack_rows("Revenue parameters", 14, 16)
```



```{r fig-results-harvest-gantt, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Total production quantity of biomass types and its industrial usage over planning horizon (in tons).\\label{fig:fig-results-harvest-gantt}", fig.width=7, fig.height=5.5, dev="pdf", out.width="\\textwidth", cache=T, cache.extra = c("ext","sites_clustered", "storages" , "customers", "milp_instance")}


# ── Labels & colours ──────────────────────────────────────────────────────────
# Source side (axis 1): biomass fractions delivered from sites → hubs (Xij)
#   product index 1 = Stems, 2 = Branches, 3 = Residues  [§2.2, f_p(t)]
SRC_LABEL <- c(
  "1" = "Stems",
  "2" = "Branches",
  "3" = "Residues"
)

# Delivery side (axis 2): consumer categories receiving from hubs (Xjk)
#   del_product 1 = Industrial (P1), 2 = Pulp & Paper (P2), 3 = Energy (P3)
DEL_LABEL <- c(
  "1" = "Industrial",
  "2" = "Pulp & Paper",
  "3" = "Energy"
)

# Fill colours for source strata — must match paper Fig. max3-growth
P_COLS_SRC <- c(
  "Stems"    = "#1b9e77",
  "Branches" = "#d95f02",
  "Residues" = "#7570b3"   
)

# ── Aggregate Xjk: src_product × del_product, summed over all periods/consumers

flow_df <- ext$Xjk %>%
  group_by(src_product, del_product) %>%
  summarise(volume = sum(value, na.rm = TRUE), .groups = "drop") %>%
  filter(volume > 0) %>%
  mutate(
    # axis1 = biomass fraction (source side)
    src = factor(SRC_LABEL[as.character(src_product)], levels = SRC_LABEL),
    # axis2 = consumer category (delivery/sink side)
    del = factor(DEL_LABEL[as.character(del_product)], levels = DEL_LABEL)
  )

# ── Plot ──────────────────────────────────────────────────────────────────────
ggplot(flow_df,
       aes(axis1 = src, axis2 = del, y = volume, fill = src)) +

  geom_alluvium(
    alpha      = 0.72,
    width      = 1 / 3,
    knot.pos   = 0.45,
    curve_type = "sigmoid"
  ) +

  geom_stratum(
    width     = 1 / 3,
    fill      = "grey94",
    colour    = "grey50",
    linewidth = 0.3
  ) +

  geom_text(
    stat       = "stratum",
    aes(label  = after_stat(stratum)),
    size       = 2.8,
    colour     = "grey15",
    lineheight = 0.88
  ) +

  # Axis header annotations — placed below the stratum columns
  # x = 1 (axis1 = Feedstock fraction), x = 2 (axis2 = End use)
  annotate("text",
    x        = c(1, 2),
    y        = -Inf,
    label    = c("Feedstock", "End use"),
    vjust    = 1.6,
    size     = 3.0,
    fontface = "bold",
    colour   = "grey30"
  ) +

  scale_fill_manual(
    name   = "Biomass fraction",
    values = P_COLS_SRC          # only 3 source entries; no ghost colours
  ) +

  # ggalluvial uses integer axis positions (1, 2, …); do NOT set
  # limits = c("src", "del") — that breaks stratum positioning
  scale_x_discrete(
    expand = expansion(add = c(0.25, 0.25))  # widen margins for stratum labels
  ) +

  scale_y_continuous(
    name   = "quantity (t fresh biomass)",
    labels = comma_format(big.mark = ".", decimal.mark = ",")
  ) +

  coord_cartesian(clip = "off") +   # needed for below-axis annotations
  theme_minimal(base_size = 9) +
  theme(
    axis.title.x       = element_blank(),
    axis.text.x        = element_blank(),
    axis.ticks.x       = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    legend.position    = "none",
    legend.key.size    = unit(0.42, "cm"),
    plot.margin        = margin(t = 6, r = 6, b = 28, l = 6)
  )
```


<!-- ```{r fig-consumer-demand, cache=F, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Total biomass demand by consumer type and grade (P1 chemical, P2 pulp \\& paper, P3 energy/biogas). Bars show aggregated annual demand in kt wood input per year across all individual plants within each consumer type.", fig.width = 5.5, fig.height = 2.8, out.width="\\textwidth"} -->

<!-- # Re-use consumers_typed from above -->
<!-- # create unique types of consumers -->

<!-- consumers_typed <- consumers %>% -->
<!--   mutate( -->
<!--     type = case_when( -->
<!--       demand_P1 >= demand_P2 & demand_P1 >= demand_P3 & demand_P1 > 0 ~ "Industrial", -->
<!--       demand_P2 >= demand_P1 & demand_P2 >= demand_P3 & demand_P2 > 0 ~ "Paper & pulp", -->
<!--       demand_P3 > 0 ~ "Biogas", -->
<!--       TRUE          ~ "Other" -->
<!--     ) -->
<!--   ) -->


<!-- # Aggregate and pivot to long -->
<!-- demand_agg <- consumers_typed %>% -->
<!--   group_by(type) %>% -->
<!--   summarise( -->
<!--     P1 = sum(demand_P1), -->
<!--     P2 = sum(demand_P2), -->
<!--     P3 = sum(demand_P3), -->
<!--     .groups = "drop" -->
<!--   ) %>% -->
<!--   pivot_longer(cols = c(P1, P2, P3), -->
<!--                names_to  = "grade", -->
<!--                values_to = "demand_kt") %>% -->
<!--   filter(demand_kt > 0) %>% -->
<!--   mutate( -->
<!--     grade = factor(grade, levels = c("P1", "P2", "P3"), -->
<!--                    labels = c("P1 – Chemical", "P2 – Pulp", "P3 – Energy")), -->
<!--     type  = factor(type, -->
<!--                    levels = c("Industrial", "Paper & pulp", "Biogas")) -->
<!--   ) -->

<!-- GRADE_COLS <- c("P1 – Chemical" = "#1b7837", -->
<!--                 "P2 – Pulp"     = "#4393c3", -->
<!--                 "P3 – Energy"   = "#d6604d") -->

<!-- ggplot(demand_agg, aes(x = type, y = demand_kt, fill = grade)) + -->
<!--   geom_col(position = "dodge", width = 0.65, colour = "white", linewidth = 0.2) + -->
<!--   geom_text(aes(label = round(demand_kt, 0)), -->
<!--             position = position_dodge(width = 0.65), -->
<!--             vjust = -0.35, size = 2.4, colour = "grey30") + -->
<!--   scale_fill_manual(values = GRADE_COLS, name = "Biomass grade") + -->
<!--   scale_y_continuous( -->
<!--     expand = expansion(mult = c(0, 0.15)), -->
<!--     name   = "Annual demand (kt wood input / yr)" -->
<!--   ) + -->
<!--   labs(x = NULL) + -->
<!--   theme_minimal(base_size = 9) + -->
<!--   theme( -->
<!--     panel.grid.major.x = element_blank(), -->
<!--     panel.grid.minor   = element_blank(), -->
<!--     legend.position    = "bottom", -->
<!--     legend.key.size    = unit(0.35, "cm"), -->
<!--     legend.text        = element_text(size = 7), -->
<!--     axis.text.x        = element_text(size = 8) -->
<!--   ) -->
<!-- ``` -->


<!-- ```{r fig-results-inventory, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="End-of-period inventory levels at storage/pre-treatment hubs, disaggregated by product type. Each panel corresponds to one active hub.", fig.width=7, fig.height=3.5, dev="pdf", out.width="\\textwidth", cache=F, cache.extra = c("ext","sites_clustered", "storages" , "customers", "milp_instance")} -->

<!-- if( nrow(ext$S) > 0){ -->
<!-- inv_ts <- ext$S %>% -->
<!--   mutate(p = factor(p, labels = c("Stem", "Branches", "Residues")[seq_len(max(p))])) %>% -->
<!--   left_join( -->
<!--     storages %>% mutate(hub_id = row_number()) %>% select(hub_id, hub_name = name), -->
<!--     by = "hub_id" -->
<!--   ) %>% -->
<!--   filter(!is.na(value), value > 0) -->

<!-- ggplot(inv_ts, aes(x = period, y = value, fill = p)) + -->
<!--   geom_area(position = "stack", alpha = 0.78, colour = "white", linewidth = 0.2) + -->
<!--   facet_wrap(~hub_name, scales = "free_y") + -->
<!--   scale_fill_manual( -->
<!--     name   = "Product", -->
<!--     values = c("Stem" = "#1b9e77", "Branches" = "#d95f02", "Residues" = "#7570b3") -->
<!--   ) + -->
<!--   scale_x_continuous(breaks = seq(1, milp_instance$n_periods, by = 4)) + -->
<!--   scale_y_continuous(name = "Inventory (t)", labels = scales::comma) + -->
<!--   labs(x = "Planning period (year)") + -->
<!--   theme_minimal(base_size = 8) + -->
<!--   theme(legend.position = "bottom", -->
<!--         panel.grid.minor = element_blank(), -->
<!--         strip.text = element_text(face = "bold", size = 7))   -->
<!-- }else{ -->
<!--   NULL -->
<!-- } -->


<!-- ``` -->



