---
title: "Agroforestry SmartAgroforest: Poplar biomass modelling"
author:
  - name: "Thomas Kirschstein"
    affiliation: 
      - "Fraunhofer IKTS, Dresden, Germany"
    email: "thomas.kirschstein@ikts.fraunhofer.de"
    correspondingauthor: true
  
journal: "Journal Name"
# elsarticle options: preprint preview style, 1p or 3p for layout
classoption: "preprint,review,1p,authoryear"
date: "`r Sys.Date()`"
linenumbers: true
numbersections: true
always_allow_html: true


keywords: 
  - Agroforestry 
  - Short-rotation coppice 
  - Biomass supply chain
  - Product cascading

abstract: |
  Short-rotation agroforestry systems are increasingly promoted as a low-carbon feedstock option for bio-based value chains, but their economic viability depends on coordinated decisions on stand establishment, multi-cycle harvesting, and product cascading in spatially distributed supply chains. This paper develops a mixed-integer linear programming (MILP) model for the integrated design and operation of an agroforestry biomass supply chain with explicit age-lag constraints that link consecutive harvests at each site. The formulation couples allometric yield functions for different biomass types with a multi-product, multi-period network design model including storage, processing, and product quality cascades. The model is applied to a stylised case study of Central European poplar-based alley-cropping systems to explore trade-offs between harvest rotation length, product portfolios, and infrastructure configuration. We outline a computational experimentation plan to test the impact of allometric parameter uncertainty, price scenarios, and policy incentives on the optimal deployment of agroforestry systems in cleaner bioeconomy pathways.

output:
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
    keep_tex: true
    latex_engine: xelatex
    number_sections: true
    

header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{hyperref}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{float}
  - \usepackage{xcolor}
  - \usepackage{eurosym}
  - \newcommand{\Tc}{\mathcal{T}}
  - \newcommand{\Pc}{\mathcal{P}}
  - \newcommand{\Qc}{\mathcal{Q}}
  - \newcommand{\Sc}{\mathcal{S}}
  - \newcommand{\Ic}{\mathcal{I}}
  - \newcommand{\Jc}{\mathcal{J}}
  - \newcommand{\Kc}{\mathcal{K}}
  - \newcommand{\problem}{AFS-SCD }
  
bibliography: references.bib
csl: elsevier-harvard.csl
editor_options: 
  markdown: 
    wrap: 125
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo        = FALSE,
  message     = FALSE,
  warning     = FALSE,
  fig.align   = "center",
  fig.path    = "figures/",
  # --- additions ---
  cache       = FALSE,          # set TRUE per-chunk for heavy computations
  out.width   = "\\textwidth",  # consistent figure width for elsevier 1p
  dev         = "pdf"           # vector output for LaTeX/XeLaTeX pipeline
)

# Suppress noisy package startup messages globally
suppressPackageStartupMessages({
  library(magick)
  library(dplyr)
  library(plotly)
  library(ggplot2)
  library(tidyr)
  library(patchwork)
  library(knitr)
  library(kableExtra)
  library(sf)
  library(tibble)
  library(stringr)
  library(leaflet)
  library(osrm)
  library(ROI)
  library(ROI.plugin.glpk)
  library(gurobi)
  library(ROI.plugin.gurobi)
  library(ggrepel)
  library(ggnewscale)
  library(rnaturalearth)
  library(rnaturalearthdata)
  library(Rcpp)
  library(RcppArmadillo)
  library(scatterpie)
  library(networkD3)
  library(htmlwidgets)
  library(webshot2)
  library(ggalluvial)
  library(scales)
  library(purrr)
  library(broom)
  library(lme4)        # mixed-effects models
  library(lmerTest)    # Satterthwaite p-values for lmer
  library(mgcv)        # GAM + GAMM
  library(gratia)      # tidy GAM diagnostics & draw()
  library(MuMIn)       # R² for mixed models (r.squaredGLMM)
  library(AICcmodavg)  # AICc comparison table
  library(performance) # model_performance(), compare_performance(
  library(broom.mixed) # tidy() for lmer objects
  library(kableExtra)
})
options(kableExtra.latex.load_packages = FALSE)
options(knitr.table.format = "latex")

```

```{r load-model-data, echo=FALSE, message=FALSE, warning=FALSE, cache=T, cache.extra = tools::md5sum(c(     "../Modell/!helper_func.r", "../Modell/!helper_instance_builder_v8a.R","../Modell/!build_AFS_milp.r",      "../Modell/build_lp_rcpp_wrapper.R","../Modell/gurobi_adapter.R","../Modell/gurobi_warmstart.R",      "../Modell/gurobi_feascheck.R","../Modell/extract_result.R","../Modell/build_lp_rcpp.cpp","afs_workspace_red.RData",      "lp_r.RDS","lp_r_result.rds","lp_instance_data.rds" ))}
source("../Modell/!afs_biomass_setup.r")
source("../Modell/!helper_func.r")
source("../Modell/!helper_instance_builder_v8a.R")
source("../Modell/!build_AFS_milp.r")
source("../Modell/build_lp_rcpp_wrapper.R")
source("../Modell/gurobi_adapter.R")
source("../Modell/gurobi_warmstart.R")
source("../Modell/gurobi_feascheck.R")
source("../Modell/extract_result.R")
#Rcpp::sourceCpp("../Modell/build_lp_rcpp.cpp")
#Rcpp::sourceCpp("../Modell/build_and_solve_afs_milp.cpp")
Rcpp::sourceCpp("../Modell/build_and_solve_afs_lp_v11.cpp")


# ── 1. Hilfsfunktion: Site-Level KPIs für ein Szenario extrahieren ──────────
extract_site_profit <- function(res, scenario_name = NA_character_) {
  
  milp_instance <- res$milp_instance
  ext           <- res$ext
  setting       <- res$setting
  
  # Szenario-Parameter
  scen_opp      <- if (!any(is.na(setting))) setting$opp      else 150
  scen_cost_log <- if (!any(is.na(setting))) setting$cost.log else 1.0   
  scen_cost_est <- if (!any(is.na(setting))) setting$cost.est else 1.0   
  scen_revenue  <- if (!any(is.na(setting))) setting$revenue  else 1.0
  
  # Aktive Standorte
  active_site_ids <- ext$z %>%
    filter(arc_type == "harvest") %>%
    pull(site_id) %>% unique()
  
  site_meta <- milp_instance$sites %>%
    mutate(site_id = row_number()) %>%
    filter(site_id %in% active_site_ids)
  
  # Preistabelle
  price_df <- as.data.frame(milp_instance$consumer_prices) %>%
    setNames(c("consumer_id", "del_product", "price")) %>%
    mutate(consumer_id = as.integer(consumer_id),
           del_product  = as.integer(del_product))
  
  # ── Distanzmatrizen ────────────────────────────────────────────────────────
  dist_ij_m <- as.matrix(milp_instance$d_ij)
  dist_jk_m <- as.matrix(milp_instance$d_jk)
  
  # Hub-Anteile pro Standort und Periode (für Kostenzurechnung)
  hub_site_share <- ext$Xij %>%
    group_by(hub_id, period) %>%
    mutate(hub_total = sum(value)) %>%
    ungroup() %>%
    mutate(share = ifelse(hub_total > 0, value / hub_total, 0)) %>%
    select(site_id, hub_id, period, share)
  
  # ── Kosten je Standort ─────────────────────────────────────────────────────
  sc_est <- ext$z %>% filter(arc_type == "establishment") %>%
    left_join(site_meta, by = "site_id") %>%
    mutate(cost = C_est * value) %>%
    group_by(site_id) %>% summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
    mutate(component = "Establishment")
  
  sc_harv <- ext$z %>% filter(arc_type == "harvest") %>%
    left_join(site_meta, by = "site_id") %>%
    mutate(cost = C_harv * value) %>%
    group_by(site_id) %>% summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
    mutate(component = "Harvesting")
  
  sc_main <- ext$z %>% filter(arc_type == "harvest") %>%
    mutate(age_len = t - s) %>%
    left_join(site_meta, by = "site_id") %>%
    mutate(cost = C_main * value * age_len) %>%
    group_by(site_id) %>% summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
    mutate(component = "Maintenance")
  
  sc_opp <- ext$z %>% filter(arc_type == "harvest") %>%
    mutate(age_len = t - s) %>%
    left_join(site_meta, by = "site_id") %>%
    mutate(cost = C_opp * value * age_len) %>%
    group_by(site_id) %>% summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
    mutate(component = "Opportunity")
  
  sc_tr_raw <- ext$Xij %>%
    mutate(dist_km = mapply(function(i, j) dist_ij_m[i, j], site_id, hub_id),
           cost    = milp_instance$c_tr_raw * dist_km * value) %>%
    group_by(site_id) %>% summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
    mutate(component = "Transport_raw")
  
  sc_tr_pre <- ext$Xjk %>%
    mutate(dist_km  = mapply(function(j, k) dist_jk_m[j, k], hub_id, consumer_id),
           cost_jkt = milp_instance$c_tr_pre * dist_km * value) %>%
    left_join(hub_site_share, by = c("hub_id", "period"), relationship = "many-to-many") %>%
    mutate(site_cost = cost_jkt * share) %>%
    group_by(site_id) %>% summarise(cost = sum(site_cost, na.rm = TRUE), .groups = "drop") %>%
    filter(!is.na(site_id)) %>%
    mutate(component = "Transport_pre")
  
  sc_stor <- ext$S %>%
    left_join(milp_instance$storages %>% mutate(hub_id = row_number()) %>% select(hub_id, c_stor),
              by = "hub_id") %>%
    mutate(stor_cost = c_stor * value) %>%
    select(hub_id, period, stor_cost) %>%
    left_join(hub_site_share, by = c("hub_id", "period"), relationship = "many-to-many") %>%
    mutate(site_cost = stor_cost * share) %>%
    group_by(site_id) %>% summarise(cost = sum(site_cost, na.rm = TRUE), .groups = "drop") %>%
    filter(!is.na(site_id)) %>%
    mutate(component = "Storage")
  
  # ── Erlöse je Standort ─────────────────────────────────────────────────────
  site_revenue <- ext$Xjk %>%
    left_join(price_df, by = c("consumer_id", "del_product")) %>%
    mutate(rev_jkt = value * coalesce(price, 0)) %>%
    left_join(hub_site_share, by = c("hub_id", "period"), relationship = "many-to-many") %>%
    mutate(site_rev = rev_jkt * share) %>%
    group_by(site_id) %>%
    summarise(revenue = sum(site_rev, na.rm = TRUE), .groups = "drop") %>%
    filter(!is.na(site_id))
  
  # ── Normierung: €/ha/a (nur aktive Produktionsjahre) ──────────────────────
  #est_p  <- ext$z %>% filter(arc_type == "establishment") %>% select(site_id, t_est = t)
  #term_p <- ext$z %>% filter(arc_type == "termination")  %>% select(site_id, t_term = s)
  active_p <- ext$z %>% filter(arc_type == "establishment") %>% 
    group_by(site_id) %>% 
    summarise(tot_value = sum(value))
# 
  est_p  <- ext$z %>% filter(arc_type == "establishment") %>% group_by(site_id) %>% summarise(t_est = mean(sum(t*value)/sum(value)))
  term_p <- ext$z %>% filter(arc_type == "termination")  %>% group_by(site_id) %>% summarise(t_term = mean(sum(s*value)/sum(value)))
 
  site_norm <- site_meta %>%
    #select(site_id, area_ha) %>%
    left_join(active_p,  by = "site_id") %>%
    left_join(est_p,  by = "site_id") %>%
    left_join(term_p, by = "site_id") %>%
    mutate(t_term       = ifelse(is.na(t_term), milp_instance$n_periods, t_term),
           active_years = pmax(t_term - t_est, 1L),
           denom        = tot_value * active_years)
  
  all_costs <- bind_rows(sc_est, sc_harv, sc_main, sc_opp,
                         sc_tr_raw, sc_tr_pre, sc_stor) %>%
    group_by(site_id) %>%
    summarise(total_cost = sum(cost, na.rm = TRUE), .groups = "drop")
  
  # ── Durchschnittliche Distanzen je Standort ────────────────────────────────
  # (a) Ø Distanz Standort → Hub (gewichtet nach Liefervolumen)
  avg_dist_to_hub <- ext$Xij %>%
    mutate(dist_km = mapply(function(i, j) dist_ij_m[i, j], site_id, hub_id)) %>%
    group_by(site_id) %>%
    summarise(avg_dist_hub_km = weighted.mean(dist_km, w = pmax(value, 0), na.rm = TRUE),
              .groups = "drop")
  
  # (b) Ø Distanz Hub → Kunden (anteilig auf Standort zurückgerechnet)
  avg_dist_to_consumer <- ext$Xjk %>%
    mutate(dist_km = mapply(function(j, k) dist_jk_m[j, k], hub_id, consumer_id)) %>%
    left_join(hub_site_share, by = c("hub_id", "period"), relationship = "many-to-many") %>%
    filter(!is.na(site_id), share > 0) %>%
    group_by(site_id) %>%
    summarise(avg_dist_consumer_km = weighted.mean(dist_km, w = pmax(share * value, 0), na.rm = TRUE),
              .groups = "drop")
  
  # ── Weitere strukturelle Variablen ─────────────────────────────────────────
  # (c) Ø Erntezyklus-Länge (Rotationsalter) je Standort
  avg_rotation <- ext$z %>%
    filter(arc_type == "harvest") %>%
    mutate(cycle_len = t - s) %>%
    group_by(site_id) %>%
    summarise(avg_rotation_yr = sum(cycle_len*value/sum(value), na.rm = TRUE),
              n_harvests       = n(),
              .groups = "drop")
  
  # (d) Produktmix-Index: Anteil hochwertiger Biomasse (P1 Stämme) an Gesamtlieferung
  p1_share <- ext$Xij %>%
    group_by(site_id) %>%
    summarise(
      vol_total = sum(value, na.rm = TRUE),
      vol_p1    = sum(value[p == 1], na.rm = TRUE),
      .groups   = "drop"
    ) %>%
    mutate(share_p1 = ifelse(vol_total > 0, vol_p1 / vol_total, NA_real_)) %>%
    select(site_id, share_p1, vol_total)
  
  # ── Zusammenführen ─────────────────────────────────────────────────────────
  result_df <- site_norm %>%
    left_join(site_revenue,          by = "site_id") %>%
    left_join(all_costs,             by = "site_id") %>%
    left_join(avg_dist_to_hub,       by = "site_id") %>%
    left_join(avg_dist_to_consumer,  by = "site_id") %>%
    left_join(avg_rotation,          by = "site_id") %>%
    left_join(p1_share,              by = "site_id") %>%
    mutate(
      profit_ha_yr         = (coalesce(revenue, 0) - coalesce(total_cost, 0)) / denom,
      scenario             = scenario_name,
      opp_cost             = scen_opp,
      cost_log_level       = scen_cost_log,   # renamed: was cost_level
      cost_est_level       = scen_cost_est,   # new
      revenue_level        = scen_revenue,
      opp_cost_site        = C_opp
    ) %>%
    select(site_id, scenario, profit_ha_yr,
           opp_cost, opp_cost_site, cost_log_level, cost_est_level, revenue_level,
           avg_dist_hub_km, avg_dist_consumer_km,
           avg_rotation_yr, n_harvests, share_p1,
           area_ha, active_years, n_sites)
  
  return(result_df)
}


#load("../Modell/afs_workspace.RData")
load("afs_workspace_red.RData")

sites_sf   <- afs_workspace$sites_sf
sites      <- afs_workspace$sites
sites_clustered      <- afs_workspace$sites_clustered
storages   <- afs_workspace$storages
consumers  <- afs_workspace$consumers

# scale down industrial demand quantities by dividing by 5
consumers <- consumers %>%
  mutate(demand_P1 = demand_P1 / 5)

# drop small Biogas plants --> compute total demand and drop all consumer with D<2
consumers <- consumers %>%
  mutate(total_demand = demand_P1 + demand_P2 + demand_P3) %>%
  filter(total_demand >= 1.5) %>%
  select(-total_demand) 


dist_jk <- afs_workspace$dist_jk[,consumers$consumer_id]

yields_by_age <- build_scenario_ts(ages = 1:20 , N = 250, C_site = 6.500, k = 0.194, t0 = 9.7, label = "Conservative (250 trees/ha)")


yields_by_age <- yields_by_age %>% 
  # drop column scenario from yields_by_age
  select(-scenario) %>% 
  # make long --> bring columns Merch. stem Merch. branch Residue to column "product" and values to "yield_ha"
  pivot_longer(-1 , names_to = "product", values_to = "yield_ha") %>% 
  # rename entries in "product" to 1:3
  mutate(product = as.integer(as.character(factor(product, labels = c(2,1,3) ))))

# scale yields to fresh biomass
yields_by_age$yield_ha <- yields_by_age$yield_ha / (1-0.5)


set.seed(123578)
# generate random vector of opportunity cost for clustered sites in range [0,600]
c.opp.vec <- rnorm(nrow(sites_clustered), 1, .75)


sites_clustered <- sites_clustered %>%
   mutate(
     C_est  = 2500,   # €/ha  establishment (Section 2.1 mid-range)
     C_harv =  150,    # €/ha  harvest cost  (Section 2.1 mid-range)
     C_main = 10 , #  €/ha  maintenance cost  (Section 2.1 mid-range)
     C_opp = 150*c.opp.vec, #  €/ha  opportunity cost  (Section 2.1 mid-range)
   ) #%>%   # change column name area_ha to area if necessary  rename(area_ha = area)


# # change storage costs
# storages$c_stor <- storages$c_stor
# # change capacities
 storages$CAP_stor <- storages$CAP_stor * 100000
 storages$CAP_proc <- storages$CAP_proc * 100000
# 
data_obj <- list(
   sites    = sites_clustered,
  #sites    = sites_filt,
   storages = storages,
   consumers = consumers,
   dist_ij  = afs_workspace$dist_ij_clust,
   #dist_ij  = dist_ij_filt$distance_matrix_km,
   dist_jk  = dist_jk,
   yields_by_age = yields_by_age
 )
# rescale demands in consumers to tons
data_obj$consumers[,c("demand_P1","demand_P2","demand_P3")]  <- data_obj$consumers[,c("demand_P1","demand_P2","demand_P3")] * 1000

# 
# # ── 3. PARAMS OBJECT — all scalar parameters ──────────────────────────────────
params_obj <- list(
   n_periods  = 40,
   max_age    = 20L,    # A_max: LRAFS upper rotation bound
   min_age    =  3L,    # A_min: SRAFS lower rotation bound
   c_tr_raw   = 0.08,   # €/(t·km)  raw biomass  site → storage
   c_tr_pre   = 0.06   # €/(t·km)  chips        storage → consumer
 )
# 
# ── 4. generate instance ─────────────────────────────────────
milp_instance <- build_optimization_instance(
    data         = data_obj,
    params       = params_obj
  ) 

#result <- build_and_solve_afs_lp_v11(milp_instance, 
#                           gurobi_params = list(TimeLimit = 300),
#                           verbose = TRUE)
# 
#saveRDS(result, "result_lp_v11.rds")

result <- readRDS("result_lp_v11.rds")

ext <- result$solution_vector

#  renaming

ext$Xjk <- ext$Xjk %>% 
  rename(src_product = p, del_product = q, period = t, hub_id = j, consumer_id = k)

ext$Xij <- ext$Xij %>% 
  rename(period = t, site_id = i, hub_id = j)

ext$z <- ext$z %>% 
  rename(arc_type = type, site_id = i )

ext$S <- ext$S %>% 
  rename(hub_id = j, period = t)

# sensitivity analysis ##########################################
# 
# scen.mat <- expand.grid(opp = c(75, 150, 300), cost.log = c(.5, 1.5, 2), cost.est = c(.5, 1.5, 2), revenue = c(.5, .75, 1.25)  )
# nb.rep <- 10
# 
# res.sens.list <- vector("list", nrow(scen.mat)*nb.rep)
# 
# for(i in seq_len(nrow(scen.mat))){
#   tmp.setting <- list(
#     opp = scen.mat$opp[i], 
#     cost.log = scen.mat$cost.log[i],
#     cost.est = scen.mat$cost.est[i],
#     revenue = scen.mat$revenue[i]
#     )
# 
#   # storage cost
#   tmp.storages <- storages
#   tmp.storages$c_stor <- tmp.storages$c_stor * tmp.setting$cost.log
#   # consumers
#   tmp.consumers <- consumers
#   tmp.consumers$P1 <- tmp.consumers$P1 * tmp.setting$revenue
#   tmp.consumers$P2 <- tmp.consumers$P2 * tmp.setting$revenue
#   tmp.consumers$P3 <- tmp.consumers$P3 * tmp.setting$revenue
# 
#   # # ── 3. PARAMS OBJECT — all scalar parameters ──────────────────────────────────
#   tmp.params_obj <- list(
#        n_periods  = 40,
#        max_age    = 20L,    # A_max: LRAFS upper rotation bound
#        min_age    =  3L,    # A_min: SRAFS lower rotation bound
#        c_tr_raw   = 0.08 * tmp.setting$cost.log,   # €/(t·km)  raw biomass  site → storage
#        c_tr_pre   = 0.06 * tmp.setting$cost.log   # €/(t·km)  chips        storage → consumer
#      )
#     #
# 
#   for(j in seq_len(nb.rep)){
#     tmp.c.opp.vec <- rnorm(nrow(sites_clustered), 1, .75)
# 
#     # opportunity cost
#     tmp.c.opp.vec <- tmp.c.opp.vec * tmp.setting$opp
# 
#     # site details
#     tmp.sites_clustered <- sites_clustered %>%
#        mutate(
#          C_est  = 2500 * tmp.setting$cost.est,   # €/ha  establishment 
#          C_harv = 150 * tmp.setting$cost.log,    # €/ha  harvest cost  
#          C_main = 10 * tmp.setting$opp/150, #  €/ha  maintenance cost  
#          C_opp  = tmp.c.opp.vec #  €/ha  opportunity cost  
#        )
#     #
#     tmp.data_obj <- list(
#        sites    = tmp.sites_clustered,
#        storages = tmp.storages,
#        consumers = tmp.consumers,
#        dist_ij  = afs_workspace$dist_ij_clust,
#        dist_jk  = dist_jk,
#        yields_by_age = yields_by_age
#      )
# 
#     # rescale demands in consumers to tons
#     tmp.data_obj$consumers[,c("demand_P1","demand_P2","demand_P3")]  <- tmp.data_obj$consumers[,c("demand_P1","demand_P2","demand_P3")] * 1000
# 
#     # ── 4. generate instance ─────────────────────────────────────
#     tmp.milp_instance <- build_optimization_instance(
#         data         = tmp.data_obj,
#         params       = tmp.params_obj
#       )
#     #
#     tmp.result <- build_and_solve_afs_lp_v11(
#        tmp.milp_instance,
#        gurobi_params = list(
#          TimeLimit = 600),
#        verbose = TRUE)
#     #
# 
#     tmp.ext <- tmp.result$solution_vector
#     #  renaming
# 
#     tmp.ext$Xjk <- tmp.ext$Xjk %>%
#       rename(src_product = p, del_product = q, period = t, hub_id = j, consumer_id = k)
# 
#     tmp.ext$Xij <- tmp.ext$Xij %>%
#       rename(period = t, site_id = i, hub_id = j)
# 
#     tmp.ext$z <- tmp.ext$z %>%
#       rename(arc_type = type, site_id = i )
# 
#     tmp.ext$S <- tmp.ext$S %>%
#       rename(hub_id = j, period = t)
# 
#     res.sens.list[[(i-1)*nb.rep + j]] <- list(setting = tmp.setting, milp_instance = tmp.milp_instance, ext = tmp.ext)
#   }
# 
# }
# 
# saveRDS(res.sens.list, "res_sens_list_lp_v11.rds")

res.sens.list <- readRDS("res_sens_list_lp_v11.rds")

# ── 2. Benchmark + alle Sensitivitätsszenarien zusammenführen ────────────────
res_bench_full <- list(
  setting       = list(opp = 150, cost.log = 1.0, cost.est = 1.0, revenue = 1.0),
  milp_instance = milp_instance,
  ext           = ext,
  result        = result
)

df_bench <- extract_site_profit(res_bench_full, "Bench.")

df_bench <- df_bench %>% mutate(opp_cost = opp_cost/150)

df_sens <- purrr::imap_dfr(
  res.sens.list,
  ~ extract_site_profit(.x, scenario_name = paste0("S", .y))
)

df_sens <- df_sens %>% mutate(opp_cost = opp_cost/150)

```

# Introduction

Agroforestry systems that integrate trees with agricultural crops
(silvoarable systems) are increasingly recognised as a promising
land-use option to reconcile climate mitigation, biodiversity
conservation, and rural income generation on the same hectare of land
[@Toensmeier2017_Carbon_Agroforestry]. Temperate agroforestry systems
(AFS) based on fast-growing species such as poplar, willow, alder, and
black locust can deliver substantial biomass yields while improving soil
quality and providing multiple ecosystem services compared to
conventional monocultures [@Huber2018_SRAFS_growth]. Biomass produced
in AFS is considered as a sustainable feedstock and can be used either
energetically, chemically or materially (e.g. in the building sector).
The latter two options are generally preferred over the energetic usage
from a cleaner production perspective as biogenic carbon is stored in
products during their life span which contributes to climate change
mitigation.

Despite substantial research on ecological advantages and pioneering
initiatives to establish AFS in practice, the share of AFS in
agricultural production systems is still small. E.g., in Germany about
1200 hectares of silvoarable AFS are established in 2025, which accounts
for about 0.02% of total agricultural land use [@DEFAF]. One of the core
challenges in establishing AFS systems at large scale in practice is a
lack of opportunities to generate substantial revenues from woody
biomasses of AFS. On the other hand, the rise of bioeconomy creates
substantial demands for woody biomass and requires a reliable and
cost-efficient feedstock supply. As woody biomasses are bulky feedstocks
with low energy density, the design of cost-efficient supply chains is
crucial to ensure the economic viability of AFS-based value chains. This
requires coordinated decisions on stand establishment, multi-cycle
harvesting, and product cascading in spatially distributed supply
chains, which can be supported by integrated optimisation models that
link stand-level growth dynamics to regional supply chain design.

In the last two decades, a large body of research has analysed biomass
supply chains for bioenergy and biorefineries using mathematical
programming models, often in the form of mixed-integer linear
programming (MILP) [@Zahraee2020_BSC_review; @Sharma2013_BSC_review].
These models typically optimise network configuration, feedstock
sourcing, and logistics under cost or profit objectives. They often
treat biomass availability as an exogenous input or rely on simplified
agricultural production decisions. Moreover, downstream processing of
biomasses is concentrated on one particular value chain e.g. bioenergy
plants consuming all produced biomasses. Recent work has started to
integrate operational details such as various biomass types, biomass
growth models and regeneration processes in multi-period supply chain
models.

The present paper contributes to this literature in three ways. First, we
propose a novel MILP formulation for AFS-based biomass supply chain
design that explicitly links establishment decisions and consecutive
harvest decisions, ensuring that harvesting decisions at each site are
consistent with minimum and maximum rotation ages. Second, we embed
allometric yield functions for different biomass types generating
age-specific yields for different woody biomass types (like stems,
branches, or leaves and twigs). Third, these different woody biomass
types can be processed in specific downstream industries, like chemical
or paper industry as well as energy sector. The model aims to provide
strategic decision support for establishing regional AFS-based
bioeconomy ecosystems connecting multiple industries and product
markets.

The remainder of the paper is structured as follows. Section
\@ref(sec:litreview) reviews the relevant literature on agroforestry
biomass systems, allometric biomass modelling, and biomass supply chain
optimisation. Section \@ref(sec:milp) summarises the MILP formulation
for the agroforestry supply chain with age-lag constraints. Section
\@ref(sec-casestudy) presents the case study instantiation, the base-case optimisation results,
and the sensitivity analysis, examining the implications for cleaner production strategies.

# Literature review {#sec:litreview}

Agroforestry biomass systems, stand-level growth models, and biomass
supply chain optimisation have each been studied extensively, but they
are rarely combined in an integrated framework that links age-dependent
stand dynamics to regional product-cascading value chains
[@Zahraee2020_BSC_review; @Sharma2013_BSC_review;
@Espinoza2017_Biorefinery_SC_review]. This section reviews (i)
agroforestry systems with emphasis on system types, economics, and the
cascading use of biomass as the primary revenue driver; (ii) biomass
growth models at tree and stand level with explicit attention to the
allometric--diameter--stand-biomass pathway; and (iii) biomass supply
chain design using mathematical programming. The section concludes by
identifying the modelling gap addressed in this paper.

## Agroforestry systems {#sec:lit-afs}

Agroforestry deliberately integrates woody perennials with crops and/or
livestock on the same management unit, with the aim of enhancing
productivity, profitability, and environmental performance
[@Nair1993_IntroAgroforestry; @Toensmeier2017_Carbon_Agroforestry]. In
the temperate zone, three distinct system types can be distinguished
based on the spatial arrangement of trees and the rotation length at
which woody biomass is harvested.

**Short-rotation coppice (SRC)** is defined as a monoculture plantation
of fast-growing woody species covering the entire management unit, with
planting densities of 8,000--20,000 cuttings\~ha$^{-1}$. After each
harvest the root system regenerates without replanting. Rotation cycles
are typically 3--7\~years and total stand lifetime is 20--25\~years
[@Trnka2008_Poplar_SRF; @Faasch2012_SRC_Economics]. SRCs are no AFSs in
the narrow sense, but it is helpful to benchmark biomass production.

**Short-rotation agroforestry systems (SRAFS)** integrate SRC-type tree
strips (typically 1--4 rows of the same fast-growing species like poplar
or willow) into an otherwise arable field in an alley-cropping
arrangement. Crop production continues in the inter-row strips
[@Huber2018_SRAFS_growth; @Graves2007_Agroforestry_Economics]. Thus,
tree strips typically cover 10--30% of the total parcel area. Rotation
lengths and total stand lifetime in SRAFS are comparable to SRCs. In
SRAFS all trees are effectively edge trees, exposed to higher light and
water availability than interior SRC trees. This boundary effect raises
individual-tree biomass accumulation but also intensifies competition
with adjacent crop rows [@Huber2018_SRAFS_growth].

**Long-rotation agroforestry systems (LRAFS)** use the same
alley-cropping spatial structure as SRAFS but with rotation cycles of
12--15\~years or longer [@Graves2007_Agroforestry_Economics;
@Nair1993_IntroAgroforestry]. At this rotation length, individual trees
develop substantial stem diameters and the proportion of high-value
timber increases markedly, enabling premium material uses (veneer,
sawlogs, engineered wood products) that are inaccessible in
short-rotation systems. Therefore, often timber-quality tree species
(such as walnut, cherry, oak, or high-grade poplar clones) are planted.

SRAFS and LRAFS are not clearly differentiated in practice. First, a
clear cut-off age is hard to define and, more importantly, the intended
use of AFS biomass is not always clear and can also change over time.
Nonetheless, when establishing AFS, the structure of AFS and its
intended use after harvest drives planting decisions and, subsequently,
economic viability. Thus, in the present paper SRAFS and LRAFS are not
considered as separate AFS systems, but the interplay of intended use,
establishment, and harvesting decisions is analysed.

## Economic rationale of landowners

The decision of a landowner to establish an AFS involves weighing
substantial up-front costs and long-term opportunity costs against a
diversified revenue stream. On the cost side, establishment requires
site preparation, planting material, and planting labour, typically
amounting to 1,500--3,500 \euro/ha for SRC or SRAFS
[@Faasch2012_SRC_Economics; @Testa2014_Poplar_Economics]. Recurrent
harvest and refitting costs (machine hire, chipping, transport to
roadside) add approximately 200--400 \euro/ha per rotation depending on
stand density and equipment availability [@Testa2014_Poplar_Economics;
@Spinelli2009_SRC_Harvesting]. During the rotation, maintenance costs of
AFS arise on a yearly basis. Next to expenses for establishment,
maintenance and harvesting, opportunity costs arise as no revenues from
crops are generated on tree strips. In regions with fertile soils and
high cereal prices, yearly crop revenues may sum up to 400--600 \euro/ha
[@Faasch2012_SRC_Economics; @Graves2007_Agroforestry_Economics]. Thus,
opportunity costs are often the dominant cost driver in SRAFS
profitability analyses.

On the **benefit side**, the primary revenue of AFS is the sale of
harvested woody biomass, whose magnitude depends critically on rotation
length, species, and site productivity. For poplar-based SRAFS in
Central Europe, typical yields range from 10 to 25 t of dry matter
biomass per ha for typical rotation lengths [@Huber2018_SRAFS_growth;
@Niemczyk2021_Poplar_Rotation]. Additionally, positive crop-yield
effects in the inter-row strips of SRAFS and LRAFS may result. Tree rows
act as windbreaks, reducing soil erosion and evapotranspiration; they
provide partial shading that moderates heat stress during drought years;
and their root systems may improve water retention in the topsoil
[@Graves2007_Agroforestry_Economics; @Grunewald2007_Robinia_AFS].
Empirical measurements in Central European alley-cropping trials
indicate that these microclimate effects can increase cereal yields in
adjacent strips by 5--15% relative to open-field controls, partially or
fully compensating the loss of arable area occupied by the tree strips
[@Graves2007_Agroforestry_Economics]. Beyond direct yield effects, AFS
deliver broader ecosystem services (like soil organic matter
accumulation, carbon sequestration, biodiversity enhancement) that may
lead to agri-environment subsidies and carbon credits
[@Toensmeier2017_Carbon_Agroforestry; @Faasch2012_SRC_Economics].

Nonetheless, economic attractiveness of AFS depends primarily on
revenues of AFS biomass. Thereby, revenues are different for the types
of biomass and in which value chain it is used. Often AFS biomasses are
used energetically, i.e. for heat and power generation, or anaerobic
digestion, which typically creates low prices per tonne of biomass.[REF]
Alternatively, woody parts of the biomass (i.e., the stem and large
branches) can also be used as pulp and paper grade material as well as
feedstock for chemical industry or as timber achieving higher prices
[@Lamers2015_Feedstock_Design] [EXTEND REFS]. Thus, fostering
AFS establishment by maximizing economic returns from AFS biomass
requires a strategic approach to product cascading that allocates
different biomass fractions to the highest-value applications, while
ensuring that the rotation length is aligned with the growth dynamics of
the stand and the market demand for different biomass types
[@Lamers2015_Feedstock_Design].

Thus, for making educated decisions on AFS establishment and harvesting,
the expected biomass quantities at harvest for stem, branches, and
residues need to be estimated. Thereby, the shares of the biomass types
change with stand age. E.g.\~the share of stem weight rises from roughly
55% at age 4 to over 75% at age 10 [@Huber2018_SRAFS_growth;
@Testa2014_Poplar_Economics]. Therefore, in Section
\@ref(sec:lit-growth) growth modelling approaches are reviewed that
capture the age-dependent growth dynamics of different biomass types in
AFS.

## Biomass growth models {#sec:lit-growth}

```{r tab-notation-growth, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr); library(kableExtra)
options(kableExtra.latex.load_packages = FALSE)
options(knitr.table.format = "latex")

notation <- tribble(
  ~Symbol, ~Description,
  # --- stand-level Gompertz ---
  "$M^{sl}(t)$",       "Total above- and belowground dry biomass per ha of AFS (t\\,DM\\,ha$^{-1}$) at stand age $t$",
  "$A(N)$",            "Stand-level Gompertz asymptote (t\\,DM\\,ha$^{-1}$); density- and site-dependent",
  "$k$",               "Gompertz growth-rate coefficient (yr$^{-1}$)",
  "$t_0$",             "Inflection age of stand biomass accumulation (yr)",
  # --- density--asymptote power law ---
  "$N$",               "Stand density (trees\\,ha$^{-1}$)",
  "$C_{\\mathrm{site}}$", "Site-quality constant (kg\\,ha$^{\\beta-1}$); conservative: 4\\,475, good site: 6\\,471",
  "$\\beta$",          "Competition-density exponent; fitted $\\hat{\\beta}=0.380$",
  # --- AGB/BGB split ---
  "$\\phi_{\\mathrm{ag}}$", "Aboveground fraction of total stand biomass; $\\phi_{\\mathrm{ag}}=0.89$",
  "$M^{ag}(t)$",       "Aboveground dry biomass per ha (t\\,DM\\,ha$^{-1}$) at age $t$",
  # --- compartment fractions ---
  "$f_p(t)$",          "asymtotic fraction of AGB attributed to compartment $p \\in \\{s,b,r\\}$ at age $t$",
  "$\\alpha_p,\\,\\delta_p$", "Intercept and slope of the linear fraction model for compartment $p$",
  # --- logistic merchantability ---
  "$q_p(t)$",          "Merchantable fraction of compartment $p$ biomass at age $t$ (logistic)",
  "$r_p$",             "Logistic growth rate for merchantability of compartment $p$",
  "$t_{50,p}$",        "Age at which 50\\% of compartment $p$ biomass is merchantable (yr)",
  # --- harvest yields ---
  "$Y_p(t)$",          "Merchantable dry biomass of compartment $p$ per ha at harvest age $t$ (t\\,DM\\,ha$^{-1}$)",
  "$\\omega$",         "Moisture content at felling (mass fraction); $\\omega = 0.55$"
)

kable(notation,
      format    = "latex", booktabs = TRUE, escape = FALSE,
      col.names = c("Symbol", "Description"),
      caption   = "Notation for the biomass growth and harvest yield models.",
      label     = "tab-notation-growth") |>
  kable_styling(font_size = 8,
                latex_options = c("hold_position", "scale_down"),
                full_width = FALSE) |>
  column_spec(1, width = "3.2cm") |>
  column_spec(2, width = "10.8cm")
```

Reliable estimation of stand-level biomass quantities and their age-dependent composition is a necessary prerequisite for any strategic
analysis of AFS-based supply chains. Individual-tree biomass is commonly approximated by species-specific allometric power laws relating shoot
dry matter to stem base diameter, and these relationships have been calibrated for Central European poplar clones under SRC and SRAFS conditions [@Huber2016_Allometric_SRAFS; @Huber2018_SRAFS_growth]. For supply chain planning, however, the relevant quantity is stand-level
biomass per hectare over a multi-decade rotation horizon, which requires a temporal growth model that captures the characteristic sigmoidal accumulation trajectory of woody biomass. Therefore, a set of variables is used summarized in \autoref{tab-notation-growth}.

### Stand-level growth model

Stand-level aboveground and belowground dry biomass per hectare,
$M(t)$, accumulates in a right-skewed sigmoidal pattern with a
relatively early inflection point, driven by rapid juvenile shoot growth
followed by self-thinning and competitive deceleration as canopy closure
intensifies [@Niemczyk2021_Poplar_Rotation;
@Testa2014_Poplar_Economics]. This trajectory is described by a Gompertz
model: \begin{equation}
  M(t) \;=\; A(N) \cdot \exp\!\bigl(-\exp(-k \cdot (t - t_0))\bigr),
  \label{eq:gomp-stand}
\end{equation} where $t_0$ is the age of maximum instantaneous biomass
increment and $k$ determines how quickly the stand approaches its
site-specific asymptote $A(N)$. For Central European poplar under
medium-rotation AFS management (10--20 yr), analysing
empirical data from @Huber2018_SRAFS_growth,
@Niemczyk2021_Poplar_Rotation, and @Jha2018_Poplar_AFS_Biomass yields parameter intervals of 
$k \in [0.17,0.2]$ and $t_0 \in [9,10]$ years, depending on local specifications of the AFS under study.
Parameter $t_0$ indicates an upper bound on the optimal rotation age
as growth slows down or diminishes entirely close to this age. 

The asymptote $A(N)$ depends largerly on planting density and can be modelled by a competition-density power law
[@Shinozaki1961_CD_Effect]: 
\begin{equation}
  A(N) \;=\; C_{\mathrm{site}} \cdot N^{1-\beta},
  \label{eq:A-density}
\end{equation} 
where $C_{\mathrm{site}}$ encodes site quality, $N$ indiactes tree density (e.g. trees per hectar), and
$\beta$ is the competition exponent. Equivalently, the per-tree
asymptote $A_{\mathrm{tree}}(N) = A(N)/N = C_{\mathrm{site}}\cdot N^{-\beta}$
decreases with density as individual trees have access to less growing
space. Parameters $C_{\mathrm{site}}$ and $\beta$ are highly site- and specicies-specific. For Poplar clones under European conditions,  $\beta \in [0.3,0.5]$ and $C_{\mathrm{site}} \in [4,6.5]$ [@Niemczyk2021_Poplar_Rotation,@Shinozaki1961_CD_Effect]. \autoref{fig:growth-asymptote-functions} shows asymptote functions according to \eqref{eq:A-density} for advantageous and disadvantageous site conditions.

```{r growth-asymptote-functions, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Tree-level and stand-level Gompertz asymptote function $A(N)$ of planting density for the advanteguous ($N=4$, $\\beta=-0.3$) and disadvantageous ($N=6.5$, $\\beta=-0.5$) scenarios.",  fig.width=6.5, fig.height=4, out.width="\\textwidth", cache=FALSE}

# ── Panel (a): A_stand vs N ─────────────────────────────────────────────────
N_seq <- seq(20, 400, by = 1)
df_a  <- bind_rows(
  tibble(N = N_seq,
         A = A_stand(N_seq, C_site = 4.000, beta=-.3),
         scenario = "advanteguous"),
  tibble(N = N_seq,
         A = A_stand(N_seq, C_site = 6.500, beta = -.5),
         scenario = "disadvantageous")
)

df_tree  <- bind_rows(
  tibble(N = N_seq,
         A = A_tree(N_seq, C_site = 4.000, beta=-.3),
         scenario = "advanteguous"),
  tibble(N = N_seq,
         A = A_tree(N_seq, C_site = 6.500, beta = -.5),
         scenario = "disadvantageous")
)


p_a <- ggplot(df_a, aes(x = N, y = A, linetype = scenario)) +
  geom_line(colour = "grey20", linewidth = 0.7) +
  scale_linetype_manual(
    name   = "Scenario",
    values = c("advanteguous" = "solid", "disadvantageous" = "dashed")) +
  labs(x = expression("Density "*italic(N)*" (trees ha"^{-1}*")"),
       y = expression(italic(A(N))*" (t DM ha"^{-1}*")")) +
  theme_minimal(base_size = 8) +
  theme(legend.position  = c(0.72, 0.82),
        legend.key.size  = unit(0.4, "cm"),
        legend.text      = element_text(size = 7),
        legend.title     = element_text(size = 7, face = "bold"),
        legend.background = element_rect(fill = "white", colour = NA),
        panel.grid.minor = element_blank())

p_b <- ggplot(df_tree, aes(x = N, y = A, linetype = scenario)) +
  geom_line(colour = "grey20", linewidth = 0.7) +
  scale_linetype_manual(
    name   = "Scenario",
    values = c("advanteguous" = "solid", "disadvantageous" = "dashed")) +
  labs(x = expression("Density "*italic(N)*" (trees ha"^{-1}*")"),
       y = expression(italic(A(N))*" (t DM per tree)")) +
  theme_minimal(base_size = 8) +
  theme(legend.position  = c(0.72, 0.82),
        legend.key.size  = unit(0.4, "cm"),
        legend.text      = element_text(size = 7),
        legend.title     = element_text(size = 7, face = "bold"),
        legend.background = element_rect(fill = "white", colour = NA),
        panel.grid.minor = element_blank())
# ── Assemble with patchwork ─────────────────────────────────────────────────
p_b / p_a
```

When assessing potential downstream application of AFS biomass, biomass fraction are to be defined meeting the requirements of downstream consumers. E.g. for timber and biochemical applications, stem diameters of AFS trees need to exceed a minimum threshold in order to qualify as a suitable feedstock. In the following three biomass fractions are distinguished: stems, branch wood, and residue. The stem fraction consists of stems with a diameter of at least 15 cm, while branch wood requires a diameter of at least 7 cm. The residue fraction comprises are smaler parts of the tree. To subdivide the total AFS biomass modelled by \eqref{eq:gomp-stand} into the three fractions depending on tree age, a logistic growth model is used: 
\begin{equation}
  q_p(t) \;=\; \frac{f_p}{1 + \exp\!\bigl(-r_p \cdot (t - t_{50,p})\bigr)},
  \label{eq:logistic-merch}
\end{equation} 
where $t_{50,p}$ is the age at which 50% of compartment $p$ biomass surpasses the diameter threshold and $f_p$ denotes the asymptotic share of component $p$ for mature trees. Using empirical data from data from @Civitarese2019_Poplar_Harvest, @Jha2018_Poplar_AFS_Biomass and @Niemczyk2021_Poplar_Rotation, $r_{stem} = 0.467$,yr$^{-1}$, $t_{50,stem} = 8.19$ years, and $f_{stem} = 0.75$ is estimated for the stem fraction (threshold $d > 15$ cm) and $r_{bran} = 0.698$ yr$^{-1}$, $t_{50,bran} = 7.55$,yr, and $f_{bran} = 0.15$ for the branch fraction (diameter at least 7 cm at insertion). \autoref{fig:afs-model-functions} illustrates the development of the fraction shares over stand age. 

```{r afs-model-functions, echo=FALSE, message=FALSE, warning=FALSE,     fig.cap="Logistic diameter-class merchantability functions $q_p(t)$ for stem  ($d > 15$\\,cm, green) and branches ($d > 7$\\,cm, orange); the 50\\% threshold ages $t_{50}$ are indicated by dashed vertical lines.", fig.width=6.5, fig.height=3, out.width="\\textwidth", cache=FALSE}

# ── Panel (b): logistic merchantability q_p(t) ─────────────────────────────
ages_c <- seq(1, 20, by = 0.1)
df_c <- tibble(
  age                    = ages_c,
  `Stem (d > 15 cm)`    = f_stem_merch(ages_c),
  `Branch (d > 7 cm)`   = f_branch_merch(ages_c)
) |>
  pivot_longer(-age, names_to = "compartment", values_to = "q") |>
  mutate(compartment = factor(compartment,
           levels = c("Stem (d > 15 cm)", "Branch (d > 7 cm)")))

t50_vals <- tibble(
  compartment = factor(c("Stem (d > 15 cm)", "Branch (d > 7 cm)"),
                       levels = c("Stem (d > 15 cm)", "Branch (d > 7 cm)")),
  t50 = c(8.19, 7.55)
)

merch_cols <- c("Stem (d > 15 cm)"  = "#1b9e77",
                "Branch (d > 7 cm)" = "#d95f02")

p_c <- ggplot(df_c, aes(x = age, y = q, colour = compartment)) +
  geom_vline(data = t50_vals,
             aes(xintercept = t50, colour = compartment),
             linetype = "dashed", linewidth = 0.45, show.legend = FALSE) +
  geom_hline(yintercept = 0.5,
             linetype = "dotted", colour = "grey60", linewidth = 0.4) +
  geom_line(linewidth = 0.8) +
  annotate("text", x = 8.19 + 0.3, y = 0.08,
           label = expression(italic(t)[50*","*s]*" = 8.2 yr"),
           size = 2.4, hjust = 0, colour = "#1b9e77") +
  annotate("text", x = 7.55 - 0.3, y = 0.20,
           label = expression(italic(t)[50*","*b]*" = 7.6 yr"),
           size = 2.4, hjust = 1, colour = "#d95f02") +
  scale_colour_manual(name = "Compartment", values = merch_cols) +
  scale_x_continuous(breaks = seq(0, 20, 2)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, max(df_c$q)*1.1)) +
  labs(x = "Stand age (yr)",
       y = expression(italic(q[p])*"(t)  [merchantable fraction]")) +
  theme_minimal(base_size = 8) +
  theme(legend.position  = c(0.15, 0.75),
        legend.key.size  = unit(0.4, "cm"),
        legend.text      = element_text(size = 7),
        legend.title     = element_text(size = 7, face = "bold"),
        legend.background = element_rect(fill = "white", colour = NA),
        panel.grid.minor = element_blank())

# ── Assemble with patchwork ─────────────────────────────────────────────────
p_c
```

Combining equations \eqref{eq:gomp-stand}--\eqref{eq:logistic-merch}, the dry biomass yield of fraction $p$ per hectare at harvest age $t$ is: \begin{equation}
  \eta_p(t) \;=\; q_p(t)\cdot M(t),
  \label{eq:yield}
\end{equation} 
with all non-merchantable fractions aggregated into the residue fraction. Fresh-weight yields at felling are obtained from dry matter yields via $\eta_p^{\mathrm{fw}} = \eta_p/(1-\omega)$, where the moisture content at felling $\omega = 0.55$ is taken from @Civitarese2019_Poplar_Harvest. \autoref{fig:afs-growth-curves} shows the fractions' stacked biomass quantities $\eta_p(t)$  depending on stand age compared with empirical observations from @Jha2018_Poplar_AFS_Biomass

```{r afs-growth-curves, echo=FALSE, message=FALSE, warning=FALSE,     fig.cap="Stand-level aboveground biomass components for poplar under AFS management with $N=250$ trees\\,ha$^{-1}$. Empirical AGB observations compiled from \\citet{Jha2018_Poplar_AFS_Biomass}, \\citet{Civitarese2019_Poplar_Harvest}, and \\citet{Niemczyk2021_Poplar_Rotation}. Grey band: typical medium-rotation harvest window (5--15 years).",    fig.width=6.5, fig.height=4.0, out.width="\\textwidth", cache=FALSE}

age.vec <- seq(0, 20, by = 0.1)

cons <- build_scenario_ts(ages = age.vec , N = 300, C_site = 6.500, k = 0.194, t0 = 9.7, label = "Conservative (250 trees/ha)")

comp_long <- bind_rows(cons) |>
  rename(stem = "Merch. stem", branch = "Merch. branch", residue = Residue) %>% 
  pivot_longer(c("stem", "branch", "residue"),
               names_to = "component", values_to = "biomass") |>
  mutate(
    component = factor(component,
      levels = c("residue", "branch", "stem")),
    scenario  = factor(scenario,
      levels = c("Conservative (113 trees/ha)", "Good site (139 trees/ha)"))
  ) %>% 
  select(-scenario)

# totals_long <- comp_long |>
#   group_by(age, scenario) |>
#   summarise(total = sum(biomass), .groups = "drop")

emp_pts <- tribble(
  ~age, ~AGB,   ~p_stem, ~p_branch,
   8,   134.3,  0.747,   0.166, # N = 400
   9,    74.0,  0.830,   0.110, # N = 500
   9,    51.5,  0.920,   0.040, # N = 500
  10,   119,    0.76,   0.210,# N = 139
  13,   152.0,  0.770,   0.220 # N = 139
) |>
  mutate(stem_cum   = AGB * p_stem,
         branch_cum = AGB * (p_stem + p_branch),
         total      = AGB)

emp_long <- bind_rows(
  emp_pts |> transmute(age, biomass = stem_cum,   series = "Empirical: stem"),
  emp_pts |> transmute(age, biomass = branch_cum, series = "Empirical: stem+branches"),
  emp_pts |> transmute(age, biomass = total,      series = "Empirical: total biomass")
) |> mutate(series = factor(series, levels = c(
  "Empirical: stem", "Empirical: stem+branches", "Empirical: total biomass")))

emp_col   <- c("Empirical: stem"          = "darkgreen",
               "Empirical: stem+branches" = "darkorange",
               "Empirical: total biomass"     = "navy")
emp_shape <- c("Empirical: stem"          = 21L,
               "Empirical: stem+branches" = 24L,
               "Empirical: total biomass"     = 23L)

ggplot(comp_long, aes(x = age, y = biomass)) +
  annotate("rect", xmin = 5, xmax = 15, ymin = 0, ymax = Inf,
           fill = "grey80", alpha = 0.25) +
  geom_area(aes(fill = component),  #, linetype = scenario
            position = "stack", alpha = 0.70, colour = "white", linewidth = 0.2) +
  # geom_line(data    = totals_long,
  #           aes(x = age, y = total, linetype = scenario),
  #           colour  = "grey30", linewidth = 0.45, inherit.aes = FALSE) +
  geom_point(data = emp_long,
             aes(x = age, y = biomass,
                 colour = series, shape = series, fill = series),
             size = 2.5, alpha = 0.90, inherit.aes = FALSE) +
  scale_fill_manual(
    name   = "Biomass fraction",
    values = c("stem"   = "#1b9e77",
               "branch" = "#d95f02",
               "residue"       = "#7570b3"),
    breaks = c("stem", "branch", "residue")) +
  scale_colour_manual(name = "Empirical data", values = emp_col) +
  scale_shape_manual (name = "Empirical data", values = emp_shape) +
  # scale_linetype_manual(
  #   name   = "Scenario",
  #   values = c("Conservative (113 trees/ha)" = "solid",
  #              "Good site (139 trees/ha)"    = "dashed")) +
  guides(
    fill     = guide_legend(order = 1,
                 override.aes = list(shape = NA, colour = NA, linetype = 0)),
    colour   = guide_legend(order = 3,
                 override.aes = list(
                   shape  = emp_shape[levels(emp_long$series)],
                   size   = 2.5,
                   fill   = emp_col[levels(emp_long$series)])),
    linetype = guide_legend(order = 2,
                 override.aes = list(fill = NA, colour = "grey30")),
    shape    = "none") +
  scale_x_continuous(breaks = seq(0, 20, 2)) +
  labs(x = "Stand age (years)",
       y = expression("biomass (t DM "*ha^{-1}*")")) +
  theme_minimal(base_size = 9) +
  theme(legend.position  = "right",
        legend.key.size  = unit(0.4, "cm"),
        legend.text      = element_text(size = 7),
        legend.title     = element_text(size = 8, face = "bold"),
        panel.grid.minor = element_blank())
```

## Biomass supply chain design {#sec:lit-sc}

The design and planning of biomass supply chains has attracted substantial research attention over the past two decades, driven by the
need to ensure cost-efficient and sustainable feedstock provision for bioenergy and biorefinery applications. Several systematic reviews
provide comprehensive overviews of this literature. @Cambero2014_ForestBSC_Review survey economic, social, and environmental
assessment and optimisation studies for forest biomass supply chains, identifying MILP as the dominant modelling approach and noting a growing
trend towards multi-objective formulations integrating life cycle assessment.

@Yue2014_BSC_Optimization_Overview review biomass-to-bioenergy and
biofuel supply chain models with a focus on multi-scale modelling,
uncertainty treatment, and sustainability metrics, and conclude that the
coupling of upstream feedstock dynamics with downstream network design
remains an underexplored area. @Malladi2018_BSC_Logistics survey
logistics-specific features of biomass supply chain models—seasonal
availability, moisture content degradation, preprocessing steps—and
highlight that most models treat biomass as a homogeneous commodity,
neglecting quality differentiation. @Zahraee2020_BSC_review provide a
recent meta-review of 140+ studies, confirming that cost minimisation is
still the dominant objective, and that multi-product cascading
formulations are rare.

The structure of biomass supply chain models typically encompasses three
decision layers: (i) strategic network design (facility location,
capacity sizing, technology selection), (ii) tactical planning (harvest
scheduling, seasonal flows, inventory), and (iii) operational routing.
The following paragraphs discuss key individual contributions and their
specific methodological advances.

@Eksioglu2009_Biorefinery_SC develop one of the earliest large-scale
MILP models for biomass-to-biorefinery supply chain design, integrating
facility location, capacity sizing, and multimodal transport for
lignocellulosic feedstocks. Their model minimises total system cost and
introduces binary facility-opening variables that have since become
standard in the field. A key limitation, however, is that biomass supply
is treated as an exogenous input, so feedstock availability is not
endogenised as a function of land-use or growth decisions.
@Mansoornejad2013_Forest_Biorefinery_SC extend the strategic design
perspective to forest biorefineries by integrating product portfolio
design with supply chain configuration in a single MILP. Their model
explicitly accounts for product conversion efficiencies along different
processing pathways, enabling joint optimisation of process technology
selection and logistics.

@DeMeyer2015_OPTIMASS introduce the OPTIMASS framework, a generic MILP
for biomass-based supply chain optimisation that explicitly models
product quality cascades: higher-quality biomass classes can substitute
for lower-quality demands downstream, whereas the reverse is infeasible.
This cascade hierarchy, formalised through ordered product sets, is
particularly relevant for woody biomass where stem, branch, and residue
fractions command different market prices. The companion t-OPTIMASS
extension [@DeMeyer2016_tOPTIMASS] incorporates temporal dynamics by
explicitly representing biomass growth and regeneration over multiple
periods, enabling simultaneous optimisation of harvest timing and
network flows. However, in both OPTIMASS formulations the growth
dynamics are represented as exogenous time-series rather than as
age-structured stand models, so rotation-length decisions and
age-dependent biomass quality are not endogenised.
@Arabi2019_Microalgae_Biobutanol_SC demonstrate multi-period MILP
extensions for bio-based supply chains with production dynamics, but
focus on microalgae rather than woody biomass and do not address product
quality cascades.

@Sharma2013_BSC_review identify supply-side uncertainty (yield
variability, seasonal availability, moisture content) as the most
critical source of risk in biomass supply chains. In response, a strand
of literature has developed stochastic and robust optimisation
extensions. @Zahraee2020_BSC_review document that two-stage stochastic
MILP and robust optimisation approaches now account for roughly 30% of
the modelling literature, with biomass yield uncertainty and price
volatility as the dominant stochastic parameters.

Biomass supply chain research focusing specifically on agroforestry or
SRC contexts is comparatively sparse.
@Espinoza2017_Biorefinery_SC_review note that most supply chain models
assume continuous feedstock availability and do not represent the
periodic harvest cycles characteristic of coppice systems.
@Lamers2015_Feedstock_Design analyse feedstock design principles for
lignocellulosic supply chains and argue that the allocation of different
biomass fractions to valorisation pathways—product cascading—is a key
lever for economic viability, yet is rarely optimised jointly with
network design decisions. The model developed in
Section \@ref(sec:milp) of this paper directly addresses this gap by
endogenising stand age dynamics through binary age-lag variables and
coupling component-specific yield coefficients with a product quality
cascade.

\autoref{tab:lit-sc-features} summarises the key modelling features
of the reviewed contributions.

```{r lit-sc-features, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
options(kableExtra.latex.load_packages = FALSE)
options(knitr.table.format = "latex")
knitr::opts_chunk$set(dev = "pdf")

lit_tab <- tribble(
  ~Study,                            ~Objective,          ~`Multi-product`,  ~`Quality cascade`, ~`Age dynamics`,    ~`Uncertainty`,    ~`Level`,
  "Eksioglu et al. (2009)",          "Cost min.",         "No",              "No",               "No",               "No",              "Strategic",
  "Sharma \\& Taaffe (2013)",        "Cost min.",         "Partial",         "No",               "No",               "No",              "Strategic/Tactical",
  "Mansoornejad et al. (2013)",      "Profit max.",       "Yes",             "No",               "No",               "No",              "Strategic",
  "Cambero \\& Sowlati (2014)",      "Cost/GHG/jobs",     "No",              "No",               "No",               "Stochastic",      "Strategic",
  "Yue et al. (2014)",               "Cost/GHG",          "Yes",             "No",               "No",               "No",              "Multi-scale",
  "De Meyer et al. (2015)",          "Cost min.",         "Yes",             "Yes",              "Exogenous",        "No",              "Strategic/Tactical",
  "De Meyer et al. (2016)",          "Cost min.",         "Yes",             "Yes",              "Exogenous TS",     "No",              "Multi-period",
  "Lamers et al. (2015)",            "Cost/quality",      "Yes",             "Partial",          "No",               "Scenario",        "Strategic",
  "Espinoza et al. (2017)",          "Sustainability",    "Yes",             "No",               "No",               "Scenario",        "Strategic",
  "Malladi \\& Sowlati (2018)",      "Cost min.",         "Partial",         "No",               "No",               "No",              "Tactical/Oper.",
  "Arabi et al. (2019)",             "Profit max.",       "Yes",             "No",               "Exogenous",        "No",              "Multi-period",
  "Zahraee et al. (2020)",           "Multi-obj.",        "Yes",             "No",               "No",               "Stochastic",      "Strategic/Tactical",
  "\\textbf{This paper}",            "NPV max.",          "Yes",             "Yes",              "\\textbf{Endogenous}", "Scenario",   "Strategic/Multi-period"
) |>
  mutate(across(everything(), ~ gsub("Yes", "\\\\checkmark", .))) |>
  mutate(across(everything(), ~ gsub("No",  "---",            .)))

kable(lit_tab,
      format    = "latex",
      booktabs  = TRUE,
      escape    = FALSE,
      caption   = "Key modelling features of selected biomass supply chain design studies.",
      label     = "lit-sc-features") |>
  kable_styling(
  font_size     = 7,
  latex_options = c("hold_position", "scale_down"),
  full_width    = FALSE
) |>
  row_spec(nrow(lit_tab), bold = TRUE) |>
  column_spec(1, width = "3.2cm") |>
  column_spec(2, width = "1.6cm") |>
  column_spec(3:7, width = "1.4cm")
```

# MILP Model for Agroforestry Supply Chain Design {#sec:milp}

Building on the biomass growth models introduced in Section
\@ref(sec:lit-growth) and the supply chain design literature reviewed in
Section \@ref(sec:lit-sc), this section describes the AFS supply chain
design problem (\problem), a mixed-integer linear programming (MILP)
model for the integrated design and operation of an agroforestry biomass
supply chain with multiple AFS biomass types (i.e., stem, branch,
and residue fractions) used as feedstocks for potential consumers in chemical, pulp and
paper as well as energy industry.

## Sets, Indices, and Time Structure

The model considers a set of agroforestry sites $\mathcal{I}$, storage/pre-treatment facilities $\mathcal{J}$, consumer sites $\mathcal{K}$, and biomass types $\mathcal{P} = \{stem, branches, residues\}$, whose growth models are outlined above. Time is discretised into annual periods $t \in \Tc = \{1,\dots,T_{\max}\}$. To reduce complexity, reflect legal definitions of SRAFS and LRAFS as well as discard implausible options (e.g. extremly long rotation cycles), $A^{\min}$ and $A^{\max}$ define minimum and maximum stand ages. Thus, harvest are possible in periods $\mathcal{T}^{\text{harv}} = \{A^{\min}+1,\dots,T_{\max}\}$. To consistently model establishment and consecutive harvest decisions, the following arc sets are defined: $\Sc^{est}$ defines potential establishments of AFS. If an AFS is established in period $t \in \Tc$, arc $(0,t)$ is selected. Harvest set $\Sc^{harv}$ summarizes all potential consecutive harvest periods within the age interval $[A^{\min},A^{\max}]$. I.e., if an AFS is established or harvested in period $s$ and next harvest is in period $t$, arc $(s,t)$ is selected. Finally, if there is no subsequent harvest after period $t$, the AFS is terminated by selecting arc $(t,T^{\max}+1)$. Termination arcs are summarized in set $\Sc^{term}$. All arc sets are formally defined as
follows: \begin{align}
  \Sc^{{est}}  &= \{(0,t) \mid t \in \{1,\dots,T^{\max}-A^{\min}\}\}, \notag\\
  \Sc^{{harv}} &= \{(s,t) \mid s,t \in \Tc \wedge t-s \in \{A^{\min},\dots,A^{\max}\}\},  \notag\\
  \Sc^{{term}} &= \{(s,T^{\max}+1) \mid s \in \{A^{\min}+1,\dots,T_{\max}\}\}, \notag\\
  \Sc               &= \Sc^{{est}} \cup \Sc^{{harv}} \cup \Sc^{{term}}. \notag
\end{align}

\autoref{fig:fig-set-S-tikz} illustrates this network structure of all arc sets for $T=8$, $A^{\min}=3$, $A^{\max}=5$.

```{tikz fig-set-S-tikz, echo=FALSE, fig.cap="Network structure of set $\\mathcal{S}$ for $T=8$, $A_{\\min}=3$, $A_{\\max}=5$. Dashed blue arcs: establishment arcs $(0,t)$, $t\\in\\{1,\\ldots,T-1\\}$. Solid green arcs above: harvest arcs $(s,t)$ with odd $s$; below: harvest arcs with even $s$. Solid red arcs:    termination arcs $(s,T+1)$, $s\\in\\{A_{\\min}+1,\\ldots,T\\}$.",    fig.align="center", out.width="\\textwidth", engine.opts=list(extra.preamble=c(      "\\usetikzlibrary{arrows.meta,bending,positioning}"   )), cache=TRUE}


\begin{tikzpicture}[
  x=1.55cm, y=1cm,
  node/.style={circle, draw, fill=#1, text=white,
               font=\small\bfseries, inner sep=2pt, minimum size=18pt},
  est/.style  ={-{Stealth[length=4pt]}, blue!70!black,  dashed, semithick},
  harva/.style={-{Stealth[length=4pt]}, green!40!black, semithick},
  harvb/.style={-{Stealth[length=4pt]}, green!40!black, semithick},
  term/.style ={-{Stealth[length=4pt]}, red!70!black,   semithick},
]

\node[node=blue!70!black]  (n0) at (0,0) {$0$};
\foreach \i in {1,...,8} {
  \node[node=gray!50!black] (n\i) at (\i,0) {$\i$};
}
\node[node=red!70!black]   (n9) at (9,0) {$9$};

\draw[est] (n0) -- (n1);
\foreach \t/\l in {2/1.3, 3/1.0, 4/0.80, 5/0.65}{
  \draw[est] (n0) to[out=55, in=125, looseness=\l] (n\t);
}

\draw[harva] (n1) to[out=55,  in=125,  looseness=0.55] (n4);
\draw[harvb] (n2) to[out=-55, in=-125, looseness=0.55] (n5);
\draw[harva] (n3) to[out=55,  in=125,  looseness=0.55] (n6);
\draw[harvb] (n4) to[out=-55, in=-125, looseness=0.55] (n7);
\draw[harva] (n5) to[out=55,  in=125,  looseness=0.55] (n8);
\draw[harva] (n1) to[out=60,  in=120,  looseness=0.42] (n5);
\draw[harvb] (n2) to[out=-60, in=-120, looseness=0.42] (n6);
\draw[harva] (n3) to[out=60,  in=120,  looseness=0.42] (n7);
\draw[harvb] (n4) to[out=-60, in=-120, looseness=0.42] (n8);
\draw[harva] (n1) to[out=65,  in=115,  looseness=0.35] (n6);
\draw[harvb] (n2) to[out=-65, in=-115, looseness=0.35] (n7);
\draw[harva] (n3) to[out=65,  in=115,  looseness=0.35] (n8);

\foreach \s/\l in {4/0.80, 5/0.65, 6/0.55, 7/0.48}{
  \draw[term] (n\s) to[out=55, in=125, looseness=\l] (n9);
}
\draw[term] (n8) -- (n9);

\node[anchor=north west] at (0.0,2.5) {%
  \begin{tikzpicture}[x=0.6cm,y=0.35cm,baseline=0pt]
    \draw[est]   (0,0)--(1,0);
    \node[font=\small, anchor=west] at (1.1,0) {Establishment arc $(0,t)$};
  \end{tikzpicture}};
\node[anchor=north west] at (3.5,2.5) {%
  \begin{tikzpicture}[x=0.6cm,y=0.35cm,baseline=0pt]
    \draw[harva] (0,0)--(1,0);
    \node[font=\small, anchor=west] at (1.1,0) {Harvest arc $(s,t)$};
  \end{tikzpicture}};
\node[anchor=north west] at (6.3,2.5) {%
  \begin{tikzpicture}[x=0.6cm,y=0.35cm,baseline=0pt]
    \draw[term]  (0,0)--(1,0);
    \node[font=\small, anchor=west] at (1.1,0) {Termination arc $(s,9)$};
  \end{tikzpicture}};

\end{tikzpicture}
```

## Decision Variables

Temporal flow Variables $z_{ist}$ indicate on which quantity of the area of site $i$ an an AFS is operated between $(s,t) \in \Sc$. Depending on the arc type, this implies establishment, harvest, or termination. Selected arcs buiild a network flow over the complete planning horizon from $t=0$ to $T^{max}+1$. \autoref{fig:fig-set-S-path} illustrates a simple network flow with establishment of an AFS in period 1 and harvests in periods 5 and 8. 

```{tikz fig-set-S-path, echo=FALSE, fig.cap="Example of a feasible path through the network for $T=8$,$A_{\\min}=3$, $A_{\\max}=5$: establishment in period~1, harvests in periods~5 and~8, termination at $T+1$.",    fig.align="center", out.width="\\textwidth", engine.opts=list(extra.preamble=c(     "\\usetikzlibrary{arrows.meta,bending,positioning}"    )), cache=TRUE}

\begin{tikzpicture}[
  x=1.55cm, y=1cm,
  node/.style    ={circle, draw, fill=#1, text=white,
                   font=\small\bfseries, inner sep=2pt, minimum size=18pt},
  inactive/.style={node=gray!25, text=gray!50},
  pathnode/.style={node=green!40!black},
  n0style/.style ={node=blue!70!black},
  n9style/.style ={node=red!70!black},
  patharc/.style ={-{Stealth[length=5pt]}, green!40!black, very thick},
  est/.style     ={patharc, blue!70!black, dashed},
  term/.style    ={patharc, red!70!black},
]

\node[n0style]   (n0) at (0,0) {$0$};
\foreach \i in {2,3,4,6,7} {
  \node[inactive] (n\i) at (\i,0) {$\i$};
}
\foreach \i in {1,5,8} {
  \node[pathnode]  (n\i) at (\i,0) {$\i$};
}
\node[n9style]   (n9) at (9,0) {$9$};

\draw[est]     (n0) -- (n1);
\draw[patharc] (n1) to[out=60, in=120, looseness=0.42] (n5);
\draw[term] (n5) to[out=-50, in=-130, looseness=0.42] (n9);
\draw[patharc] (n5) to[out=55, in=125, looseness=0.55] (n8);
\draw[term]    (n8) -- (n9);

\node[font=\footnotesize, text=blue!70!black]   at (0.5, 0.28) {est.};
\node[font=\footnotesize, text=blue!70!black]   at (0.5, -0.3) {$100$};
\node[font=\footnotesize, text=green!40!black]  at (3.0, 1.1)  {$100$};
\node[font=\footnotesize, text=green!40!black]  at (3.0, 0.6)  {$t-s=4$};
\node[font=\footnotesize, text=green!40!black]  at (6.5, 1) {$70$};
\node[font=\footnotesize, text=red!70!black]  at (7, -1.0) {$30$};
\node[font=\footnotesize, text=red!70!black]  at (8.5, -0.3) {$70$};
\node[font=\footnotesize, text=green!40!black]  at (6.5, 0.55) {$t-s=3$};
\node[font=\footnotesize, text=red!70!black]    at (8.5, 0.28) {term.};

\end{tikzpicture}
```

Spatial flow variables $X_{ijpt}$ and $X_{jkpp't}$ represent transport of products from AFS sites to storage/pre-treatment facilities and from there to consumers, respectively, while $S_{jpt}$ captures end-of-period inventory at facility $j$. \autoref{tab:tab-notation-milp} summarises all sets, decision variables, and parameters used in the MILP formulation, while $\eta_{pa}$ refers to \eqref{eq:yield}.

```{r tab-notation-milp, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
options(kableExtra.latex.load_packages = FALSE)
options(knitr.table.format = "latex")

math_break <- function(line1, line2) {
  paste0("\\shortstack[l]{$", line1, "$\\\\$", line2, "$}")
}

notation <- tribble(
  ~Symbol, ~Description,
  # --- SETS & INDICES ---
  "$i \\in \\mathcal{I}$",   "Agroforestry sites",
  "$j \\in \\mathcal{J}$",   "Storage / pre-treatment facilities",
  "$k \\in \\mathcal{K}$",   "Consumer sites",
  "$p \\in \\mathcal{P}$",   "Product types (stem, branches, residues)",
  "$(p,p') \\in \\mathcal{Q}$",   "Admissible product-substitution pairs (quality cascade)",
  "$\\mathcal{T}=\\{1,\\dots,T\\}$",
    "Planning periods (years)",
  "$\\mathcal{T}^{\\text{harv}}=\\{A^{\\min}+1,\\dots,T\\}$",
    "Feasible harvest periods",
  "$\\mathcal{S}^{\\text{est}}=\\{(0,t)\\mid t\\in\\{1,\\dots,T-A^{\\min}\\}\\}$",
    "Establishment arcs: dummy source to period $t$",
  math_break(
    "\\mathcal{S}^{\\text{harv}}=\\{(s,t)\\mid s,t\\in \\mathcal{T},",
    "t-s\\in\\{A^{\\min},\\dots,A^{\\max}\\}\\}"),
    "Harvest cycles within age-lag bounds",
  math_break(
    "\\mathcal{S}^{\\text{term}}=\\{(s,T+1)\\mid",
    "s\\in\\{A^{\\min}+1,\\dots,T\\}\\}"),
    "Termination arcs to dummy sink $T+1$",
  "$\\mathcal{S} = \\mathcal{S}^{\\text{est}} \\cup \\mathcal{S}^{\\text{harv}} \\cup\\; \\mathcal{S}^{\\text{term}}$",
    "Age-lag arc set (establishment $\\cup$ harvest $\\cup$ termination)",
  # --- DECISION VARIABLES ---
  "$z_{ist}\\in\\{0,AREA_i\\}$",
    "Area of site $i$ with AFS operation between $s$ and $t$",
  "$X_{ijpt}\\ge 0$",
    "Transport flow (t) of $p$ from site $i$ to facility $j$ in $t$",
  "$X_{jkpp't}\\ge 0$",
    "Flow of $p$ from facility $j$ to consumer $k$ satisfying demand $p'$ in $t$",
  "$S_{jpt}\\ge 0$",
    "Inventory (t) of product $p$ at facility $j$ end of period $t$",
  # --- PARAMETERS ---
  "$T$",              "Length of planning horizon (years)",
  "$A^{\\min}$, $A^{\\max}$", "Minimum / maximum stand age at harvest (years)",
  "$\\text{AREA}_i$",         "Area of site $i$ (ha)",
  "$\\eta_{pa}$",
    "Yield coefficient (t/ha) of fraction $p$ at stand age $a$, derived from \\eqref{eq:yield}",
  "$R_{kp}$",                 "Revenue (\\euro/t) for product $p$ at consumer $k$",
  "$C^{\\text{est}}_i$",      "Establishment cost (\\euro/ha) at site $i$",
  "$C^{\\text{opp}}_i$",      "Annual opportunity cost (\\euro/ha) at site $i$",
  "$C^{\\text{harv}}_i$",     "Harvest \\& refitting cost (\\euro/ha) at site $i$",
  "$c^{\\text{tr-raw}}$",     "Transport cost rate (\\euro/t\\,km) for raw biomass",
  "$c^{\\text{tr-pre}}$",     "Transport cost rate (\\euro/t\\,km) for pre-treated biomass",
  "$c^{\\text{stor}}_j$",     "Storage holding cost (\\euro/t) at facility $j$",
  "$d_{ij}$, $d_{jk}$",       "Distance (km) between nodes",
  "$\\text{CAP}^{\\text{stor}}_j/\\text{CAP}^{\\text{proc}}_j$", "Storage and processing capacity (t) at facility $j$",
  "$D^{\\max}_{kpt}$",        "Maximum demand (t) of fraction $p$ at consumer $k$ in period $t$"
)

tbl <- kable(notation,
      format    = "latex",
      booktabs  = TRUE,
      escape    = FALSE,
      col.names = NULL,
      caption   = "Notation for the MILP model.",
      label     = "tab-notation-milp") |>
  kable_styling(
    font_size     = 8,
    latex_options = c("hold_position", "scale_down", "striped"),
    full_width    = FALSE
  ) |>
  column_spec(1, width = "5cm") |>
   column_spec(2, width = "9cm") |>
   pack_rows("Sets \\& Indices",    1,  11,
             bold = TRUE, italic = FALSE, hline_before = TRUE, escape = FALSE) |>
   pack_rows("Decision Variables", 12,  15,
             bold = TRUE, italic = FALSE, hline_before = TRUE) |>
   pack_rows("Parameters",         16,  29,
             bold = TRUE, italic = FALSE, hline_before = TRUE)

tbl <- gsub(
  "p\\{(5\\.8cm|8\\.5cm)\\}",
  ">\\{\\\\raggedright\\\\arraybackslash\\}p\\{\\1\\}",
  tbl
)

tbl
```

## Objective Function

As discussed in Section \@ref(sec:lit-afs), the economic attractiveness of AFS supply chains is modelled from a global perspective taking into account all processes, activities, and participants involved in finally serving customer markets with AFS-based feedstocks. Therefore, revenues for the different biomass fractions and are weighed against costs occurring at all value-adding stages of the supply chains. Revenues created by selling AFS-based feedstocks to different industrial consumers are counterweighted by establishment, opportunity, harvest, transport, and storage costs of landowners, agricultural firms as well as biomass logistics companies. The model therefore maximises total profit of all participants in the AFS supply chain over the planning horizon as follows: 
\begin{align}
\max\; Z \;=\;&
  \sum_{k \in \Kc}\sum_{p \in \Pc} R_{kp} \cdot \sum_{j \in \Jc, }\sum_{(p',p) \in \Qc} \sum_{t \in \Tc}  X_{jkp'pt}  \label{eq:obj-revenue} \\
&- \sum_{i \in \Ic} c^{{est}}_i \cdot  \cdot\sum_{(0,t) \in \Sc^{{est}}} z_{i0t} \notag \\
% &- \sum_{i \in \Ic} c^{{opp}}_i \cdot \left(\sum_{(s,T+1) \in \Sc^{term}} (s+1) \cdot z_{is(T+1)}  -  \sum_{(0,t) \in \Sc^{{est}}}  t \cdot z_{i0t} \right) \notag \\ 
&- \sum_{i \in \Ic} (c^{main}_{i} + c^{{opp}}_i) \cdot (t-s) \cdot \sum_{(s,t) \in \Sc^{{harv}}} z_{ist} \notag \\
&- \sum_{i \in \Ic} c^{{harv}}_i  \cdot \sum_{(s,t) \in \Sc^{{harv}}} z_{ist} \notag \\
&- \sum_{i \in \Ic}\sum_{j \in \Jc}\sum_{p \in \Pc}\sum_{t \in \Tc} c^{\text{tr-raw}}_p \cdot d_{ij} \cdot X_{ijpt} \notag \\
&- \sum_{j \in \Jc}\sum_{k \in \Kc}\sum_{(p,p') \in \Qc} \sum_{t \in \Tc} c^{\text{tr-pre}}_{p'} \cdot d_{jk} \cdot X_{jkpp't} \notag \\
&- \sum_{j \in \Jc}\sum_{p \in \Pc}\sum_{t \in \Tc}  c^{{stor}}_j \cdot S_{jpt}  \notag
\end{align}

Here, $R_{kp}$ are product- and consumer-specific revenues for each ton of biomass fraction (\euro{}/t). The opportunity cost term penalises the land use when AFS are established. Thereby, $c^{{opp}}_i$ can be negative or positive depending on the site-specific conditions. It comprises effect of lost field crop revenues due to AFS as well as higher yields of remaining field crop areas e.g. due to wind and water erosion protection by AFS. While the assessment of lost field crop revenues is usually well predictable, the economic evaluation of positive ecological effects of AFS (e.g. on neighboring crop yields) is more difficult. Thus, $c^{{opp}}_i$ can also contain regulatory subsidies to support establishing and maintaining AFS. Transport cost terms mirror the two-stage flow structure of the AFS supply chain with differentiated rates for raw and pre-treated biomass depending on the biomass type. Cost rates differ because as e.g. the density and water content of the shipped biomasses differ or because specialized equippment (e.g. for log transports) is used [REFERENCES]. Moreover, site-specific establishment cost for planting the AFS trees is considered, next to yearly maintenance and harvesting costs. All these costs scale with the size of a site ($AREA_i$) measured in hectar. Maintenance cost comprise e.g. costs for trimming, replanting, and pruning unnecessary shoots. Storage costs are considered if biomasses are stored for more than a year. Storage cost rate $c^{{stor}}_j$ measures the costs for storing a ton of biomass for one year and comprises direct storage cost (e.g. for storage site maintenance) as well as opportunity cost due to potential material losses and capital commitment. Biomasses harvested and consumed in the same year incur no storage cost, as usually AFS are harvested at the beginning of a year. Harvested biomass has to dry during spring and summer such that the biomasses can be used in autumn or winter the same year. As under-year drying is necessary for most industrial applications and beneficial for transport economics, drying storage in the year of harvest is not associated with additional cost.

## Constraints

Constraints \eqref{eq:C1} limits the area dedicated to AFS operations to the maximum size of site $i$. Constraints \eqref{eq:C2} enforce flow conservation in each period. If AFS systems of size $z_{ist}$ are established or harvested in period $t \in \Tc$, there needs to be a succeeding activity (harvest or termination) of this area. This way, \eqref{eq:C2} guarantee valid (possibly empty) AFS activity  over the complete time horizon for each site. Constraints \eqref{eq:C3} links
the binary harvest-cycle variables to physical yield quantities via the
age-dependent yield coefficients $\eta_{p(t-s)}$. These coefficients are
pre-computed by evaluating the component-specific Gompertz functions
\eqref{eq:biomass-partial} and \eqref{eq:component-share} at stand age
$a = t - s$ and scaled by site area $\text{AREA}_i$. Thus, implicitly we
assume constant yields per hectar for each site $i$. Note that this
simplification can be dropped and site-specific yields $\eta_{ip(t-s)}$
could be introduced without increasing numerical complexity of the
\problem. Constraints \eqref{eq:C4}--\eqref{eq:C5} link harvest
quantities to transport flows and ensure inventory balance at each
facility, following the multi-period network-flow structure of OPTIMASS
[@DeMeyer2015_OPTIMASS; @DeMeyer2016_tOPTIMASS]. Thereby, harvest
quantities can be smaller than available AFS biomasses allowing partial
harvesting. Constraints \eqref{eq:C6}--\eqref{eq:C7} enforce storage and
processing capacity limits at each facility $j$. Constraints
\eqref{eq:C8} implement product usage at multiple consumption sites $k$.
Thereby, $(p',p) \in \Qc$ defines a product mapping allowing
high-quality product $p'$ (e.g. stems) satisfying demand of a
lower-quality products $p$ up to maximum demand $D^{\max}_{kpt}$. This
way, the cascading principle is operationalized as a central economic
rationale.

\begin{align}
  % --- C1: Establishment (at most once per site) ---
  \sum_{(0,t) \in \Sc^{{est}}} z_{i0t} &\leq AREA_i && \forall\; i \in \Ic \label{eq:C1} \\[2pt]
  %
  % --- C2: Path connectivity (flow conservation) ---
  \sum_{(s,t) \in \Sc} z_{ist} &= \sum_{(t,u) \in \Sc} z_{itu} && \forall\; i \in \Ic,\; t \in \Tc \label{eq:C2} \\[2pt]
  %
  % --- C3: Biomass yield from Gompertz coefficients ---
  \sum_{j \in \Jc} X_{ijpt}    &\leq \sum_{(s,t) \in \Sc^{{harv}}} \eta_{p(t-s)} \cdot z_{ist} && \forall\; i \in \Ic,\; p \in \Pc,\; t \in \Tc^{{harv}}     \label{eq:C3} \\[2pt]
  %
  % --- C5: Inventory balance ---
  S_{jpt}    = S_{jp,t-1} + &\sum_{i \in \Ic} X_{ijpt} - \sum_{k \in \Kc,\,(p,p') \in \Qc} X_{jkpp't}   && \forall\; j \in \Jc,\; p \in \Pc,\; t \in \Tc^{harv}  \label{eq:C5} \\[2pt]
  %
  % --- C6: Storage capacity ---
  \sum_{p \in \Pc} S_{jpt}    &\leq \text{CAP}^{{stor}}_j    && \forall\; j \in \Jc,\; t \in \Tc     \label{eq:C6} \\[2pt]
  %
  % --- C7: Processing capacity ---
  \sum_{i \in \Ic,\, p \in \Pc} X_{ijpt}  &\leq \text{CAP}^{{proc}}_j && \forall\; j \in \Jc,\; t \in \Tc^{harv} \label{eq:C7} \\[2pt]
  %
  % --- C8: Demand satisfaction with quality cascade ---
  \sum_{j \in \Jc,\,(p',p) \in \Qc} X_{jkp'pt} &\leq D^{\max}_{kpt} && \forall\; k \in \Kc,\; p \in \Pc,\; t \in \Tc^{harv}     \label{eq:C8} \\[2pt]
  %
  % --- Domain: continuous non-negativity ---
  z_{ist},\;X_{ijpt},\; &X_{jkp'pt},\; S_{jpt} \geq 0 && \forall\; i,j,k,p,p',t     \label{eq:D2}
  %
\end{align}

The yield coefficients $\eta_{pa}$ in \eqref{eq:C3} are the critical
bridge between the stand-level growth model of Chapter\~2 and the supply
chain optimisation. For stand age $a = t - s$ and product component $p$,
they are computed as: 
where $M^{sl}(a)$ follows from
\eqref{eq:total-stand-biomass} and $f_p(a)$ follows from
\eqref{eq:component-share}. For poplar clone Max\~3 under SRAFS
conditions, the component-specific Gompertz parameterisations from
Section \@ref(sec:lit-growth) imply that $\eta_{\text{stem},a}$ rises
steeply between ages\~4 and 10, while $\eta_{\text{branch},a}$ and
$\eta_{\text{residue},a}$ plateau earlier, as illustrated in
Figure \@ref(fig:max3-growth). This age-dependence of product fractions
is the core mechanism through which rotation-length decisions ---
governed by $A^{\min}$ and $A^{\max}$ and the arc set $\mathcal{S}$ ---
determine achievable product portfolios and, hence, revenues in the
objective function \eqref{eq:obj-revenue}.

# Case Study {#sec-casestudy}

The proposed MILP model is intended to support the design of
agroforestry biomass supply chains for cleaner production in specific
regional contexts. In a forthcoming case study, we focus on poplar-based
alley-cropping systems in a temperate European setting, informed by
empirical data from SRAFS trials in Southern Germany and SRC plantations
in Central Europe [@Huber2018_SRAFS_growth; @Trnka2008_Poplar_SRF;
@Testa2014_Poplar_Economics].

```{r fig-supply-chain-network, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="AFS supply chain network in Saxony-Anhalt. Grey lines: German federal state boundaries. Blue lines: major rivers. Coloured polygons: potential KUP/AFS sites grouped by HAC super-site (colours encode cluster membership; no legend). Orange symbols: pre-treatment facilities (circle = sawmill, triangle = logistics hub, size proportional to processing capacity). Coloured squares: consumer sites (violet: P1 chemical/pulp, blue: P2 pulp/paper, red: P3 energy/biogas, size proportional to total demand).", fig.width=6.5, fig.height=5.5, out.width="\\textwidth", cache=F}

crs_map <- 25833

# ── Colour palette constants (identical to fig-results-sc-map) ────────────────
COL_SAW <- "#e65c00"
COL_HUB <- "#c05000"
COL_P1  <- "#6a0dad"
COL_P2  <- "#08519c"
COL_P3  <- "#a50026"

# ── Map extent: defined in WGS84, transformed to projected CRS ───────────────
bbox_wgs84 <- st_as_sfc(
  st_bbox(c(xmin = 10.6, xmax = 13.2, ymin = 50.9, ymax = 52.8),
          crs = st_crs(4326))
)
bbox_map <- st_transform(bbox_wgs84, crs_map)
bbox_num <- st_bbox(bbox_map)

# ── Base-map layers ────────────────────────────────────────────────────────────
de_states <- ne_states(country = "Germany", returnclass = "sf") %>%
  st_transform(crs_map) %>%
  st_intersection(bbox_map)

rivers_eu <- ne_download(
  scale    = 10,
  type     = "rivers_lake_centerlines",
  category = "physical",
  returnclass = "sf"
) %>%
  st_transform(crs_map) %>%
  st_intersection(bbox_map)

# ── Cluster colour palette (no legend printed) ────────────────────────────────
cluster_assig <- afs_workspace$site_cluster_assig %>%
  select(site_id, hac_cluster) %>%
  mutate(hac_cluster = as.factor(hac_cluster))

n_clusters <- length(unique(cluster_assig$hac_cluster))

cluster_pal <- setNames(
  if (n_clusters <= 12) {
    colorRampPalette(RColorBrewer::brewer.pal(min(n_clusters, 12), "Set3"))(n_clusters)
  } else {
    scales::viridis_pal(option = "turbo")(n_clusters)
  },
  levels(cluster_assig$hac_cluster)
)

# ── Site polygons: join cluster assignment, project ───────────────────────────
sites_sf_clust <- sites_sf %>%
  mutate(site_id = seq_len(nrow(sites_sf))) %>%
  left_join(cluster_assig, by = "site_id") %>%
  st_transform(crs_map)

# merge opportunity cost of 
tmp.opp <- sites_clustered %>% 
  select(site_id, C_opp) %>% 
  rename(hac_cluster = site_id) %>% 
  mutate(hac_cluster = as.factor(hac_cluster))

sites_sf_clust <- sites_sf_clust %>% 
  left_join(tmp.opp,by = "hac_cluster")

# ── Consumers & storages ───────────────────────────────────────────────────────
consumers_plt <- consumers %>%
  mutate(
    total_demand = demand_P1 + demand_P2 + demand_P3,
    kategorie = case_when(
      demand_P1 >= demand_P2 & demand_P1 >= demand_P3 & demand_P1 > 0 ~ "Chemical / Pulp (P1)",
      demand_P2 >= demand_P1 & demand_P2 >= demand_P3 & demand_P2 > 0 ~ "Pulp / Paper (P2)",
      demand_P3 > 0                                                     ~ "Energy / Biogas (P3)",
      TRUE                                                              ~ "Other"
    ),
    pt_size = pmin(4.5, pmax(1.5, 1.5 + log1p(total_demand) * 0.55))
  )

storages_plt <- storages %>%
  mutate(
    #subtype = if_else(type == "Hub", "Logistics Hub", "Sawmill"),
    subtype = if_else(type == "Hub", "Hub", "Hub"),
    pt_size = 3.5
  )

stor_sf <- st_as_sf(storages_plt, coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(crs_map)

cons_sf <- st_as_sf(consumers_plt, coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(crs_map)

# ── Plot ───────────────────────────────────────────────────────────────────────
ggplot() +

  # Base map
  geom_sf(data = de_states,
          fill = "grey96", colour = "grey55", linewidth = 0.35) +
  geom_sf(data = rivers_eu,
          colour = "#4292c6", linewidth = 0.45, alpha = 0.7) +

  # KUP site polygons coloured by cluster — legend suppressed via guide = "none"
  geom_sf(data      = sites_sf_clust,
          aes(fill  = C_opp),
          colour    = "grey30",
          linewidth = 0.05,
          alpha     = 0.65,
          show.legend = FALSE) +
  scale_fill_gradient2(
  name     = "Opportunity cost (€/ha)",
  low      = "#1a9641",   # green  → low / negative C_opp
  mid      = "#ffffbf",   # pale yellow → midpoint
  high     = "#d7191c",   # red    → high / positive C_opp
  midpoint = 0,           # centre the diverging scale on zero
  na.value = "grey",
  guide    = guide_colourbar(
    barwidth       = 0.5,
    barheight      = 4,
    title.position = "top"
  ) 
  ) +
  # scale_fill_manual(
  #   values = cluster_pal,
  #   guide  = "none"          # colours visible on map; no legend entry printed
  # ) +

  # Pre-treatment facilities — new colour + shape scale after cluster fill
  ggnewscale::new_scale_colour() +
  ggnewscale::new_scale("shape") +
  ggnewscale::new_scale_fill() +
  geom_sf(data = stor_sf,
          aes(colour = subtype, shape = subtype, size = pt_size),
          stroke = 1.2, alpha = 0.92, show.legend = F) +
  scale_colour_manual(
    name   = "Pre-treatment",
    values = c("Sawmill" = COL_SAW, "Hub" = COL_HUB)
  ) +
  # scale_shape_manual(
  #   name   = "Pre-treatment",
  #   values = c("Sawmill" = 16, "Logistics Hub" = 17)
  # ) +

  # Consumer sites — fill scale registered after new_scale_fill() above
  geom_sf(data  = cons_sf,
          aes(fill = kategorie, size = pt_size),
          shape = 22, colour = "white", stroke = 0.45,
          alpha = 0.88, show.legend = TRUE) +
  scale_fill_manual(
    name   = "Consumer type",
    values = c(
      "Chemical / Pulp (P1)" = COL_P1,
      "Pulp / Paper (P2)"    = COL_P2,
      "Energy / Biogas (P3)" = COL_P3,
      "Other"                = "grey60"
    )
  ) +

  scale_size_continuous(range = c(1.5, 5), guide = "none") +
  guides(
    fill   = guide_legend(order = 1,
                          override.aes = list(size = 3.5, stroke = 0.4,
                                              colour = "white")),
    colour = guide_legend(order = 2,
                          override.aes = list(size = 4, stroke = 1.5)),
    shape  = guide_legend(order = 2,
                          override.aes = list(size = 4, stroke = 1.5))
  ) +

  coord_sf(
    xlim   = c(bbox_num["xmin"], bbox_num["xmax"]),
    ylim   = c(bbox_num["ymin"], bbox_num["ymax"]),
    expand = FALSE
  ) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid      = element_line(colour = "grey88", linewidth = 0.25),
    legend.position = "right",
    legend.key.size = unit(0.40, "cm"),
    legend.text     = element_text(size = 6),
    legend.title    = element_text(size = 7.5, face = "bold"),
    axis.text       = element_text(size = 7)
  )
```

## Study Region and Supply Chain Configuration

The proposed \problem model is instantiated on a real-world biomass
supply chain located in the tri-state region of Saxony-Anhalt, northern
Thuringia, and northern Saxony, Germany (Figure
\@ref(fig:fig-supply-chain-network)). This region combines the highest
density of registered short-rotation coppice (SRC) Feldblöcke in eastern
Germany with a cluster of major wood-processing industries that
collectively represent one of the largest forest-based biomass demand
centres in Central Europe. The spatial extent of the study region covers
approximately $250 \times 200$ km, with road distances between potential
KUP sites and consumer locations ranging from under 10 km to over 180
km.

The supply chain comprises three node types: ($i$) KUP production sites
$\mathcal{I}$, ($ii$) pre-treatment and intermediate storage facilities
$\mathcal{J}$, and ($iii$) consumer sites $\mathcal{K}$. Candidate KUP
sites are drawn from the official Feldblock register of Saxony-Anhalt,
filtered to parcels with a minimum area of 1 ha and classified as
potentially suitable for SRC under the InVeKoS land-use scheme.
Following the two-stage spatial clustering procedure described in
Section \@ref(sec-casestudy), the
$|\mathcal{I}| = `r nrow(afs_workspace$sites)`$ raw Feldblöcke are
grouped into $K$ super-sites using DBSCAN ($\varepsilon = 15$ km,
$\mathit{minPts} = 3$) and Ward.D2 hierarchical clustering (maximum
within-cluster radius 15 km), yielding a computationally tractable MILP
instance while preserving the total eligible area. The HAC cluster
assignments, visible as distinct colour zones in Figure
\@ref(fig:fig-supply-chain-network), confirm a spatially coherent
partitioning: dense KUP corridors along the Elbe and Saale floodplains
form contiguous super-sites, while isolated parcels in the Harz
foothills constitute smaller or singleton clusters.

Two types of intermediate infrastructure nodes are considered:
conventional sawmills and logistics hubs. Sawmills perform primary
mechanical processing (debarking, chipping) and can serve product
streams P1 and P2; logistics hubs serve as consolidation and temporary
storage points for all three product grades and are characterised by
lower capital intensity but limited processing capacity. The
$|\mathcal{J}| = `r nrow(afs_workspace$storages)`$ candidate
storage/pre-treatment locations are sourced from publicly registered
forestry enterprises and existing logistic nodes in the study region.
Road distances $d_{ij}$ (site $i$ to storage $j$) and $d_{jk}$ (storage
$j$ to consumer $k$) are computed via the OSRM open-source routing
engine using the German road network, providing realistic transport cost
estimates that are embedded directly in the MILP objective.

## Consumer Demand and Product Classification

Nine consumer sites are included in the base case ($|\mathcal{K}| = 9$),
covering the full quality cascade from high-value chemical/pulp
applications down to energy use (Table \@ref(tab:tab-consumers)).
Consumer demands $D_{kpt}^{\max}$ are derived from publicly reported
annual throughput figures, production capacities, and regulatory
filings, and are held constant across periods
($D_{kpt}^{\max} = D_{kp}^{\max}\ \forall\, t \in \mathcal{T}$) to
reflect stable long-term offtake contracts.

Three biomass product grades are distinguished, reflecting the
cascading-use principle embedded in the \problem objective:

- **P1 – Chemical/pulp grade** ($p = 1$): long-fibre hardwood chips
  suitable for chemical pulping and biorefinery conversion. Dominated by
  Mercer Stendal GmbH (Arneburg, 1{,}500 kt/yr) and Zellstoff
  Rosenthal/Mercer (Blankenstein, 1{,}000 kt/yr), which together account
  for `r round(100*(1500+1000)/sum(consumers$demand_P1), 0)`% of total
  P1 demand. UPM Biochemicals (Leuna) contributes 500 kt/yr of
  high-purity beech-fraction demand for biochemical conversion. Gate
  prices for P1 range from 80 to 120 €/t wood input.

- **P2 – Sawlog/pulpwood grade** ($p = 2$): roundwood and chipped
  material suitable for mechanical processing in sawmills. Demand is
  distributed across six smaller facilities with individual annual
  demands of 5–300 kt/yr. Gate prices range from 55 to 75 €/t.

- **P3 – Energy/pellets grade** ($p = 3$): residual wood fractions
  (chips, bark, off-cuts) used for heat and electricity generation or
  pellet production. This grade absorbs material that cannot meet P1 or
  P2 quality thresholds, operationalising the cascade constraint in the
  MILP. Gate prices range from 30 to 40 €/t.

The demand derivation follows a three-step procedure. First, publicly
available capacity data (kt wood input per year) are collected for each
facility from official sources, industry reports, and corporate
disclosures [@mercerStendal2024; @upmBiochemicals2024]. Second, the
total annual capacity is assigned to the product grade that corresponds
to the primary feedstock of each facility (P1 for chemical/biorefinery
processes, P2 for mechanical sawmilling, P3 for thermal conversion and
pellet lines). Third, a minor P3 component is added to sawmill
facilities to capture by-product streams (bark, sawdust), consistent
with observed industrial practice. The resulting demand matrix
$D_{kp}^{\max}$ (kt/yr) is shown in Figure
\@ref(fig:fig-consumer-demand) and summarised in Table
\@ref(tab:tab-consumers). Aggregate demand across the study region
totals `r sum(consumers$demand_P1)` kt/yr for P1,
`r sum(consumers$demand_P2)` kt/yr for P2, and
`r sum(consumers$demand_P3)` kt/yr for P3.

```{r tab-consumers, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}

consumers %>%
  mutate(name_pseudo = paste("consumer ", seq_len(nrow(consumers))) ) %>% 
  select(name_pseudo, demand_P1, demand_P2, demand_P3, P1, P2, P3) %>%
  mutate(across(c(demand_P1, demand_P2, demand_P3), ~ ifelse(.x == 0, "–", format(.x, big.mark = ",")))) %>%
  mutate(across(c(P1, P2, P3), ~ ifelse(.x == 0, "–", paste0(.x)))) %>%
  kbl(
    col.names = c("Consumer", "P1 (kt/yr)", "P2 (kt/yr)", "P3 (kt/yr)",
                  "P1 (€/t)", "P2 (€/t)", "P3 (€/t)"),
    caption   = "Consumer demand quantities and gate prices by product grade. P1: chemical/pulp; P2: sawlog/pulpwood; P3: energy/pellets. Demands in kt wood input per year; prices in €/t wood input at the factory gate.",
    booktabs  = TRUE,
    align     = c("l", rep("r", 6)),
    label     = "tab-consumers"
  ) %>%
  kable_styling(latex_options = c("striped", "hold_position", "scale_down"),
                font_size = 8) %>%
  add_header_above(c(" " = 1, "Annual demand" = 3, "Gate price" = 3))
```

```{r fig-consumer-demand, cache=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Total biomass demand by consumer type and grade (P1 chemical, P2 pulp \\& paper, P3 energy/biogas). Bars show aggregated annual demand in kt wood input per year across all individual plants within each consumer type.", fig.width = 5.5, fig.height = 2.8, out.width="\\textwidth"}

# Re-use consumers_typed from above
# create unique types of consumers

consumers_typed <- consumers %>%
  mutate(
    type = case_when(
      demand_P1 >= demand_P2 & demand_P1 >= demand_P3 & demand_P1 > 0 ~ "Industrial",
      demand_P2 >= demand_P1 & demand_P2 >= demand_P3 & demand_P2 > 0 ~ "Paper & pulp",
      demand_P3 > 0 ~ "Biogas",
      TRUE          ~ "Other"
    )
  )


# Aggregate and pivot to long
demand_agg <- consumers_typed %>%
  group_by(type) %>%
  summarise(
    P1 = sum(demand_P1),
    P2 = sum(demand_P2),
    P3 = sum(demand_P3),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(P1, P2, P3),
               names_to  = "grade",
               values_to = "demand_kt") %>%
  filter(demand_kt > 0) %>%
  mutate(
    grade = factor(grade, levels = c("P1", "P2", "P3"),
                   labels = c("P1 – Chemical", "P2 – Pulp", "P3 – Energy")),
    type  = factor(type,
                   levels = c("Industrial", "Paper & pulp", "Biogas"))
  )

GRADE_COLS <- c("P1 – Chemical" = "#1b7837",
                "P2 – Pulp"     = "#4393c3",
                "P3 – Energy"   = "#d6604d")

ggplot(demand_agg, aes(x = type, y = demand_kt, fill = grade)) +
  geom_col(position = "dodge", width = 0.65, colour = "white", linewidth = 0.2) +
  geom_text(aes(label = round(demand_kt, 0)),
            position = position_dodge(width = 0.65),
            vjust = -0.35, size = 2.4, colour = "grey30") +
  scale_fill_manual(values = GRADE_COLS, name = "Biomass grade") +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.15)),
    name   = "Annual demand (kt wood input / yr)"
  ) +
  labs(x = NULL) +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    legend.position    = "bottom",
    legend.key.size    = unit(0.35, "cm"),
    legend.text        = element_text(size = 7),
    axis.text.x        = element_text(size = 8)
  )
```

## MILP Model Parameters

Table \@ref(tab:tab-milp-params) summarises all scalar parameters of the
\problem instance. The planning horizon covers $|\mathcal{T}| = 15$
periods, corresponding to one full poplar rotation (years 1–15), with a
minimum harvest age $a^{\min} = 4$ and maximum productive stand age
$a^{\max} = 15$ [@lehmkuhl2023]. Biomass yield functions $\eta_{pa}$
(odt ha$^{-1}$ yr$^{-1}$) follow the age-structured Gompertz
profile calibrated in Section \@ref(sec:lit-growth), where distinct yield
curves are assigned to each product grade to reflect the fraction of
stem, branch, and foliage biomass that qualifies under each quality
threshold.

Transport costs are disaggregated into two components reflecting the
two-echelon logistics structure. Raw material transport from site $i$ to
pre-treatment facility $j$ incurs a variable cost
$c^{tr\text{-}raw} \cdot d_{ij}$ (€ t$^{-1}$ km$^{-1}$), while processed
material transport from $j$ to consumer $k$ is costed at
$c^{tr\text{-}pre} \cdot d_{jk}$. The pre-treatment rate
$c^{tr\text{-}pre}$ is set lower than $c^{tr\text{-}raw}$ to reflect the
higher load density and standardised packaging of chipped and pelletised
material compared to loose roundwood. An opportunity cost parameter
$c^{opp}$ (€ ha$^{-1}$ yr$^{-1}$) accounts for foregone agricultural
revenue on converted parcels, calibrated to the average arable land
rental rate in Saxony-Anhalt [@destatis2023]. Storage capacity at each
facility $j$ is bounded by $\overline{S}_j$, and processing capacity by
$\overline{C}_j$ (kt yr$^{-1}$), both drawn from publicly available
facility data.

```{r tab-milp-params, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
params_df <- data.frame(
  Symbol          = c("$|\\mathcal{T}|$", "$|\\mathcal{I}|$", "$|\\mathcal{J}|$",
                      "$|\\mathcal{K}|$", "$|\\mathcal{P}|$",
                      "$a^{\\min}$", "$a^{\\max}$",
                      "$c^{tr\\text{-}raw}$", "$c^{tr\\text{-}pre}$",
                      "$c^{opp}$",
                      "$\\bar{R}_{P1}$", "$\\bar{R}_{P2}$", "$\\bar{R}_{P3}$"),
  Description     = c("Planning periods", "KUP super-sites (post-clustering)",
                      "Pre-treatment / storage nodes", "Consumer sites", "Product grades",
                      "Minimum harvest age (years)", "Maximum stand age (years)",
                      "Raw transport cost", "Pre-processed transport cost",
                      "Opportunity cost",
                      "Mean gate price P1", "Mean gate price P2", "Mean gate price P3"),
  Value           = c(milp_instance$n_periods, milp_instance$n_sites, nrow(afs_workspace$storages),
                      milp_instance$n_consumers, "3",
                      milp_instance$min_age, milp_instance$max_age,
                      milp_instance$c_tr_raw, milp_instance$c_tr_pre,
                      round(mean(milp_instance$sites$C_opp),0),
                      round(mean(milp_instance$consumers$P1[milp_instance$consumers$P1>0]),0),
                      round(mean(milp_instance$consumers$P2[milp_instance$consumers$P2>0]),0),
                      round(mean(milp_instance$consumers$P3[milp_instance$consumers$P3>0]),0)),
  Unit            = c("yr", "–", "–", "–", "–",
                      "yr", "yr",
                      "€ t$^{-1}$ km$^{-1}$", "€ t$^{-1}$ km$^{-1}$",
                      "€ ha$^{-1}$ yr$^{-1}$",
                      "€ t$^{-1}$", "€ t$^{-1}$", "€ t$^{-1}$"),
  stringsAsFactors = FALSE
)

params_df %>%
  kbl(
    col.names = c("Parameter", "Description", "Base-case value", "Unit"),
    caption   = "\\problem base-case parameter values for the Saxony-Anhalt case study. Transport costs refer to road distance (OSRM). Gate prices are weighted averages across consumer sites.",
    booktabs  = TRUE,
    escape    = FALSE,
    align     = c("l", "l", "r", "l"),
    label     = "tab-milp-params"
  ) %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 8) %>%
  pack_rows("Index sets", 1, 5) %>%
  pack_rows("Biomass dynamics", 6, 7) %>%
  pack_rows("Cost parameters", 8, 10) %>%
  pack_rows("Revenue parameters", 11, 13)
```

## Optimization results

The AFS-SCD model is solved for the Saxony-Anhalt base-case instance
described in Section \@ref(sec-casestudy). All results presented in
this section refer to the base-case parametrisation summarised in
Table \@ref(tab:tab-milp-params). The model selects a subset of KUP
super-sites for establishment and identifies the optimal harvest
schedule, intermediate facility utilisation, and product allocation
across the 15-period planning horizon.

Figure \@ref(fig:fig-results-sc-map) shows the spatial configuration of
the optimal supply chain. Active super-sites are concentrated in the
Elbe and Saale floodplains, where high biomass yields and proximity to
major processing facilities create the most favourable economics. Sites
in the Harz foothills and peripheral areas with high opportunity costs
are not selected. The pie charts overlaid on each active site indicate
the product-mix composition: sites closer to P1 consumers (chemical/pulp
grade) allocate a larger stem fraction, while more remote sites rely
more heavily on P3 (energy) grade biomass.

Figures \@ref(fig:fig-results-alluvial)
and \@ref(fig:fig-results-deliveries) report the biomass flows over the
planning horizon. Total annual biomass throughput rises steeply in the
first five periods as newly established sites reach minimum harvest age,
stabilises during the middle of the horizon, and undergoes a partial
decline towards the end as rotation cycles exhaust the planning window.
The product cascade is clearly visible: stem biomass (P1) dominates the
early harvest periods when young stands deliver the highest merchantable
stem fraction, while branch and residue shares increase as stands mature
and are harvested at longer rotation lengths.

The distribution of harvesting cycle lengths at active sites is
illustrated in Figure \@ref(fig:fig-results-harvest-cycles). The
majority of harvest events occur at stand ages between the minimum age
$a^{\min} = 4$ years and approximately 8--10 years, reflecting the
trade-off between cumulative yield growth and the opportunity cost of
delayed harvesting. A secondary peak at longer rotation lengths
(12--15 years) corresponds to sites with lower opportunity costs and
higher P1 demand in their vicinity.

Figure \@ref(fig:fig-results-costs-profit) presents the aggregate
cost--profit waterfall over the full horizon. Total revenue is dominated
by P1 deliveries to the two large pulp mills. Opportunity costs and
transport costs are the two largest cost drivers; establishment costs
are one-time and comparatively modest. The site-level revenue--cost
decomposition in Figure \@ref(fig:fig-results-site-profit) confirms that
profitability is highly heterogeneous: a small number of large sites
with low opportunity costs and short distances to P1 consumers account
for the bulk of total net profit, while several marginal sites operate
near break-even. Figure \@ref(fig:fig-results-opp-dist) reveals a clear
negative relationship between opportunity cost rate and site
profitability, with sites exceeding 300 \euro\ ha$^{-1}$ yr$^{-1}$ rarely
selected in the optimal solution.

```{r fig-results-sc-map, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Optimal agroforestry supply chain network. Green polygons: active super-sites with biomass harvest; grey polygons: non-selected sites. Pie charts at active sites show the product mix of harvested biomass (stem = P1, branches = P2, residues = P3), scaled by total harvested volume. Orange symbols: pre-treatment facilities (circle = sawmill, triangle = logistics hub). Coloured squares: consumer sites by dominant demand category.", fig.width=8, fig.height=7, dev="pdf", out.width="\\textwidth", cache=T, cache.extra = c("ext","sites_clustered", "storages" , "customers", "milp_instance")}


crs_map <- 25833

# ── Colour palette constants ─────────────────────────────────────────────────
COL_SAW       <- "#e65c00"
COL_HUB       <- "#c05000"
COL_P1        <- "#6a0dad"
COL_P2        <- "#08519c"
COL_P3        <- "#a50026"
COL_SITE_NONE <- "#bdbdbd"

pal_products <- c(
  p1 = "#1b9e77",
  p2 = "#d95f02",
  p3 = "#7570b3"
)

# ── Map extent ────────────────────────────────────────────────────────────────
bbox_wgs84 <- st_as_sfc(
  st_bbox(c(xmin = 10.6, xmax = 13.2, ymin = 50.9, ymax = 52.8), crs = st_crs(4326))
)
bbox_map <- st_transform(bbox_wgs84, crs_map)
bbox_num <- st_bbox(bbox_map)

# ── Base-map layers ───────────────────────────────────────────────────────────
de_states <- ne_states(country = "Germany", returnclass = "sf") %>%
  st_transform(crs_map) %>%
  st_intersection(bbox_map)

rivers_eu <- ne_download(
  scale = 10,
  type = "rivers_lake_centerlines",
  category = "physical",
  returnclass = "sf"
) %>%
  st_transform(crs_map) %>%
  st_intersection(bbox_map)

# ── Aggregate harvest volumes per site over all periods ──────────────────────
site_harvest <- ext$Xij %>%
  group_by(site_id, p) %>%
  summarise(vol = sum(value, na.rm = TRUE), .groups = "drop") %>%
  tidyr::pivot_wider(
    names_from   = p,
    values_from  = vol,
    values_fill  = 0,
    names_prefix = "p"
  )


for (col in c("p1", "p2", "p3")) {
  if (!col %in% names(site_harvest)) site_harvest[[col]] <- 0
}

site_harvest <- site_harvest %>%
  mutate(total_vol = p1 + p2 + p3) %>%
  filter(total_vol > 0)

active_site_ids <- unique(site_harvest$site_id)

# ── Site polygons ─────────────────────────────────────────────────────────────
sites_sf_map <- sites_sf %>%
  mutate(site_id = seq_len(nrow(sites_sf))) %>%
  inner_join(afs_workspace$site_cluster_assig, "site_id") %>%
  st_transform(crs_map) %>%
  mutate(active = hac_cluster %in% active_site_ids)

tmp.opp <- sites_clustered %>%
  select(site_id, C_opp) %>%
  rename(hac_cluster = site_id)

sites_sf_map <- sites_sf_map %>%
  left_join(tmp.opp, by = "hac_cluster")

# Split into active and inactive subsets for separate fill scales
# CHANGED: split polygons so inactive uses fixed grey fill and
#          active uses a continuous C_opp colour scale
sites_inactive <- sites_sf_map %>% filter(!active)
sites_active   <- sites_sf_map %>% filter(active)

# ── Pie-chart data ────────────────────────────────────────────────────────────
sites_pie_sf <- milp_instance$sites %>%
  inner_join(site_harvest, by = "site_id") %>%
  st_as_sf(coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(crs_map)

coords_sites <- st_coordinates(sites_pie_sf)

sites_pie <- sites_pie_sf %>%
  st_drop_geometry() %>%
  mutate(
    x = coords_sites[, 1],
    y = coords_sites[, 2]
  )

vol_med <- median(sites_pie$total_vol[sites_pie$total_vol > 0], na.rm = TRUE)
if (is.na(vol_med) || vol_med <= 0) vol_med <- 1

sites_pie <- sites_pie %>%
  mutate(r = 3500 * sqrt(total_vol / vol_med))

# ── Consumers & storages ──────────────────────────────────────────────────────
consumers_plt <- consumers %>%
  mutate(
    total_demand = demand_P1 + demand_P2 + demand_P3,
    kategorie = case_when(
      demand_P1 >= demand_P2 & demand_P1 >= demand_P3 & demand_P1 > 0 ~ "Chemical / Pulp (P1)",
      demand_P2 >= demand_P1 & demand_P2 >= demand_P3 & demand_P2 > 0 ~ "Pulp / Paper (P2)",
      demand_P3 > 0 ~ "Energy / Biogas (P3)",
      TRUE ~ "Other"
    ),
    pt_size = pmin(4.5, pmax(1.5, 1.5 + log1p(total_demand) * 0.55))
  )

storages_plt <- storages %>%
  mutate(
    subtype = if_else(type == "Hub", "Logistics Hub", "Sawmill"),
    pt_size = pmin(3.5, pmax(1.2, 1.2 + log1p(CAP_proc) * 0.38))
  )

stor_sf <- st_as_sf(storages_plt, coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(crs_map)

cons_sf <- st_as_sf(consumers_plt, coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(crs_map)

# ── Plot ──────────────────────────────────────────────────────────────────────
ggplot() +
  # Base map
  geom_sf(data = de_states,
          fill = "grey96", colour = "grey55", linewidth = 0.35) +
  geom_sf(data = rivers_eu,
          colour = "#4292c6", linewidth = 0.35, alpha = 0.7) +

  # CHANGED: Inactive site polygons — fixed grey fill, no legend scale needed
  geom_sf(data = sites_inactive,
          fill = COL_SITE_NONE,
          colour = NA, alpha = 0.55, show.legend = FALSE) +

  # CHANGED: Active site polygons — continuous fill mapped to C_opp
  geom_sf(data = sites_active,
          aes(fill = C_opp),
          colour = NA, alpha = 0.75, show.legend = TRUE) +
 scale_fill_gradient2(
  name     = "Opportunity cost(€/ha)",
  low      = "#1a9641",   # green  → low / negative C_opp
  mid      = "#ffffbf",   # pale yellow → midpoint
  high     = "#d7191c",   # red    → high / positive C_opp
  midpoint = 0,           # centre the diverging scale on zero
  na.value = COL_SITE_NONE,
  guide    = guide_colourbar(
    barwidth       = 0.5,
    barheight      = 4,
    title.position = "top"
  ) 
  ) +

  # Pre-treatment facilities
  ggnewscale::new_scale_colour() +
  ggnewscale::new_scale("shape") +
  geom_sf(data = stor_sf,
          aes(colour = subtype, shape = subtype, size = pt_size),
          stroke = 1.0, alpha = 0.92, show.legend = TRUE) +
  scale_colour_manual(
    name   = "Pre-treatment",
    values = c("Sawmill" = COL_SAW, "Logistics Hub" = COL_HUB)
  ) +
  scale_shape_manual(
    name   = "Pre-treatment",
    values = c("Sawmill" = 16, "Logistics Hub" = 17)
  ) +

  # Consumer sites
  ggnewscale::new_scale_fill() +
  geom_sf(data = cons_sf,
          aes(fill = kategorie, size = pt_size),
          shape = 22, colour = "white", stroke = 0.45,
          alpha = 0.9, show.legend = TRUE) +
  scale_fill_manual(
    name   = "Consumer type",
    values = c(
      "Chemical / Pulp (P1)" = COL_P1,
      "Pulp / Paper (P2)"    = COL_P2,
      "Energy / Biogas (P3)" = COL_P3,
      "Other"                = "grey60"
    )
  ) +

  # Pie charts at active sites
  ggnewscale::new_scale_fill() +
  scatterpie::geom_scatterpie(
    data         = sites_pie,
    aes(x = x, y = y, r = r),
    cols         = c("p1", "p2", "p3"),
    colour       = "white",
    linewidth    = 0.15,
    donut_radius = .5,
    alpha        = 0.7,
    show.legend  = TRUE
  ) +
  scale_fill_manual(
    name   = "Harvested biomass",
    values = pal_products,
    labels = c(p1 = "Stem", p2 = "Branches", p3 = "Residues")
  ) +

  scale_size_continuous(range = c(1.5, 5), guide = "none") +
  guides(
    fill   = guide_legend(order = 1),
    colour = guide_legend(order = 2,
                          override.aes = list(size = 4, stroke = 1.2)),
    shape  = guide_legend(order = 2,
                          override.aes = list(size = 4, stroke = 1.2))
  ) +

  coord_sf(
    xlim   = c(bbox_num["xmin"], bbox_num["xmax"]),
    ylim   = c(bbox_num["ymin"], bbox_num["ymax"]),
    expand = FALSE
  ) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid      = element_line(colour = "grey88", linewidth = 0.25),
    legend.position = "right",
    legend.key.size = unit(0.40, "cm"),
    legend.text     = element_text(size = 6),
    legend.title    = element_text(size = 7.5, face = "bold"),
    axis.text       = element_text(size = 7)
  )
```

```{r fig-results-harvest-gantt, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Total production quantity of biomass types and its industrial usage over planning horizon (in tons).", fig.width=7, fig.height=5.5, dev="pdf", out.width="\\textwidth", cache=T, cache.extra = c("ext","sites_clustered", "storages" , "customers", "milp_instance")}


# ── Labels & colours ──────────────────────────────────────────────────────────
# Source side (axis 1): biomass fractions delivered from sites → hubs (Xij)
#   product index 1 = Stems, 2 = Branches, 3 = Residues  [§2.2, f_p(t)]
SRC_LABEL <- c(
  "1" = "Stems",
  "2" = "Branches",
  "3" = "Residues"
)

# Delivery side (axis 2): consumer categories receiving from hubs (Xjk)
#   del_product 1 = Industrial (P1), 2 = Pulp & Paper (P2), 3 = Energy (P3)
DEL_LABEL <- c(
  "1" = "Industrial",
  "2" = "Pulp & Paper",
  "3" = "Energy"
)

# Fill colours for source strata — must match paper Fig. max3-growth
P_COLS_SRC <- c(
  "Stems"    = "#1b9e77",
  "Branches" = "#d95f02",
  "Residues" = "#7570b3"   
)

# ── Aggregate Xjk: src_product × del_product, summed over all periods/consumers

flow_df <- ext$Xjk %>%
  group_by(src_product, del_product) %>%
  summarise(volume = sum(value, na.rm = TRUE), .groups = "drop") %>%
  filter(volume > 0) %>%
  mutate(
    # axis1 = biomass fraction (source side)
    src = factor(SRC_LABEL[as.character(src_product)], levels = SRC_LABEL),
    # axis2 = consumer category (delivery/sink side)
    del = factor(DEL_LABEL[as.character(del_product)], levels = DEL_LABEL)
  )

# ── Plot ──────────────────────────────────────────────────────────────────────
ggplot(flow_df,
       aes(axis1 = src, axis2 = del, y = volume, fill = src)) +

  geom_alluvium(
    alpha      = 0.72,
    width      = 1 / 3,
    knot.pos   = 0.45,
    curve_type = "sigmoid"
  ) +

  geom_stratum(
    width     = 1 / 3,
    fill      = "grey94",
    colour    = "grey50",
    linewidth = 0.3
  ) +

  geom_text(
    stat       = "stratum",
    aes(label  = after_stat(stratum)),
    size       = 2.8,
    colour     = "grey15",
    lineheight = 0.88
  ) +

  # Axis header annotations — placed below the stratum columns
  # x = 1 (axis1 = Feedstock fraction), x = 2 (axis2 = End use)
  annotate("text",
    x        = c(1, 2),
    y        = -Inf,
    label    = c("Feedstock", "End use"),
    vjust    = 1.6,
    size     = 3.0,
    fontface = "bold",
    colour   = "grey30"
  ) +

  scale_fill_manual(
    name   = "Biomass fraction",
    values = P_COLS_SRC          # only 3 source entries; no ghost colours
  ) +

  # ggalluvial uses integer axis positions (1, 2, …); do NOT set
  # limits = c("src", "del") — that breaks stratum positioning
  scale_x_discrete(
    expand = expansion(add = c(0.25, 0.25))  # widen margins for stratum labels
  ) +

  scale_y_continuous(
    name   = "Volume (t dry matter)",
    labels = comma_format(big.mark = ".", decimal.mark = ",")
  ) +

  coord_cartesian(clip = "off") +   # needed for below-axis annotations
  theme_minimal(base_size = 9) +
  theme(
    axis.title.x       = element_blank(),
    axis.text.x        = element_blank(),
    axis.ticks.x       = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    legend.position    = "none",
    legend.key.size    = unit(0.42, "cm"),
    plot.margin        = margin(t = 6, r = 6, b = 28, l = 6)
  )
```

```{r fig-results-product-mix, echo=FALSE, message=FALSE, warning=FALSE,     fig.cap="Total biomass delivered to consumers by product type over the planning horizon. Stacked areas show period-level volumes of stem (P1), branches (P2), and residues (P3) flowing from storage hubs to all consumers. Dashed line: total harvested raw biomass shipped from sites to hubs.", fig.width=6.5, fig.height=4, dev="pdf", out.width="\\textwidth", cache=T, cache.extra = c("ext","sites_clustered", "storages" , "customers", "milp_instance")}

prod_labels <- c("1" = "Stem (P1)", "2" = "Branches (P2)", "3" = "Residues (P3)")


# Deliveries aggregated by period × del_product
delivery_ts <- ext$Xjk %>%
  group_by(period, product = factor(del_product)) %>%
  summarise(volume = sum(value), .groups = "drop")

# Raw biomass shipped (Xij) aggregated by period — shown as dashed reference
shipped_ts <- ext$Xij %>%
  group_by(period) %>%
  summarise(shipped = sum(value), .groups = "drop")

ggplot(delivery_ts, aes(x = period, y = volume, fill = product)) +
  geom_area(position = "stack", alpha = 0.80, colour = "white", linewidth = 0.2) +
  geom_line(data = shipped_ts,
            aes(x = period, y = shipped),
            inherit.aes = FALSE,
            colour = "#636363", linetype = "dashed", linewidth = 0.7) +
  scale_fill_manual(
    name   = "Product",
    values = c("1" = "#1b9e77", "2" = "#d95f02", "3" = "#7570b3"),
    labels = prod_labels
  ) +
  scale_x_continuous(name = "Planning period (year)",
                     breaks = seq(1, milp_instance$n_periods, by = 2)) +
  scale_y_continuous(name = "Volume (t dry matter)",
                     labels = scales::comma) +
  annotate("text", x = max(shipped_ts$period) - 1,
           y = max(shipped_ts$shipped) * 0.95,
           label = "Raw shipped\n(site→hub)", size = 2.5,
           colour = "#636363", hjust = 1) +
  theme_minimal(base_size = 9) +
  theme(legend.position = "right",
        panel.grid.minor = element_blank())
```

```{r fig-results-harvest-cycle-dist, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of harvesting cycle lengths at active agroforestry sites. Each bar represents the number of individual harvest events (over all periods and active sites) whose cycle length (stand age at harvest, i.e. $t - s$) falls in that year bin. Vertical dashed line: mean cycle length. Active sites are defined as sites with at least one selected harvest arc in the optimal solution.", fig.width=5.5, fig.height=3.5, dev="pdf", out.width="\\textwidth", cache=T, cache.extra = c("ext", "milp_instance")}

# --- Extract harvest-arc cycle lengths from binary z variables ---------------
# ext$z contains one row per selected arc; columns: site_id, s, t, arc_type
# Harvest arcs have arc_type == "harvest" and cycle_length = t - s

cycle_df <- ext$z %>%
  filter(arc_type == "harvest") %>%
  mutate(cycle_length = t - s)

# Restrict to active sites (at least one harvest arc selected)
active_sites <- unique(cycle_df$site_id)

# Summary statistics for annotation
mean_cl  <- mean(cycle_df$cycle_length)
med_cl   <- median(cycle_df$cycle_length)
n_events <- nrow(cycle_df)
n_sites  <- length(active_sites)


# --- Plot --------------------------------------------------------------------
ggplot(cycle_df, aes(x = cycle_length)) +
  geom_bar(fill = "#1b9e77", colour = "white", linewidth = 0.25, width = 0.75) +
  geom_vline(xintercept = mean_cl,
             linetype = "dashed", colour = "grey30", linewidth = 0.6) +
  annotate("text",
           x     = mean_cl + 0.25,
           y     = Inf,
           label = paste0("Mean = ", round(mean_cl, 1), " yr"),
           hjust = 0, vjust = 1.5, size = 2.8, colour = "grey20") +
  scale_x_continuous(
    name   = "Harvesting cycle length (years)",
    breaks = seq(params_obj$min_age, params_obj$max_age, by = 1),
    limits = c(params_obj$min_age - 0.5, params_obj$max_age + 0.5)
  ) +
  scale_y_continuous(
    name   = "Number of harvest events",
    expand = expansion(mult = c(0, 0.12))
  ) +
  labs(
    caption = paste0(
      n_events, " harvest events across ", n_sites,
      " active sites; median = ", med_cl, " yr"
    )
  ) +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    plot.caption       = element_text(size = 7, colour = "grey40", hjust = 0)
  )
```

```{r fig-results-cost-profit-breakdown, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Total cost–profit breakdown across all active sites over the planning horizon. The left panel (waterfall) decomposes the objective value into revenue and individual cost categories; the right panel shows the period-level trajectory of revenues and stacked costs. Cultivation costs are split into establishment, harvesting, maintenance, and opportunity costs; logistics costs cover raw biomass transport (site$\\to$hub) and pre-processed transport (hub$\\to$consumer); storage costs reflect hub inventory holding charges.", fig.width = 7, fig.height = 4.5, dev = "pdf", out.width = "\\textwidth", cache = T, cache.extra = c("ext", "milp_instance", "sites_clustered", "storages", "params_obj")}

# ── 0. Colour palette (consistent with paper) ──────────────────────────────
COL_REV   <- "#1b9e77"   # revenue: green
COL_EST   <- "#d95f02"   # establishment
COL_HARV  <- "#7570b3"   # harvesting
COL_MAIN  <- "#e7298a"   # maintenance
COL_OPP   <- "#66a61e"   # opportunity
COL_TR_R  <- "#e6ab02"   # transport raw
COL_TR_P  <- "#a6761d"   # transport pre-processed
COL_STOR  <- "#666666"   # storage

# ── 1. Site meta: area_ha and cost parameters for active sites only ─────────
active_site_ids <- ext$z %>%
  filter(arc_type == "harvest") %>%
  pull(site_id) %>% unique()

site_meta <- milp_instance$sites %>%
  mutate(site_id = row_number()) %>%
  filter(site_id %in% active_site_ids)

# ── 2. Cultivation costs (derived from arc decisions in ext$z) ──────────────
cost_est <- ext$z %>%
  filter(arc_type == "establishment") %>%
  #select(-area_ha) %>%
  left_join(site_meta, by = "site_id") %>%
  mutate(cost = C_est * value, period = t) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Establishment")

cost_harv <- ext$z %>%
  filter(arc_type == "harvest") %>%
  #select(-area_ha) %>%
  left_join(site_meta, by = "site_id") %>%
  mutate(cost = C_harv * value, period = t) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Harvesting")

cost_main <- ext$z %>%
  filter(arc_type == "harvest") %>%
  #select(-area_ha) %>%
  mutate(age_len = t-s) %>% 
  left_join(site_meta, by = "site_id") %>%
  mutate(cost = C_main * value * age_len, period = t) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Maintenance")


cost_opp <- ext$z %>%
  filter(arc_type == "harvest") %>%
  #select(-area_ha) %>%
  left_join(site_meta, by = "site_id") %>%
  mutate(age_len = t-s) %>% 
  mutate(cost = C_opp * value * age_len, period = t) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Opportunity")


# ── 3. Transport costs ───────────────────────────────────────────────────────
dist_ij_m <- as.matrix(milp_instance$d_ij)

cost_tr_raw <- ext$Xij %>%
  mutate(
    dist_km = mapply(function(i, j) dist_ij_m[i, j], site_id, hub_id),
    cost    = params_obj$c_tr_raw * dist_km * value
  ) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Transport (raw)")

dist_jk_m <- as.matrix(milp_instance$d_jk)

cost_tr_pre <- ext$Xjk %>%
  mutate(
    dist_km = mapply(function(j, k) dist_jk_m[j, k], hub_id, consumer_id),
    cost    = params_obj$c_tr_pre * dist_km * value
  ) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Transport (pre-proc.)")

# ── 4. Storage costs ─────────────────────────────────────────────────────────
cost_stor <- ext$S %>%
  left_join(
    milp_instance$storages %>%
      mutate(hub_id = row_number()) %>%
      select(hub_id, c_stor),
    by = "hub_id"
  ) %>%
  mutate(cost = c_stor * value) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Storage")

# ── 5. Revenue per period ─────────────────────────────────────────────────────
price_df <- as.data.frame(milp_instance$consumer_prices) %>%
  setNames(c("consumer_id", "del_product", "price")) %>%
  mutate(consumer_id = as.integer(consumer_id),
         del_product  = as.integer(del_product))

revenue_ts <- ext$Xjk %>%
  left_join(price_df, by = c("consumer_id", "del_product")) %>%
  mutate(rev = value * coalesce(price, 0)) %>%
  group_by(period) %>%
  summarise(rev = sum(rev, na.rm = TRUE), .groups = "drop")

# ── 6. Assemble long-format time series ─────────────────────────────────────
all_periods <- tibble(period = seq_len(milp_instance$n_periods))

cost_ts <- bind_rows(
  cost_est, cost_harv, cost_main, cost_opp,
  cost_tr_raw, cost_tr_pre, cost_stor
) %>%
  right_join(crossing(period    = seq_len(milp_instance$n_periods),
                      component = unique(.$component)),
             by = c("period", "component")) %>%
  replace_na(list(cost = 0)) %>%
  mutate(component = factor(component, levels = c(
    "Establishment", "Harvesting", "Maintenance", "Opportunity",
    "Transport (raw)", "Transport (pre-proc.)", "Storage"
  )))

rev_ts <- all_periods %>%
  left_join(revenue_ts, by = "period") %>%
  replace_na(list(rev = 0))

# ── 7. Waterfall totals (left panel) ─────────────────────────────────────────
totals <- cost_ts %>%
  group_by(component) %>%
  summarise(value = -sum(cost), .groups = "drop") %>%
  bind_rows(tibble(component = "Revenue", value = sum(rev_ts$rev))) %>%
  mutate(component = factor(component, levels = c(
    "Revenue", "Establishment", "Harvesting", "Maintenance", "Opportunity",
    "Transport (raw)", "Transport (pre-proc.)", "Storage"
  ))) %>%
  arrange(component) %>%
  mutate(
    end      = cumsum(ifelse(component == "Revenue", value, value)),
    start    = lag(end, default = 0),
    bar_low  = ifelse(component == "Revenue", 0,     end),
    bar_high = ifelse(component == "Revenue", value, start)
  )

col_map <- c(
  "Revenue"               = COL_REV,
  "Establishment"         = COL_EST,
  "Harvesting"            = COL_HARV,
  "Maintenance"           = COL_MAIN,
  "Opportunity"           = COL_OPP,
  "Transport (raw)"       = COL_TR_R,
  "Transport (pre-proc.)" = COL_TR_P,
  "Storage"               = COL_STOR
)

p_waterfall <- ggplot(totals, aes(x = component, fill = component)) +
  geom_rect(aes(xmin = as.integer(component) - 0.38,
                xmax = as.integer(component) + 0.38,
                ymin = bar_low / 1e6, ymax = bar_high / 1e6),
            colour = "white", linewidth = 0.25) +
  geom_hline(yintercept = 0, colour = "grey40", linewidth = 0.35) +
  scale_fill_manual(values = col_map, guide = "none") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  scale_y_continuous(name   = "Cumulative value (M€)",
                     labels = scales::label_number(suffix = " M")) +
  labs(x = NULL, title = "Waterfall") +
  theme_minimal(base_size = 8) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor   = element_blank(),
        axis.text.x        = element_text(size = 6.5, angle = 30, hjust = 1),
        plot.title         = element_text(size = 8, face = "bold", hjust = 0.5))

# ── 8. Cumulative cash flow (right panel) ────────────────────────────────────
# Period net cash flow = revenue - total costs in that period
total_cost_ts <- cost_ts %>%
  group_by(period) %>%
  summarise(total_cost = sum(cost, na.rm = TRUE), .groups = "drop")

cashflow_ts <- all_periods %>%
  left_join(rev_ts,        by = "period") %>%
  left_join(total_cost_ts, by = "period") %>%
  replace_na(list(rev = 0, total_cost = 0)) %>%
  mutate(
    net_period  = rev - total_cost,
    cum_rev     = cumsum(rev),
    cum_cost    = cumsum(total_cost),
    cum_net     = cumsum(net_period)
  )

# Identify break-even period (first period where cumulative net turns positive)
breakeven_period <- cashflow_ts %>%
  filter(cum_net >= 0) %>%
  slice_min(period, n = 1) %>%
  pull(period)

p_cashflow <- ggplot(cashflow_ts, aes(x = period)) +
  # Shaded area between cumulative cost and cumulative revenue
  geom_ribbon(aes(ymin = pmin(cum_rev, cum_cost) / 1e6,
                  ymax = pmax(cum_rev, cum_cost) / 1e6,
                  fill = cum_net >= 0),
              alpha = 0.18, show.legend = FALSE) +
  scale_fill_manual(values = c("TRUE" = COL_REV, "FALSE" = COL_EST)) +
  # Cumulative revenue line
  geom_line(aes(y = cum_rev  / 1e6),
            colour = COL_REV, linewidth = 0.9, linetype = "solid") +
  # Cumulative cost line
  geom_line(aes(y = cum_cost / 1e6),
            colour = "grey30", linewidth = 0.9, linetype = "dashed") +
  # Cumulative net cash flow line
  geom_line(aes(y = cum_net  / 1e6),
            colour = "grey10", linewidth = 0.65, linetype = "dotted") +
  geom_hline(yintercept = 0, colour = "grey50", linewidth = 0.35) +
  # Break-even marker
  {if (length(breakeven_period) == 1)
    list(
      geom_vline(xintercept = breakeven_period,
                 colour = "grey40", linewidth = 0.45, linetype = "dotdash"),
      annotate("text",
               x = breakeven_period + 0.4, y = 0,
               label = paste0("BEP: t=", breakeven_period),
               size = 2.4, hjust = 0, vjust = -0.4,
               colour = "grey30", fontface = "italic")
    )
  } +
  # Line labels at right margin
  annotate("text",
           x     = milp_instance$n_periods - 0.3,
           y     = last(cashflow_ts$cum_rev)  / 1e6,
           label = "Cum. revenue", size = 2.4, hjust = 1, vjust = -0.4,
           colour = COL_REV, fontface = "italic") +
  annotate("text",
           x     = milp_instance$n_periods - 0.3,
           y     = last(cashflow_ts$cum_cost) / 1e6,
           label = "Cum. cost",    size = 2.4, hjust = 1, vjust =  1.3,
           colour = "grey30",      fontface = "italic") +
  annotate("text",
           x     = milp_instance$n_periods - 0.3,
           y     = last(cashflow_ts$cum_net)  / 1e6,
           label = "Cum. net CF",  size = 2.4, hjust = 1, vjust = -0.4,
           colour = "grey10",      fontface = "italic") +
  scale_x_continuous(name   = "Planning period (year)",
                     breaks = seq(1, milp_instance$n_periods, by = 2)) +
  scale_y_continuous(name   = "Cumulative value (M€)",
                     labels = scales::label_number(suffix = " M")) +
  labs(title = "Cumulative cash flow") +
  theme_minimal(base_size = 8) +
  theme(panel.grid.minor = element_blank(),
        plot.title       = element_text(size = 8, face = "bold", hjust = 0.5))

# ── 9. Assemble with patchwork ───────────────────────────────────────────────
p_waterfall + p_cashflow +
  plot_layout(widths = c(1, 1.8)) +
  plot_annotation(
    caption = paste0(
      "Active sites: ", length(active_site_ids),
      "  |  Total revenue: ",
      format(round(sum(rev_ts$rev)   / 1e6, 2), nsmall = 2), " M€",
      "  |  Net margin: ",
      format(round((sum(rev_ts$rev) - sum(cost_ts$cost)) / 1e6, 2),
             nsmall = 2), " M€",
      if (length(breakeven_period) == 1)
        paste0("  |  Break-even: period ", breakeven_period)
      else
        "  |  No break-even within horizon"
    ),
    theme = theme(plot.caption = element_text(size = 7, colour = "grey40"))
  )
```

```{r fig-results-site-breakdown, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Total revenue and cost breakdown per active agroforestry site over the full planning horizon. Each bar represents one active site; revenue (green, above zero axis) and cost components (stacked below) are expressed in thousand euros per hectare for comparability across sites of different size. Sites are ordered by descending net margin.", fig.width = 7, fig.height = 5, dev = "pdf", out.width = "\\textwidth", cache = T, cache.extra = c("ext", "milp_instance", "sites_clustered", "params_obj")}


# ── 0. Colour palette (reuse from previous chunk) ───────────────────────────
COL_REV  <- "#1b9e77"
COL_EST  <- "#d95f02"
COL_HARV <- "#7570b3"
COL_MAIN <- "#e7298a"
COL_OPP  <- "#aff315"
COL_TR_R <- "#e6ab02"
COL_TR_P <- "#a6761d"
COL_STOR <- "#666666"

col_map <- c(
  "Revenue"               = COL_REV,
  "Establishment"         = COL_EST,
  "Harvesting"            = COL_HARV,
  "Maintenance"           = COL_MAIN,
  "Opportunity"           = COL_OPP,
  "Transport (raw)"       = COL_TR_R,
  "Transport (pre-proc.)" = COL_TR_P,
  "Storage"               = COL_STOR
)

# ── 1. Site meta ─────────────────────────────────────────────────────────────
active_site_ids <- ext$z %>%
  filter(arc_type == "harvest") %>%
  pull(site_id) %>% unique()

site_meta <- milp_instance$sites %>%
  mutate(site_id = row_number()) %>%
  filter(site_id %in% active_site_ids)

# ── 2. Per-site cultivation costs ────────────────────────────────────────────
sc_est <- ext$z %>%
  filter(arc_type == "establishment") %>%
  left_join(site_meta, by = "site_id") %>%
  mutate(cost = C_est * value) %>%
  group_by(site_id) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Establishment")

sc_harv <- ext$z %>%
  filter(arc_type == "harvest") %>%
  left_join(site_meta, by = "site_id") %>%
  mutate(cost = C_harv * value) %>%
  group_by(site_id) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Harvesting")

sc_main <- ext$z %>%
  filter(arc_type == "harvest") %>%
  mutate(age_len = t-s) %>% 
  left_join(site_meta, by = "site_id") %>%
  mutate(cost = C_main * value * age_len) %>%
  group_by(site_id) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Maintenance")

sc_opp <- ext$z %>%
  filter(arc_type == "harvest") %>%
  mutate(age_len = t-s) %>% 
  left_join(site_meta, by = "site_id") %>%
  mutate(cost = C_opp * value * age_len) %>%
  group_by(site_id) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Opportunity")

 active_p <- ext$z %>% filter(arc_type == "establishment") %>% 
   group_by(site_id) %>% 
   summarise(tot_value = sum(value))
# 
 est_p  <- ext$z %>% filter(arc_type == "establishment") %>% group_by(site_id) %>% summarise(t_est = mean(sum(t*value)/sum(value)))
 term_p <- ext$z %>% filter(arc_type == "termination")  %>% group_by(site_id) %>% summarise(t_term = mean(sum(s*value)/sum(value)))
# 
# sc_opp <- est_p %>%
#   left_join(term_p, by = "site_id") %>%
#   left_join(active_p, by = "site_id") %>%
#   left_join(site_meta, by = "site_id") %>%
#   mutate(t_term = ifelse(is.na(t_term), milp_instance$n_periods, t_term)) %>%
#   mutate(cost = C_opp * tot_value * (t_term - t_est)) %>%
#   group_by(site_id) %>%
#   summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
#   mutate(component = "Opportunity")

# ── 3. Per-site transport costs ───────────────────────────────────────────────
dist_ij_m <- as.matrix(milp_instance$d_ij)
dist_jk_m <- as.matrix(milp_instance$d_jk)

sc_tr_raw <- ext$Xij %>%
  mutate(
    dist_km = mapply(function(i, j) dist_ij_m[i, j], site_id, hub_id),
    cost    = params_obj$c_tr_raw * dist_km * value
  ) %>%
  group_by(site_id) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Transport (raw)")

hub_site_share <- ext$Xij %>%
  group_by(hub_id, period) %>%
  mutate(hub_total = sum(value)) %>%
  ungroup() %>%
  mutate(share = ifelse(hub_total > 0, value / hub_total, 0)) %>%
  select(site_id, hub_id, period, share)

sc_tr_pre <- ext$Xjk %>%
  mutate(
    dist_km  = mapply(function(j, k) dist_jk_m[j, k], hub_id, consumer_id),
    cost_jkt = params_obj$c_tr_pre * dist_km * value
  ) %>%
  left_join(hub_site_share, by = c("hub_id", "period"),
            relationship = "many-to-many") %>%
  mutate(site_cost = cost_jkt * share) %>%
  group_by(site_id) %>%
  summarise(cost = sum(site_cost, na.rm = TRUE), .groups = "drop") %>%
  filter(!is.na(site_id)) %>%
  mutate(component = "Transport (pre-proc.)")

# ── 4. Storage costs: apportion to sites via Xij share ───────────────────────
hub_stor_cost <- ext$S %>%
  left_join(
    milp_instance$storages %>% mutate(hub_id = row_number()) %>% select(hub_id, c_stor),
    by = "hub_id"
  ) %>%
  mutate(stor_cost = c_stor * value) %>%
  select(hub_id, period, stor_cost)

sc_stor <- hub_stor_cost %>%
  left_join(hub_site_share, by = c("hub_id", "period"),
            relationship = "many-to-many") %>%
  mutate(site_cost = stor_cost * share) %>%
  group_by(site_id) %>%
  summarise(cost = sum(site_cost, na.rm = TRUE), .groups = "drop") %>%
  filter(!is.na(site_id)) %>%
  mutate(component = "Storage")

# ── 5. Per-site revenue ───────────────────────────────────────────────────────
price_df <- as.data.frame(milp_instance$consumer_prices) %>%
  setNames(c("consumer_id", "del_product", "price")) %>%
  mutate(consumer_id = as.integer(consumer_id),
         del_product  = as.integer(del_product))

site_revenue <- ext$Xjk %>%
  left_join(price_df, by = c("consumer_id", "del_product")) %>%
  mutate(rev_jkt = value * coalesce(price, 0)) %>%
  left_join(hub_site_share, by = c("hub_id", "period"),
            relationship = "many-to-many") %>%
  mutate(site_rev = rev_jkt * share) %>%
  group_by(site_id) %>%
  summarise(revenue = sum(site_rev, na.rm = TRUE), .groups = "drop") %>%
  filter(!is.na(site_id))

# ── 6. Active production years per site & normalise to € ha yr ──────────
# Active years = t_term - t_est; sites still running at horizon end use n_periods
site_active_years <- est_p %>%
  left_join(term_p, by = "site_id") %>%
  mutate(
    t_term       = ifelse(is.na(t_term), milp_instance$n_periods, t_term),
    active_years = pmax(t_term - t_est, 1L)   # guard against zero-length edge case
  ) %>%
  select(site_id, active_years)

# Normalisation denominator: area_ha * active_years  →  € ha yr
site_norm_denom <- active_p %>%
  left_join(site_active_years, by = "site_id") %>%
  mutate(denom = tot_value * active_years)

all_costs_site <- bind_rows(sc_est, sc_harv, sc_main, sc_opp,
                             sc_tr_raw, sc_tr_pre, sc_stor) %>%
  mutate(component = factor(component, levels = c(
    "Establishment", "Harvesting", "Maintenance", "Opportunity",
    "Transport (raw)", "Transport (pre-proc.)", "Storage"
  )))

cost_site_norm <- all_costs_site %>%
  left_join(site_norm_denom, by = "site_id") %>%
  mutate(cost_norm = cost / denom)          # € ha yr

rev_site_norm <- site_revenue %>%
  left_join(site_norm_denom, by = "site_id") %>%
  mutate(rev_norm = revenue / denom)        # € ha yr

# Site ordering by descending net margin (€ ha yr)
total_cost_site <- cost_site_norm %>%
  group_by(site_id) %>%
  summarise(total_cost_norm = sum(cost_norm, na.rm = TRUE), .groups = "drop")

site_order <- rev_site_norm %>%
  left_join(total_cost_site, by = "site_id") %>%
  mutate(net_margin = rev_norm - total_cost_norm) %>%
  arrange(desc(net_margin)) %>%
  pull(site_id)

cost_site_norm <- cost_site_norm %>%
  mutate(site_id = factor(site_id, levels = site_order))

rev_site_norm <- rev_site_norm %>%
  mutate(site_id = factor(site_id, levels = site_order))

total_cost_site <- total_cost_site %>%
  mutate(site_id = factor(site_id, levels = site_order))

# ── 7. Plot ───────────────────────────────────────────────────────────────────
ggplot() +
  geom_col(data = rev_site_norm,
           aes(x = site_id, y = rev_norm),
           fill = COL_REV, alpha = 0.88, width = 0.65) +
  geom_col(data = cost_site_norm,
           aes(x = site_id, y = -cost_norm, fill = component),
           position = "stack", width = 0.65, alpha = 0.88) +
  geom_point(
    data = rev_site_norm %>%
      left_join(total_cost_site, by = "site_id") %>%
      mutate(net = rev_norm - total_cost_norm),
    aes(x = site_id, y = net),
    shape = 18, size = 1.8, colour = "grey15"
  ) +
  geom_hline(yintercept = 0, colour = "grey40", linewidth = 0.4) +
  scale_fill_manual(
    name   = "Cost component",
    values = col_map[levels(cost_site_norm$component)],
    guide  = guide_legend(ncol = 1, keyheight = unit(0.38, "cm"),
                          override.aes = list(alpha = 1))
  ) +
  scale_y_continuous(
    name   = "Value (€ ha yr)",
    labels = scales::label_number(big.mark = ",")
  ) +
  scale_x_discrete(
    name   = "Active site (ordered by net margin)",
    labels = function(x) paste0("S", x)
  ) +
  annotate("text", x = 0.7, y = max(rev_site_norm$rev_norm) * 0.92,
           label = "Revenue", size = 2.4, hjust = 0,
           colour = COL_REV, fontface = "italic") +
  annotate("text", x = 0.7, y = -max(total_cost_site$total_cost_norm) * 0.92,
           label = "Costs", size = 2.4, hjust = 0,
           colour = "grey30", fontface = "italic") +
  labs(
    caption = paste0(
      length(active_site_ids), " active sites  |  = net margin  |  ",
      "normalised to € ha yr (active production years only)"
    )
  ) +
  theme_minimal(base_size = 8) +
  theme(
    axis.text.x        = element_text(size = 6, angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    legend.position    = "right",
    legend.text        = element_text(size = 6.5),
    legend.title       = element_text(size = 7, face = "bold"),
    plot.caption       = element_text(size = 7, colour = "grey40", hjust = 0)
  )
```

```{r fig-results-site-scatter, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Scatterplot of active agroforestry sites: opportunity cost rate (€ ha yr) on the x-axis versus demand-weighted average delivery distance (km) on the y-axis. Point size is proportional to the site's net margin (€ ha yr); sites with negative net margin are shown as open circles. A LOESS smoothing line indicates the joint trend.", fig.width = 5.5, fig.height = 4, dev = "pdf", out.width = "0.72\\textwidth", cache = T, cache.extra = c("ext", "milp_instance", "params_obj")}


# ── 1. Opportunity cost rate per site (€ ha yr) ─────────────────────────
# C_opp is already a per-ha-per-year rate stored in site_meta
opp_rate <- site_meta %>%
  select(site_id, opp_rate = C_opp, area_ha)   # unit: € ha yr

# ── 2. Demand-weighted average distance to customers per site ─────────────────
# Step 1: for each (site → hub) shipment, find the downstream (hub → consumer)
#         volume routed onward in the same period — use hub_site_share to
#         attribute the hub→consumer distance back to the originating site.
dist_jk_m <- as.matrix(milp_instance$d_jk)

# Demand-weighted mean distance over all periods:
#   d̄_i = Σ_{j,k,t}  share_{i,j,t} * X_{jk,t} * d_{jk}
#          ─────────────────────────────────────────────
#          Σ_{j,k,t}  share_{i,j,t} * X_{jk,t}
site_avg_dist <- ext$Xjk %>%
  mutate(dist_km = mapply(function(j, k) dist_jk_m[j, k], hub_id, consumer_id)) %>%
  left_join(hub_site_share, by = c("hub_id", "period"),
            relationship = "many-to-many") %>%
  filter(!is.na(site_id)) %>%
  mutate(
    weighted_dist = share * value * dist_km,
    weight        = share * value
  ) %>%
  group_by(site_id) %>%
  summarise(
    avg_dist_km = sum(weighted_dist, na.rm = TRUE) /
                  pmax(sum(weight,        na.rm = TRUE), 1e-9),
    .groups = "drop"
  )

# ── 3. Net margin per site (€ ha yr) — reuse from previous chunk ─────────
# (site_norm_denom, rev_site_norm, total_cost_site already computed above)
net_margin_site <- rev_site_norm %>%
  left_join(total_cost_site, by = "site_id") %>%
  mutate(net_margin = rev_norm - total_cost_norm) %>%
  select(site_id, net_margin)

# ── 4. Assemble scatter data ──────────────────────────────────────────────────
scatter_df <- opp_rate %>%
  left_join(site_avg_dist,   by = "site_id") %>%
  left_join(net_margin_site %>% mutate(site_id = as.integer(as.character(site_id))), by = "site_id") %>%
  mutate(
    profitable   = net_margin >= 0,
    # size scale: use absolute net margin; negatives shown as open circles
    abs_margin   = abs(net_margin)
  )

# ── 5. Plot ───────────────────────────────────────────────────────────────────
ggplot(scatter_df,
       aes(x = opp_rate, y = abs_margin)) +
  # LOESS trend (all sites)
  geom_smooth(method = "lm", formula = y ~ x,
              colour = "grey60", fill = "grey88",
              linewidth = 0.55, alpha = 0.35, se = TRUE) +
  # Profitable sites: filled circles
  geom_point(data = filter(scatter_df,  profitable),
             aes(size = area_ha),
             shape = 16, colour = "#1b9e77", alpha = 0.75) +
  # Loss-making sites: open circles
  geom_point(data = filter(scatter_df, !profitable),
             aes(size = area_ha),
             shape = 1, colour = "#d95f02", stroke = 0.6, alpha = 0.85) +
  # Size legend — both shapes share the same scale
   scale_size_area(
     name     = "site size (ha)",
     max_size = 8,
     breaks   = scales::breaks_extended(n = 4),
     labels   = scales::label_number(big.mark = ",")
   ) +
  scale_x_continuous(
    name   = "Opportunity cost rate (€ ha yr)",
    labels = scales::label_number(big.mark = ",")
  ) +
  scale_y_continuous(
    name   = "|Net margin| (€ ha yr)",
    labels = scales::label_number(accuracy = 1)
  ) +
  # Colour legend annotation (manual, avoids a redundant colour scale)
  annotate("point", x = -Inf, y = Inf, size = 0, alpha = 0) +  # spacer
  labs(
    caption = paste0(
      length(active_site_ids), " active sites  |  ",
      "point size ~ site size (ha)  |  ",
      "green filled = profitable, orange open = loss-making"
    )
  ) +
  theme_minimal(base_size = 8) +
  theme(
    panel.grid.minor  = element_blank(),
    legend.position   = "right",
    legend.text       = element_text(size = 6.5),
    legend.title      = element_text(size = 7, face = "bold"),
    plot.caption      = element_text(size = 7, colour = "grey40", hjust = 0)
  )
```

# Sensitivity analysis

To assess the robustness of the baseline results and to identify the
principal drivers of supply chain profitability, a factorial sensitivity
analysis is conducted. The analysis follows a full-factorial design over
three scenario dimensions: opportunity cost level
(\euro\ ha$^{-1}$ yr$^{-1}$), a cost multiplier applied to all variable
cost parameters (establishment, harvest, transport, and storage), and a
revenue multiplier applied to all gate prices. The resulting
$2 \times 2 \times 2 = 8$ scenarios are solved independently in
addition to the baseline, yielding nine comparable instances
(Table \@ref(tab:tab-scenario-kpis)).

Table \@ref(tab:tab-scenario-kpis) summarises the key performance
indicators across all scenarios. Total profit per hectare per year is
most sensitive to the revenue multiplier: elevated gate prices
(revenue multiplier = 1.5) substantially improve profitability, while
reduced prices (multiplier = 0.5) push marginal sites into loss
territory. The supply chain remains profitable in all scenarios with
full or elevated revenue even when opportunity costs are high. Scenarios
combining reduced revenue (multiplier = 0.5) with high opportunity costs
consistently yield negative net profits, indicating that the economic
viability of AFS-based supply chains is critically conditional on
achieving sufficient product-quality premiums from P1 offtakers.
The number of active sites and total wood throughput respond
monotonically to the revenue multiplier, confirming that higher gate
prices incentivise the activation of marginal sites.

The OLS regression reported in Figure \@ref(fig:fig-effects-ols) and
Table \@ref(tab:tab-lm-profit) decomposes profit variation across
scenarios and sites into effects of four explanatory variables:
opportunity cost rate, cost level, revenue level, and log-transformed
site area. The revenue multiplier carries the largest absolute
coefficient, followed by opportunity cost rate; site area exerts a
positive but diminishing marginal effect consistent with economies of
scale in logistics and harvesting. Rotation length is positively
associated with profit in low-opportunity-cost settings but negatively
so when foregone agricultural revenue is high, confirming the
rotation-length trade-off in the AFS-SCD formulation
(Section \@ref(sec:milp)).

```{r lm-site-profit-sensitivity, echo=FALSE, message=FALSE, warning=FALSE, cache=T}

df_lm <- bind_rows(df_bench, df_sens) %>%
  filter(!is.na(profit_ha_yr),
         !is.na(avg_dist_hub_km),
         !is.na(avg_dist_consumer_km))

# ── 1. Data preparation ───────────────────────────────────────────────────────
# Ensure grouping factors are proper factors
df_mod <- df_lm %>%
  mutate(
    site_id_f    = factor(site_id),
    scenario_f   = factor(scenario),
    # Centre continuous predictors for numerical stability and interpretability
    opp_cost_c   = scale(opp_cost_site,       center = TRUE, scale = FALSE)[,1],
    rot_c        = scale(avg_rotation_yr, center = TRUE, scale = FALSE)[,1],
    log_area     = log(area_ha),
    log_area_c   = scale(log(area_ha),   center = TRUE, scale = FALSE)[,1],
    avg_rotation_yr_c   = scale(avg_rotation_yr,   center = TRUE, scale = FALSE)[,1]
  )

# # ── 3. Lineares Modell ────────────────────────────────────────────────────────
# # Vollständiges Modell inkl. Interaktion opp_cost × cost_level
# lm_full <- lm(
#   profit_ha_yr ~
#     opp_cost_site +            # Opportunitätskosten (€/ha)
#     cost_log_level +      
#     cost_est_level + 
#     revenue_level +         # Preisniveau (Szenario-Multiplikator)
#     avg_rotation_yr +     # Ø Rotationslänge (Erntealter in Jahren)
#     log_area,             # Standortgröße (log-transformiert wegen Skaleneffekten)
#   data = df_mod
# )
# 
# 
# 
# sum.lm <- summary(lm_full)
# 
# tmp.lm.adj.r2 <- round(sum.lm$adj.r.squared*100,0)
# tmp.lm.nobs <- sum.lm$df[2]
# tmp.lm.sigma <- round(sum.lm$sigma,0)
# 
# # Fixed effects mirror lm_full but use centred predictors
# lmm_full <- lmer(
#   profit_ha_yr ~
#     opp_cost_site + 
#     cost_log_level +      
#     cost_est_level + 
#     revenue_level +
#     avg_rotation_yr +
#     log_area +
# #    (1 | site_id_f) +     # site-level random intercept
#     (1 | scenario_f),     # scenario-level random intercept
#   data    = df_mod,
#   REML    = TRUE,         # REML for variance component estimation
#   control = lmerControl(optimizer = "bobyqa",
#                         optCtrl   = list(maxfun = 2e5))
# )


# ── 4. Generalised Additive Model (GAM) ──────────────────────────────────────
# Smooth terms for:
#   s(opp_cost_c)   : nonlinear opportunity cost effect (expected threshold behaviour)
#   s(rot_c)        : nonlinear rotation-age effect (Gompertz upstream → nonlinear profit)
#   s(log_area_c)   : nonlinear scale economies
# Parametric terms retain scenario multipliers (discrete factor-like variables)
# bs="tp" (thin-plate regression splines) — default, scale-invariant
# k=5 chosen conservatively given the scenario grid resolution

gam_full <- gam(
  profit_ha_yr ~
    s(opp_cost_site,  bs = "tp", k = 5) +
    s(avg_rotation_yr,       bs = "tp", k = 5) +
    s(log_area,  bs = "tp", k = 5) +
    cost_log_level +      
    cost_est_level + 
    revenue_level ,       # parametric interaction (discrete multipliers)
  data   = df_mod,
  method = "REML",                    # REML smoothness selection
  family = gaussian()
)


# ── 6. Model Comparison ───────────────────────────────────────────────────────
# 
# # 6a. Information criteria table
# # Note: AIC for lmer uses ML (not REML) — refit with ML for fair comparison
# lmm_full_ml  <- refitML(lmm_full)
# #lmm_slope_ml <- refitML(lmm_slope)
# 
# aic_tab <- data.frame(
#   Model = c("OLS (baseline)", "LMM (random intercepts)",
#  "GAM"),
#   K     = c(
#     length(coef(lm_full)),
#     attr(logLik(lmm_full_ml),  "df"),
#     sum(gam_full$edf) + 1
#   ),
#   LogLik = c(
#     as.numeric(logLik(lm_full)),
#     as.numeric(logLik(lmm_full_ml)),
#     as.numeric(logLik(gam_full))
#     ),
#   AIC = c(
#     AIC(lm_full),
#     AIC(lmm_full_ml),
#     AIC(gam_full)
#   ),
#   BIC = c(
#     BIC(lm_full),
#     BIC(lmm_full_ml),       # approximate
#     BIC(gam_full)
#   )
# ) %>%
#   mutate(
#     delta_AIC = AIC - min(AIC),
#     AIC_weight = exp(-0.5 * delta_AIC) / sum(exp(-0.5 * delta_AIC)),
#     across(c(LogLik, AIC, BIC, delta_AIC), ~ round(.x, 2)),
#     AIC_weight = round(AIC_weight, 3),
#     K          = round(K, 1)
#   ) %>%
#   arrange(AIC)
# 
# # 6b. R² comparison
# r2_ols  <- summary(lm_full)$r.squared
# r2_lmm  <- MuMIn::r.squaredGLMM(lmm_full)   # returns marginal + conditional R²
# r2_gam  <- summary(gam_full)$r.sq
# 
# r2_tab <- data.frame(
#   Model         = c("OLS", "LMM", "GAM"),
#   R2_marginal   = c(r2_ols,
#                     r2_lmm[1, "R2m"],
#                     r2_gam),
#   R2_conditional = c(r2_ols,              # OLS has no random structure
#                      r2_lmm[1, "R2c"],
#                      NA_real_)
# ) %>%
#   mutate(across(where(is.numeric), ~ round(.x, 3)))

# 6c. Likelihood ratio tests for LMM random structure
#lrt_slope_vs_intercept <- anova(lmm_full_ml, lmm_slope_ml)

# 
# # ── 7. Output: AIC comparison table ──────────────────────────────────────────
# aic_tab %>%
#   kbl(
#     format    = "latex",
#     booktabs  = TRUE,
#     escape    = FALSE,
#     caption   = paste0(
#       "Model comparison by information criteria. $K$: effective degrees of freedom ",
#       "(edf for GAM/GAMM); $\\Delta$AIC: AIC difference relative to best model; ",
#       "AIC weight: Akaike weight. Models ranked by ascending AIC."
#     ),
#     col.names = c("Model", "$K$", "Log-Lik.", "AIC", "BIC",
#                   "$\\Delta$AIC", "AIC weight"),
#     label     = "tab-model-comparison",
#     align     = c("l", "r", "r", "r", "r", "r", "r")
#   ) %>%
#   kable_styling(
#     font_size     = 8,
#     latex_options = c("hold_position", "scale_down")
#   ) %>%
#   row_spec(1, bold = TRUE)   # highlight best model
# 
# 
# # ── 8. Output: R² table ───────────────────────────────────────────────────────
# r2_tab %>%
#   kbl(
#     format    = "latex",
#     booktabs  = TRUE,
#     escape    = FALSE,
#     caption   = paste0(
#       "$R^2$ comparison across models. For the LMM, marginal $R^2$ reflects ",
#       "variance explained by fixed effects only; conditional $R^2$ includes ",
#       "random effects. GAM/GAMM report deviance-based $R^2$."
#     ),
#     col.names = c("Model", "$R^2$ marginal", "$R^2$ conditional"),
#     label     = "tab-r2-comparison",
#     align     = c("l", "r", "r")
#   ) %>%
#   kable_styling(font_size = 8, latex_options = "hold_position")
# 

```

```{r fig-boxplot-scenario-profit, echo=FALSE, message=FALSE, warning=FALSE,    fig.cap="Verteilung des Standortprofits (€ ha yr) über die Ausprägungen der Szenarienparameter Opportunitätskosten (opp\\_cost), Kostenniveau (cost\\_level) und Preisniveau (revenue\\_level). Jeder Boxplot fasst alle aktiven Standorte aller Replikationen der jeweiligen Parameterkombination zusammen.",  fig.width = 7, fig.height = 5.5, dev = "pdf", out.width = "\\textwidth", cache = FALSE}

df_lm %>%
  mutate(
    opp_cost_lbl     = factor(
      paste0("Opp. cost: ", opp_cost, " €/ha"),
      levels = paste0("Opp. cost: ", sort(unique(opp_cost)), " €/ha")
    ),
    cost_level_lbl   = factor(
      paste0("Cost ×", cost_level),
      levels = paste0("Cost ×", sort(unique(cost_level)))
    ),
    revenue_level_lbl = factor(
      paste0("Rev. ×", revenue_level),
      levels = paste0("Rev. ×", sort(unique(revenue_level)))
    )
  ) %>%
  ggplot(aes(x = revenue_level_lbl, y = profit_ha_yr, fill = cost_level_lbl)) +
  geom_hline(yintercept = 0, colour = "grey50", linetype = "dashed", linewidth = 0.4) +
  geom_boxplot(
    outlier.size   = 0.8,
    outlier.alpha  = 0.5,
    linewidth      = 0.4,
    width          = 0.65,
    position       = position_dodge(width = 0.75)
  ) +
  facet_wrap(~ opp_cost_lbl, ncol = 2, scales = "fixed") +
  scale_fill_manual(
    name   = "Cost level",
    values = c("Cost ×0.5" = "#74c476", "Cost ×2"   = "#fd8d3c")
  ) +
  scale_y_continuous(
    name   = "Profit (€ ha yr)",
    labels = scales::label_number(big.mark = ",")
  ) +
  labs(x = "Revenue level") +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    strip.text         = element_text(face = "bold", size = 8),
    legend.position    = "bottom",
    legend.key.size    = unit(0.4, "cm"),
    legend.text        = element_text(size = 7),
    legend.title       = element_text(size = 7.5, face = "bold")
  )
```


```{r fig-boxplot-scenario-area, echo=FALSE, message=FALSE, warning=FALSE,    fig.cap="Verteilung der bewirtschafteten AFS-Fläche (ha) über die Ausprägungen der Szenarienparameter Opportunitätskosten (opp\\_cost), Kostenniveau (cost\\_level) und Preisniveau (revenue\\_level). Jeder Boxplot fasst die Summe der aktiven Standortflächen je Replikation und Parameterkombination zusammen. Höhere Werte zeigen eine breitere räumliche Aktivierung des AFS-Potenzials an.", fig.width = 7, fig.height = 5.5, dev = "pdf", out.width = "\\textwidth", cache = FALSE}

df_lm %>%
  # Aggregiere pro Szenario-Replikation: Gesamtfläche aller aktiven Standorte
  group_by(scenario, opp_cost, cost_level, revenue_level) %>%
  summarise(
    total_area_ha = sum(area_ha, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    opp_cost_lbl      = factor(
      paste0("Opp. cost: ", opp_cost, " €/ha"),
      levels = paste0("Opp. cost: ", sort(unique(opp_cost)), " €/ha")
    ),
    cost_level_lbl    = factor(
      paste0("Cost ×", cost_level),
      levels = paste0("Cost ×", sort(unique(cost_level)))
    ),
    revenue_level_lbl = factor(
      paste0("Rev. ×", revenue_level),
      levels = paste0("Rev. ×", sort(unique(revenue_level)))
    )
  ) %>%
  ggplot(aes(x = revenue_level_lbl, y = total_area_ha, fill = cost_level_lbl)) +
  geom_boxplot(
    outlier.size  = 0.8,
    outlier.alpha = 0.5,
    linewidth     = 0.4,
    width         = 0.65,
    position      = position_dodge(width = 0.75)
  ) +
  geom_jitter(
    aes(colour = cost_level_lbl),
    position      = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.75),
    size          = 1.0,
    alpha         = 0.55,
    show.legend   = FALSE
  ) +
  facet_wrap(~ opp_cost_lbl, ncol = 2, scales = "fixed") +
  scale_fill_manual(
    name   = "Cost level",
    values = c("Cost ×0.5" = "#74c476", "Cost ×2" = "#fd8d3c")
  ) +
  scale_colour_manual(
    values = c("Cost ×0.5" = "#41ab5d", "Cost ×2" = "#e6550d")
  ) +
  scale_y_continuous(
    name   = "Bewirtschaftete AFS-Fläche (ha)",
    labels = scales::label_number(big.mark = ",")
  ) +
  labs(x = "Revenue level") +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    strip.text         = element_text(face = "bold", size = 8),
    legend.position    = "bottom",
    legend.key.size    = unit(0.4, "cm"),
    legend.text        = element_text(size = 7),
    legend.title       = element_text(size = 7.5, face = "bold")
  )
```



```{r lm-table, echo=F, message=FALSE, warning=FALSE, cache=F, fig.cap="Effect plot of explaining variables of OLS regression model.", fig.width = 5.5, fig.height = 4, dev = "pdf", out.width = "0.72\\textwidth"}

# ── 4. Ergebnisausgabe ────────────────────────────────────────────────────────
tidy(lm_full, conf.int = TRUE) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
   # Sonderzeichen in Variablennamen für LaTeX entschärfen
  mutate(term = term %>%
    stringr::str_replace_all("_", "\\\\_") %>%          # Unterstriche escapen
    stringr::str_replace_all(":", " $\\\\times$ ") %>%  # Interaktion als ×
    stringr::str_replace_all("\\(", "\\\\(") %>%        # Klammern escapen
    stringr::str_replace_all("\\)", "\\\\)")
  ) %>%
  kbl(
    format    = "latex",
    booktabs  = TRUE,
    escape    = FALSE,
    caption   = paste("Overview on OLS regression result (adj. R$^2$:", tmp.lm.adj.r2, "\\%, \\# obervations: ", tmp.lm.nobs, ", $\\sigma$: ",tmp.lm.sigma, ")", sep=""),
    col.names = c("variable", "estimate", "std. error", "t value", "p value",
                  "CI 2.5\\%", "CI 97.5\\%"),
    label     = "tab-lm-profit"
  ) %>%
  kable_styling(font_size = 8,
                latex_options = c("hold_position", "scale_down"))
```

```{r lm-efffect-plot, echo=F, message=FALSE, warning=FALSE}
# ── 6. Koeffizientenplot ──────────────────────────────────────────────────────
tidy(lm_full, conf.int = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(term = forcats::fct_reorder(term, estimate)) %>%
  ggplot(aes(x = estimate, y = term,
             xmin = conf.low, xmax = conf.high,
             colour = p.value < 0.05)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_errorbarh(height = 0.25, linewidth = 0.5) +
  geom_point(size = 2) +
  scale_colour_manual(
   name   = "significant (p < 0.05)",
     values = c("TRUE" = "#1b9e77", "FALSE" = "#d95f02")
   ) +
  labs(x = "regression coeffizient (€ ha yr)",
       y = NULL) +
  theme_minimal(base_size = 9) +
  theme(legend.position  = "bottom",
        panel.grid.minor = element_blank())
```


# Conclusion

This paper has developed and applied the AFS-SCD model, a MILP
formulation for the integrated design and operation of agroforestry
biomass supply chains. The model endogenises stand age dynamics through
binary arc-based age-lag variables, couples component-specific Gompertz
yield coefficients for stem, branch, and residue fractions with
multi-echelon logistics decisions, and embeds a product quality cascade
that allocates biomass grades to chemical, pulp, and energy markets
within a unified optimisation framework. The formulation is, to the best
of our knowledge, the first supply chain model to simultaneously
optimise AFS establishment timing, consecutive harvest scheduling under
rotation-age constraints, and product-grade allocation across spatially
heterogeneous consumer markets.

The case study for the Saxony-Anhalt tri-state region demonstrates the
practical relevance of the approach. The optimal supply chain
concentrates AFS establishments in the Elbe and Saale floodplains,
where high biomass productivity and proximity to large pulp mills create
the most favourable economics. The sensitivity analysis confirms that
profitability depends primarily on achievable gate prices for
high-quality (P1) biomass grades: scenarios combining reduced revenue
with high opportunity costs yield negative net profits, whereas
scenarios with full or elevated prices remain viable under cost pressure.
Opportunity cost is the dominant site-level discriminator, with parcels
exceeding approximately 300 \euro\ ha$^{-1}$ yr$^{-1}$ rarely selected
in the optimal solution. These findings underscore the importance of
product cascading as a strategic lever for AFS viability: maximising
the share of stem biomass directed to high-value chemical and pulp
applications substantially improves profitability compared to
energy-only offtake strategies.

Methodologically, the AFS-SCD model bridges the gap between stand-level
growth modelling and regional supply chain optimisation identified in
the literature [@Zahraee2020_BSC_review; @Lamers2015_Feedstock_Design].
The age-lag constraint structure generalises to other perennial biomass
crops with periodic coppice cycles, such as willow or black locust, and
to long-rotation silvoarable systems in which rotation length is a
strategic variable. The model is extensible to stochastic settings by
replacing deterministic yield coefficients with scenario-weighted
counterparts or by incorporating robust optimisation over yield
uncertainty intervals [@Sharma2013_BSC_review].

Several limitations point to directions for future research. First,
consumer demand is treated as exogenous and time-invariant; endogenising
demand responses to biomass price changes would require a
partial-equilibrium extension. Second, the case study uses a single
allometric scenario calibrated to Central European poplar; propagating
Gompertz parameter uncertainty through to supply chain outcomes would
strengthen the robustness analysis. Third, environmental co-benefits of
AFS --- including carbon sequestration, nitrate leaching reduction, and
biodiversity enhancement --- are not monetised in the current objective.
Integrating life cycle assessment indicators as additional objectives
or constraints would support a more comprehensive sustainability
evaluation, consistent with the cleaner production framing of this work.


# Appendix

<!-- ```{r fig-results-inventory, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="End-of-period inventory levels at storage/pre-treatment hubs, disaggregated by product type. Each panel corresponds to one active hub.", fig.width=7, fig.height=3.5, dev="pdf", out.width="\\textwidth", cache=F, cache.extra = c("ext","sites_clustered", "storages" , "customers", "milp_instance")} -->

<!-- if( nrow(ext$S) > 0){ -->
<!-- inv_ts <- ext$S %>% -->
<!--   mutate(p = factor(p, labels = c("Stem", "Branches", "Residues")[seq_len(max(p))])) %>% -->
<!--   left_join( -->
<!--     storages %>% mutate(hub_id = row_number()) %>% select(hub_id, hub_name = name), -->
<!--     by = "hub_id" -->
<!--   ) %>% -->
<!--   filter(!is.na(value), value > 0) -->

<!-- ggplot(inv_ts, aes(x = period, y = value, fill = p)) + -->
<!--   geom_area(position = "stack", alpha = 0.78, colour = "white", linewidth = 0.2) + -->
<!--   facet_wrap(~hub_name, scales = "free_y") + -->
<!--   scale_fill_manual( -->
<!--     name   = "Product", -->
<!--     values = c("Stem" = "#1b9e77", "Branches" = "#d95f02", "Residues" = "#7570b3") -->
<!--   ) + -->
<!--   scale_x_continuous(breaks = seq(1, milp_instance$n_periods, by = 4)) + -->
<!--   scale_y_continuous(name = "Inventory (t)", labels = scales::comma) + -->
<!--   labs(x = "Planning period (year)") + -->
<!--   theme_minimal(base_size = 8) + -->
<!--   theme(legend.position = "bottom", -->
<!--         panel.grid.minor = element_blank(), -->
<!--         strip.text = element_text(face = "bold", size = 7))   -->
<!-- }else{ -->
<!--   NULL -->
<!-- } -->


<!-- ``` -->



```{r tab-scenario-kpis, echo=FALSE, message=FALSE, warning=FALSE}
# Helper to compute KPIs for one scenario result ------------------------------
compute_kpis <- function(res, scenario_name = NA_character_) {
  
  milp_instance <- res$milp_instance
  ext           <- res$ext
  #result        <- res$result
  obj.val       <- result$objective_value

  # [NEU] Szenario-Definitions-Parameter aus res$setting extrahieren ----------
  setting       <- res$setting
  scen_opp      <- if (!any(is.na(setting)))  setting$opp      else 150   # €/ha Opportunitätskosten-Basiswert
  scen_cost     <- if (!any(is.na(setting)))  setting$cost     else 1.0   # Kostenmultiplikator (1 = Basis)
  scen_revenue  <- if (!any(is.na(setting)))  setting$revenue  else 1.0   # Erlösmultiplikator  (1 = Basis)
  # ---------------------------------------------------------------------------

  # 1) Total profit over horizon (revenue – all costs)
  price_df <- as.data.frame(milp_instance$consumer_prices) %>%
    setNames(c("consumer_id", "del_product", "price")) %>%
    mutate(
      consumer_id = as.integer(consumer_id),
      del_product = as.integer(del_product)
    )

  rev_ts <- ext$Xjk %>%
    left_join(price_df, by = c("consumer_id", "del_product")) %>%
    mutate(rev = value * coalesce(price, 0)) %>%
    group_by(period) %>%
    summarise(rev = sum(rev, na.rm = TRUE), .groups = "drop")

  # Site meta & active sites
  active_site_ids <- ext$z %>%
    filter(arc_type == "harvest") %>%
    pull(site_id) %>% unique()

  site_meta <- milp_instance$sites %>%
    mutate(site_id = row_number()) %>%
    filter(site_id %in% active_site_ids)

  # Establishment, harvest, maintenance, opportunity costs
  cost_est <- ext$z %>%
    filter(arc_type == "establishment") %>%
    left_join(site_meta, by = "site_id") %>%
    mutate(cost = C_est * value, period = t) %>%
    group_by(period) %>%
    summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop")

  cost_harv <- ext$z %>%
    filter(arc_type == "harvest") %>%
    mutate(age_len = t - s) %>%
    left_join(site_meta, by = "site_id") %>%
    mutate(cost = C_harv * value, period = t) %>%
    group_by(period) %>%
    summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop")

  cost_main <- ext$z %>%
    filter(arc_type == "harvest") %>%
    mutate(age_len = t - s) %>%
    left_join(site_meta, by = "site_id") %>%
    mutate(cost = C_main * value * age_len, period = t) %>%
    group_by(period) %>%
    summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop")

  cost_opp <- ext$z %>%
    filter(arc_type == "harvest") %>%
    mutate(age_len = t - s) %>%
    left_join(site_meta, by = "site_id") %>%
    mutate(cost = C_opp * value * age_len, period = t) %>%
    group_by(period) %>%
    summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop")

  # Transport costs
  dist_ij_m <- as.matrix(milp_instance$d_ij)
  dist_jk_m <- as.matrix(milp_instance$d_jk)

  cost_tr_raw <- ext$Xij %>%
    mutate(
      dist_km = mapply(function(i, j) dist_ij_m[i, j], site_id, hub_id),
      cost    = milp_instance$c_tr_raw * dist_km * value
    ) %>%
    group_by(period) %>%
    summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop")

  cost_tr_pre <- ext$Xjk %>%
    mutate(
      dist_km = mapply(function(j, k) dist_jk_m[j, k], hub_id, consumer_id),
      cost    = milp_instance$c_tr_pre * dist_km * value
    ) %>%
    group_by(period) %>%
    summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop")

  # Storage cost
  cost_stor <- ext$S %>%
    left_join(
      milp_instance$storages %>%
        mutate(hub_id = row_number()) %>%
        select(hub_id, c_stor),
      by = "hub_id"
    ) %>%
    mutate(cost = c_stor * value) %>%
    group_by(period) %>%
    summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop")

  # Combine all cost components
  cost_ts <- bind_rows(
    cost_est    %>% mutate(component = "Establishment"),
    cost_harv   %>% mutate(component = "Harvesting"),
    cost_main   %>% mutate(component = "Maintenance"),
    cost_opp    %>% mutate(component = "Opportunity"),
    cost_tr_raw %>% mutate(component = "Transport (raw)"),
    cost_tr_pre %>% mutate(component = "Transport (pre-proc.)"),
    cost_stor   %>% mutate(component = "Storage")
  )

  total_rev    <- sum(rev_ts$rev,   na.rm = TRUE)
  total_cost   <- sum(cost_ts$cost, na.rm = TRUE)
  total_profit <- total_rev - total_cost

  # 2) Number of active sites
  n_active_sites    <- length(active_site_ids)
  area_active_sites <- sum(site_meta$area_ha)

  # 3) Total quantity of wood produced
  total_wood <- ext$Xij %>%
    group_by(p) %>%
    summarise(prod_total = sum(value, na.rm = TRUE), .groups = "drop") %>%
    rename(product = p)

  # 4) Share of total customer demand satisfied
  demand_df <- milp_instance$consumers %>%
    select(starts_with("demand")) %>%
    colSums() %>%
    as.data.frame() %>%
    rownames_to_column("product")

  colnames(demand_df)[2] <- "demand"

  demand_df <- demand_df %>%
    mutate(
      product = as.integer(as.character(factor(product, labels = 1:3))),
      demand  = demand * milp_instance$n_periods
    )

  total_delivered <- ext$Xjk %>%
    group_by(del_product) %>%
    summarise(total_consume = sum(value, na.rm = TRUE), .groups = "drop") %>%
    rename(product = del_product)

  dem_prod_df <- demand_df %>%
    left_join(total_wood,      by = "product") %>%
    left_join(total_delivered, by = "product") %>%
    mutate(
      share_demand = total_consume / demand,
      share_prod   = total_consume / prod_total
    )

  Share_demand_satisfied <- sum(dem_prod_df$total_consume) / sum(dem_prod_df$demand)

  tibble(
    Scenario                   = scenario_name,
    Opp_cost_EUR_ha            = scen_opp,       # Opportunitätskosten €/ha
    Cost_multiplier            = scen_cost,       # Kostenmultiplikator  (Basis = 1)
    Revenue_multiplier         = scen_revenue,    # Erlösmultiplikator   (Basis = 1)
    Rep_Total_profit_MEUR      = obj.val / 1e6,
    Total_profit_MEUR          = total_profit / 1e6,
    rel_profit                 = total_profit / area_active_sites / milp_instance$n_periods,
    Active_sites               = n_active_sites,
    Total_wood_units_kt        = sum(total_wood$prod_total) / milp_instance$n_periods / 1000,
    Share_demand_satisfied_P1  = dem_prod_df$share_demand[2],
    Share_demand_satisfied_P2  = dem_prod_df$share_demand[1],
    Share_demand_satisfied_P3  = dem_prod_df$share_demand[3],
    Share_demand_satisfied     = Share_demand_satisfied
  )
}

# ------------------------------------------------------------------
#  Compute KPIs for benchmark and all sensitivity scenarios
# ------------------------------------------------------------------
res_bench <- list(setting = NA, milp_instance = milp_instance, ext = ext, result = result)
kpi_bench <- compute_kpis(res_bench, scenario_name = "Bench.")

names(res.sens.list) <- 1:length(res.sens.list)

kpi_sens <- purrr::imap_dfr(
  res.sens.list,
  ~ compute_kpis(.x, scenario_name = .y)
)

kpi_all <- bind_rows(kpi_bench, kpi_sens) %>%
  group_by(Opp_cost_EUR_ha, Cost_multiplier, Revenue_multiplier) %>% 
  summarize(
    av_rel_profit = mean(rel_profit) , 
    av_active_sites = mean(Active_sites), 
    av_wood = mean(Total_wood_units_kt), 
    av_dem_sat = mean(Share_demand_satisfied)) %>% 
  mutate(
    av_rel_profit      = round(av_rel_profit, 0),
    av_active_sites      = round(av_active_sites, 0),
    av_wood        = round(av_wood, 0),
    Opp_cost_EUR_ha            = round(Opp_cost_EUR_ha, 0),
    Cost_multiplier            = round(Cost_multiplier, 2),
    Revenue_multiplier         = round(Revenue_multiplier, 2),
    av_dem_sat     = scales::percent(av_dem_sat,    accuracy = 0.1)
  )

kpi_all %>%
  knitr::kable(
    format   = "latex",
    booktabs = TRUE,
    align    = c("l", "r", "r", "r", "r", "r", "r", "r"),
    col.names = c(
      #"Scen.",
      "Opp. cost (€/ha)",   # [NEU]
      "Cost mult.",         # [NEU]
      "Rev. mult.",         # [NEU]
      "Profit (€/ha·a)",
      "Act. sites",
      "Wood (kt/a)",
      "Demand sat. (%)"
    ),
    caption = "Key performance indicators for benchmark and sensitivity scenarios.
               Opp.~cost: opportunity cost per hectare; Cost/Rev.~mult.: multipliers
               relative to the baseline scenario (1.0 = baseline).
               Total profit is net revenue minus all costs over the full planning
               horizon; demand satisfaction reports the share of total customer
               demand served."
  ) %>%
  kable_styling(font_size = 8, latex_options = c("striped", "scale_down"))
```




```{r lm-diagnostics, fig.cap="OLS-Diagnostik: Residuen vs. Fitted, Q-Q-Plot, Scale-Location und Cook's Distance.", fig.width=7, fig.height=5.5}

par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))
plot(lm_full, which = 1:4, cex = 0.5, col = "grey40")
```



<!-- ```{r lm-marginal-effects, fig.cap="Partielle Marginaleffekte der kontinuierlichen Prädiktoren auf den Standortprofit. Alle anderen Variablen werden auf ihrem Mittelwert gehalten. Grau: 95\\%-Konfidenzband.", fig.width=7, fig.height=5} -->

<!-- library(marginaleffects) -->

<!-- vars_continuous <- c("opp_cost", "cost_level", "revenue_level", -->
<!--                      "avg_dist_hub_km", "avg_dist_consumer_km", -->
<!--                      "avg_rotation_yr", "share_p1") -->

<!-- plots_list <- lapply(vars_continuous, function(v) { -->
<!--   plot_predictions(lm_full, condition = v) + -->
<!--     labs(x = v, y = "Profit (€ ha yr)") + -->
<!--     theme_minimal(base_size = 8) + -->
<!--     theme(panel.grid.minor = element_blank()) -->
<!-- }) -->

<!-- patchwork::wrap_plots(plots_list, ncol = 3) -->
<!-- ``` -->


<!-- ```{r lm-avplots, fig.cap="Added-Variable Plots (partielle Regressionen): isolierter Effekt jedes Prädiktors auf den Standortprofit, kontrolliert für alle anderen Variablen.", fig.width=7, fig.height=6} -->

<!-- library(car) -->
<!-- avPlots(lm_full, cex = 0.5, col = "grey40", -->
<!--         col.lines = "#1b9e77", lwd = 1.5) -->
<!-- ``` -->

