---
title: "Agroforestry SmartAgroforest: Poplar biomass modelling"
author:
  - name: "Thomas Kirschstein"
    affiliation: 
      - "Fraunhofer IKTS, Dresden, Germany"
    email: "thomas.kirschstein@ikts.fraunhofer.de"
    correspondingauthor: true
  
journal: "Journal Name"
# elsarticle options: preprint preview style, 1p or 3p for layout
classoption: "preprint,review,1p,authoryear"
date: "`r Sys.Date()`"
linenumbers: true
numbersections: true
always_allow_html: true


keywords: 
  - Agroforestry 
  - Short-rotation coppice 
  - Biomass supply chain
  - Product cascading

abstract: |
  Short-rotation agroforestry systems are increasingly promoted as a low-carbon feedstock option for bio-based value chains, but their economic viability depends on coordinated decisions on stand establishment, multi-cycle harvesting, and product cascading in spatially distributed supply chains. This paper develops a mixed-integer linear programming (MILP) model for the integrated design and operation of an agroforestry biomass supply chain with explicit age-lag constraints that link consecutive harvests at each site. The formulation couples allometric yield functions for different biomass types with a multi-product, multi-period network design model including storage, processing, and product quality cascades. The model is applied to a stylised case study of Central European poplar-based alley-cropping systems to explore trade-offs between harvest rotation length, product portfolios, and infrastructure configuration. We outline a computational experimentation plan to test the impact of allometric parameter uncertainty, price scenarios, and policy incentives on the optimal deployment of agroforestry systems in cleaner bioeconomy pathways.

output:
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
    keep_tex: true
    latex_engine: xelatex
    number_sections: true
    

header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{hyperref}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{float}
  - \usepackage{xcolor}
  - \usepackage{eurosym}
  - \newcommand{\Tc}{\mathcal{T}}
  - \newcommand{\Pc}{\mathcal{P}}
  - \newcommand{\Qc}{\mathcal{Q}}
  - \newcommand{\Sc}{\mathcal{S}}
  - \newcommand{\Ic}{\mathcal{I}}
  - \newcommand{\Jc}{\mathcal{J}}
  - \newcommand{\Kc}{\mathcal{K}}
  - \newcommand{\problem}{AFS-SCD }
  
bibliography: references.bib
csl: elsevier-harvard.csl
editor_options: 
  markdown: 
    wrap: 125
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo        = FALSE,
  message     = FALSE,
  warning     = FALSE,
  fig.align   = "center",
  fig.path    = "figures/",
  # --- additions ---
  cache       = FALSE,          # set TRUE per-chunk for heavy computations
  out.width   = "\\textwidth",  # consistent figure width for elsevier 1p
  dev         = "pdf"           # vector output for LaTeX/XeLaTeX pipeline
)

# Suppress noisy package startup messages globally
suppressPackageStartupMessages({
  library(magick)
  library(dplyr)
  library(plotly)
  library(ggplot2)
  library(tidyr)
  library(patchwork)
  library(knitr)
  library(kableExtra)
  library(sf)
  library(tibble)
  library(stringr)
  library(leaflet)
  library(osrm)
  library(ROI)
  library(ROI.plugin.glpk)
  library(gurobi)
  library(ROI.plugin.gurobi)
  library(ggrepel)
  library(ggnewscale)
  library(rnaturalearth)
  library(rnaturalearthdata)
  library(Rcpp)
  library(RcppArmadillo)
  library(scatterpie)
  library(networkD3)
  library(htmlwidgets)
  library(webshot2)
  library(ggalluvial)
  library(scales)
  library(purrr)
  library(broom)
  library(lme4)        # mixed-effects models
  library(lmerTest)    # Satterthwaite p-values for lmer
  library(mgcv)        # GAM + GAMM
  library(gratia)      # tidy GAM diagnostics & draw()
  library(MuMIn)       # R² for mixed models (r.squaredGLMM)
  library(AICcmodavg)  # AICc comparison table
  library(performance) # model_performance(), compare_performance(
  library(broom.mixed) # tidy() for lmer objects
  library(kableExtra)
})
options(kableExtra.latex.load_packages = FALSE)
options(knitr.table.format = "latex")

```

```{r load-model-data, echo=FALSE, message=FALSE, warning=FALSE, cache=T, cache.extra = tools::md5sum(c(     "../Modell/!helper_func.r", "../Modell/!helper_instance_builder_v8a.R","../Modell/!build_AFS_milp.r",      "../Modell/build_lp_rcpp_wrapper.R","../Modell/gurobi_adapter.R","../Modell/gurobi_warmstart.R",      "../Modell/gurobi_feascheck.R","../Modell/extract_result.R","../Modell/build_lp_rcpp.cpp","afs_workspace_red.RData",      "lp_r.RDS","lp_r_result.rds","lp_instance_data.rds" ))}
source("../Modell/!afs_biomass_setup.r")
source("../Modell/!helper_func.r")
source("../Modell/!helper_instance_builder_v8a.R")
source("../Modell/!build_AFS_milp.r")
source("../Modell/build_lp_rcpp_wrapper.R")
source("../Modell/gurobi_adapter.R")
source("../Modell/gurobi_warmstart.R")
source("../Modell/gurobi_feascheck.R")
source("../Modell/extract_result.R")
#Rcpp::sourceCpp("../Modell/build_lp_rcpp.cpp")
#Rcpp::sourceCpp("../Modell/build_and_solve_afs_milp.cpp")
Rcpp::sourceCpp("../Modell/build_and_solve_afs_lp_v11.cpp")


# ── 1. Hilfsfunktion: Site-Level KPIs für ein Szenario extrahieren ──────────
extract_site_profit <- function(res, scenario_name = NA_character_) {
  
  milp_instance <- res$milp_instance
  ext           <- res$ext
  setting       <- res$setting
  
  # Szenario-Parameter
  scen_opp      <- if (!any(is.na(setting))) setting$opp      else 150
  scen_cost_log <- if (!any(is.na(setting))) setting$cost.log else 1.0   
  scen_cost_est <- if (!any(is.na(setting))) setting$cost.est else 1.0   
  scen_revenue  <- if (!any(is.na(setting))) setting$revenue  else 1.0
  
  # no open sites
  if(nrow(ext$z) == 0){
    
    result_df <- data.frame( 
      scenario = scenario_name, 
      opp_cost = scen_opp,
      cost_log_level       = scen_cost_log,   
      cost_est_level       = scen_cost_est,   
      revenue_level        = scen_revenue,
       profit_ha_yr = NA,
      opp_cost = NA, 
      opp_cost_site = NA, 
      avg_dist_hub_km = NA, 
      avg_dist_consumer_km = NA,
      avg_rotation_yr = NA, 
      n_harvests = NA, 
      share_p1 = NA,
      area_ha =NA, 
      area_afs = NA, 
      active_years = NA, 
      n_sites = NA)
    
  }
  else{
    # Aktive Standorte
  active_site_ids <- ext$z %>%
    filter(arc_type == "harvest") %>%
    pull(site_id) %>% unique()
  
  site_meta <- milp_instance$sites %>%
    mutate(site_id = row_number()) %>%
    filter(site_id %in% active_site_ids)
  
  # Preistabelle
  price_df <- as.data.frame(milp_instance$consumer_prices) %>%
    setNames(c("consumer_id", "del_product", "price")) %>%
    mutate(consumer_id = as.integer(consumer_id),
           del_product  = as.integer(del_product))
  
  # ── Distanzmatrizen ────────────────────────────────────────────────────────
  dist_ij_m <- as.matrix(milp_instance$d_ij)
  dist_jk_m <- as.matrix(milp_instance$d_jk)
  
  # Hub-Anteile pro Standort und Periode (für Kostenzurechnung)
  hub_site_share <- ext$Xij %>%
    group_by(hub_id, period) %>%
    mutate(hub_total = sum(value)) %>%
    ungroup() %>%
    mutate(share = ifelse(hub_total > 0, value / hub_total, 0)) %>%
    select(site_id, hub_id, period, share)
  
  # ── Kosten je Standort ─────────────────────────────────────────────────────
  sc_est <- ext$z %>% filter(arc_type == "establishment") %>%
    left_join(site_meta, by = "site_id") %>%
    mutate(cost = C_est * value) %>%
    group_by(site_id) %>% summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
    mutate(component = "Establishment")
  
  sc_harv <- ext$z %>% filter(arc_type == "harvest") %>%
    left_join(site_meta, by = "site_id") %>%
    mutate(cost = C_harv * value) %>%
    group_by(site_id) %>% summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
    mutate(component = "Harvesting")
  
  sc_main <- ext$z %>% filter(arc_type == "harvest") %>%
    mutate(age_len = t - s) %>%
    left_join(site_meta, by = "site_id") %>%
    mutate(cost = C_main * value * age_len) %>%
    group_by(site_id) %>% summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
    mutate(component = "Maintenance")
  
  sc_opp <- ext$z %>% filter(arc_type == "harvest") %>%
    mutate(age_len = t - s) %>%
    left_join(site_meta, by = "site_id") %>%
    mutate(cost = C_opp * value * age_len) %>%
    group_by(site_id) %>% summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
    mutate(component = "Opportunity")
  
  sc_tr_raw <- ext$Xij %>%
    mutate(dist_km = mapply(function(i, j) dist_ij_m[i, j], site_id, hub_id),
           cost    = milp_instance$c_tr_raw * dist_km * value) %>%
    group_by(site_id) %>% summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
    mutate(component = "Transport_raw")
  
  sc_tr_pre <- ext$Xjk %>%
    mutate(dist_km  = mapply(function(j, k) dist_jk_m[j, k], hub_id, consumer_id),
           cost_jkt = milp_instance$c_tr_pre * dist_km * value) %>%
    left_join(hub_site_share, by = c("hub_id", "period"), relationship = "many-to-many") %>%
    mutate(site_cost = cost_jkt * share) %>%
    group_by(site_id) %>% summarise(cost = sum(site_cost, na.rm = TRUE), .groups = "drop") %>%
    filter(!is.na(site_id)) %>%
    mutate(component = "Transport_pre")
  
  sc_stor <- ext$S %>%
    left_join(milp_instance$storages %>% mutate(hub_id = row_number()) %>% select(hub_id, c_stor),
              by = "hub_id") %>%
    mutate(stor_cost = c_stor * value) %>%
    select(hub_id, period, stor_cost) %>%
    left_join(hub_site_share, by = c("hub_id", "period"), relationship = "many-to-many") %>%
    mutate(site_cost = stor_cost * share) %>%
    group_by(site_id) %>% summarise(cost = sum(site_cost, na.rm = TRUE), .groups = "drop") %>%
    filter(!is.na(site_id)) %>%
    mutate(component = "Storage")
  
  # ── Erlöse je Standort ─────────────────────────────────────────────────────
  site_revenue <- ext$Xjk %>%
    left_join(price_df, by = c("consumer_id", "del_product")) %>%
    mutate(rev_jkt = value * coalesce(price, 0)) %>%
    left_join(hub_site_share, by = c("hub_id", "period"), relationship = "many-to-many") %>%
    mutate(site_rev = rev_jkt * share) %>%
    group_by(site_id) %>%
    summarise(revenue = sum(site_rev, na.rm = TRUE), .groups = "drop") %>%
    filter(!is.na(site_id))
  
  # ── Normierung: €/ha/a (nur aktive Produktionsjahre) ──────────────────────
  #est_p  <- ext$z %>% filter(arc_type == "establishment") %>% select(site_id, t_est = t)
  #term_p <- ext$z %>% filter(arc_type == "termination")  %>% select(site_id, t_term = s)
  active_p <- ext$z %>% filter(arc_type == "establishment") %>% 
    group_by(site_id) %>% 
    summarise(tot_value = sum(value))
# 
  est_p  <- ext$z %>% filter(arc_type == "establishment") %>% group_by(site_id) %>% summarise(t_est = mean(sum(t*value)/sum(value)))
  term_p <- ext$z %>% filter(arc_type == "termination")  %>% group_by(site_id) %>% summarise(t_term = mean(sum(s*value)/sum(value)))
 
  site_norm <- site_meta %>%
    #select(site_id, area_ha) %>%
    left_join(active_p,  by = "site_id") %>%
    left_join(est_p,  by = "site_id") %>%
    left_join(term_p, by = "site_id") %>%
    mutate(t_term       = ifelse(is.na(t_term), milp_instance$n_periods, t_term),
           active_years = pmax(t_term - t_est, 1L),
           denom        = tot_value * active_years)
  
  all_costs <- bind_rows(sc_est, sc_harv, sc_main, sc_opp,
                         sc_tr_raw, sc_tr_pre, sc_stor) %>%
    group_by(site_id) %>%
    summarise(total_cost = sum(cost, na.rm = TRUE), .groups = "drop")
  
  # ── Durchschnittliche Distanzen je Standort ────────────────────────────────
  # (a) Ø Distanz Standort → Hub (gewichtet nach Liefervolumen)
  avg_dist_to_hub <- ext$Xij %>%
    mutate(dist_km = mapply(function(i, j) dist_ij_m[i, j], site_id, hub_id)) %>%
    group_by(site_id) %>%
    summarise(avg_dist_hub_km = weighted.mean(dist_km, w = pmax(value, 0), na.rm = TRUE),
              .groups = "drop")
  
  # (b) Ø Distanz Hub → Kunden (anteilig auf Standort zurückgerechnet)
  avg_dist_to_consumer <- ext$Xjk %>%
    mutate(dist_km = mapply(function(j, k) dist_jk_m[j, k], hub_id, consumer_id)) %>%
    left_join(hub_site_share, by = c("hub_id", "period"), relationship = "many-to-many") %>%
    filter(!is.na(site_id), share > 0) %>%
    group_by(site_id) %>%
    summarise(avg_dist_consumer_km = weighted.mean(dist_km, w = pmax(share * value, 0), na.rm = TRUE),
              .groups = "drop")
  
  # ── Weitere strukturelle Variablen ─────────────────────────────────────────
  # (c) Ø Erntezyklus-Länge (Rotationsalter) je Standort
  avg_rotation <- ext$z %>%
    filter(arc_type == "harvest") %>%
    mutate(cycle_len = t - s) %>%
    group_by(site_id) %>%
    summarise(avg_rotation_yr = sum(cycle_len*value/sum(value), na.rm = TRUE),
              n_harvests       = n(),
              .groups = "drop")
  
  # (d) Produktmix-Index: Anteil hochwertiger Biomasse (P1 Stämme) an Gesamtlieferung
  p1_share <- ext$Xij %>%
    group_by(site_id) %>%
    summarise(
      vol_total = sum(value, na.rm = TRUE),
      vol_p1    = sum(value[p == 1], na.rm = TRUE),
      .groups   = "drop"
    ) %>%
    mutate(share_p1 = ifelse(vol_total > 0, vol_p1 / vol_total, NA_real_)) %>%
    select(site_id, share_p1, vol_total)
  
  # ── Zusammenführen ─────────────────────────────────────────────────────────
  result_df <- site_norm %>%
    left_join(site_revenue,          by = "site_id") %>%
    left_join(all_costs,             by = "site_id") %>%
    left_join(avg_dist_to_hub,       by = "site_id") %>%
    left_join(avg_dist_to_consumer,  by = "site_id") %>%
    left_join(avg_rotation,          by = "site_id") %>%
    left_join(p1_share,              by = "site_id") %>%
    mutate(
      profit_ha_yr         = (coalesce(revenue, 0) - coalesce(total_cost, 0)) / denom,
      scenario             = scenario_name,
      opp_cost             = scen_opp,
      cost_log_level       = scen_cost_log,   # renamed: was cost_level
      cost_est_level       = scen_cost_est,   # new
      revenue_level        = scen_revenue,
      opp_cost_site        = C_opp,
      area_afs             = tot_value
    ) %>%
    select(site_id, scenario, profit_ha_yr,
           opp_cost, opp_cost_site, cost_log_level, cost_est_level, revenue_level,
           avg_dist_hub_km, avg_dist_consumer_km,
           avg_rotation_yr, n_harvests, share_p1,
           area_ha, area_afs, active_years, n_sites)
    
  }
  
  
  
  return(result_df)
}


#load("../Modell/afs_workspace.RData")
load("afs_workspace_red.RData")

sites_sf   <- afs_workspace$sites_sf
sites      <- afs_workspace$sites
sites_clustered      <- afs_workspace$sites_clustered
storages   <- afs_workspace$storages
consumers  <- afs_workspace$consumers

# scale down industrial demand quantities by dividing by 5
consumers <- consumers %>%
  mutate(demand_P1 = demand_P1 / 5)

# scale down prices to 40-70 €/t chemical, 30-40 €/t pulp + 15-20 €/t
consumers <- consumers %>% 
  mutate(P1 = P1/2, P2 = P2/2, P3 = P3*.6 )

# drop small Biogas plants --> compute total demand and drop all consumer with D<2
consumers <- consumers %>%
  mutate(total_demand = demand_P1 + demand_P2 + demand_P3) %>%
  filter(total_demand >= 1.5) %>%
  select(-total_demand) 


dist_jk <- afs_workspace$dist_jk[,consumers$consumer_id]

yields_by_age <- build_scenario_ts(ages = 1:20 , N = 250, C_site = 6.500, k = 0.194, t0 = 9.7, label = "Conservative (250 trees/ha)")


yields_by_age <- yields_by_age %>% 
  # drop column scenario from yields_by_age
  select(-scenario) %>% 
  # make long --> bring columns Merch. stem Merch. branch Residue to column "product" and values to "yield_ha"
  pivot_longer(-1 , names_to = "product", values_to = "yield_ha") %>% 
  # rename entries in "product" to 1:3
  mutate(product = as.integer(as.character(factor(product, labels = c(2,1,3) ))))

# scale yields to fresh biomass
yields_by_age$yield_ha <- yields_by_age$yield_ha / (1-0.5)


set.seed(123578)
# generate random vector of opportunity cost for clustered sites in range [0,600]
c.opp.vec <- rnorm(nrow(sites_clustered), 1, .3)


sites_clustered <- sites_clustered %>%
   mutate(
     C_est  = 2500,   # €/ha  establishment (Section 2.1 mid-range)
     C_harv =  150,    # €/ha  harvest cost  (Section 2.1 mid-range)
     C_main = 10 , #  €/ha  maintenance cost  (Section 2.1 mid-range)
     C_opp = 500*c.opp.vec, #  €/ha  opportunity cost  (Section 2.1 mid-range)
   ) #%>%   # change column name area_ha to area if necessary  rename(area_ha = area)


# # change storage costs
# storages$c_stor <- storages$c_stor
# # change capacities
 storages$CAP_stor <- storages$CAP_stor * 100000
 storages$CAP_proc <- storages$CAP_proc * 100000
# 
data_obj <- list(
   sites    = sites_clustered,
  #sites    = sites_filt,
   storages = storages,
   consumers = consumers,
   dist_ij  = afs_workspace$dist_ij_clust,
   #dist_ij  = dist_ij_filt$distance_matrix_km,
   dist_jk  = dist_jk,
   yields_by_age = yields_by_age
 )
# rescale demands in consumers to tons
data_obj$consumers[,c("demand_P1","demand_P2","demand_P3")]  <- data_obj$consumers[,c("demand_P1","demand_P2","demand_P3")] * 1000

# 
# # ── 3. PARAMS OBJECT — all scalar parameters ──────────────────────────────────
params_obj <- list(
   n_periods  = 40,
   max_age    = 20L,    # A_max: LRAFS upper rotation bound
   min_age    =  3L,    # A_min: SRAFS lower rotation bound
   c_tr_raw   = 0.08,   # €/(t·km)  raw biomass  site → storage
   c_tr_pre   = 0.06   # €/(t·km)  chips        storage → consumer
 )
# 
# ── 4. generate instance ─────────────────────────────────────
milp_instance <- build_optimization_instance(
    data         = data_obj,
    params       = params_obj
  ) 

#result <- build_and_solve_afs_lp_v11(milp_instance, 
#                            gurobi_params = list(TimeLimit = 300),
#                            verbose = TRUE)
 
#saveRDS(result, "result_lp_v11.rds")

result <- readRDS("result_lp_v11.rds")

ext <- result$solution_vector

#  renaming

ext$Xjk <- ext$Xjk %>% 
  rename(src_product = p, del_product = q, period = t, hub_id = j, consumer_id = k)

ext$Xij <- ext$Xij %>% 
  rename(period = t, site_id = i, hub_id = j)

ext$z <- ext$z %>% 
  rename(arc_type = type, site_id = i )

ext$S <- ext$S %>% 
  rename(hub_id = j, period = t)

# sensitivity analysis ##########################################
#
# scen.mat <- expand.grid(opp = c(150, 300, 600), cost.log = c(.5, 1.5, 2), cost.est = c(.5, 1.5, 2), revenue = c(.5, .75, 1.25)  )
# nb.rep <- 10
# 
# res.sens.list <- vector("list", nrow(scen.mat)*nb.rep)
# 
# for(i in seq_len(nrow(scen.mat))){
#   tmp.setting <- list(
#     opp = scen.mat$opp[i],
#     cost.log = scen.mat$cost.log[i],
#     cost.est = scen.mat$cost.est[i],
#     revenue = scen.mat$revenue[i]
#     )
# 
#   # storage cost
#   tmp.storages <- storages
#   tmp.storages$c_stor <- tmp.storages$c_stor * tmp.setting$cost.log
#   # consumers
#   tmp.consumers <- consumers
#   tmp.consumers$P1 <- tmp.consumers$P1 * tmp.setting$revenue
#   tmp.consumers$P2 <- tmp.consumers$P2 * tmp.setting$revenue
#   tmp.consumers$P3 <- tmp.consumers$P3 * tmp.setting$revenue
# 
#   # # ── 3. PARAMS OBJECT — all scalar parameters ──────────────────────────────────
#   tmp.params_obj <- list(
#        n_periods  = 40,
#        max_age    = 20L,    # A_max: LRAFS upper rotation bound
#        min_age    =  3L,    # A_min: SRAFS lower rotation bound
#        c_tr_raw   = 0.08 * tmp.setting$cost.log,   # €/(t·km)  raw biomass  site → storage
#        c_tr_pre   = 0.06 * tmp.setting$cost.log   # €/(t·km)  chips        storage → consumer
#      )
#     #
# 
#   for(j in seq_len(nb.rep)){
#     tmp.c.opp.vec <- rnorm(nrow(sites_clustered), 1, .3)
# 
#     # opportunity cost
#     tmp.c.opp.vec <- tmp.c.opp.vec * tmp.setting$opp
# 
#     # site details
#     tmp.sites_clustered <- sites_clustered %>%
#        mutate(
#          C_est  = 2500 * tmp.setting$cost.est,   # €/ha  establishment
#          C_harv = 150 * tmp.setting$cost.log,    # €/ha  harvest cost
#          C_main = 10 * tmp.setting$opp/150, #  €/ha  maintenance cost
#          C_opp  = tmp.c.opp.vec #  €/ha  opportunity cost
#        )
#     #
#     tmp.data_obj <- list(
#        sites    = tmp.sites_clustered,
#        storages = tmp.storages,
#        consumers = tmp.consumers,
#        dist_ij  = afs_workspace$dist_ij_clust,
#        dist_jk  = dist_jk,
#        yields_by_age = yields_by_age
#      )
# 
#     # rescale demands in consumers to tons
#     tmp.data_obj$consumers[,c("demand_P1","demand_P2","demand_P3")]  <- tmp.data_obj$consumers[,c("demand_P1","demand_P2","demand_P3")] * 1000
# 
#     # ── 4. generate instance ─────────────────────────────────────
#     tmp.milp_instance <- build_optimization_instance(
#         data         = tmp.data_obj,
#         params       = tmp.params_obj
#       )
#     #
#     tmp.result <- build_and_solve_afs_lp_v11(
#        tmp.milp_instance,
#        gurobi_params = list(
#          TimeLimit = 600),
#        verbose = TRUE)
#     #
# 
#     tmp.ext <- tmp.result$solution_vector
#     #  renaming
# 
#     tmp.ext$Xjk <- tmp.ext$Xjk %>%
#       rename(src_product = p, del_product = q, period = t, hub_id = j, consumer_id = k)
# 
#     tmp.ext$Xij <- tmp.ext$Xij %>%
#       rename(period = t, site_id = i, hub_id = j)
# 
#     tmp.ext$z <- tmp.ext$z %>%
#       rename(arc_type = type, site_id = i )
# 
#     tmp.ext$S <- tmp.ext$S %>%
#       rename(hub_id = j, period = t)
# 
#     res.sens.list[[(i-1)*nb.rep + j]] <- list(setting = tmp.setting, milp_instance = tmp.milp_instance, ext = tmp.ext)
#   }
# 
# }
# 
# saveRDS(res.sens.list, "res_sens_list_lp_v11.rds")

res.sens.list <- readRDS("res_sens_list_lp_v11.rds")

# ── 2. Benchmark + alle Sensitivitätsszenarien zusammenführen ────────────────
res_bench_full <- list(
  setting       = list(opp = 500, cost.log = 1.0, cost.est = 1.0, revenue = 1.0),
  milp_instance = milp_instance,
  ext           = ext,
  result        = result
)

df_bench <- extract_site_profit(res_bench_full, "Bench.")

df_bench <- df_bench %>% mutate(opp_cost = opp_cost/500)

df_sens <- purrr::imap_dfr(
  res.sens.list,
  ~ extract_site_profit(.x, scenario_name = paste0("S", .y))
)

df_sens <- df_sens %>% mutate(opp_cost = opp_cost/500)

```

# Introduction

Agroforestry systems that integrate trees with agricultural crops
(silvoarable systems) are increasingly recognised as a promising
land-use option to reconcile climate mitigation, biodiversity
conservation, and rural income generation on the same hectare of land
[@Toensmeier2017_Carbon_Agroforestry]. Temperate agroforestry systems
(AFS) based on fast-growing species such as poplar, willow, alder, and
black locust can deliver substantial biomass yields while improving soil
quality and providing multiple ecosystem services compared to
conventional monocultures [@Huber2018_SRAFS_growth]. Biomass produced
in AFS is considered as a sustainable feedstock and can be used either
energetically, chemically or materially (e.g. in the building sector).
The latter two options are generally preferred over the energetic usage
from a cleaner production perspective as biogenic carbon is stored in
products during their life span which contributes to climate change
mitigation.

Despite substantial research on ecological advantages and pioneering
initiatives to establish AFS in practice, the share of AFS in
agricultural production systems is still small. E.g., in Germany about
1200 hectares of silvoarable AFS are established in 2025, which accounts
for about 0.02% of total agricultural land use [@DEFAF]. One of the core
challenges in establishing AFS systems at large scale in practice is a
lack of opportunities to generate substantial revenues from woody
biomasses of AFS. On the other hand, the rise of bioeconomy creates
substantial demands for woody biomass and requires a reliable and
cost-efficient feedstock supply. As woody biomasses are bulky feedstocks
with low energy density, the design of cost-efficient supply chains is
crucial to ensure the economic viability of AFS-based value chains. This
requires coordinated decisions on stand establishment, multi-cycle
harvesting, and product cascading in spatially distributed supply
chains, which can be supported by integrated optimisation models that
link stand-level growth dynamics to regional supply chain design.

In the last two decades, a large body of research has analysed biomass
supply chains for bioenergy and biorefineries using mathematical
programming models, often in the form of mixed-integer linear
programming (MILP) [@Zahraee2020_BSC_review; @Sharma2013_BSC_review].
These models typically optimise network configuration, feedstock
sourcing, and logistics under cost or profit objectives. They often
treat biomass availability as an exogenous input or rely on simplified
agricultural production decisions. Moreover, downstream processing of
biomasses is concentrated on one particular value chain e.g. bioenergy
plants consuming all produced biomasses. Recent work has started to
integrate operational details such as various biomass types, biomass
growth models and regeneration processes in multi-period supply chain
models.

The present paper contributes to this literature in three ways. First, we
propose a novel MILP formulation for AFS-based biomass supply chain
design that explicitly links establishment decisions and consecutive
harvest decisions, ensuring that harvesting decisions at each site are
consistent with minimum and maximum rotation ages. Second, we embed
allometric yield functions for different biomass types generating
age-specific yields for different woody biomass types (like stems,
branches, or leaves and twigs). Third, these different woody biomass
types can be processed in specific downstream industries, like chemical
or paper industry as well as energy sector. The model aims to provide
strategic decision support for establishing regional AFS-based
bioeconomy ecosystems connecting multiple industries and product
markets.

The remainder of the paper is structured as follows. Section
\@ref(sec:litreview) reviews the relevant literature on agroforestry
biomass systems, allometric biomass modelling, and biomass supply chain
optimisation. Section \@ref(sec:milp) summarises the MILP formulation
for the agroforestry supply chain with age-lag constraints. Section
\@ref(sec-casestudy) presents the case study instantiation, the base-case optimisation results,
and the sensitivity analysis, examining the implications for cleaner production strategies.

# Literature review {#sec:litreview}

Agroforestry biomass systems, stand-level growth models, and biomass
supply chain optimisation have each been studied extensively, but they
are rarely combined in an integrated framework that links age-dependent
stand dynamics to regional product-cascading value chains
[@Zahraee2020_BSC_review; @Sharma2013_BSC_review;
@Espinoza2017_Biorefinery_SC_review]. This section reviews (i)
agroforestry systems with emphasis on system types, economics, and the
cascading use of biomass as the primary revenue driver; (ii) biomass
growth models at tree and stand level with explicit attention to the
allometric--diameter--stand-biomass pathway; and (iii) biomass supply
chain design using mathematical programming. The section concludes by
identifying the modelling gap addressed in this paper.

## Agroforestry systems {#sec:lit-afs}

Agroforestry deliberately integrates woody perennials with crops and/or
livestock on the same management unit, with the aim of enhancing
productivity, profitability, and environmental performance
[@Nair1993_IntroAgroforestry; @Toensmeier2017_Carbon_Agroforestry]. In
the temperate zone, three distinct system types can be distinguished
based on the spatial arrangement of trees and the rotation length at
which woody biomass is harvested.

**Short-rotation coppice (SRC)** is defined as a monoculture plantation
of fast-growing woody species covering the entire management unit, with
planting densities of 8,000--20,000 cuttings\~ha$^{-1}$. After each
harvest the root system regenerates without replanting. Rotation cycles
are typically 3--7\~years and total stand lifetime is 20--25\~years
[@Trnka2008_Poplar_SRF; @Faasch2012_SRC_Economics]. SRCs are no AFSs in
the narrow sense, but it is helpful to benchmark biomass production.

**Short-rotation agroforestry systems (SRAFS)** integrate SRC-type tree
strips (typically 1--4 rows of the same fast-growing species like poplar
or willow) into an otherwise arable field in an alley-cropping
arrangement. Crop production continues in the inter-row strips
[@Huber2018_SRAFS_growth; @Graves2007_Agroforestry_Economics]. Thus,
tree strips typically cover 10--30% of the total parcel area. Rotation
lengths and total stand lifetime in SRAFS are comparable to SRCs. In
SRAFS all trees are effectively edge trees, exposed to higher light and
water availability than interior SRC trees. This boundary effect raises
individual-tree biomass accumulation but also intensifies competition
with adjacent crop rows [@Huber2018_SRAFS_growth].

**Long-rotation agroforestry systems (LRAFS)** use the same
alley-cropping spatial structure as SRAFS but with rotation cycles of
12--15\~years or longer [@Graves2007_Agroforestry_Economics;
@Nair1993_IntroAgroforestry]. At this rotation length, individual trees
develop substantial stem diameters and the proportion of high-value
timber increases markedly, enabling premium material uses (veneer,
sawlogs, engineered wood products) that are inaccessible in
short-rotation systems. Therefore, often timber-quality tree species
(such as walnut, cherry, oak, or high-grade poplar clones) are planted.

SRAFS and LRAFS are not clearly differentiated in practice. First, a
clear cut-off age is hard to define and, more importantly, the intended
use of AFS biomass is not always clear and can also change over time.
Nonetheless, when establishing AFS, the structure of AFS and its
intended use after harvest drives planting decisions and, subsequently,
economic viability. Thus, in the present paper SRAFS and LRAFS are not
considered as separate AFS systems, but the interplay of intended use,
establishment, and harvesting decisions is analysed.

## Economic rationale of landowners

The decision of a landowner to establish an AFS involves weighing
substantial up-front costs and long-term opportunity costs against a
diversified revenue stream. On the cost side, establishment requires
site preparation, planting material, and planting labour, typically
amounting to 1,500--3,500 \euro/ha for SRC or SRAFS
[@Faasch2012_SRC_Economics; @Testa2014_Poplar_Economics]. Recurrent
harvest and refitting costs (machine hire, chipping, transport to
roadside) add approximately 200--400 \euro/ha per rotation depending on
stand density and equipment availability [@Testa2014_Poplar_Economics;
@Spinelli2009_SRC_Harvesting]. During the rotation, maintenance costs of
AFS arise on a yearly basis. Next to expenses for establishment,
maintenance and harvesting, opportunity costs arise as no revenues from
crops are generated on tree strips. In regions with fertile soils and
high cereal prices, yearly crop revenues may sum up to 400--600 \euro/ha
[@Faasch2012_SRC_Economics; @Graves2007_Agroforestry_Economics]. Thus,
opportunity costs are often the dominant cost driver in SRAFS
profitability analyses.

On the **benefit side**, the primary revenue of AFS is the sale of
harvested woody biomass, whose magnitude depends critically on rotation
length, species, and site productivity. For poplar-based SRAFS in
Central Europe, typical yields range from 10 to 25 t of dry matter
biomass per ha for typical rotation lengths [@Huber2018_SRAFS_growth;
@Niemczyk2021_Poplar_Rotation]. Additionally, positive crop-yield
effects in the inter-row strips of SRAFS and LRAFS may result. Tree rows
act as windbreaks, reducing soil erosion and evapotranspiration; they
provide partial shading that moderates heat stress during drought years;
and their root systems may improve water retention in the topsoil
[@Graves2007_Agroforestry_Economics; @Grunewald2007_Robinia_AFS].
Empirical measurements in Central European alley-cropping trials
indicate that these microclimate effects can increase cereal yields in
adjacent strips by 5--15% relative to open-field controls, partially or
fully compensating the loss of arable area occupied by the tree strips
[@Graves2007_Agroforestry_Economics]. Beyond direct yield effects, AFS
deliver broader ecosystem services (like soil organic matter
accumulation, carbon sequestration, biodiversity enhancement) that may
lead to agri-environment subsidies and carbon credits
[@Toensmeier2017_Carbon_Agroforestry; @Faasch2012_SRC_Economics].

Nonetheless, economic attractiveness of AFS depends primarily on
revenues of AFS biomass. Thereby, revenues are different for the types
of biomass and in which value chain it is used. Often AFS biomasses are
used energetically, i.e. for heat and power generation, or anaerobic
digestion, which typically creates low prices per tonne of biomass.[REF]
Alternatively, woody parts of the biomass (i.e., the stem and large
branches) can also be used as pulp and paper grade material as well as
feedstock for chemical industry or as timber achieving higher prices
[@Lamers2015_Feedstock_Design] [EXTEND REFS]. Thus, fostering
AFS establishment by maximizing economic returns from AFS biomass
requires a strategic approach to product cascading that allocates
different biomass fractions to the highest-value applications, while
ensuring that the rotation length is aligned with the growth dynamics of
the stand and the market demand for different biomass types
[@Lamers2015_Feedstock_Design].

Thus, for making educated decisions on AFS establishment and harvesting,
the expected biomass quantities at harvest for stem, branches, and
residues need to be estimated. Thereby, the shares of the biomass types
change with stand age. E.g.\~the share of stem weight rises from roughly
55% at age 4 to over 75% at age 10 [@Huber2018_SRAFS_growth;
@Testa2014_Poplar_Economics]. Therefore, in Section
\@ref(sec:lit-growth) growth modelling approaches are reviewed that
capture the age-dependent growth dynamics of different biomass types in
AFS.

## Biomass growth models {#sec:lit-growth}

```{r tab-notation-growth, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr); library(kableExtra)
options(kableExtra.latex.load_packages = FALSE)
options(knitr.table.format = "latex")

notation <- tribble(
  ~Symbol, ~Description,
  # --- stand-level Gompertz ---
  "$M^{sl}(t)$",       "Total above- and belowground dry biomass per ha of AFS (t\\,DM\\,ha$^{-1}$) at stand age $t$",
  "$A(N)$",            "Stand-level Gompertz asymptote (t\\,DM\\,ha$^{-1}$); density- and site-dependent",
  "$k$",               "Gompertz growth-rate coefficient (yr$^{-1}$)",
  "$t_0$",             "Inflection age of stand biomass accumulation (yr)",
  # --- density--asymptote power law ---
  "$N$",               "Stand density (trees\\,ha$^{-1}$)",
  "$C_{\\mathrm{site}}$", "Site-quality constant (kg\\,ha$^{\\beta-1}$); conservative: 4\\,475, good site: 6\\,471",
  "$\\beta$",          "Competition-density exponent; fitted $\\hat{\\beta}=0.380$",
  # --- AGB/BGB split ---
  "$\\phi_{\\mathrm{ag}}$", "Aboveground fraction of total stand biomass; $\\phi_{\\mathrm{ag}}=0.89$",
  "$M^{ag}(t)$",       "Aboveground dry biomass per ha (t\\,DM\\,ha$^{-1}$) at age $t$",
  # --- compartment fractions ---
  "$f_p(t)$",          "asymtotic fraction of AGB attributed to compartment $p \\in \\{s,b,r\\}$ at age $t$",
  "$\\alpha_p,\\,\\delta_p$", "Intercept and slope of the linear fraction model for compartment $p$",
  # --- logistic merchantability ---
  "$q_p(t)$",          "Merchantable fraction of compartment $p$ biomass at age $t$ (logistic)",
  "$r_p$",             "Logistic growth rate for merchantability of compartment $p$",
  "$t_{50,p}$",        "Age at which 50\\% of compartment $p$ biomass is merchantable (yr)",
  # --- harvest yields ---
  "$Y_p(t)$",          "Merchantable dry biomass of compartment $p$ per ha at harvest age $t$ (t\\,DM\\,ha$^{-1}$)",
  "$\\omega$",         "Moisture content at felling (mass fraction); $\\omega = 0.55$"
)

kable(notation,
      format    = "latex", booktabs = TRUE, escape = FALSE,
      col.names = c("Symbol", "Description"),
      caption   = "Notation for the biomass growth and harvest yield models.",
      label     = "tab-notation-growth") |>
  kable_styling(font_size = 8,
                latex_options = c("hold_position", "scale_down", "striped"),
                full_width = FALSE) |>
  column_spec(1, width = "3.2cm") |>
  column_spec(2, width = "10.8cm")
```

Reliable estimation of stand-level biomass quantities and their age-dependent composition is a necessary prerequisite for any strategic
analysis of AFS-based supply chains. Individual-tree biomass is commonly approximated by species-specific allometric power laws relating shoot
dry matter to stem base diameter, and these relationships have been calibrated for Central European poplar clones under SRC and SRAFS conditions [@Huber2016_Allometric_SRAFS; @Huber2018_SRAFS_growth]. For supply chain planning, however, the relevant quantity is stand-level
biomass per hectare over a multi-decade rotation horizon, which requires a temporal growth model that captures the characteristic sigmoidal accumulation trajectory of woody biomass. Therefore, a set of variables is used summarized in \autoref{tab-notation-growth}.

### Stand-level growth model

Stand-level aboveground and belowground dry biomass per hectare,
$M(t)$, accumulates in a right-skewed sigmoidal pattern with a
relatively early inflection point, driven by rapid juvenile shoot growth
followed by self-thinning and competitive deceleration as canopy closure
intensifies [@Niemczyk2021_Poplar_Rotation;
@Testa2014_Poplar_Economics]. This trajectory is described by a Gompertz
model: \begin{equation}
  M(t) \;=\; A(N) \cdot \exp\!\bigl(-\exp(-k \cdot (t - t_0))\bigr),
  \label{eq:gomp-stand}
\end{equation} where $t_0$ is the age of maximum instantaneous biomass
increment and $k$ determines how quickly the stand approaches its
site-specific asymptote $A(N)$. For Central European poplar under
medium-rotation AFS management (10--20 yr), analysing
empirical data from @Huber2018_SRAFS_growth,
@Niemczyk2021_Poplar_Rotation, and @Jha2018_Poplar_AFS_Biomass yields parameter intervals of 
$k \in [0.17,0.2]$ and $t_0 \in [9,10]$ years, depending on local specifications of the AFS under study.
Parameter $t_0$ indicates an upper bound on the optimal rotation age
as growth slows down or diminishes entirely close to this age. 

The asymptote $A(N)$ depends largerly on planting density and can be modelled by a competition-density power law
[@Shinozaki1961_CD_Effect]: 
\begin{equation}
  A(N) \;=\; C_{\mathrm{site}} \cdot N^{1-\beta},
  \label{eq:A-density}
\end{equation} 
where $C_{\mathrm{site}}$ encodes site quality, $N$ indiactes tree density (e.g. trees per hectar), and
$\beta$ is the competition exponent. Equivalently, the per-tree
asymptote $A_{\mathrm{tree}}(N) = A(N)/N = C_{\mathrm{site}}\cdot N^{-\beta}$
decreases with density as individual trees have access to less growing
space. Parameters $C_{\mathrm{site}}$ and $\beta$ are highly site- and specicies-specific. For Poplar clones under European conditions,  $\beta \in [0.3,0.5]$ and $C_{\mathrm{site}} \in [4,6.5]$ [@Niemczyk2021_Poplar_Rotation,@Shinozaki1961_CD_Effect]. \autoref{fig:growth-asymptote-functions} shows asymptote functions according to \eqref{eq:A-density} for advantageous and disadvantageous site conditions.

```{r growth-asymptote-functions, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Tree-level and stand-level Gompertz asymptote function $A(N)$ of planting density for the advanteguous ($N=4$, $\\beta=-0.3$) and disadvantageous ($N=6.5$, $\\beta=-0.5$) scenarios.",  fig.width=6.5, fig.height=4, out.width="\\textwidth", cache=FALSE}

# ── Panel (a): A_stand vs N ─────────────────────────────────────────────────
N_seq <- seq(20, 400, by = 1)
df_a  <- bind_rows(
  tibble(N = N_seq,
         A = A_stand(N_seq, C_site = 4.000, beta=-.3),
         scenario = "advanteguous"),
  tibble(N = N_seq,
         A = A_stand(N_seq, C_site = 6.500, beta = -.5),
         scenario = "disadvantageous")
)

df_tree  <- bind_rows(
  tibble(N = N_seq,
         A = A_tree(N_seq, C_site = 4.000, beta=-.3),
         scenario = "advanteguous"),
  tibble(N = N_seq,
         A = A_tree(N_seq, C_site = 6.500, beta = -.5),
         scenario = "disadvantageous")
)


p_a <- ggplot(df_a, aes(x = N, y = A, linetype = scenario)) +
  geom_line(colour = "grey20", linewidth = 0.7) +
  scale_linetype_manual(
    name   = "Scenario",
    values = c("advanteguous" = "solid", "disadvantageous" = "dashed")) +
  labs(x = expression("Density "*italic(N)*" (trees ha"^{-1}*")"),
       y = expression(italic(A(N))*" (t DM ha"^{-1}*")")) +
  theme_minimal(base_size = 8) +
  theme(legend.position  = c(0.72, 0.82),
        legend.key.size  = unit(0.4, "cm"),
        legend.text      = element_text(size = 7),
        legend.title     = element_text(size = 7, face = "bold"),
        legend.background = element_rect(fill = "white", colour = NA),
        panel.grid.minor = element_blank())

p_b <- ggplot(df_tree, aes(x = N, y = A, linetype = scenario)) +
  geom_line(colour = "grey20", linewidth = 0.7) +
  scale_linetype_manual(
    name   = "Scenario",
    values = c("advanteguous" = "solid", "disadvantageous" = "dashed")) +
  labs(x = expression("Density "*italic(N)*" (trees ha"^{-1}*")"),
       y = expression(italic(A(N))*" (t DM per tree)")) +
  theme_minimal(base_size = 8) +
  theme(legend.position  = c(0.72, 0.82),
        legend.key.size  = unit(0.4, "cm"),
        legend.text      = element_text(size = 7),
        legend.title     = element_text(size = 7, face = "bold"),
        legend.background = element_rect(fill = "white", colour = NA),
        panel.grid.minor = element_blank())
# ── Assemble with patchwork ─────────────────────────────────────────────────
p_b / p_a
```

When assessing potential downstream application of AFS biomass, biomass fraction are to be defined meeting the requirements of downstream consumers. E.g. for timber and biochemical applications, stem diameters of AFS trees need to exceed a minimum threshold in order to qualify as a suitable feedstock. In the following three biomass fractions are distinguished: stems, branch wood, and residue. The stem fraction consists of stems with a diameter of at least 15 cm, while branch wood requires a diameter of at least 7 cm. The residue fraction comprises are smaler parts of the tree. To subdivide the total AFS biomass modelled by \eqref{eq:gomp-stand} into the three fractions depending on tree age, a logistic growth model is used: 
\begin{equation}
  q_p(t) \;=\; \frac{f_p}{1 + \exp\!\bigl(-r_p \cdot (t - t_{50,p})\bigr)},
  \label{eq:logistic-merch}
\end{equation} 
where $t_{50,p}$ is the age at which 50% of compartment $p$ biomass surpasses the diameter threshold and $f_p$ denotes the asymptotic share of component $p$ for mature trees. Using empirical data from data from @Civitarese2019_Poplar_Harvest, @Jha2018_Poplar_AFS_Biomass and @Niemczyk2021_Poplar_Rotation, $r_{stem} = 0.467$,yr$^{-1}$, $t_{50,stem} = 8.19$ years, and $f_{stem} = 0.75$ is estimated for the stem fraction (threshold $d > 15$ cm) and $r_{bran} = 0.698$ yr$^{-1}$, $t_{50,bran} = 7.55$,yr, and $f_{bran} = 0.15$ for the branch fraction (diameter at least 7 cm at insertion). \autoref{fig:afs-model-functions} illustrates the development of the fraction shares over stand age. 

```{r afs-model-functions, echo=FALSE, message=FALSE, warning=FALSE,     fig.cap="Logistic diameter-class merchantability functions $q_p(t)$ for stem  ($d > 15$\\,cm, green) and branches ($d > 7$\\,cm, orange); the 50\\% threshold ages $t_{50}$ are indicated by dashed vertical lines.", fig.width=6.5, fig.height=3, out.width="\\textwidth", cache=FALSE}

# ── Panel (b): logistic merchantability q_p(t) ─────────────────────────────
ages_c <- seq(1, 20, by = 0.1)
df_c <- tibble(
  age                    = ages_c,
  `Stem (d > 15 cm)`    = f_stem_merch(ages_c),
  `Branch (d > 7 cm)`   = f_branch_merch(ages_c)
) |>
  pivot_longer(-age, names_to = "compartment", values_to = "q") |>
  mutate(compartment = factor(compartment,
           levels = c("Stem (d > 15 cm)", "Branch (d > 7 cm)")))

t50_vals <- tibble(
  compartment = factor(c("Stem (d > 15 cm)", "Branch (d > 7 cm)"),
                       levels = c("Stem (d > 15 cm)", "Branch (d > 7 cm)")),
  t50 = c(8.19, 7.55)
)

merch_cols <- c("Stem (d > 15 cm)"  = "#1b9e77",
                "Branch (d > 7 cm)" = "#d95f02")

p_c <- ggplot(df_c, aes(x = age, y = q, colour = compartment)) +
  geom_vline(data = t50_vals,
             aes(xintercept = t50, colour = compartment),
             linetype = "dashed", linewidth = 0.45, show.legend = FALSE) +
  geom_hline(yintercept = 0.5,
             linetype = "dotted", colour = "grey60", linewidth = 0.4) +
  geom_line(linewidth = 0.8) +
  annotate("text", x = 8.19 + 0.3, y = 0.08,
           label = expression(italic(t)[50*","*s]*" = 8.2 yr"),
           size = 2.4, hjust = 0, colour = "#1b9e77") +
  annotate("text", x = 7.55 - 0.3, y = 0.20,
           label = expression(italic(t)[50*","*b]*" = 7.6 yr"),
           size = 2.4, hjust = 1, colour = "#d95f02") +
  scale_colour_manual(name = "Compartment", values = merch_cols) +
  scale_x_continuous(breaks = seq(0, 20, 2)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, max(df_c$q)*1.1)) +
  labs(x = "Stand age (yr)",
       y = expression(italic(q[p])*"(t)  [merchantable fraction]")) +
  theme_minimal(base_size = 8) +
  theme(legend.position  = c(0.15, 0.75),
        legend.key.size  = unit(0.4, "cm"),
        legend.text      = element_text(size = 7),
        legend.title     = element_text(size = 7, face = "bold"),
        legend.background = element_rect(fill = "white", colour = NA),
        panel.grid.minor = element_blank())

# ── Assemble with patchwork ─────────────────────────────────────────────────
p_c
```

Combining equations \eqref{eq:gomp-stand}--\eqref{eq:logistic-merch}, the dry biomass yield of fraction $p$ per hectare at harvest age $t$ is: \begin{equation}
  \eta_p(t) \;=\; q_p(t)\cdot M(t),
  \label{eq:yield}
\end{equation} 
with all non-merchantable fractions aggregated into the residue fraction. Fresh-weight yields at felling are obtained from dry matter yields via $\eta_p^{\mathrm{fw}}(t) = \eta_p(t)/(1-\omega)$, where the moisture content at felling $\omega = 0.55$ is taken from @Civitarese2019_Poplar_Harvest. \autoref{fig:afs-growth-curves} shows the fractions' stacked biomass quantities $\eta_p(t)$  depending on stand age compared with empirical observations from @Jha2018_Poplar_AFS_Biomass

```{r afs-growth-curves, echo=FALSE, message=FALSE, warning=FALSE,     fig.cap="Stand-level aboveground biomass components for poplar under AFS management with $N=250$ trees\\,ha$^{-1}$. Empirical AGB observations compiled from \\citet{Jha2018_Poplar_AFS_Biomass}, \\citet{Civitarese2019_Poplar_Harvest}, and \\citet{Niemczyk2021_Poplar_Rotation}. Grey band: typical medium-rotation harvest window (5--15 years).",    fig.width=6.5, fig.height=4.0, out.width="\\textwidth", cache=FALSE}

age.vec <- seq(0, 20, by = 0.1)

cons <- build_scenario_ts(ages = age.vec , N = 300, C_site = 6.500, k = 0.194, t0 = 9.7, label = "Conservative (250 trees/ha)")

comp_long <- bind_rows(cons) |>
  rename(stem = "Merch. stem", branch = "Merch. branch", residue = Residue) %>% 
  pivot_longer(c("stem", "branch", "residue"),
               names_to = "component", values_to = "biomass") |>
  mutate(
    component = factor(component,
      levels = c("residue", "branch", "stem")),
    scenario  = factor(scenario,
      levels = c("Conservative (113 trees/ha)", "Good site (139 trees/ha)"))
  ) %>% 
  select(-scenario)

# totals_long <- comp_long |>
#   group_by(age, scenario) |>
#   summarise(total = sum(biomass), .groups = "drop")

emp_pts <- tribble(
  ~age, ~AGB,   ~p_stem, ~p_branch,
   8,   134.3,  0.747,   0.166, # N = 400
   9,    74.0,  0.830,   0.110, # N = 500
   9,    51.5,  0.920,   0.040, # N = 500
  10,   119,    0.76,   0.210,# N = 139
  13,   152.0,  0.770,   0.220 # N = 139
) |>
  mutate(stem_cum   = AGB * p_stem,
         branch_cum = AGB * (p_stem + p_branch),
         total      = AGB)

emp_long <- bind_rows(
  emp_pts |> transmute(age, biomass = stem_cum,   series = "Empirical: stem"),
  emp_pts |> transmute(age, biomass = branch_cum, series = "Empirical: stem+branches"),
  emp_pts |> transmute(age, biomass = total,      series = "Empirical: total biomass")
) |> mutate(series = factor(series, levels = c(
  "Empirical: stem", "Empirical: stem+branches", "Empirical: total biomass")))

emp_col   <- c("Empirical: stem"          = "darkgreen",
               "Empirical: stem+branches" = "darkorange",
               "Empirical: total biomass"     = "navy")
emp_shape <- c("Empirical: stem"          = 21L,
               "Empirical: stem+branches" = 24L,
               "Empirical: total biomass"     = 23L)

ggplot(comp_long, aes(x = age, y = biomass)) +
  annotate("rect", xmin = 5, xmax = 15, ymin = 0, ymax = Inf,
           fill = "grey80", alpha = 0.25) +
  geom_area(aes(fill = component),  #, linetype = scenario
            position = "stack", alpha = 0.70, colour = "white", linewidth = 0.2) +
  # geom_line(data    = totals_long,
  #           aes(x = age, y = total, linetype = scenario),
  #           colour  = "grey30", linewidth = 0.45, inherit.aes = FALSE) +
  geom_point(data = emp_long,
             aes(x = age, y = biomass,
                 colour = series, shape = series, fill = series),
             size = 2.5, alpha = 0.90, inherit.aes = FALSE) +
  scale_fill_manual(
    name   = "Biomass fraction",
    values = c("stem"   = "#1b9e77",
               "branch" = "#d95f02",
               "residue"       = "#7570b3"),
    breaks = c("stem", "branch", "residue")) +
  scale_colour_manual(name = "Empirical data", values = emp_col) +
  scale_shape_manual (name = "Empirical data", values = emp_shape) +
  # scale_linetype_manual(
  #   name   = "Scenario",
  #   values = c("Conservative (113 trees/ha)" = "solid",
  #              "Good site (139 trees/ha)"    = "dashed")) +
  guides(
    fill     = guide_legend(order = 1,
                 override.aes = list(shape = NA, colour = NA, linetype = 0)),
    colour   = guide_legend(order = 3,
                 override.aes = list(
                   shape  = emp_shape[levels(emp_long$series)],
                   size   = 2.5,
                   fill   = emp_col[levels(emp_long$series)])),
    linetype = guide_legend(order = 2,
                 override.aes = list(fill = NA, colour = "grey30")),
    shape    = "none") +
  scale_x_continuous(breaks = seq(0, 20, 2)) +
  labs(x = "Stand age (years)",
       y = expression("biomass (t DM "*ha^{-1}*")")) +
  theme_minimal(base_size = 9) +
  theme(legend.position  = "right",
        legend.key.size  = unit(0.4, "cm"),
        legend.text      = element_text(size = 7),
        legend.title     = element_text(size = 8, face = "bold"),
        panel.grid.minor = element_blank())
```

## Biomass supply chain design {#sec:lit-sc}

The design and planning of biomass supply chains has attracted substantial research attention over the past two decades, driven by the
need to ensure cost-efficient and sustainable feedstock provision for bioenergy and biorefinery applications. Several systematic reviews
provide comprehensive overviews of this literature. @Cambero2014_ForestBSC_Review survey economic, social, and environmental
assessment and optimisation studies for forest biomass supply chains, identifying MILP as the dominant modelling approach and noting a growing
trend towards multi-objective formulations integrating life cycle assessment.

@Yue2014_BSC_Optimization_Overview review biomass-to-bioenergy and
biofuel supply chain models with a focus on multi-scale modelling,
uncertainty treatment, and sustainability metrics, and conclude that the
coupling of upstream feedstock dynamics with downstream network design
remains an underexplored area. @Malladi2018_BSC_Logistics survey
logistics-specific features of biomass supply chain models—seasonal
availability, moisture content degradation, preprocessing steps—and
highlight that most models treat biomass as a homogeneous commodity,
neglecting quality differentiation. @Zahraee2020_BSC_review provide a
recent meta-review of 140+ studies, confirming that cost minimisation is
still the dominant objective, and that multi-product cascading
formulations are rare.

The structure of biomass supply chain models typically encompasses three
decision layers: (i) strategic network design (facility location,
capacity sizing, technology selection), (ii) tactical planning (harvest
scheduling, seasonal flows, inventory), and (iii) operational routing.
The following paragraphs discuss key individual contributions and their
specific methodological advances.

@Eksioglu2009_Biorefinery_SC develop one of the earliest large-scale
MILP models for biomass-to-biorefinery supply chain design, integrating
facility location, capacity sizing, and multimodal transport for
lignocellulosic feedstocks. Their model minimises total system cost and
introduces binary facility-opening variables that have since become
standard in the field. A key limitation, however, is that biomass supply
is treated as an exogenous input, so feedstock availability is not
endogenised as a function of land-use or growth decisions.
@Mansoornejad2013_Forest_Biorefinery_SC extend the strategic design
perspective to forest biorefineries by integrating product portfolio
design with supply chain configuration in a single MILP. Their model
explicitly accounts for product conversion efficiencies along different
processing pathways, enabling joint optimisation of process technology
selection and logistics.

@DeMeyer2015_OPTIMASS introduce the OPTIMASS framework, a generic MILP
for biomass-based supply chain optimisation that explicitly models
product quality cascades: higher-quality biomass classes can substitute
for lower-quality demands downstream, whereas the reverse is infeasible.
This cascade hierarchy, formalised through ordered product sets, is
particularly relevant for woody biomass where stem, branch, and residue
fractions command different market prices. The companion t-OPTIMASS
extension [@DeMeyer2016_tOPTIMASS] incorporates temporal dynamics by
explicitly representing biomass growth and regeneration over multiple
periods, enabling simultaneous optimisation of harvest timing and
network flows. However, in both OPTIMASS formulations the growth
dynamics are represented as exogenous time-series rather than as
age-structured stand models, so rotation-length decisions and
age-dependent biomass quality are not endogenised.
@Arabi2019_Microalgae_Biobutanol_SC demonstrate multi-period MILP
extensions for bio-based supply chains with production dynamics, but
focus on microalgae rather than woody biomass and do not address product
quality cascades.

@Sharma2013_BSC_review identify supply-side uncertainty (yield
variability, seasonal availability, moisture content) as the most
critical source of risk in biomass supply chains. In response, a strand
of literature has developed stochastic and robust optimisation
extensions. @Zahraee2020_BSC_review document that two-stage stochastic
MILP and robust optimisation approaches now account for roughly 30% of
the modelling literature, with biomass yield uncertainty and price
volatility as the dominant stochastic parameters.

Biomass supply chain research focusing specifically on agroforestry or
SRC contexts is comparatively sparse.
@Espinoza2017_Biorefinery_SC_review note that most supply chain models
assume continuous feedstock availability and do not represent the
periodic harvest cycles characteristic of coppice systems.
@Lamers2015_Feedstock_Design analyse feedstock design principles for
lignocellulosic supply chains and argue that the allocation of different
biomass fractions to valorisation pathways—product cascading—is a key
lever for economic viability, yet is rarely optimised jointly with
network design decisions. The model developed in
Section \@ref(sec:milp) of this paper directly addresses this gap by
endogenising stand age dynamics through binary age-lag variables and
coupling component-specific yield coefficients with a product quality
cascade.

\autoref{tab:lit-sc-features} summarises the key modelling features
of the reviewed contributions.

```{r lit-sc-features, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
options(kableExtra.latex.load_packages = FALSE)
options(knitr.table.format = "latex")
knitr::opts_chunk$set(dev = "pdf")

lit_tab <- tribble(
  ~Study,                            ~Objective,          ~`Multi-product`,  ~`Quality cascade`, ~`Age dynamics`,    ~`Uncertainty`,    ~`Level`,
  "Eksioglu et al. (2009)",          "Cost min.",         "No",              "No",               "No",               "No",              "Strategic",
  "Sharma \\& Taaffe (2013)",        "Cost min.",         "Partial",         "No",               "No",               "No",              "Strategic/Tactical",
  "Mansoornejad et al. (2013)",      "Profit max.",       "Yes",             "No",               "No",               "No",              "Strategic",
  "Cambero \\& Sowlati (2014)",      "Cost/GHG/jobs",     "No",              "No",               "No",               "Stochastic",      "Strategic",
  "Yue et al. (2014)",               "Cost/GHG",          "Yes",             "No",               "No",               "No",              "Multi-scale",
  "De Meyer et al. (2015)",          "Cost min.",         "Yes",             "Yes",              "Exogenous",        "No",              "Strategic/Tactical",
  "De Meyer et al. (2016)",          "Cost min.",         "Yes",             "Yes",              "Exogenous TS",     "No",              "Multi-period",
  "Lamers et al. (2015)",            "Cost/quality",      "Yes",             "Partial",          "No",               "Scenario",        "Strategic",
  "Espinoza et al. (2017)",          "Sustainability",    "Yes",             "No",               "No",               "Scenario",        "Strategic",
  "Malladi \\& Sowlati (2018)",      "Cost min.",         "Partial",         "No",               "No",               "No",              "Tactical/Oper.",
  "Arabi et al. (2019)",             "Profit max.",       "Yes",             "No",               "Exogenous",        "No",              "Multi-period",
  "Zahraee et al. (2020)",           "Multi-obj.",        "Yes",             "No",               "No",               "Stochastic",      "Strategic/Tactical",
  "\\textbf{This paper}",            "NPV max.",          "Yes",             "Yes",              "\\textbf{Endogenous}", "Scenario",   "Strategic/Multi-period"
) |>
  mutate(across(everything(), ~ gsub("Yes", "\\\\checkmark", .))) |>
  mutate(across(everything(), ~ gsub("No",  "---",            .)))

kable(lit_tab,
      format    = "latex",
      booktabs  = TRUE,
      escape    = FALSE,
      caption   = "Key modelling features of selected biomass supply chain design studies.",
      label     = "lit-sc-features") |>
  kable_styling(
  font_size     = 7,
  latex_options = c("hold_position", "scale_down", "striped"),
  full_width    = FALSE
) |>
  row_spec(nrow(lit_tab), bold = TRUE) |>
  column_spec(1, width = "3.2cm") |>
  column_spec(2, width = "1.6cm") |>
  column_spec(3:7, width = "1.45cm")
```

# MILP Model for Agroforestry Supply Chain Design {#sec:milp}

Building on the biomass growth models introduced in Section
\@ref(sec:lit-growth) and the supply chain design literature reviewed in
Section \@ref(sec:lit-sc), this section describes the AFS supply chain
design problem (\problem), a mixed-integer linear programming (MILP)
model for the integrated design and operation of an agroforestry biomass
supply chain with multiple AFS biomass types (i.e., stem, branch,
and residue fractions) used as feedstocks for potential consumers in chemical, pulp and
paper as well as energy industry.

## Sets, Indices, and Time Structure

The model considers a set of agroforestry sites $\mathcal{I}$, storage/pre-treatment facilities $\mathcal{J}$, consumer sites $\mathcal{K}$, and biomass types $\mathcal{P} = \{stem, branches, residues\}$, whose growth models are outlined above. Time is discretised into annual periods $t \in \Tc = \{1,\dots,T_{\max}\}$. To reduce complexity, reflect legal definitions of SRAFS and LRAFS as well as discard implausible options (e.g. extremly long rotation cycles), $A^{\min}$ and $A^{\max}$ define minimum and maximum stand ages. Thus, harvest are possible in periods $\mathcal{T}^{\text{harv}} = \{A^{\min}+1,\dots,T_{\max}\}$. To consistently model establishment and consecutive harvest decisions, the following arc sets are defined: $\Sc^{est}$ defines potential establishments of AFS. If an AFS is established in period $t \in \Tc$, arc $(0,t)$ is selected. Harvest set $\Sc^{harv}$ summarizes all potential consecutive harvest periods within the age interval $[A^{\min},A^{\max}]$. I.e., if an AFS is established or harvested in period $s$ and next harvest is in period $t$, arc $(s,t)$ is selected. Finally, if there is no subsequent harvest after period $t$, the AFS is terminated by selecting arc $(t,T^{\max}+1)$. Termination arcs are summarized in set $\Sc^{term}$. All arc sets are formally defined as
follows: \begin{align}
  \Sc^{{est}}  &= \{(0,t) \mid t \in \{1,\dots,T^{\max}-A^{\min}\}\}, \notag\\
  \Sc^{{harv}} &= \{(s,t) \mid s,t \in \Tc \wedge t-s \in \{A^{\min},\dots,A^{\max}\}\},  \notag\\
  \Sc^{{term}} &= \{(s,T^{\max}+1) \mid s \in \{A^{\min}+1,\dots,T_{\max}\}\}, \notag\\
  \Sc               &= \Sc^{{est}} \cup \Sc^{{harv}} \cup \Sc^{{term}}. \notag
\end{align}

\autoref{fig:fig-set-S-tikz} illustrates this network structure of all arc sets for $T=8$, $A^{\min}=3$, $A^{\max}=5$.

```{tikz fig-set-S-tikz, echo=FALSE, fig.cap="Network structure of set $\\mathcal{S}$ for $T=8$, $A_{\\min}=3$, $A_{\\max}=5$. Dashed blue arcs: establishment arcs $(0,t)$, $t\\in\\{1,\\ldots,T-1\\}$. Solid green arcs above: harvest arcs $(s,t)$ with odd $s$; below: harvest arcs with even $s$. Solid red arcs:    termination arcs $(s,T+1)$, $s\\in\\{A_{\\min}+1,\\ldots,T\\}$.",    fig.align="center", out.width="\\textwidth", engine.opts=list(extra.preamble=c(      "\\usetikzlibrary{arrows.meta,bending,positioning}"   )), cache=TRUE}


\begin{tikzpicture}[
  x=1.55cm, y=1cm,
  node/.style={circle, draw, fill=#1, text=white,
               font=\small\bfseries, inner sep=2pt, minimum size=18pt},
  est/.style  ={-{Stealth[length=4pt]}, blue!70!black,  dashed, semithick},
  harva/.style={-{Stealth[length=4pt]}, green!40!black, semithick},
  harvb/.style={-{Stealth[length=4pt]}, green!40!black, semithick},
  term/.style ={-{Stealth[length=4pt]}, red!70!black,   semithick},
]

\node[node=blue!70!black]  (n0) at (0,0) {$0$};
\foreach \i in {1,...,8} {
  \node[node=gray!50!black] (n\i) at (\i,0) {$\i$};
}
\node[node=red!70!black]   (n9) at (9,0) {$9$};

\draw[est] (n0) -- (n1);
\foreach \t/\l in {2/1.3, 3/1.0, 4/0.80, 5/0.65}{
  \draw[est] (n0) to[out=55, in=125, looseness=\l] (n\t);
}

\draw[harva] (n1) to[out=55,  in=125,  looseness=0.55] (n4);
\draw[harvb] (n2) to[out=-55, in=-125, looseness=0.55] (n5);
\draw[harva] (n3) to[out=55,  in=125,  looseness=0.55] (n6);
\draw[harvb] (n4) to[out=-55, in=-125, looseness=0.55] (n7);
\draw[harva] (n5) to[out=55,  in=125,  looseness=0.55] (n8);
\draw[harva] (n1) to[out=60,  in=120,  looseness=0.42] (n5);
\draw[harvb] (n2) to[out=-60, in=-120, looseness=0.42] (n6);
\draw[harva] (n3) to[out=60,  in=120,  looseness=0.42] (n7);
\draw[harvb] (n4) to[out=-60, in=-120, looseness=0.42] (n8);
\draw[harva] (n1) to[out=65,  in=115,  looseness=0.35] (n6);
\draw[harvb] (n2) to[out=-65, in=-115, looseness=0.35] (n7);
\draw[harva] (n3) to[out=65,  in=115,  looseness=0.35] (n8);

\foreach \s/\l in {4/0.80, 5/0.65, 6/0.55, 7/0.48}{
  \draw[term] (n\s) to[out=55, in=125, looseness=\l] (n9);
}
\draw[term] (n8) -- (n9);

\node[anchor=north west] at (0.0,2.5) {%
  \begin{tikzpicture}[x=0.6cm,y=0.35cm,baseline=0pt]
    \draw[est]   (0,0)--(1,0);
    \node[font=\small, anchor=west] at (1.1,0) {Establishment arc $(0,t)$};
  \end{tikzpicture}};
\node[anchor=north west] at (3.5,2.5) {%
  \begin{tikzpicture}[x=0.6cm,y=0.35cm,baseline=0pt]
    \draw[harva] (0,0)--(1,0);
    \node[font=\small, anchor=west] at (1.1,0) {Harvest arc $(s,t)$};
  \end{tikzpicture}};
\node[anchor=north west] at (6.3,2.5) {%
  \begin{tikzpicture}[x=0.6cm,y=0.35cm,baseline=0pt]
    \draw[term]  (0,0)--(1,0);
    \node[font=\small, anchor=west] at (1.1,0) {Termination arc $(s,9)$};
  \end{tikzpicture}};

\end{tikzpicture}
```

## Decision Variables

Temporal flow Variables $z_{ist}$ indicate on which quantity of the area of site $i$ an an AFS is operated between $(s,t) \in \Sc$. Depending on the arc type, this implies establishment, harvest, or termination. Selected arcs buiild a network flow over the complete planning horizon from $t=0$ to $T^{max}+1$. \autoref{fig:fig-set-S-path} illustrates a simple network flow with establishment of an AFS in period 1 and harvests in periods 5 and 8. 

```{tikz fig-set-S-path, echo=FALSE, fig.cap="Example of a feasible path through the network for $T=8$,$A_{\\min}=3$, $A_{\\max}=5$: establishment in period~1, harvests in periods~5 and~8, termination at $T+1$.",    fig.align="center", out.width="\\textwidth", engine.opts=list(extra.preamble=c(     "\\usetikzlibrary{arrows.meta,bending,positioning}"    )), cache=TRUE}

\begin{tikzpicture}[
  x=1.55cm, y=1cm,
  node/.style    ={circle, draw, fill=#1, text=white,
                   font=\small\bfseries, inner sep=2pt, minimum size=18pt},
  inactive/.style={node=gray!25, text=gray!50},
  pathnode/.style={node=green!40!black},
  n0style/.style ={node=blue!70!black},
  n9style/.style ={node=red!70!black},
  patharc/.style ={-{Stealth[length=5pt]}, green!40!black, very thick},
  est/.style     ={patharc, blue!70!black, dashed},
  term/.style    ={patharc, red!70!black},
]

\node[n0style]   (n0) at (0,0) {$0$};
\foreach \i in {2,3,4,6,7} {
  \node[inactive] (n\i) at (\i,0) {$\i$};
}
\foreach \i in {1,5,8} {
  \node[pathnode]  (n\i) at (\i,0) {$\i$};
}
\node[n9style]   (n9) at (9,0) {$9$};

\draw[est]     (n0) -- (n1);
\draw[patharc] (n1) to[out=60, in=120, looseness=0.42] (n5);
\draw[term] (n5) to[out=-50, in=-130, looseness=0.42] (n9);
\draw[patharc] (n5) to[out=55, in=125, looseness=0.55] (n8);
\draw[term]    (n8) -- (n9);

\node[font=\footnotesize, text=blue!70!black]   at (0.5, 0.28) {est.};
\node[font=\footnotesize, text=blue!70!black]   at (0.5, -0.3) {$100$};
\node[font=\footnotesize, text=green!40!black]  at (3.0, 1.1)  {$100$};
\node[font=\footnotesize, text=green!40!black]  at (3.0, 0.6)  {$t-s=4$};
\node[font=\footnotesize, text=green!40!black]  at (6.5, 1) {$70$};
\node[font=\footnotesize, text=red!70!black]  at (7, -1.0) {$30$};
\node[font=\footnotesize, text=red!70!black]  at (8.5, -0.3) {$70$};
\node[font=\footnotesize, text=green!40!black]  at (6.5, 0.55) {$t-s=3$};
\node[font=\footnotesize, text=red!70!black]    at (8.5, 0.28) {term.};

\end{tikzpicture}
```

Spatial flow variables $X_{ijpt}$ and $X_{jkpp't}$ represent transport of products from AFS sites to storage/pre-treatment facilities and from there to consumers, respectively, while $S_{jpt}$ captures end-of-period inventory at facility $j$. \autoref{tab:tab-notation-milp} summarises all sets, decision variables, and parameters used in the MILP formulation, while $\eta_{pa}$ refers to \eqref{eq:yield}.

```{r tab-notation-milp, echo=FALSE, message=FALSE, warning=FALSE, cache=T}
library(knitr)
library(kableExtra)
options(kableExtra.latex.load_packages = FALSE)
options(knitr.table.format = "latex")

math_break <- function(line1, line2) {
  paste0("\\shortstack[l]{$", line1, "$\\\\$", line2, "$}")
}

notation <- tribble(
  ~Symbol, ~Description,
  # --- SETS & INDICES ---
  "$i \\in \\mathcal{I}$",   "Agroforestry sites",
  "$j \\in \\mathcal{J}$",   "Storage / pre-treatment facilities",
  "$k \\in \\mathcal{K}$",   "Consumer sites",
  "$p \\in \\mathcal{P}$",   "Product types (stem, branches, residues)",
  "$(p,p') \\in \\mathcal{Q}$",   "Admissible product-substitution pairs (quality cascade)",
  "$\\mathcal{T}=\\{1,\\dots,T\\}$",
    "Planning periods (years)",
  "$\\mathcal{T}^{\\text{harv}}=\\{A^{\\min}+1,\\dots,T\\}$",
    "Feasible harvest periods",
  "$\\mathcal{S}^{\\text{est}}=\\{(0,t)\\mid t\\in\\{1,\\dots,T-A^{\\min}\\}\\}$",
    "Establishment arcs: dummy source to period $t$",
  math_break(
    "\\mathcal{S}^{\\text{harv}}=\\{(s,t)\\mid s,t\\in \\mathcal{T},",
    "t-s\\in\\{A^{\\min},\\dots,A^{\\max}\\}\\}"),
    "Harvest cycles within age-lag bounds",
  math_break(
    "\\mathcal{S}^{\\text{term}}=\\{(s,T+1)\\mid",
    "s\\in\\{A^{\\min}+1,\\dots,T\\}\\}"),
    "Termination arcs to dummy sink $T+1$",
  "$\\mathcal{S} = \\mathcal{S}^{\\text{est}} \\cup \\mathcal{S}^{\\text{harv}} \\cup\\; \\mathcal{S}^{\\text{term}}$",
    "Age-lag arc set (establishment $\\cup$ harvest $\\cup$ termination)",
  # --- DECISION VARIABLES ---
  "$z_{ist} \\geq 0$",
    "Area of site $i$ with AFS operation between $s$ and $t$",
  "$X_{ijpt}\\ge 0$",
    "Transport flow (t) of $p$ from site $i$ to facility $j$ in $t$",
  "$X_{jkpp't}\\ge 0$",
    "Flow of $p$ from facility $j$ to consumer $k$ satisfying demand $p'$ in $t$",
  "$S_{jpt}\\ge 0$",
    "Inventory (t) of product $p$ at facility $j$ end of period $t$",
  # --- PARAMETERS ---
  "$T$",              "Length of planning horizon (years)",
  "$A^{\\min}$, $A^{\\max}$", "Minimum / maximum stand age at harvest (years)",
  "$\\text{AREA}_i$",         "Area of site $i$ (ha)",
  "$\\eta_{pa}$",
    "Yield coefficient (t/ha) of fraction $p$ at stand age $a$, derived from \\eqref{eq:yield}",
  "$R_{kp}$",                 "Revenue (\\euro/t) for product $p$ at consumer $k$",
  "$C^{\\text{est}}_i$",      "Establishment cost (\\euro/ha) at site $i$",
  "$C^{\\text{opp}}_i$",      "Annual opportunity cost (\\euro/ha) at site $i$",
  "$C^{\\text{harv}}_i$",     "Harvest \\& refitting cost (\\euro/ha) at site $i$",
  "$c^{\\text{tr-raw}}$",     "Transport cost rate (\\euro/t\\,km) for raw biomass",
  "$c^{\\text{tr-pre}}$",     "Transport cost rate (\\euro/t\\,km) for pre-treated biomass",
  "$c^{\\text{stor}}_j$",     "Storage holding cost (\\euro/t) at facility $j$",
  "$d_{ij}$, $d_{jk}$",       "Distance (km) between nodes",
  "$\\text{CAP}^{\\text{stor}}_j/\\text{CAP}^{\\text{proc}}_j$", "Storage and processing capacity (t) at facility $j$",
  "$D^{\\max}_{kpt}$",        "Maximum demand (t) of fraction $p$ at consumer $k$ in period $t$"
)

tbl <- kable(notation,
      format    = "latex",
      booktabs  = TRUE,
      escape    = FALSE,
      col.names = NULL,
      caption   = "Notation for the MILP model.",
      label     = "tab-notation-milp") |>
  kable_styling(
    font_size     = 8,
    latex_options = c("hold_position", "scale_down", "striped"),
    full_width    = FALSE
  ) |>
  column_spec(1, width = "5cm") |>
   column_spec(2, width = "9cm") |>
   pack_rows("Sets \\& Indices",    1,  11,
             bold = TRUE, italic = FALSE, hline_before = TRUE, escape = FALSE) |>
   pack_rows("Decision Variables", 12,  15,
             bold = TRUE, italic = FALSE, hline_before = TRUE) |>
   pack_rows("Parameters",         16,  29,
             bold = TRUE, italic = FALSE, hline_before = TRUE)

tbl <- gsub(
  "p\\{(5\\.8cm|8\\.5cm)\\}",
  ">\\{\\\\raggedright\\\\arraybackslash\\}p\\{\\1\\}",
  tbl
)

tbl
```

## Objective Function

As discussed in Section \@ref(sec:lit-afs), the economic attractiveness of AFS supply chains is modelled from a global perspective taking into account all processes, activities, and participants involved in finally serving customer markets with AFS-based feedstocks. Therefore, revenues for the different biomass fractions and are weighed against costs occurring at all value-adding stages of the supply chains. Revenues created by selling AFS-based feedstocks to different industrial consumers are counterweighted by establishment, opportunity, harvest, transport, and storage costs of landowners, agricultural firms as well as biomass logistics companies. The model therefore maximizes total profit of all participants in the AFS supply chain over the planning horizon as follows: 
\begin{align}
\max\; Z \;=\;&
  \sum_{k \in \Kc}\sum_{p \in \Pc} R_{kp} \cdot \sum_{j \in \Jc, }\sum_{(p',p) \in \Qc} \sum_{t \in \Tc}  X_{jkp'pt}  \label{eq:obj-revenue} \\
&- \sum_{i \in \Ic} c^{{est}}_i \cdot \sum_{(0,t) \in \Sc^{{est}}} z_{i0t} \notag \\
% &- \sum_{i \in \Ic} c^{{opp}}_i \cdot \left(\sum_{(s,T+1) \in \Sc^{term}} (s+1) \cdot z_{is(T+1)}  -  \sum_{(0,t) \in \Sc^{{est}}}  t \cdot z_{i0t} \right) \notag \\ 
&- \sum_{i \in \Ic} (c^{main}_{i} + c^{{opp}}_i) \cdot (t-s) \cdot \sum_{(s,t) \in \Sc^{{harv}}} z_{ist} \notag \\
&- \sum_{i \in \Ic} c^{{harv}}_i  \cdot \sum_{(s,t) \in \Sc^{{harv}}} z_{ist} \notag \\
&- \sum_{i \in \Ic}\sum_{j \in \Jc}\sum_{p \in \Pc}\sum_{t \in \Tc} c^{\text{tr-raw}}_p \cdot d_{ij} \cdot X_{ijpt} \notag \\
&- \sum_{j \in \Jc}\sum_{k \in \Kc}\sum_{(p,p') \in \Qc} \sum_{t \in \Tc} c^{\text{tr-pre}}_{p'} \cdot d_{jk} \cdot X_{jkpp't} \notag \\
&- \sum_{j \in \Jc}\sum_{p \in \Pc}\sum_{t \in \Tc}  c^{{stor}}_j \cdot S_{jpt}  \notag
\end{align}

Here, $R_{kp}$ are product- and consumer-specific revenues for each ton of biomass fraction (\euro{}/t). The opportunity cost term penalises the land use when AFS are established. Thereby, $c^{{opp}}_i$ can be negative or positive depending on the site-specific conditions. It comprises effect of lost field crop revenues due to AFS as well as higher yields of remaining field crop areas e.g. due to wind and water erosion protection by AFS. While the assessment of lost field crop revenues is usually well predictable, the economic evaluation of positive ecological effects of AFS (e.g. on neighboring crop yields) is more difficult. Thus, $c^{{opp}}_i$ can also contain regulatory subsidies to support establishing and maintaining AFS. Transport cost terms mirror the two-stage flow structure of the AFS supply chain with differentiated rates for raw and pre-treated biomass depending on the biomass type. Cost rates differ because as e.g. the density and water content of the shipped biomass types differ or because specialized equipment (e.g. for log transports) is used [REFERENCES]. Moreover, site-specific establishment cost for planting the AFS trees is considered, next to yearly maintenance and harvesting costs. All these costs scale with the size of a site ($AREA_i$) measured in hectar. Maintenance cost comprise e.g. costs for trimming, replanting, and pruning unnecessary shoots. Storage costs are considered if biomass types are stored for more than a year. Storage cost rate $c^{{stor}}_j$ measures the costs for storing a ton of biomass for one year and comprises direct storage cost (e.g. for storage site maintenance) as well as opportunity cost due to potential material losses and capital commitment. Biomass types harvested and consumed in the same year incur no storage cost, as usually AFS are harvested at the beginning of a year. Harvested biomass has to dry during spring and summer such that the biomass can be used in autumn or winter the same year. As under-year drying is necessary for some industrial applications and beneficial for transport economics, drying storage in the year of harvest is not associated with additional cost in this study.

## Constraints

Constraints \eqref{eq:C1} limits the area dedicated to AFS operations to the maximum size of site $i$. Constraints \eqref{eq:C2} enforce flow conservation in each period. If AFS systems of size $z_{ist}$ are established or harvested in period $t \in \Tc$, there needs to be a succeeding activity (harvest or termination) of this area. This way, \eqref{eq:C2} guarantee valid (possibly empty) AFS activity  over the complete time horizon for each site. Constraints \eqref{eq:C3} links the harvest-cycle variables $z_{ist}$ to physical yield quantities $\eta^{fw}_{p(t-s)}$ for age $t-s$, which are pre-computed using \eqref{eq:gomp-stand}--\eqref{eq:yield} scaled by moisture content $\omega$. Thus, we assume constant yields per hectar for each site $i$. Note that this simplification can be dropped and site-specific yields $\eta^{fw}_{ip(t-s)}$ could be introduced without increasing numerical complexity of the \problem. Constraints \eqref{eq:C5} link harvest quantities to transport flows and ensure inventory balance at each facility. Thereby, harvest quantities can be smaller than available AFS biomass allowing partial harvesting. Constraints \eqref{eq:C6}--\eqref{eq:C7} enforce storage and processing capacity limits at each facility $j$. Constraints \eqref{eq:C8} implement product usage at multiple consumption sites $k$. Thereby, $(p',p) \in \Qc$ defines a product mapping allowing high-quality product $p'$ (e.g. stems) satisfying demand of a lower-quality products $p$ up to maximum demand $D^{\max}_{kpt}$.

\begin{align}
  \sum_{(0,t) \in \Sc^{{est}}} z_{i0t} &\leq AREA_i && \forall\; i \in \Ic \label{eq:C1} \\[1pt]
  %
  \sum_{(s,t) \in \Sc} z_{ist} &= \sum_{(t,u) \in \Sc} z_{itu} && \forall\; i \in \Ic,\; t \in \Tc \label{eq:C2} \\[1pt]
  %
  \sum_{j \in \Jc} X_{ijpt}    &\leq \sum_{(s,t) \in \Sc^{{harv}}} \eta^{fw}_{p(t-s)} \cdot z_{ist} && \forall\; i \in \Ic,\; p \in \Pc,\; t \in \Tc^{{harv}}     \label{eq:C3} \\[1pt]
  %
  S_{jpt}    = S_{jp,t-1} + &\sum_{i \in \Ic} X_{ijpt} - \sum_{k \in \Kc,\,(p,p') \in \Qc}  X_{jkpp't}   && \forall\; j \in \Jc,\; p \in \Pc,\; t \in \Tc^{harv}  \label{eq:C5} \\[1pt]
  %
  \sum_{p \in \Pc} S_{jpt}    &\leq \text{CAP}^{{stor}}_j    && \forall\; j \in \Jc,\; t \in \Tc     \label{eq:C6} \\[1pt]
  %
  \sum_{i \in \Ic,\, p \in \Pc} X_{ijpt}  &\leq \text{CAP}^{{proc}}_j && \forall\; j \in \Jc,\; t \in \Tc^{harv} \label{eq:C7} \\[1pt]
  %
  \sum_{j \in \Jc,\,(p',p) \in \Qc} X_{jkp'pt} &\leq D^{\max}_{kpt} && \forall\; k \in \Kc,\; p \in \Pc,\; t \in \Tc^{harv}     \label{eq:C8} \\[1pt]
  %
  z_{ist},\;X_{ijpt},\; &X_{jkp'pt},\; S_{jpt} \geq 0 && \forall\; i,j,k,p,p',t     \label{eq:D2}
  %
\end{align}

In summary, \problem constitutes are large-scaled LP which allows to decide on which fraction of a site, AFS are established and how the AFS operations are scheduled in a multi-year planning horizon. The \problem takes a supply chain perspective and maximizes total profit of all actors involved in these processes. Thus, it does not explain how the profits are split e.g. between farmers, AFS service providers and logistics service providers. The continuous modelling of AFS areas in $z_{ist}$ leads to a comparatively easy to solve problem. On the other side, it is unclear which part of a site's total area is chosen for AFS operations if $z_{ist} < AREA_i$, which holds for establishment as well as harvesting and termination.

# Case Study {#sec-casestudy}

The \problem is intended to support the design of regional agroforestry biomass supply chains fostering regional bioeconomy clusters thereby connecting multiple industrial sectors. The proposed \problem model is instantiated on a real-world biomass supply chain case study located in the tri-state region of Saxony-Anhalt, northern Thuringia, and northern Saxony, Germany. This region is characterized by a high density of agricultural land usage as well as cluster of paper and wood as well as chemical industry. The regional chemical industry relies on fossil ressources so fa, but is under substantial transformative pressure which led to investments in sustainable production processes requiring sustainable regional bio-based feedstocks. \autoref{fig:fig-supply-chain-network} shows agricultural production sites along with the set of potential wood-processing hubs, and the regional users of AFS biomasses from chemical, paper and wood processing industry as well as the energy sector. 

```{r fig-supply-chain-network, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="AFS supply chain network in Saxony-Anhalt. Grey lines: German federal state boundaries. Blue lines: major rivers. Coloured polygons: potential AFS parcels grouped by site (colours encode cluster membership; no legend). Orange symbols: pre-treatment facilities (circle = sawmill, triangle = logistics hub, size proportional to processing capacity). Coloured squares: consumer sites (violet: P1 chemical/pulp, blue: P2 pulp/paper, red: P3 energy/biogas, size proportional to total demand).", fig.width=6.5, fig.height=5.5, out.width="\\textwidth", cache=T}

crs_map <- 25833

# ── Colour palette constants (identical to fig-results-sc-map) ────────────────
COL_SAW <- "#e65c00"
COL_HUB <- "grey40"
COL_P1  <- "#6a0dad"
COL_P2  <- "#08519c"
COL_P3  <- "#a50026"

# ── Map extent: defined in WGS84, transformed to projected CRS ───────────────
bbox_wgs84 <- st_as_sfc(
  st_bbox(c(xmin = 10.6, xmax = 13.2, ymin = 50.9, ymax = 52.8),
          crs = st_crs(4326))
)
bbox_map <- st_transform(bbox_wgs84, crs_map)
bbox_num <- st_bbox(bbox_map)

# ── Base-map layers ────────────────────────────────────────────────────────────
de_states <- ne_states(country = "Germany", returnclass = "sf") %>%
  st_transform(crs_map) %>%
  st_intersection(bbox_map)

rivers_eu <- ne_download(
  scale    = 10,
  type     = "rivers_lake_centerlines",
  category = "physical",
  returnclass = "sf"
) %>%
  st_transform(crs_map) %>%
  st_intersection(bbox_map)

# ── Cluster colour palette (no legend printed) ────────────────────────────────
cluster_assig <- afs_workspace$site_cluster_assig %>%
  select(site_id, hac_cluster) %>%
  mutate(hac_cluster = as.factor(hac_cluster))

n_clusters <- length(unique(cluster_assig$hac_cluster))

cluster_pal <- setNames(
  if (n_clusters <= 12) {
    colorRampPalette(RColorBrewer::brewer.pal(min(n_clusters, 12), "Set3"))(n_clusters)
  } else {
    scales::viridis_pal(option = "turbo")(n_clusters)
  },
  levels(cluster_assig$hac_cluster)
)

# ── Site polygons: join cluster assignment, project ───────────────────────────
sites_sf_clust <- sites_sf %>%
  mutate(site_id = seq_len(nrow(sites_sf))) %>%
  left_join(cluster_assig, by = "site_id") %>%
  st_transform(crs_map)

# merge opportunity cost of 
tmp.opp <- sites_clustered %>% 
  select(site_id, C_opp) %>% 
  rename(hac_cluster = site_id) %>% 
  mutate(hac_cluster = as.factor(hac_cluster))

sites_sf_clust <- sites_sf_clust %>% 
  left_join(tmp.opp,by = "hac_cluster")

# ── Consumers & storages ───────────────────────────────────────────────────────
consumers_plt <- consumers %>%
  mutate(
    total_demand = demand_P1 + demand_P2 + demand_P3,
    kategorie = case_when(
      demand_P1 >= demand_P2 & demand_P1 >= demand_P3 & demand_P1 > 0 ~ "P1: stem/chemical",
      demand_P2 >= demand_P1 & demand_P2 >= demand_P3 & demand_P2 > 0 ~ "P2: branch/pulp",
      demand_P3 > 0                                                     ~ "P3: residue/energy",
      TRUE                                                              ~ "Other"
    ),
    pt_size = pmin(4.5, pmax(1.5, 1.5 + log1p(total_demand) * 0.55))
  )

storages_plt <- storages %>%
  mutate(
    #subtype = if_else(type == "Hub", "Logistics Hub", "Sawmill"),
    subtype = if_else(type == "Hub", "Hub", "Hub"),
    pt_size = 1.5
  )

stor_sf <- st_as_sf(storages_plt, coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(crs_map)

cons_sf <- st_as_sf(consumers_plt, coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(crs_map)

# color range C_opp
copp_range <- range(sites_sf_clust$C_opp, na.rm = TRUE)
copp_lim   <- c(min(copp_range[1], -1), max(copp_range[2], 1))

# ── Plot ───────────────────────────────────────────────────────────────────────
ggplot() +

  # Base map
  geom_sf(data = de_states,
          fill = "grey96", colour = "grey55", linewidth = 0.35) +
  geom_sf(data = rivers_eu,
          colour = "#4292c6", linewidth = 0.45, alpha = 0.7) +

  # KUP site polygons coloured by cluster — legend suppressed via guide = "none"
  geom_sf(data      = sites_sf_clust,
          aes(fill  = C_opp),
          colour    = "grey30",
          linewidth = 0.05,
          alpha     = 0.65,
          show.legend = FALSE) +
  scale_fill_gradient2(
  name     = "Opportunity cost (€/ha)",
  low      = "#1a9641",   # green  → low / negative C_opp
  mid      = "#ffffbf",   # pale yellow → midpoint
  high     = "#d7191c",   # red    → high / positive C_opp
  midpoint = 0,           # centre the diverging scale on zero
  na.value = "grey",
  limits   = copp_lim,        
  oob      = scales::squish,  
  guide    = guide_colourbar(
    barwidth       = 0.5,
    barheight      = 4,
    title.position = "top"
  ) 
  ) +
  # scale_fill_manual(
  #   values = cluster_pal,
  #   guide  = "none"          # colours visible on map; no legend entry printed
  # ) +

  # Pre-treatment facilities — new colour + shape scale after cluster fill
  ggnewscale::new_scale_colour() +
  ggnewscale::new_scale("shape") +
  ggnewscale::new_scale_fill() +
  geom_sf(data = stor_sf,
          aes(size = pt_size, colour = subtype),
          stroke = 1.2, alpha = 0.92, show.legend = T) +
   scale_colour_manual(
    name   = "Logistics hubs",
      values = c("Hub" = COL_HUB)
    ) +
  # scale_shape_manual(
  #   name   = "Pre-treatment",
  #   values = c("Hub" = 17)
  # ) +

  # Consumer sites — fill scale registered after new_scale_fill() above
  geom_sf(data  = cons_sf,
          aes(fill = kategorie, size = pt_size),
          shape = 22, colour = "white", stroke = 0.45,
          alpha = 0.88, show.legend = TRUE) +
  scale_fill_manual(
    name   = "Consumers",
    values = c(
      "P1: stem/chemical" = COL_P1,
      "P2: branch/pulp"    = COL_P2,
      "P3: residue/energy" = COL_P3,
      "Other"                = "grey60"
    )
  ) +

  scale_size_continuous(range = c(1.5, 5), guide = "none") +
  guides(
    fill   = guide_legend(order = 1,
                          override.aes = list(size = 3.5, stroke = 0.4,
                                              colour = "white")),
  colour = guide_legend(order = 2,
                        override.aes = list(shape  = 16,      
                                            size   = 4,
                                            stroke = 0,       
                                            colour = COL_HUB)),
    shape  = guide_legend(order = 2,
                          override.aes = list(size = 4, stroke = 1.5))
  ) +

  coord_sf(
    xlim   = c(bbox_num["xmin"], bbox_num["xmax"]),
    ylim   = c(bbox_num["ymin"], bbox_num["ymax"]),
    expand = FALSE
  ) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid      = element_line(colour = "grey88", linewidth = 0.25),
    legend.position = "right",
    legend.key.size = unit(0.40, "cm"),
    legend.text     = element_text(size = 6),
    legend.title    = element_text(size = 7.5, face = "bold"),
    axis.text       = element_text(size = 7)
  )
```

The supply chain comprises `r nrow(afs_workspace$sites)` agricultural production parcels which are grouped into  $|\mathcal{I}|=$ `r nrow(afs_workspace$sites_clustered)` potential AFS production sites. Clustering is based on geographical distances between parcel centroid using a  k-means approach in order to reduce problem size. For distance calculations, for each site a centroid is calculated. Pre-treatment and intermediate storage facilities $\mathcal{J}$ are located on existing wood processing facilities as well as potential new log or wood chip yards. All harvested biomass quantities at AFS sites are transported to logistics hubs first. At the hubs sorting and transshipment processes take place before biomass types are forwarded to the consumer facilities. Storing and transshipment capacities at the hibs are assumed to be constant in time and universally set to `r round(mean(storages$CAP_stor)/1000,0)` kt per year.  Finally, $|\mathcal{K}|=$ `r nrow(consumers)` consumer sites are included. Each consumer site belongs to an existing facility from chemical, paper/wood processing industry, or the energy sector. Distances between the facilities $d_{ij}$ and $d_{jk}$ are calculated as road distances between the facilities' coordinates. For each consumer facility, an yearly upper demand capacity for each of the three distinguished biomass product types are given along with consumer-specific prices. The following biomass types are distinguished:

- **P1: stem wood – chemical/paper grade**: long-fibre wood chips from stems suitable for conversion in a local biorefinery and paper factory. In total, a yearly demand of `r round(sum(consumers$demand_P1), 0)` kt and an average prices of `r round(mean(milp_instance$consumers$P1[milp_instance$consumers$P1>0]),0)` € is assumed. 

- **P2: branch wood – sawmill/pulp grade**: medium/short-fibre wood chips from larger branches suitable for mechanical processing in sawmills and lower-grade pulping. Demand is distributed across `r sum(consumers$demand_P2 > 0)` smaller facilities and accumulates to `r round(sum(consumers$demand_P2), 0)` kt/year. Gate prices range from `r round(min(milp_instance$consumers$P2[milp_instance$consumers$P2>0]),0)`-`r round(max(milp_instance$consumers$P2[milp_instance$consumers$P2>0]),0)` €/t.

- **P3: residue – energy grade**: residual wood fractions (chips, bark, off-cuts) used for heat and electricity generation or lower-grade pellet production. Demand is accumulates among `r sum(consumers$demand_P3 > 0)` consumers to `r round(sum(consumers$demand_P3), 0)` kt/year. Gate prices range from `r round(min(milp_instance$consumers$P3[milp_instance$consumers$P3>0]),0)`-`r round(max(milp_instance$consumers$P3[milp_instance$consumers$P3>0]),0)` €/t.

\autoref{tab:tab-consumers} in the appendix summarizes the consumer-specific demand data. It is assumed that the biomass product types show a quality cascade which means that the demand of pulp grade (P2) can also be satisfied by chemical/paper grade biomass (P1). Likewise, energy grade demand (P3) can be satisfied by all biomass types. Note that constant demands over the planning horizon are assumed, although an increasing biomass demand can be expected due to the green transition in those industries. From this perspective, a rather conservative market demand scenario is assumed.

Growth model introduced in  \autoref{sec:lit-growth} is used to model biomass quantities depending on age and the area dedicated to AFS. The growth model is parameterized for poplar clones at central European conditions as introduced above with a planting density of $N=400$ trees per hectar. If AFS are established, they have to be harvested within $A^{\\min}=$ `r params_obj$min_age` to $A^{\\max}=$ `r params_obj$max_age` years allowing a wide range of potential harvesting strategies between short-rotation and long-rotation AFS. As the total planning horizon is $|\mathcal{T}|=$ `r params_obj$n_periods` years multiple harvesting cycles can be established. The sites do not differ with respect to growth conditions as the climatic and soil quality variation within the comparatively small region is rather limited. However, the sites differ with respect to size (given in hectar) and opportunity cost rates (given in € per hectar and year). Site-specific opportunity cost rates $c^{opp}_i$ is  normally distributed with mean `r round(mean(sites_clustered$C_opp),0)` and standard deviation `r round(0.75*mean(sites_clustered$C_opp),0)`. For most sites, substantial opportunity costs are assumed reflecting the loss in field-crop revenue, risk premiums of landowners as well as higher costs for field crop maintenance. The `r length(sites_clustered$C_opp<0)` sites with negative opportunity cost rates reflect sites where AFS lead to an increase in field crop productivity on the remaining field area. The variation in opportunity cost rates to study the effect of opportunity cost on site selection for AFS and gives valuable insights particularly for communication with farmers and politics. Subsidies for establishing AFS are one of the core instruments to foster AFS. The height and effectiveness of those subsidies, however, is controversially disputed. Remaining cost parameters are universally set for all sites and comprise maintenance cost rates $c^{main}_i$ (€ per year and hectar), establishment cost $c^{est}_i$ (€ per hectar) and harvesting cost $c^{harv}_i$ (€ per hectar). Likewise, transport cost rates for harvested biomasses $c^{tr-raw}_p$ and pre-treated/sorted biomasses $c^{tr-raw}_{p'}$ are set universally. All scalar parameters are shown in \autoref{tab:tab-milp-params}.

```{r tab-milp-params, echo=FALSE, message=FALSE, warning=FALSE, cache=F}
params_df <- data.frame(
  Symbol          = c("$|\\mathcal{T}|$", "$|\\mathcal{I}|$", "$|\\mathcal{J}|$",
                      "$|\\mathcal{K}|$", "$|\\mathcal{P}|$",
                      "$A^{\\min}$", "$A^{\\max}$",
                      "$c^{tr\\text{-}raw}$", "$c^{tr\\text{-}pre}$",
                      "$c^{est}_i$",
                      "$c^{harv}_i$",
                      "$c^{main}_i$",
                      "$c_i^{opp}$",
                      "$R_{P1}$", "$R_{P2}$", "$R_{P3}$"),
  Description     = c("Planning periods", "Sites",
                      "Pre-treatment / storage nodes", "Consumers", "Product grades",
                      "Minimum harvest age (years)", "Maximum stand age (years)",
                      "Raw transport cost", "Pre-processed transport cost",
                      "Establishment cost rate",
                      "harvest cost rate",
                      "Maintenance cost rate",
                      "aver. opportunity cost",
                      "Mean gate price P1", "Mean gate price P2", "Mean gate price P3"),
  Value           = c(milp_instance$n_periods, milp_instance$n_sites, nrow(afs_workspace$storages),
                      milp_instance$n_consumers, "3",
                      milp_instance$min_age, milp_instance$max_age,
                      milp_instance$c_tr_raw, milp_instance$c_tr_pre,
                      round(mean(milp_instance$sites$C_est)),
                      round(mean(milp_instance$sites$C_harv)),
                      round(mean(milp_instance$sites$C_main)),
                      round(mean(milp_instance$sites$C_opp),0),
                      round(mean(milp_instance$consumers$P1[milp_instance$consumers$P1>0]),0),
                      round(mean(milp_instance$consumers$P2[milp_instance$consumers$P2>0]),0),
                      round(mean(milp_instance$consumers$P3[milp_instance$consumers$P3>0]),0)),
  Unit            = c("yr", "–", "–", "–", "–",
                      "yr", "yr",
                      "€ t$^{-1}$ km$^{-1}$", "€ t$^{-1}$ km$^{-1}$",
                      "€ ha$^{-1}$ yr$^{-1}$",
                      "€ ha$^{-1}$ yr$^{-1}$",
                      "€ ha$^{-1}$ yr$^{-1}$",
                      "€ ha$^{-1}$ yr$^{-1}$",
                      "€ t$^{-1}$", "€ t$^{-1}$", "€ t$^{-1}$"),
  stringsAsFactors = FALSE
)

params_df %>%
  kbl(
    col.names = c("Parameter", "Description", "Base-case value", "Unit"),
    caption   = "Parameter values for the \\problem case study. Transport costs refer to road distance",
    booktabs  = TRUE,
    escape    = FALSE,
    align     = c("l", "l", "r", "l"),
    label     = "tab-milp-params"
  ) %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 8) %>%
  pack_rows("Index sets", 1, 5) %>%
  pack_rows("Biomass dynamics", 6, 7) %>%
  pack_rows("Cost parameters", 8, 13) %>%
  pack_rows("Revenue parameters", 14, 16)
```

## Optimization results

The AFS-SCD model is solved for the instance data described above with Gurobi. The model selects a subset of sites for AFS establishment and identifies the optimal harvest schedules, intermediate facility utilisation, and product allocation across the planning horizon.

\autoref{fig:fig-results-sc-map} shows the spatial configuration of the profit-optimal supply chain. 

```{r fig-results-sc-map, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Site selection and biomass production in optimal supply chain configuration. Green polygons: active sites with biomass harvest; grey polygons: non-selected sites. Pie charts at active sites show the product mix of harvested biomass, scaled by total harvested volume.", fig.width=8, fig.height=7, dev="pdf", out.width=".9\\textwidth", cache=T, cache.extra = c("ext","sites_clustered", "storages" , "customers", "milp_instance")}


crs_map <- 25833

# ── Colour palette constants ─────────────────────────────────────────────────
COL_SAW       <- "#e65c00"
COL_HUB       <- "grey40"
COL_P1        <- "#6a0dad"
COL_P2        <- "#08519c"
COL_P3        <- "#a50026"
COL_SITE_NONE <- "#bdbdbd"

pal_products <- c(
  p1 = "#1b9e77",
  p2 = "#d95f02",
  p3 = "#7570b3"
)

# ── Map extent ────────────────────────────────────────────────────────────────
bbox_wgs84 <- st_as_sfc(
  st_bbox(c(xmin = 10.6, xmax = 13.2, ymin = 50.9, ymax = 52.8), crs = st_crs(4326))
)
bbox_map <- st_transform(bbox_wgs84, crs_map)
bbox_num <- st_bbox(bbox_map)

# ── Base-map layers ───────────────────────────────────────────────────────────
de_states <- ne_states(country = "Germany", returnclass = "sf") %>%
  st_transform(crs_map) %>%
  st_intersection(bbox_map)

rivers_eu <- ne_download(
  scale = 10,
  type = "rivers_lake_centerlines",
  category = "physical",
  returnclass = "sf"
) %>%
  st_transform(crs_map) %>%
  st_intersection(bbox_map)

# ── Aggregate harvest volumes per site over all periods ──────────────────────
site_harvest <- ext$Xij %>%
  group_by(site_id, p) %>%
  summarise(vol = sum(value, na.rm = TRUE), .groups = "drop") %>%
  tidyr::pivot_wider(
    names_from   = p,
    values_from  = vol,
    values_fill  = 0,
    names_prefix = "p"
  )


for (col in c("p1", "p2", "p3")) {
  if (!col %in% names(site_harvest)) site_harvest[[col]] <- 0
}

site_harvest <- site_harvest %>%
  mutate(total_vol = p1 + p2 + p3) %>%
  filter(total_vol > 0)

active_site_ids <- unique(site_harvest$site_id)

# ── Site polygons ─────────────────────────────────────────────────────────────
sites_sf_map <- sites_sf %>%
  mutate(site_id = seq_len(nrow(sites_sf))) %>%
  inner_join(afs_workspace$site_cluster_assig, "site_id") %>%
  st_transform(crs_map) %>%
  mutate(active = hac_cluster %in% active_site_ids)

tmp.opp <- sites_clustered %>%
  select(site_id, C_opp) %>%
  rename(hac_cluster = site_id)

sites_sf_map <- sites_sf_map %>%
  left_join(tmp.opp, by = "hac_cluster")

# Split into active and inactive subsets for separate fill scales
# CHANGED: split polygons so inactive uses fixed grey fill and
#          active uses a continuous C_opp colour scale
sites_inactive <- sites_sf_map %>% filter(!active)
sites_active   <- sites_sf_map %>% filter(active)

# ── Pie-chart data ────────────────────────────────────────────────────────────
sites_pie_sf <- milp_instance$sites %>%
  inner_join(site_harvest, by = "site_id") %>%
  st_as_sf(coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(crs_map)

coords_sites <- st_coordinates(sites_pie_sf)

sites_pie <- sites_pie_sf %>%
  st_drop_geometry() %>%
  mutate(
    x = coords_sites[, 1],
    y = coords_sites[, 2]
  )

vol_med <- median(sites_pie$total_vol[sites_pie$total_vol > 0], na.rm = TRUE)
if (is.na(vol_med) || vol_med <= 0) vol_med <- 1

sites_pie <- sites_pie %>%
  mutate(r = 3500 * sqrt(total_vol / vol_med))

# ── Consumers & storages ──────────────────────────────────────────────────────
consumers_plt <- consumers %>%
  mutate(
    total_demand = demand_P1 + demand_P2 + demand_P3,
    kategorie = case_when(
      demand_P1 >= demand_P2 & demand_P1 >= demand_P3 & demand_P1 > 0 ~ "P1: stem/chemical",
      demand_P2 >= demand_P1 & demand_P2 >= demand_P3 & demand_P2 > 0 ~ "P2: branch/pulp",
      demand_P3 > 0 ~ "P3: residue/energy",
      TRUE ~ "Other"
    ),
    pt_size = pmin(4.5, pmax(1.5, 1.5 + log1p(total_demand) * 0.55))
  )

storages_plt <- storages %>%
  mutate(
    subtype = if_else(type == "Hub", "Hub", "Hub"),
    pt_size = 1.5
  )

stor_sf <- st_as_sf(storages_plt, coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(crs_map)

cons_sf <- st_as_sf(consumers_plt, coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(crs_map)

# Gemeinsamen Bereich aus ALLEN Sites berechnen (nicht nur sitesactive!)
copp_range <- range(sites_sf_map$C_opp, na.rm = TRUE)
copp_lim   <- c(min(copp_range[1], -1), max(copp_range[2], 1))

# ── Plot ──────────────────────────────────────────────────────────────────────
ggplot() +
  # Base map
  geom_sf(data = de_states,
          fill = "grey96", colour = "grey55", linewidth = 0.35) +
  geom_sf(data = rivers_eu,
          colour = "#4292c6", linewidth = 0.35, alpha = 0.7) +

  # CHANGED: Inactive site polygons — fixed grey fill, no legend scale needed
  geom_sf(data = sites_inactive,
          fill = COL_SITE_NONE,
          colour = NA, alpha = 0.55, show.legend = FALSE) +

  # CHANGED: Active site polygons — continuous fill mapped to C_opp
  geom_sf(data = sites_active,
          aes(fill = C_opp),
          colour = NA, alpha = 0.75, show.legend = TRUE) +
 scale_fill_gradient2(
  name     = "Opportunity cost (€/ha)",
  low      = "#1a9641",   # green  → low / negative C_opp
  mid      = "#ffffbf",   # pale yellow → midpoint
  high     = "#d7191c",   # red    → high / positive C_opp
  midpoint = 0,           # centre the diverging scale on zero
  limits   = copp_lim,        # <-- NEU: expliziter, identischer Bereich
  oob      = scales::squish,  # <-- NEU: Werte außerhalb → Endfarbe statt NA
  na.value = COL_SITE_NONE,
  guide    = guide_colourbar(
    barwidth       = 0.5,
    barheight      = 4,
    title.position = "top"
  ) 
  ) +

  # Pre-treatment facilities
  ggnewscale::new_scale_colour() +
  ggnewscale::new_scale("shape") +
  geom_sf(data = stor_sf,
          aes(colour = subtype, size = pt_size),
          stroke = 1.0, alpha = 0.92, show.legend = TRUE) +
  scale_colour_manual(
    name   = "Logistics hubs",
    values = c("Hub" = COL_HUB)
  ) +
  scale_shape_manual(
    name   = "Logistics hubs",
    values = c("Hub" = 16)
  ) +

  # Consumer sites
  ggnewscale::new_scale_fill() +
  geom_sf(data = cons_sf,
          aes(fill = kategorie, size = pt_size),
          shape = 22, colour = "white", stroke = 0.45,
          alpha = 0.9, show.legend = TRUE) +
  scale_fill_manual(
    name   = "Consumers",
    values = c(
     "P1: stem/chemical" = COL_P1,
      "P2: branch/pulp"    = COL_P2,
      "P3: residue/energy" = COL_P3,
      "Other"                = "grey60"
    ),
  guide = guide_legend(
    order        = 2,
    override.aes = list(
      shape  = 22,          # gefülltes Quadrat
      size   = 4,
      stroke = 0.4,
      colour = "white"      # Rand wie im geom_sf
    )
  )
  ) +

  # Pie charts at active sites
  ggnewscale::new_scale_fill() +
  scatterpie::geom_scatterpie(
    data         = sites_pie,
    aes(x = x, y = y, r = r),
    cols         = c("p1", "p2", "p3"),
    colour       = "white",
    linewidth    = 0.15,
    donut_radius = .5,
    alpha        = 0.7,
    show.legend  = TRUE
  ) +
  scale_fill_manual(
    name   = "Harvested biomass",
    values = pal_products,
    labels = c(p1 = "Stem", p2 = "Branches", p3 = "Residues"),
  guide  = guide_legend(
    order        = 3,
    override.aes = list(shape = 21, size = 4, stroke = 0.3)
  )
  ) +

  scale_size_continuous(range = c(1.5, 5), guide = "none") +
  guides(
  colour = guide_legend(
    order        = 1,
    override.aes = list(shape = 16, size = 4, stroke = 0, colour = COL_HUB)
  ),
  shape  = "none"   # shape-Skala unterdrücken (colour-Skala reicht für Hubs)
) +

  coord_sf(
    xlim   = c(bbox_num["xmin"], bbox_num["xmax"]),
    ylim   = c(bbox_num["ymin"], bbox_num["ymax"]),
    expand = FALSE
  ) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid      = element_line(colour = "grey88", linewidth = 0.25),
    legend.position = "right",
    legend.key.size = unit(0.40, "cm"),
    legend.text     = element_text(size = 6),
    legend.title    = element_text(size = 7.5, face = "bold"),
    axis.text       = element_text(size = 7)
  )
```

In total, on `r length(active_site_ids)` sites  AFS are established which corresponds to about `r round(length(active_site_ids)/milp_instance$n_sites*100,0)` % of all sites and `r round(sum(ext$z[ext$z$arc_type=="establishment","value"])/sum(milp_instance$sites$area_ha)*100,0)` % of the total farm area. Over the planning horizon, `r  round(sum(site_harvest$total_vol)/1000)` kt of AFS biomasses are produced on these sites which suffices to satisfy `r round(sum(site_harvest$total_vol)/sum(milp_instance$demand$D_max)*100,0)` % of total demand. \autoref{fig:fig-results-product-mix} shows the harvested biomasses over time along with total demand of all consumer sites. 

```{r fig-results-product-mix, echo=FALSE, message=FALSE, warning=FALSE,     fig.cap="Total biomass delivered to consumers by product type over the planning horizon. Stacked areas show period-level volumes of stem (P1), branches (P2), and residues (P3) flowing from storage hubs to all consumers. Dashed line: total harvested raw biomass shipped from sites to hubs.", fig.width=6.5, fig.height=3, dev="pdf", out.width="\\textwidth", cache=F, cache.extra = c("ext","sites_clustered", "storages" , "customers", "milp_instance")}

prod_labels <- c("1" = "Stem (P1)", "2" = "Branches (P2)", "3" = "Residues (P3)")
prod_colors <- c("1" = "#1b9e77", "2" = "#d95f02", "3" = "#7570b3")


# Deliveries aggregated by period × del_product
delivery_ts <- ext$Xjk %>%
  group_by(period, product = factor(del_product)) %>%
  summarise(volume = sum(value), .groups = "drop")

# Raw biomass shipped (Xij) aggregated by period — shown as dashed reference
shipped_ts <- ext$Xij %>%
  group_by(period) %>%
  summarise(shipped = sum(value), .groups = "drop")

# Maximum total consumer demand per product type
# Assumes milp_instance$demand is a data frame with columns: customer, product, period, volume
# Adjust column names to match your actual milp_instance structure
demand_max <- milp_instance$demand %>%
  group_by(product = factor(product, levels = c("3", "2", "1"))) %>%
  summarise(max_demand = sum(D_max)/milp_instance$n_periods, .groups = "drop") %>% 
  mutate(max_demand = cumsum(max_demand))

ggplot(delivery_ts, aes(x = period, y = volume, fill = product)) +
  geom_area(position = "stack", alpha = 0.80, colour = "white", linewidth = 0.2) +
  # Raw shipped biomass reference (dashed grey)
  geom_line(data = shipped_ts,
            aes(x = period, y = shipped),
            inherit.aes = FALSE,
            colour = "#636363", linetype = "dashed", linewidth = 0.7) +
  # Max consumer demand per product (dotted, colour-matched to product)
  geom_hline(data = demand_max,
             aes(yintercept = max_demand, colour = product),
             linetype = "dashed", linewidth = 1.0, inherit.aes = FALSE) +
  scale_fill_manual(
    name   = "Product",
    values = prod_colors,
    labels = prod_labels
  ) +
  scale_colour_manual(
    name   = "Max. demand",
    values = prod_colors,
    labels = prod_labels
  ) +
  scale_x_continuous(name = "Planning period (year)",
                     breaks = seq(1, milp_instance$n_periods, by = 2)) +
  scale_y_continuous(name = "Biomass quantity (t fresh biomass)",
                     labels = scales::comma) +
  annotate("text", x = max(shipped_ts$period) - 1,
           y = max(shipped_ts$shipped) * 0.95,
           label = "Raw shipped\n(site-hub)", size = 2.5,
           colour = "#636363", hjust = 1) +
  theme_minimal(base_size = 9) +
  theme(legend.position = "right",
        panel.grid.minor = element_blank())
```

During the first 10 years no AFS biomasses are harvested as trees of established AFS in first years grow up to a certain size before first harvest. Thus, about 25\% of consumers' total biomass demand cannot be satisfied due to this grow-up period. Afterwards, a cyclical pattern appears with a harvesting peak around year 20 and a breakdown in year 28 before plateau is reached from year 35-40. Thereby, the harvesting quantities of stems and branches show a similar pattern while quantities of harvested energy grade biomass are rather stable and close to the demand limit over the planning horizon. Thus, rather long-rotation AFS are established as \autoref{fig:fig-results-harvest-cycle-dist} confirms showing the histogram of harvesting ages across all sites.

```{r fig-results-harvest-cycle-dist, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of harvesting cycle lengths at active agroforestry sites. Each bar represents the number of individual harvest events (over all periods and active sites) whose cycle length (stand age at harvest, i.e. $t - s$) falls in that year bin.", fig.width=7, fig.height=3, dev="pdf", out.width="\\textwidth", cache=T, cache.extra = c("ext", "milp_instance")}

# --- Extract harvest-arc cycle lengths from binary z variables ---------------
# ext$z contains one row per selected arc; columns: site_id, s, t, arc_type
# Harvest arcs have arc_type == "harvest" and cycle_length = t - s

cycle_df <- ext$z %>%
  filter(arc_type == "harvest") %>%
  mutate(cycle_length = t - s)

# Restrict to active sites (at least one harvest arc selected)
active_sites <- unique(cycle_df$site_id)

# Summary statistics for annotation
mean_cl  <- mean(cycle_df$cycle_length)
med_cl   <- median(cycle_df$cycle_length)
n_events <- nrow(cycle_df)
n_sites  <- length(active_sites)


# --- Plot --------------------------------------------------------------------
ggplot(cycle_df, aes(x = cycle_length)) +
  geom_bar(fill = "#1b9e77", colour = "white", linewidth = 0.25, width = 0.75) +
  geom_vline(xintercept = mean_cl,
             linetype = "dashed", colour = "grey30", linewidth = 0.6) +
  annotate("text",
           x     = mean_cl + 0.25,
           y     = Inf,
           label = paste0("Mean = ", round(mean_cl, 1), " yr"),
           hjust = 0, vjust = 1.5, size = 2.8, colour = "grey20") +
  scale_x_continuous(
    name   = "Harvesting cycle length (years)",
    breaks = seq(params_obj$min_age, params_obj$max_age, by = 1),
    limits = c(params_obj$min_age - 0.5, params_obj$max_age + 0.5)
  ) +
  scale_y_continuous(
    name   = "Number of harvest events",
    expand = expansion(mult = c(0, 0.12))
  ) +
  labs(
    caption = paste0(
      n_events, " harvest events across ", n_sites,
      " active sites; median = ", med_cl, " yr"
    )
  ) +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    plot.caption       = element_text(size = 7, colour = "grey40", hjust = 0)
  )
```

The histogram illustrates that majority of harvest events occur at stand ages between 14--18 years, indicating that long-rotation AFS seem to be more profitable in this setting. Short-rotation cycles (<10 years) occur only in marginal frequencies. This optimal rotation length might be caused either due to the limited demand of energy-grade biomass or the price gaps between biomass types or a combination of effects. The quantity of energy-grade biomass created in long-rotation AFS suffices to cover total energy-grade biomass demand almost completely. Rotation length is focused on production of chemical-grade biomass where highest prices are assumed. Focusing on chemical industry supply requires a production of trees with a comparatively large stems. This strategy maximizes revenue, but in the same time increases opportunity costs as wood harvests appear in larger time intervals. \autoref{fig:fig-results-cost-profit-breakdown} presents the aggregate cost--profit waterfall over all periods and actors. 

```{r fig-results-cost-profit-breakdown, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Total cost–profit breakdown across all actors and years. The left panel (waterfall) decomposes the objective value into revenue and individual cost categories; the right panel shows the period-level trajectory of revenues and stacked costs.", fig.width = 6, fig.height = 3, dev = "pdf", out.width = "\\textwidth", cache = T, cache.extra = c("ext", "milp_instance", "sites_clustered", "storages", "params_obj")}

# ── 0. Colour palette (consistent with paper) ──────────────────────────────
COL_REV   <- "#1b9e77"   # revenue: green
COL_EST   <- "#d95f02"   # establishment
COL_HARV  <- "#7570b3"   # harvesting
COL_MAIN  <- "#e7298a"   # maintenance
COL_OPP   <- "#66a61e"   # opportunity
COL_TR_R  <- "#e6ab02"   # transport raw
COL_TR_P  <- "#a6761d"   # transport pre-processed
COL_STOR  <- "#666666"   # storage
COL_P1_REV <- "#1b9e77"   # P1 Stem
COL_P2_REV <- "#d95f02"   # P2 Branches
COL_P3_REV <- "#7570b3"   # P3 Residue

prod_labels <- c("1" = "P1 Stem", "2" = "P2 Branches", "3" = "P3 Residue")
prod_colors_rev <- c("1" = COL_P1_REV, "2" = COL_P2_REV, "3" = COL_P3_REV)

# ── Colour maps — revenues and costs clearly separated ───────────────────────
prod_colors_rev <- c(
  "Revenue P1 (Stem)"     = "#1b9e77",   # teal-green
  "Revenue P2 (Branches)" = "#d95f02",   # orange
  "Revenue P3 (Residue)"  = "#7570b3"    # purple
)

cost_colors <- c(
  "Establishment"         = "#b2182b",   # dark red
  "Harvesting"            = "#ef8aa2",   # salmon
  "Maintenance"           = "#fddbc7",   # pale peach
  "Opportunity"           = "#67001f",   # deep burgundy
  "Transport (raw)"       = "#2166ac",   # medium blue
  "Transport (pre-proc.)" = "#92c5de",   # light blue
  "Storage"               = "#4d4d4d"    # dark grey
)

col_map <- c(prod_colors_rev, cost_colors)


# ── 1. Site meta: area_ha and cost parameters for active sites only ─────────
active_site_ids <- ext$z %>%
  filter(arc_type == "harvest") %>%
  pull(site_id) %>% unique()

site_meta <- milp_instance$sites %>%
  mutate(site_id = row_number()) %>%
  filter(site_id %in% active_site_ids)

# ── 2. Cultivation costs (derived from arc decisions in ext$z) ──────────────
cost_est <- ext$z %>%
  filter(arc_type == "establishment") %>%
  #select(-area_ha) %>%
  left_join(site_meta, by = "site_id") %>%
  mutate(cost = C_est * value, period = t) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Establishment")

cost_harv <- ext$z %>%
  filter(arc_type == "harvest") %>%
  #select(-area_ha) %>%
  left_join(site_meta, by = "site_id") %>%
  mutate(cost = C_harv * value, period = t) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Harvesting")

cost_main <- ext$z %>%
  filter(arc_type == "harvest") %>%
  #select(-area_ha) %>%
  mutate(age_len = t-s) %>% 
  left_join(site_meta, by = "site_id") %>%
  mutate(cost = C_main * value * age_len, period = t) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Maintenance")


cost_opp <- ext$z %>%
  filter(arc_type == "harvest") %>%
  #select(-area_ha) %>%
  left_join(site_meta, by = "site_id") %>%
  mutate(age_len = t-s) %>% 
  mutate(cost = C_opp * value * age_len, period = t) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Opportunity")


# ── 3. Transport costs ───────────────────────────────────────────────────────
dist_ij_m <- as.matrix(milp_instance$d_ij)

cost_tr_raw <- ext$Xij %>%
  mutate(
    dist_km = mapply(function(i, j) dist_ij_m[i, j], site_id, hub_id),
    cost    = params_obj$c_tr_raw * dist_km * value
  ) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Transport (raw)")

dist_jk_m <- as.matrix(milp_instance$d_jk)

cost_tr_pre <- ext$Xjk %>%
  mutate(
    dist_km = mapply(function(j, k) dist_jk_m[j, k], hub_id, consumer_id),
    cost    = params_obj$c_tr_pre * dist_km * value
  ) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Transport (pre-proc.)")

# ── 4. Storage costs ─────────────────────────────────────────────────────────
cost_stor <- ext$S %>%
  left_join(
    milp_instance$storages %>%
      mutate(hub_id = row_number()) %>%
      select(hub_id, c_stor),
    by = "hub_id"
  ) %>%
  mutate(cost = c_stor * value) %>%
  group_by(period) %>%
  summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Storage")

# ── 5. Revenue per period ─────────────────────────────────────────────────────
price_df <- as.data.frame(milp_instance$consumer_prices) %>%
  setNames(c("consumer_id", "del_product", "price")) %>%
  mutate(consumer_id = as.integer(consumer_id),
         del_product  = as.integer(del_product))

revenue_by_product_ts <- ext$Xjk %>%
  left_join(price_df, by = c("consumer_id", "del_product")) %>%
  mutate(rev = value * coalesce(price, 0)) %>%
  group_by(period, del_product) %>%
  summarise(rev = sum(rev, na.rm = TRUE), .groups = "drop") %>%
  mutate(del_product = as.character(del_product))

# Total revenue per period (for cumulative cashflow calc)
revenue_ts <- revenue_by_product_ts %>%
  group_by(period) %>%
  summarise(rev = sum(rev), .groups = "drop")

# ── 6. Assemble long-format time series (unchanged) ──────────────────────────
all_periods <- tibble(period = seq_len(milp_instance$n_periods))

cost_ts <- bind_rows(
  cost_est, cost_harv, cost_main, cost_opp,
  cost_tr_raw, cost_tr_pre, cost_stor
) %>%
  right_join(crossing(period    = seq_len(milp_instance$n_periods),
                      component = unique(.$component)),
             by = c("period", "component")) %>%
  replace_na(list(cost = 0)) %>%
  mutate(component = factor(component, levels = c(
    "Establishment", "Harvesting", "Maintenance", "Opportunity",
    "Transport (raw)", "Transport (pre-proc.)", "Storage"
  )))

rev_ts <- all_periods %>%
  left_join(revenue_ts, by = "period") %>%
  replace_na(list(rev = 0))

# ── 7. Waterfall totals — Revenue split into 3 product bars ─────────────────
rev_totals_by_product <- revenue_by_product_ts %>%
  group_by(del_product) %>%
  summarise(value = sum(rev), .groups = "drop") %>%
  mutate(
    component = recode(del_product,
      "1" = "Revenue P1 (Stem)",
      "2" = "Revenue P2 (Branches)",
      "3" = "Revenue P3 (Residue)"
    )
  ) %>%
  select(component, value)

cost_totals <- cost_ts %>%
  group_by(component) %>%
  summarise(value = -sum(cost), .groups = "drop")

totals <- bind_rows(rev_totals_by_product, cost_totals) %>%
  mutate(component = factor(component, levels = c(
    "Revenue P1 (Stem)", "Revenue P2 (Branches)", "Revenue P3 (Residue)",
    "Establishment", "Harvesting", "Maintenance", "Opportunity",
    "Transport (raw)", "Transport (pre-proc.)", "Storage"
  ))) %>%
  arrange(component) %>%
  mutate(
    end      = cumsum(value),
    start    = lag(end, default = 0),
    bar_low  = pmin(start, end),
    bar_high = pmax(start, end)
  )

col_map <- c(
  "Revenue P1 (Stem)"      = COL_P1_REV,
  "Revenue P2 (Branches)"  = COL_P2_REV,
  "Revenue P3 (Residue)"   = COL_P3_REV,
  "Establishment"          = COL_EST,
  "Harvesting"             = COL_HARV,
  "Maintenance"            = COL_MAIN,
  "Opportunity"            = COL_OPP,
  "Transport (raw)"        = COL_TR_R,
  "Transport (pre-proc.)"  = COL_TR_P,
  "Storage"                = COL_STOR
)

# Revenue rows: one row per product, all starting from 0 (stacked bar)
rev_waterfall <- rev_totals_by_product %>%
  arrange(component) %>%                     # consistent stacking order
  mutate(
    x_pos    = 1L,                             # all revenue in column 1
    bar_low  = cumsum(lag(value, default = 0)),
    bar_high = cumsum(value),
    component = factor(component, levels = c(
      "Revenue P1 (Stem)", "Revenue P2 (Branches)", "Revenue P3 (Residue)"
    ))
  )

# Cost rows: floating waterfall segments starting where previous bar ended
# Revenue total = anchor for cost cascade
rev_total <- sum(rev_totals_by_product$value)

cost_waterfall <- cost_totals %>%             # negative values
  arrange(component) %>%
  mutate(
    x_pos    = row_number() + 1L,             # columns 2, 3, 4 …
    end      = rev_total + cumsum(value),
    start    = lag(end, default = rev_total),
    bar_low  = pmin(start, end),
    bar_high = pmax(start, end)
  )

# X-axis labels: "Revenue" for col 1, then cost component names
x_labels <- c("Revenue\n(stacked)", as.character(cost_waterfall$component))

# ── Add a grouping column so we can split the legend ─────────────────────────
rev_waterfall <- rev_waterfall %>%
  mutate(legend_group = "Revenue by product")

cost_waterfall <- cost_waterfall %>%
  mutate(legend_group = "Costs")

# ── 8. Right panel: Cumulative cash flow — stacked revenue by product ────────
rev_product_wide <- revenue_by_product_ts %>%
  right_join(crossing(period = seq_len(milp_instance$n_periods),
                      del_product = c("1", "2", "3")),
             by = c("period", "del_product")) %>%
  replace_na(list(rev = 0)) %>%
  mutate(del_product = factor(del_product, levels = c("3", "2", "1"),
                               labels = c("P3 Residue", "P2 Branches", "P1 Stem")))

total_cost_ts <- cost_ts %>%
  group_by(period) %>%
  summarise(total_cost = sum(cost, na.rm = TRUE), .groups = "drop")

cashflow_ts <- all_periods %>%
  left_join(rev_ts, by = "period") %>%
  left_join(total_cost_ts, by = "period") %>%
  replace_na(list(rev = 0, total_cost = 0)) %>%
  mutate(
    net_period = rev - total_cost,
    cum_cost   = cumsum(total_cost),
    cum_net    = cumsum(net_period)
  )

breakeven_period <- cashflow_ts %>%
  filter(cum_net >= 0) %>%
  slice_min(period, n = 1) %>%
  pull(period)


# ── Compute shared y range ────────────────────────────────────────────────────
y_min <- min(
  min(cost_waterfall$bar_low  / 1e6, na.rm = TRUE),
  min(cashflow_ts$cum_net     / 1e6, na.rm = TRUE),
  0
)
y_max <- max(
  max(rev_waterfall$bar_high  / 1e6, na.rm = TRUE),
  max(cashflow_ts$cum_rev_prod/ 1e6, na.rm = TRUE),   # top of stacked revenue area
  0
)

# add 5% padding
y_range <- c(y_min * 1.05, y_max * 1.05)


p_waterfall <- ggplot() +

  # — stacked revenue bar —
  geom_rect(
    data = rev_waterfall,
    aes(xmin = x_pos - 0.38, xmax = x_pos + 0.38,
        ymin = bar_low  / 1e6,
        ymax = bar_high / 1e6,
        fill = component),
    colour = "white", linewidth = 0.25
  ) +

  # — floating cost bars —
  geom_rect(
    data = cost_waterfall,
    aes(xmin = x_pos - 0.38, xmax = x_pos + 0.38,
        ymin = bar_low  / 1e6,
        ymax = bar_high / 1e6,
        fill = component),
    colour = "white", linewidth = 0.25
  ) +

  # — connector lines —
  geom_segment(
    data = cost_waterfall,
    aes(x    = x_pos - 0.5,   xend = x_pos - 0.38,
        y    = bar_low / 1e6, yend = bar_low / 1e6),
    colour = "grey60", linewidth = 0.3, linetype = "dotted",
    inherit.aes = FALSE
  ) +

  geom_hline(yintercept = 0, colour = "grey40", linewidth = 0.35) +

  # — split legend via two scale calls using ggnewscale —
  scale_fill_manual(
    name   = "Revenue by product",
    values = prod_colors_rev,
    breaks = names(prod_colors_rev),
    guide  = guide_legend(
      order     = 1,
      ncol      = 1,
      keywidth  = unit(0.35, "cm"),
      keyheight = unit(0.35, "cm")
    )
  ) +

  ggnewscale::new_scale_fill() +

  geom_rect(                                   # dummy layer to register cost fill
    data = cost_waterfall,
    aes(xmin = x_pos - 0.38, xmax = x_pos + 0.38,
        ymin = bar_low  / 1e6,
        ymax = bar_high / 1e6,
        fill = component),
    colour = "white", linewidth = 0.25
  ) +

  scale_fill_manual(
    name   = "Costs",
    values = cost_colors,
    breaks = names(cost_colors),
    guide  = guide_legend(
      order     = 2,
      ncol      = 1,
      keywidth  = unit(0.35, "cm"),
      keyheight = unit(0.35, "cm")
    )
  ) +

  scale_x_continuous(
    breaks = seq_along(x_labels),
    labels = x_labels,
    expand = expansion(add = 0.6)
  ) +

  scale_y_continuous(
    name   = "Cumulative value (M€)",
    labels = scales::label_number(suffix = " M"),
    limits = y_range
  ) +

  labs(x = NULL, title = "Waterfall") +

  theme_minimal(base_size = 8) +
  theme(
    panel.grid.major.x  = element_blank(),
    panel.grid.minor    = element_blank(),
    axis.text.x         = element_blank(),# element_text(size = 6.5, angle = 30, hjust = 1),
    plot.title          = element_text(size = 8, face = "bold", hjust = 0.5),
    legend.position     = "right",
    legend.text         = element_text(size = 6),
    legend.title        = element_text(size = 6.5, face = "bold"),
    legend.spacing.y    = unit(0.4, "cm")   # gap between the two legend groups
  )

# 2nd plot -----------------------------------------------------------------------

p_cashflow <- ggplot() +
  # Stacked area of cumulative revenue by product type
  geom_area(
    data = rev_product_wide %>%
      arrange(period, del_product) %>%
      group_by(del_product) %>%
      mutate(cum_rev_prod = cumsum(rev)) %>%
      ungroup(),
    aes(x = period, y = cum_rev_prod / 1e6, fill = del_product),
    position = "stack", alpha = 0.75, colour = "white", linewidth = 0.15
  ) +
  scale_fill_manual(
    name   = "Revenue by product",
    values = c("P1 Stem"     = COL_P1_REV,
               "P2 Branches" = COL_P2_REV,
               "P3 Residue"  = COL_P3_REV)
  ) +
  # Cumulative cost line
  geom_line(data = cashflow_ts,
            aes(x = period, y = cum_cost / 1e6),
            colour = "grey30", linewidth = 0.9, linetype = "dashed",
            inherit.aes = FALSE) +
  # Cumulative net cash flow
  geom_line(data = cashflow_ts,
            aes(x = period, y = cum_net / 1e6),
            colour = "grey10", linewidth = 0.65, linetype = "dotted",
            inherit.aes = FALSE) +
  geom_hline(yintercept = 0, colour = "grey50", linewidth = 0.35) +
  {if (length(breakeven_period) == 1)
    list(
      geom_vline(xintercept = breakeven_period,
                 colour = "grey40", linewidth = 0.45, linetype = "dotdash"),
      annotate("text",
               x = breakeven_period + 0.4, y = 0,
               label = paste0("BEP: t=", breakeven_period),
               size = 2.4, hjust = 0, vjust = -0.4,
               colour = "grey30", fontface = "italic")
    )
  } +
  annotate("text",
           x = milp_instance$n_periods - 0.3,
           y = last(cashflow_ts$cum_cost) / 1e6,
           label = "Cum. cost", size = 2.4, hjust = 1, vjust = 1.3,
           colour = "grey30", fontface = "italic") +
  annotate("text",
           x = milp_instance$n_periods - 0.3,
           y = last(cashflow_ts$cum_net) / 1e6,
           label = "Cum. net CF", size = 2.4, hjust = 1, vjust = -0.4,
           colour = "grey10", fontface = "italic") +
  scale_x_continuous(name   = "Planning period (year)",
                     breaks = seq(1, milp_instance$n_periods, by = 2)) +
  scale_y_continuous(name   = "Cumulative value (M€)",
                     labels = scales::label_number(suffix = " M"),
                     limits = y_range) +
  labs(title = "Cumulative cash flow by product") +
  theme_minimal(base_size = 8) +
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(size = 8, face = "bold", hjust = 0.5),
        legend.position = "none",
        #legend.key.size = unit(0.35, "cm"),
        #legend.text = element_text(size = 6.5)
        )

(p_waterfall + p_cashflow +
  plot_layout(
    widths  = c(1, 1.6),            # waterfall narrower, cashflow wider
    guides  = "collect"                # keep separate legends per panel
  ) #+
  #plot_annotation(
    #title   = "Cost–profit breakdown",
    #caption = "Left: cumulative waterfall; Right: period-by-period cash flow by product type.",
    #theme   = theme(
    #  plot.title   = element_text(size = 9, face = "bold", hjust = 0.5),
    #  plot.caption = element_text(size = 6, colour = "grey50", hjust = 0)
    #)
  #)
)

```

It appears that revenue is dominated by selling chemical-grade biomass and dominating cost component are opportunity cost although active sites are those sites with a comparatively small opportunity cost rate. Next to opportunity cost, establishment and transport costs are the most relevant cost drivers. Analyzing total cashflow over time shows that break-even is reached after about 20 years, i.e. after the first large-scale harvest occurred. Afterwards, cashflow is continuously increasing indicating that establishing AFS is a strategic change in agricultural practice and pays only if this strategy is implemented and maintained over a comparatively long time spans. This underpins the importance of strategic buying contracts with industrial partners in order to reduce revenue risks of farmers when producing AFS biomass. 

Although profitability of the total AFS suply chain is given for the case study data, profitability of AFS operations on site-level are more heterogeneous. \autoref{fig:fig-results-site-breakdown} shows revenues and costs split by component and normalized to AFS size and active years of AFS.

```{r fig-results-site-breakdown, echo=FALSE, message=FALSE, warning=FALSE,     fig.cap="Revenues, costs and net profit for sites with AFS operations split by type and expressed in € per hectare per year of AFS operation.",   fig.width = 6, fig.height = 3, dev = "pdf", out.width = "\\textwidth",    cache = T, cache.extra = c("ext", "milp_instance", "sites_clustered", "params_obj")}

# ── 0. Colour palette — identical to fig-results-cost-profit-breakdown ───────
prod_colors_rev <- c(
  "Revenue P1 (Stem)"     = "#1b9e77",
  "Revenue P2 (Branches)" = "#d95f02",
  "Revenue P3 (Residue)"  = "#7570b3"
)
cost_colors <- c(
  "Establishment"         = "#b2182b",
  "Harvesting"            = "#ef8a62",
  "Maintenance"           = "#fddbc7",
  "Opportunity"           = "#67001f",
  "Transport (raw)"       = "#2166ac",
  "Transport (pre-proc.)" = "#92c5de",
  "Storage"               = "#4d4d4d"
)

# ── 1. Site meta ─────────────────────────────────────────────────────────────
active_site_ids_sb <- ext$z %>%
  filter(arc_type == "harvest") %>%
  pull(site_id) %>% unique()

site_meta_sb <- milp_instance$sites %>%
  mutate(site_id = row_number()) %>%
  filter(site_id %in% active_site_ids_sb)

# ── 2. Hub-site share ────────────────────────────────────────────────────────
hub_site_share_sb <- ext$Xij %>%
  group_by(hub_id, period) %>%
  mutate(hub_total = sum(value)) %>%
  ungroup() %>%
  mutate(share = ifelse(hub_total > 0, value / hub_total, 0)) %>%
  select(site_id, hub_id, period, share)

# ── 3. Price table ───────────────────────────────────────────────────────────
price_df_sb <- as.data.frame(milp_instance$consumer_prices) %>%
  setNames(c("consumer_id", "del_product", "price")) %>%
  mutate(consumer_id = as.integer(consumer_id),
         del_product  = as.integer(del_product))

# ── 4. Normalisation denominator: area_afs × active_years (ha·yr) ────────────
active_p_sb <- ext$z %>%
  filter(arc_type == "establishment") %>%
  group_by(site_id) %>%
  summarise(tot_value = sum(value), .groups = "drop")

est_p_sb <- ext$z %>%
  filter(arc_type == "establishment") %>%
  group_by(site_id) %>%
  summarise(t_est = sum(t * value) / sum(value), .groups = "drop")

term_p_sb <- ext$z %>%
  filter(arc_type == "termination") %>%
  group_by(site_id) %>%
  summarise(t_term = sum(s * value) / sum(value), .groups = "drop")

site_denom_sb <- site_meta_sb %>%
  left_join(active_p_sb, by = "site_id") %>%
  left_join(est_p_sb,    by = "site_id") %>%
  left_join(term_p_sb,   by = "site_id") %>%
  mutate(
    t_term       = ifelse(is.na(t_term), milp_instance$n_periods, t_term),
    active_years = pmax(t_term - t_est, 1L),
    denom        = tot_value * active_years   # ha·yr
  ) %>%
  select(site_id, denom)

# ── 5. Revenue per site × product type, normalised to €/ha/yr ────────────────
site_rev_by_product <- ext$Xjk %>%
  left_join(price_df_sb, by = c("consumer_id", "del_product")) %>%
  mutate(rev_jkt = value * coalesce(price, 0)) %>%
  left_join(hub_site_share_sb, by = c("hub_id", "period"),
            relationship = "many-to-many") %>%
  mutate(site_rev = rev_jkt * share) %>%
  filter(!is.na(site_id)) %>%
  group_by(site_id, del_product) %>%
  summarise(value_raw = sum(site_rev, na.rm = TRUE), .groups = "drop") %>%
  left_join(site_denom_sb, by = "site_id") %>%
  mutate(
    value     = value_raw / denom,
    component = recode(as.character(del_product),
      "1" = "Revenue P1 (Stem)",
      "2" = "Revenue P2 (Branches)",
      "3" = "Revenue P3 (Residue)"
    )
  ) %>%
  select(site_id, component, value)

# ── 6. Costs per site × component, normalised to €/ha/yr (negative) ──────────
dist_ij_m_sb <- as.matrix(milp_instance$d_ij)
dist_jk_m_sb <- as.matrix(milp_instance$d_jk)

normalise_cost <- function(df, denom_df) {
  df %>%
    left_join(denom_df, by = "site_id") %>%
    mutate(value = value / denom) %>%
    select(site_id, component, value)
}

sc_est_sb <- ext$z %>% filter(arc_type == "establishment") %>%
  left_join(site_meta_sb, by = "site_id") %>%
  mutate(cost = C_est * value) %>%
  group_by(site_id) %>%
  summarise(value = -sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Establishment") %>%
  normalise_cost(site_denom_sb)

sc_harv_sb <- ext$z %>% filter(arc_type == "harvest") %>%
  left_join(site_meta_sb, by = "site_id") %>%
  mutate(cost = C_harv * value) %>%
  group_by(site_id) %>%
  summarise(value = -sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Harvesting") %>%
  normalise_cost(site_denom_sb)

sc_main_sb <- ext$z %>% filter(arc_type == "harvest") %>%
  mutate(age_len = t - s) %>%
  left_join(site_meta_sb, by = "site_id") %>%
  mutate(cost = C_main * value * age_len) %>%
  group_by(site_id) %>%
  summarise(value = -sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Maintenance") %>%
  normalise_cost(site_denom_sb)

sc_opp_sb <- ext$z %>% filter(arc_type == "harvest") %>%
  mutate(age_len = t - s) %>%
  left_join(site_meta_sb, by = "site_id") %>%
  mutate(cost = C_opp * value * age_len) %>%
  group_by(site_id) %>%
  summarise(value = -sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Opportunity") %>%
  normalise_cost(site_denom_sb)

sc_tr_raw_sb <- ext$Xij %>%
  mutate(dist_km = mapply(function(i, j) dist_ij_m_sb[i, j], site_id, hub_id),
         cost    = milp_instance$c_tr_raw * dist_km * value) %>%
  group_by(site_id) %>%
  summarise(value = -sum(cost, na.rm = TRUE), .groups = "drop") %>%
  mutate(component = "Transport (raw)") %>%
  normalise_cost(site_denom_sb)

sc_tr_pre_sb <- ext$Xjk %>%
  mutate(dist_km  = mapply(function(j, k) dist_jk_m_sb[j, k], hub_id, consumer_id),
         cost_jkt = milp_instance$c_tr_pre * dist_km * value) %>%
  left_join(hub_site_share_sb, by = c("hub_id", "period"),
            relationship = "many-to-many") %>%
  mutate(site_cost = cost_jkt * share) %>%
  group_by(site_id) %>%
  summarise(value = -sum(site_cost, na.rm = TRUE), .groups = "drop") %>%
  filter(!is.na(site_id)) %>%
  mutate(component = "Transport (pre-proc.)") %>%
  normalise_cost(site_denom_sb)

sc_stor_sb <- ext$S %>%
  left_join(milp_instance$storages %>%
              mutate(hub_id = row_number()) %>%
              select(hub_id, c_stor),
            by = "hub_id") %>%
  mutate(stor_cost = c_stor * value) %>%
  select(hub_id, period, stor_cost) %>%
  left_join(hub_site_share_sb, by = c("hub_id", "period"),
            relationship = "many-to-many") %>%
  mutate(site_cost = stor_cost * share) %>%
  group_by(site_id) %>%
  summarise(value = -sum(site_cost, na.rm = TRUE), .groups = "drop") %>%
  filter(!is.na(site_id)) %>%
  mutate(component = "Storage") %>%
  normalise_cost(site_denom_sb)

# ── 7. Combine & order sites by descending net profit/ha/yr ──────────────────
site_breakdown_df <- bind_rows(
  site_rev_by_product,
  sc_est_sb, sc_harv_sb, sc_main_sb, sc_opp_sb,
  sc_tr_raw_sb, sc_tr_pre_sb, sc_stor_sb
)

site_order <- site_breakdown_df %>%
  group_by(site_id) %>%
  summarise(profit_ha_yr = sum(value, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(profit_ha_yr)) %>%
  pull(site_id)

# ── 7b. Net profit per site (for point overlay) ───────────────────────────────
site_profit_df <- site_breakdown_df %>%
  group_by(site_id) %>%
  summarise(net = sum(value, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    site_id  = factor(site_id, levels = site_order),
    label    = "Net profit"
  )

site_profit_df_real <- site_breakdown_df  %>% 
  group_by(site_id) %>% 
  filter(component %in% c("Revenue P1 (Stem)","Revenue P2 (Branches)", "Revenue P3 (Residue)","Opportunity") ) %>% 
  summarize(net = sum(value))

# ── 8. Factor levels ─────────────────────────────────────────────────────────
component_levels <- c(
  "Revenue P1 (Stem)", "Revenue P2 (Branches)", "Revenue P3 (Residue)",
  "Establishment", "Harvesting", "Maintenance", "Opportunity",
  "Transport (raw)", "Transport (pre-proc.)", "Storage"
)

site_breakdown_df <- site_breakdown_df %>%
  mutate(
    site_id   = factor(site_id, levels = site_order),
    component = factor(component, levels = component_levels)
  )

# ── 9. Plot ───────────────────────────────────────────────────────────────────
ggplot() +

  # Revenue stacked bars (positive, €/ha/yr)
  geom_col(
    data = site_breakdown_df %>%
      filter(component %in% names(prod_colors_rev)),
    aes(x = site_id, y = value, fill = component),
    position = "stack", width = 0.7,
    colour = "white", linewidth = 0.15
  ) +
  scale_fill_manual(
    name   = "Revenue by product",
    values = prod_colors_rev,
    breaks = names(prod_colors_rev),
    guide  = guide_legend(
      order     = 1,
      ncol      = 1,
      keywidth  = unit(0.35, "cm"),
      keyheight = unit(0.35, "cm")
    )
  ) +

  ggnewscale::new_scale_fill() +

  # Cost stacked bars (negative, €/ha/yr)
  geom_col(
    data = site_breakdown_df %>%
      filter(component %in% names(cost_colors)),
    aes(x = site_id, y = value, fill = component),
    position = "stack", width = 0.7,
    colour = "white", linewidth = 0.15
  ) +
  scale_fill_manual(
    name   = "Costs",
    values = cost_colors,
    breaks = names(cost_colors),
    guide  = guide_legend(
      order     = 2,
      ncol      = 1,
      keywidth  = unit(0.35, "cm"),
      keyheight = unit(0.35, "cm")
    )
  ) +

  geom_hline(yintercept = 0, colour = "grey40", linewidth = 0.35) +

  # ── Net profit points with legend entry via colour aesthetic ─────────────
  ggnewscale::new_scale_colour() +

  geom_point(
    data = site_profit_df,
    aes(x = site_id, y = net, colour = label),
    shape = 18, size = 1.8
  ) +
  scale_colour_manual(
    name   = NULL,
    values = c("Net profit" = "grey15"),
    guide  = guide_legend(
      order     = 3,
      ncol      = 1,
      keywidth  = unit(0.35, "cm"),
      keyheight = unit(0.35, "cm"),
      override.aes = list(shape = 18, size = 1.8)
    )
  ) +

  scale_y_continuous(
    name   = "€ per hectare and year",
    labels = scales::label_number(big.mark = ",")
  ) +
  labs(x = "Site ID") +
  theme_minimal(base_size = 8) +
  theme(
    axis.text.x        = element_text(size = 5.5, angle = 60, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    legend.position    = "right",
    legend.key.size    = unit(0.35, "cm"),
    legend.text        = element_text(size = 6),
    legend.title       = element_text(size = 6.5, face = "bold"),
    legend.spacing.y   = unit(0.4, "cm")
  )
```

It appears that net profitability of AFS ranges between `r round(max(site_profit_df$net),0)` and `r round(min(site_profit_df$net),0)` € per hectare and active year. Note that this value constitutes additional profit after subtracting operational cost as well as imputed opportunity cost of lost profit from field crop. I.e., for calculating the realized net margin, opportunity can be added such that realized yearly net profit ranges between `r round(max(site_profit_df_real$net),0)` and `r round(min(site_profit_df_real$net),0)` € per hectare. Still, relative net profit is comparatively low compared to ordinary agricultural practices with profitable field crops (like winter wheat). The revenue splits of active sites indicates that at some sites medium-grade and low-grade biomass types contribute substantially or even solely to total revenue (see e.g. site 12, 135, 145, and 112). This indicates that site-specific AFS strategies might differ from the overall supply chain strategy. In other words, there seem to be strategic niches w.r.t. the consumer segments to be supplied. This might also be reflected in the rotation length and, although not discussed here, the selection of trees and clones for the AFS. Moreover, profitability of AFS sites seems to depend on various factors like local opportunity costs as well as closeness to consumers. \autoref{fig:fig-results-site-scatter} shows net margins and opportunity cost rates of sites with active AFS. 

```{r fig-results-site-scatter, echo=FALSE, message=FALSE, warning=FALSE,     fig.cap="Scatterplot of opportunity cost rate and net margin (both in € per ha and year) for each site with active AFS. Size of AFS reflected in point size.", fig.width = 7, fig.height = 3, dev = "pdf", out.width = "\\textwidth", cache = T, cache.extra = c("ext", "milp_instance", "params_obj")}


ggplot(df_bench, aes(x = opp_cost_site, y = profit_ha_yr)) +
  geom_point(aes(size = area_afs),
             colour = "grey30", fill = "steelblue", alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, colour = "firebrick", linewidth = 0.7) +
  scale_shape_manual(values = c(`TRUE` = 21, `FALSE` = 1),
                     guide = "none") +
  scale_size_area(
     name     = "AFS size (ha)",
     max_size = 8,
     breaks   = scales::breaks_extended(n = 8),
     labels   = scales::label_number(big.mark = ",")
   ) +
  #scale_size_continuous(name = "AFS size (ha)", range = c(1, 8)) +
  labs(
    x = "Opportunity cost (€ per hectare and year)",
    y = "| Net margin |\n(€ per hectare and year)"
  ) +
  theme_minimal(base_size = 9) +
  theme(legend.position  = "right",
        panel.grid.minor = element_blank())
```

As expected net margins of AFS depend heavily on a site's opportunity cost rate showing a clear negative trend. Moreover, it appears that AFS sizes vary quite heavily between `r round(min(df_bench$area_afs))` and `r round(max(df_bench$area_afs))` hectare. Probably very small sites won't operate efficiently because cost for establishment, harvesting, and maintenance are assumed to scale with size, but are rather fixed below a certain critical threshold.

# Sensitivity analysis

To assess the robustness of the baseline results and to identify the
principal drivers of supply chain profitability, a factorial sensitivity
analysis is conducted. The analysis follows a full-factorial design over
three scenario dimensions: opportunity cost level
(\euro\ ha$^{-1}$ yr$^{-1}$), a cost multiplier applied to all variable
cost parameters (establishment, harvest, transport, and storage), and a
revenue multiplier applied to all gate prices. The resulting
$2 \times 2 \times 2 = 8$ scenarios are solved independently in
addition to the baseline, yielding nine comparable instances
(Table \@ref(tab:tab-scenario-kpis)).

Table \@ref(tab:tab-scenario-kpis) summarises the key performance
indicators across all scenarios. Total profit per hectare per year is
most sensitive to the revenue multiplier: elevated gate prices
(revenue multiplier = 1.5) substantially improve profitability, while
reduced prices (multiplier = 0.5) push marginal sites into loss
territory. The supply chain remains profitable in all scenarios with
full or elevated revenue even when opportunity costs are high. Scenarios
combining reduced revenue (multiplier = 0.5) with high opportunity costs
consistently yield negative net profits, indicating that the economic
viability of AFS-based supply chains is critically conditional on
achieving sufficient product-quality premiums from P1 offtakers.
The number of active sites and total wood throughput respond
monotonically to the revenue multiplier, confirming that higher gate
prices incentivise the activation of marginal sites.

The OLS regression reported in Figure \@ref(fig:fig-effects-ols) and
Table \@ref(tab:tab-lm-profit) decomposes profit variation across
scenarios and sites into effects of four explanatory variables:
opportunity cost rate, cost level, revenue level, and log-transformed
site area. The revenue multiplier carries the largest absolute
coefficient, followed by opportunity cost rate; site area exerts a
positive but diminishing marginal effect consistent with economies of
scale in logistics and harvesting. Rotation length is positively
associated with profit in low-opportunity-cost settings but negatively
so when foregone agricultural revenue is high, confirming the
rotation-length trade-off in the AFS-SCD formulation
(Section \@ref(sec:milp)).

```{r lm-site-profit-sensitivity, echo=FALSE, message=FALSE, warning=FALSE, cache=F}

# ── 0. Prepare modelling data ─────────────────────────────────────────────────
# Combine benchmark and sensitivity runs; add normalised scenario-level area
df_model <- bind_rows(df_bench , df_sens) %>%
  # site-level area as share of its super-site (area_ha already in extract_site_profit output)
  mutate(
    # log-transform skewed continuous predictors for better GAM smoothing
    log_area_afs          = log1p(area_afs),
    log_dist_hub         = log1p(avg_dist_hub_km),
    log_dist_consumer    = log1p(avg_dist_consumer_km),
    # ensure numeric scenario parameters are present (from updated extract_site_profit)
    opp_cost             = as.numeric(opp_cost),
    cost_log_level       = as.numeric(cost_log_level),
    cost_est_level       = as.numeric(cost_est_level),
    revenue_level        = as.numeric(revenue_level)
  ) %>%
  filter(
    is.finite(profit_ha_yr),
    is.finite(avg_dist_hub_km),
    is.finite(avg_rotation_yr),
    area_ha > 0
  )

# ── 1. GAM: relative profit per site (€/ha/yr) ───────────────────────────────
# Smooth terms for all scenario parameters and key structural site variables.
# cost is now split into cost_log_level (logistics) and cost_est_level (establishment).
# te() tensor-product interaction captures the joint logistics × establishment surface.

gam_profit <- mgcv::gam(
  profit_ha_yr ~
    s(opp_cost_site,        k = 5, bs = "cr") +   # opportunity cost scaler (relative)
    cost_log_level +   # logistics cost scaler
    cost_est_level +   # establishment cost scaler
    revenue_level +   # revenue scaler
    s(avg_dist_hub_km, k = 5, bs = "cr") +   # site → hub distance
    s(avg_rotation_yr, k = 5, bs = "cr") +   # mean harvest cycle length
    s(log_area_afs,     k = 4, bs = "cr"),    # site size (log-scale)
  data   = df_model,
  method = "REML",
  family = gaussian()
)

# ── 2. GAM: total AFS area dedicated per scenario replicate ──────────────────
# Aggregate to scenario level: sum active area across all sites per scenario run.
# Each row in df_model is one site × scenario; sum area_ha to scenario level,
# join back the (constant-within-scenario) parameter values.

df_scen_area <- df_model %>%
  group_by(scenario, opp_cost, cost_log_level, cost_est_level, revenue_level) %>%
  summarise(
    total_area_afs  = sum(area_afs, na.rm = TRUE),  # weighted by activity
    n_active_sites = n_distinct(site_id),
    ave_opp_cost  = mean(opp_cost_site),
    .groups        = "drop"
  ) %>%
  filter(is.finite(total_area_afs), total_area_afs > 0) %>%
  mutate(log_total_area = log1p(total_area_afs))

# GAM on scenario-level total AFS area.
# A Gamma(link="log") family is appropriate: area is strictly positive and
# right-skewed across scenarios.
gam_area <- mgcv::gam(
  total_area_afs ~
    #s(ave_opp_cost,        k = 2, bs = "cr") +
    s(opp_cost, k = 2, bs = "cr") +
    s(cost_log_level, k = 2, bs = "cr") +
    s(cost_est_level, k = 2, bs = "cr") +
    #te(cost_log_level, cost_est_level, k = 4) +
    s(revenue_level, k = 2, bs = "cr"),
  data   = df_scen_area,
  method = "REML",
  family = Gamma(link = "log")
)

# ── 3. Model summaries ────────────────────────────────────────────────────────
# Profit GAM
gam_profit_summary <- broom::tidy(gam_profit, parametric = FALSE) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  rename(Term = term, `edf` = edf, `Ref.df` = ref.df,
         `F statistic` = statistic, `p-value` = p.value)

# Area GAM
gam_area_summary <- broom::tidy(gam_area, parametric = FALSE) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  rename(Term = term, `edf` = edf, `Ref.df` = ref.df,
         `F statistic` = statistic, `p-value` = p.value)

# ── 4. R² and model fit ───────────────────────────────────────────────────────
r2_profit <- summary(gam_profit)$r.sq
r2_area   <- summary(gam_area)$r.sq
dev_profit <- round(summary(gam_profit)$dev.expl * 100, 1)
dev_area   <- round(summary(gam_area)$dev.expl   * 100, 1)

# ── 5. Smooth effects plots (gratia) ─────────────────────────────────────────
p_profit_smooths <- gratia::draw(gam_profit,
                                  residuals  = FALSE,
                                  rug        = FALSE,
                                  ci_level   = 0.95,
                                  nrow       = 3) &
  theme_minimal(base_size = 8) &
  theme(panel.grid.minor = element_blank())


p_area_smooths <- gratia::draw(gam_area,
                                residuals  = FALSE,
                                rug        = FALSE,
                                ci_level   = 0.95,
                                nrow       = 2) &
  theme_minimal(base_size = 8) &
  theme(panel.grid.minor = element_blank())

```


```{r fig-boxplot-scenario-profit, echo=FALSE, message=FALSE, warning=FALSE,   fig.cap="Distribution of site-level profit (\\euro~ha$^{-1}$~yr$^{-1}$) across scenario parameter levels. Each panel row shows one cost dimension (logistics cost scaler \\textit{cost.log}, top; establishment cost scaler \\textit{cost.est}, bottom); columns show opportunity cost level (\\textit{opp}, relative to baseline 150~\\euro~ha$^{-1}$~yr$^{-1}$). Fill colour encodes the revenue scaler. Each box summarises all active sites across all 10 stochastic replications of that parameter combination. The dashed horizontal line marks zero profit.",   fig.width = 7, fig.height = 6, dev = "pdf", out.width = "\\textwidth", cache = FALSE}

# ── Colour palette for revenue_level ─────────────────────────────────────────
rev_cols <- c(
  "0.5"  = "#d7191c",
  "0.75" = "#fdae61",
  "1.25" = "#2c7bb6"
)
rev_labels <- c("0.5"  = "Revenue ×0.5",
                "0.75" = "Revenue ×0.75",
                "1.25" = "Revenue ×1.25")

# ── Prepare data: label factors clearly ──────────────────────────────────────
df_plot_profit <- df_sens %>%
  filter(is.finite(profit_ha_yr)) %>%
  mutate(
    opp_lbl      = factor(paste0("Opp. ×", opp_cost),
                          levels = paste0("Opp. ×", sort(unique(opp_cost)))),
    revenue_lbl  = factor(as.character(revenue_level),
                          levels = c("0.5", "0.75", "1.25")),
    cost_log_lbl = factor(paste0("×", cost_log_level),
                          levels = paste0("×", sort(unique(cost_log_level)))),
    cost_est_lbl = factor(paste0("×", cost_est_level),
                          levels = paste0("×", sort(unique(cost_est_level))))
  )

# ── Panel A: profit by logistics cost (cost_log_level) ───────────────────────
p_log <- ggplot(df_plot_profit,
                aes(x     = cost_log_lbl,
                    y     = profit_ha_yr,
                    fill  = revenue_lbl)) +
  geom_hline(yintercept = 0, linetype = "dashed",
             colour = "grey40", linewidth = 0.45) +
  geom_boxplot(
    outlier.size   = 0.35,
    outlier.alpha  = 0.35,
    linewidth      = 0.35,
    position       = position_dodge(width = 0.78),
    width          = 0.68
  ) +
  facet_wrap(~ opp_lbl, nrow = 1, labeller = label_value) +
  scale_fill_manual(values = rev_cols, labels = rev_labels,
                    name = "Revenue scaler") +
  scale_y_continuous(labels = scales::label_number(big.mark = ",")) +
  labs(
    x     = "Logistics cost scaler (cost.log)",
    y     = expression("Profit ("*euro*" ha"^{-1}*" yr"^{-1}*")"),
    title = "(a) Logistics cost dimension"
  ) +
  theme_minimal(base_size = 8) +
  theme(
    strip.text       = element_text(size = 7.5, face = "bold"),
    legend.position  = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  )

# ── Panel B: profit by establishment cost (cost_est_level) ───────────────────
p_est <- ggplot(df_plot_profit,
                aes(x    = cost_est_lbl,
                    y    = profit_ha_yr,
                    fill = revenue_lbl)) +
  geom_hline(yintercept = 0, linetype = "dashed",
             colour = "grey40", linewidth = 0.45) +
  geom_boxplot(
    outlier.size   = 0.35,
    outlier.alpha  = 0.35,
    linewidth      = 0.35,
    position       = position_dodge(width = 0.78),
    width          = 0.68
  ) +
  facet_wrap(~ opp_lbl, nrow = 1, labeller = label_value) +
  scale_fill_manual(values = rev_cols, labels = rev_labels,
                    name = "Revenue scaler") +
  scale_y_continuous(labels = scales::label_number(big.mark = ",")) +
  labs(
    x     = "Establishment cost scaler (cost.est)",
    y     = expression("Profit ("*euro*" ha"^{-1}*" yr"^{-1}*")"),
    title = "(b) Establishment cost dimension"
  ) +
  theme_minimal(base_size = 8) +
  theme(
    strip.text         = element_text(size = 7.5, face = "bold"),
    legend.position    = "bottom",
    legend.key.size    = unit(0.38, "cm"),
    legend.text        = element_text(size = 7),
    legend.title       = element_text(size = 7.5, face = "bold"),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank()
  )

# ── Assemble ──────────────────────────────────────────────────────────────────
p_log / p_est + plot_layout(heights = c(1, 1))
```

```{r fig-boxplot-scenario-area, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of total AFS area dedicated per scenario replicate (ha) across parameter levels. Each panel row shows one cost dimension (logistics cost scaler \\textit{cost.log}, top; establishment cost scaler \\textit{cost.est}, bottom); columns show opportunity cost level. Fill colour encodes the revenue scaler. Each box summarises the 10 stochastic replications of that parameter combination. Higher values indicate broader spatial activation of the AFS potential.", fig.width = 7, fig.height = 6, dev = "pdf", out.width = "\\textwidth", cache = FALSE}

# ── Aggregate to scenario level: total active AFS area per replicate ──────────
# area_afs = tot_value (established ha) from extract_site_profit; sum over sites
df_scen_area_plot <- df_sens %>%
  group_by(scenario, opp_cost, cost_log_level, cost_est_level, revenue_level) %>%
  summarise(total_area_ha = sum(area_afs, na.rm = TRUE),
            .groups = "drop") %>%
  filter(is.finite(total_area_ha)) %>%
  mutate(
    opp_lbl      = factor(paste0("Opp. ×", opp_cost),
                          levels = paste0("Opp. ×", sort(unique(opp_cost)))),
    revenue_lbl  = factor(as.character(revenue_level),
                          levels = c("0.5", "0.75", "1.25")),
    cost_log_lbl = factor(paste0("×", cost_log_level),
                          levels = paste0("×", sort(unique(cost_log_level)))),
    cost_est_lbl = factor(paste0("×", cost_est_level),
                          levels = paste0("×", sort(unique(cost_est_level))))
  )

# ── Panel A: area by logistics cost ──────────────────────────────────────────
p_area_log <- ggplot(df_scen_area_plot,
                     aes(x    = cost_log_lbl,
                         y    = total_area_ha,
                         fill = revenue_lbl)) +
  geom_boxplot(
    outlier.size  = 0.5,
    outlier.alpha = 0.45,
    linewidth     = 0.35,
    position      = position_dodge(width = 0.78),
    width         = 0.68
  ) +
  facet_wrap(~ opp_lbl, nrow = 1, labeller = label_value) +
  scale_fill_manual(values = rev_cols, labels = rev_labels,
                    name = "Revenue scaler") +
  scale_y_continuous(labels = scales::label_number(big.mark = ",")) +
  labs(
    x     = "Logistics cost scaler (cost.log)",
    y     = "Total AFS area (ha)",
    title = "(a) Logistics cost dimension"
  ) +
  theme_minimal(base_size = 8) +
  theme(
    strip.text         = element_text(size = 7.5, face = "bold"),
    legend.position    = "none",
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank()
  )

# ── Panel B: area by establishment cost ──────────────────────────────────────
p_area_est <- ggplot(df_scen_area_plot,
                     aes(x    = cost_est_lbl,
                         y    = total_area_ha,
                         fill = revenue_lbl)) +
  geom_boxplot(
    outlier.size  = 0.5,
    outlier.alpha = 0.45,
    linewidth     = 0.35,
    position      = position_dodge(width = 0.78),
    width         = 0.68
  ) +
  facet_wrap(~ opp_lbl, nrow = 1, labeller = label_value) +
  scale_fill_manual(values = rev_cols, labels = rev_labels,
                    name = "Revenue scaler") +
  scale_y_continuous(labels = scales::label_number(big.mark = ",")) +
  labs(
    x     = "Establishment cost scaler (cost.est)",
    y     = "Total AFS area (ha)",
    title = "(b) Establishment cost dimension"
  ) +
  theme_minimal(base_size = 8) +
  theme(
    strip.text         = element_text(size = 7.5, face = "bold"),
    legend.position    = "bottom",
    legend.key.size    = unit(0.38, "cm"),
    legend.text        = element_text(size = 7),
    legend.title       = element_text(size = 7.5, face = "bold"),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank()
  )

# ── Assemble ──────────────────────────────────────────────────────────────────
p_area_log / p_area_est + plot_layout(heights = c(1, 1))
```


```{r tab-gam-fit, echo=FALSE, message=FALSE, warning=FALSE}

# ── Parametric coefficients for gam_profit (cost scalers entered linearly) ───
par_profit <- broom::tidy(gam_profit, parametric = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    Term = dplyr::recode(term,
      cost_log_level = "Logistics cost scaler (linear)",
      cost_est_level = "Establishment cost scaler (linear)",
      revenue_level  = "Revenue scaler (linear)"
    ),
    Model   = "Profit GAM",
    `Coeff.` = round(estimate, 2),
    `Std. err.` = round(std.error, 2),
    `$t$`   = round(statistic, 2),
    `$p$`   = dplyr::case_when(
      p.value < 0.001 ~ "$<$0.001",
      p.value < 0.01  ~ as.character(round(p.value, 3)),
      TRUE            ~ as.character(round(p.value, 3))
    )
  ) %>%
  select(Model, Term, `Coeff.`, `Std. err.`, `$t$`, `$p$`)

# ── Smooth terms: profit GAM ──────────────────────────────────────────────────
smo_profit <- gam_profit_summary %>%
  mutate(
    Term = dplyr::recode(Term,
      "s(opp_cost_site)"     = "s(site opp.~cost)",
      "s(avg_dist_hub_km)"   = "s(dist.~site--hub, km)",
      "s(avg_rotation_yr)"   = "s(avg.~rotation, yr)",
      "s(log_area_afs)"      = "s(log AFS area)"
    ),
    Model = "Profit GAM",
    `$p$` = dplyr::case_when(
      `p-value` < 0.001 ~ "$<$0.001",
      TRUE              ~ as.character(`p-value`)
    )
  ) %>%
  select(Model, Term, edf, `Ref.df`, `F statistic`, `$p$`)

# ── Smooth terms: area GAM ────────────────────────────────────────────────────
smo_area <- gam_area_summary %>%
  mutate(
    Term = dplyr::recode(Term,
      "s(opp_cost)"                        = "s(opp.~cost scaler)",
      "s(cost_log_level)"                  = "s(logistics cost scaler)",
      "s(cost_est_level)"                  = "s(establishment cost scaler)",
      "te(cost_log_level,cost_est_level)"  = "te(logistics $\\times$ establishment)",
      "s(revenue_level)"                   = "s(revenue scaler)"
    ),
    Model = "Area GAM",
    `$p$` = dplyr::case_when(
      `p-value` < 0.001 ~ "$<$0.001",
      TRUE              ~ as.character(`p-value`)
    )
  ) %>%
  select(Model, Term, edf, `Ref.df`, `F statistic`, `$p$`)

# ── Assemble combined smooth-term table ───────────────────────────────────────
tab_smooths <- bind_rows(smo_profit, smo_area) %>%
  rename(
    `Eff.~df`   = edf,
    `Ref.~df`   = `Ref.df`,
    `$F$`       = `F statistic`
  )

# ── Model-level fit statistics footer row ─────────────────────────────────────
fit_stats <- tibble::tribble(
  ~Model,       ~Statistic, ~Value,
  "Profit GAM", "$R^2$",    round(r2_profit, 3),
  "Profit GAM", "$n$",       nrow(df_model),
  "Area GAM",   "$R^2$",    round(r2_area, 3),
  "Area GAM",   "$n$",       nrow(df_scen_area)
)

kable(
  tab_smooths,
  format    = "latex", booktabs = TRUE, escape = FALSE,
  digits    = 3,
  caption   = paste0(
    "Approximate significance of smooth and tensor-product terms in the ",
    "site-profit GAM (Gaussian, REML) and the scenario-level total AFS area ",
    "GAM (Gamma, log link, REML). ",
    "Logistics and establishment cost scalers enter the profit GAM as linear ",
    "parametric terms (see Table~\\ref{tab:tab-gam-parametric}); ",
    "both appear as smooths in the area GAM, with an additional tensor-product ",
    "interaction surface. ",
    "$R^2$ and deviance explained: ",
    "profit GAM $R^2 = ", round(r2_profit, 3), "$, dev.~", dev_profit, "\\%; ",
    "area GAM $R^2 = ",   round(r2_area,   3), "$, dev.~", dev_area,   "\\%."
  ),
  label = "tab-gam-fit"
) %>%
  kable_styling(
    font_size     = 8,
    latex_options = c("hold_position", "scale_down"),
    full_width    = FALSE
  ) %>%
  pack_rows("Site-profit GAM — smooth terms", 1, nrow(smo_profit)) %>%
  pack_rows("AFS area GAM — smooth terms",
            nrow(smo_profit) + 1, nrow(tab_smooths)) %>%
  column_spec(1, width = "2.2cm") %>%
  column_spec(2, width = "5.0cm") %>%
  column_spec(3:6, width = "1.5cm")
```

```{r tab-gam-parametric, echo=FALSE, message=FALSE, warning=FALSE}

par_tab <- broom::tidy(gam_profit, parametric = TRUE) %>%
  mutate(
    Term = dplyr::recode(term,
      "(Intercept)"  = "Intercept",
      cost_log_level = "Logistics cost scaler",
      cost_est_level = "Establishment cost scaler",
      revenue_level  = "Revenue scaler"
    ),
    `$p$` = dplyr::case_when(
      p.value < 0.001 ~ "$<$0.001",
      p.value < 0.01  ~ as.character(round(p.value, 3)),
      TRUE            ~ as.character(round(p.value, 3))
    )
  ) %>%
  select(
    Predictor   = Term,
    `$\\hat{\\beta}$` = estimate,
    `SE`        = std.error,
    `$t$`       = statistic,
    `$p$`
  ) %>%
  mutate(across(c(`$\\hat{\\beta}$`, SE, `$t$`), ~ round(.x, 2)))

kable(
  par_tab,
  format    = "latex", booktabs = TRUE, escape = FALSE,
  caption   = paste0(
    "Parametric (linear) coefficients in the site-profit GAM. ",
    "The three scenario-level cost and revenue scalers are entered as linear ",
    "terms; a unit increase in each scaler corresponds to the stated ",
    "change in profit (\\euro\\,ha$^{-1}$\\,yr$^{-1}$). ",
    "The intercept is the expected profit at scaler $= 0$ and ",
    "all smooth terms at their mean."
  ),
  label = "tab-gam-parametric"
) %>%
  kable_styling(
    font_size     = 9,
    latex_options = c("hold_position"),
    full_width    = FALSE
  ) %>%
  column_spec(1, width = "5.2cm") %>%
  column_spec(2:5, width = "1.8cm") %>%
  row_spec(which(par_tab$Predictor == "Intercept"), italic = TRUE)
```


```{r fig-gam-profit-smooths, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = paste0( "Partial effects (centred smooths) of the four non-linear terms in the ", "site-profit GAM: site-level opportunity cost, hub distance, ", "average harvest rotation length, and log-transformed AFS strip area. ", "Shaded bands: 95\\,\\% simultaneous credible intervals. ", "Deviance explained: ", dev_profit, "\\,\\%; adjusted $R^2 = ", round(r2_profit, 3), "$." ), fig.width = 6.5, fig.height = 4.5, out.width = "\\textwidth"}

# p_profit_smooths is produced in lm-site-profit-sensitivity via gratia::draw()
# Apply consistent label overrides to the gratia panels
p_profit_smooths +
  patchwork::plot_annotation(
    tag_levels  = "a",
    tag_prefix  = "(",
    tag_suffix  = ")"
  ) &
  theme_minimal(base_size = 8) &
  theme(
    panel.grid.minor  = element_blank(),
    strip.text        = element_text(size = 7, face = "bold"),
    axis.title        = element_text(size = 7),
    plot.tag          = element_text(size = 7, face = "bold")
  )
```



```{r fig-gam-area-smooths, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = paste0("Partial effects of scenario-parameter smooths in the scenario-level ", "total AFS area GAM (Gamma family, log link). ", "Panel (e) shows the tensor-product interaction surface of logistics ", "and establishment cost scalers; contours indicate predicted total AFS ", "area on the response scale (ha). ", "Shaded bands in panels (a)--(d): 95\\,\\% simultaneous credible intervals. ", "Deviance explained: ", dev_area, "\\,\\%; adjusted $R^2 = ", round(r2_area, 3), "$." ), fig.width = 6.5, fig.height = 5.0, out.width = "\\textwidth"}

p_area_smooths +
  patchwork::plot_annotation(
    tag_levels  = "a",
    tag_prefix  = "(",
    tag_suffix  = ")"
  ) &
  theme_minimal(base_size = 8) &
  theme(
    panel.grid.minor  = element_blank(),
    strip.text        = element_text(size = 7, face = "bold"),
    axis.title        = element_text(size = 7),
    plot.tag          = element_text(size = 7, face = "bold")
  )
```


# Conclusion

This paper has developed and applied the AFS-SCD model, a MILP
formulation for the integrated design and operation of agroforestry
biomass supply chains. The model endogenises stand age dynamics through
binary arc-based age-lag variables, couples component-specific Gompertz
yield coefficients for stem, branch, and residue fractions with
multi-echelon logistics decisions, and embeds a product quality cascade
that allocates biomass grades to chemical, pulp, and energy markets
within a unified optimisation framework. The formulation is, to the best
of our knowledge, the first supply chain model to simultaneously
optimise AFS establishment timing, consecutive harvest scheduling under
rotation-age constraints, and product-grade allocation across spatially
heterogeneous consumer markets.

The case study for the Saxony-Anhalt tri-state region demonstrates the
practical relevance of the approach. The optimal supply chain
concentrates AFS establishments in the Elbe and Saale floodplains,
where high biomass productivity and proximity to large pulp mills create
the most favourable economics. The sensitivity analysis confirms that
profitability depends primarily on achievable gate prices for
high-quality (P1) biomass grades: scenarios combining reduced revenue
with high opportunity costs yield negative net profits, whereas
scenarios with full or elevated prices remain viable under cost pressure.
Opportunity cost is the dominant site-level discriminator, with parcels
exceeding approximately 300 \euro\ ha$^{-1}$ yr$^{-1}$ rarely selected
in the optimal solution. These findings underscore the importance of
product cascading as a strategic lever for AFS viability: maximising
the share of stem biomass directed to high-value chemical and pulp
applications substantially improves profitability compared to
energy-only offtake strategies.

Methodologically, the AFS-SCD model bridges the gap between stand-level
growth modelling and regional supply chain optimisation identified in
the literature [@Zahraee2020_BSC_review; @Lamers2015_Feedstock_Design].
The age-lag constraint structure generalises to other perennial biomass
crops with periodic coppice cycles, such as willow or black locust, and
to long-rotation silvoarable systems in which rotation length is a
strategic variable. The model is extensible to stochastic settings by
replacing deterministic yield coefficients with scenario-weighted
counterparts or by incorporating robust optimisation over yield
uncertainty intervals [@Sharma2013_BSC_review].

Several limitations point to directions for future research. First,
consumer demand is treated as exogenous and time-invariant; endogenising
demand responses to biomass price changes would require a
partial-equilibrium extension. Second, the case study uses a single
allometric scenario calibrated to Central European poplar; propagating
Gompertz parameter uncertainty through to supply chain outcomes would
strengthen the robustness analysis. Third, environmental co-benefits of
AFS --- including carbon sequestration, nitrate leaching reduction, and
biodiversity enhancement --- are not monetised in the current objective.
Integrating life cycle assessment indicators as additional objectives
or constraints would support a more comprehensive sustainability
evaluation, consistent with the cleaner production framing of this work.


# Appendix



```{r tab-consumers, echo=FALSE, message=FALSE, warning=FALSE, cache=F}

consumers %>%
  mutate(name_pseudo = paste("consumer ", seq_len(nrow(consumers))) ) %>% 
  select(name_pseudo, demand_P1, demand_P2, demand_P3, P1, P2, P3) %>%
  mutate(across(c(demand_P1, demand_P2, demand_P3), ~ ifelse(.x == 0, "–", format(.x, big.mark = ",")))) %>%
  mutate(across(c(P1, P2, P3), ~ ifelse(.x == 0, "–", paste0(.x)))) %>%
  kbl(
    col.names = c("Consumer", "P1 (kt/yr)", "P2 (kt/yr)", "P3 (kt/yr)",
                  "P1 (€/t)", "P2 (€/t)", "P3 (€/t)"),
    caption   = "Consumer demand quantities and gate prices by product grade. P1: chemical/pulp; P2: sawlog/pulpwood; P3: energy/pellets. Demands in kt wood input per year; prices in €/t wood input at the factory gate.",
    booktabs  = TRUE,
    align     = c("l", rep("r", 6)),
    label     = "tab-consumers"
  ) %>%
  kable_styling(latex_options = c("striped", "hold_position", "scale_down"),
                font_size = 8) %>%
  add_header_above(c(" " = 1, "Annual demand" = 3, "Gate price" = 3))
```

```{r fig-results-harvest-gantt, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Total production quantity of biomass types and its industrial usage over planning horizon (in tons).", fig.width=7, fig.height=5.5, dev="pdf", out.width="\\textwidth", cache=F, cache.extra = c("ext","sites_clustered", "storages" , "customers", "milp_instance")}


# ── Labels & colours ──────────────────────────────────────────────────────────
# Source side (axis 1): biomass fractions delivered from sites → hubs (Xij)
#   product index 1 = Stems, 2 = Branches, 3 = Residues  [§2.2, f_p(t)]
SRC_LABEL <- c(
  "1" = "Stems",
  "2" = "Branches",
  "3" = "Residues"
)

# Delivery side (axis 2): consumer categories receiving from hubs (Xjk)
#   del_product 1 = Industrial (P1), 2 = Pulp & Paper (P2), 3 = Energy (P3)
DEL_LABEL <- c(
  "1" = "Industrial",
  "2" = "Pulp & Paper",
  "3" = "Energy"
)

# Fill colours for source strata — must match paper Fig. max3-growth
P_COLS_SRC <- c(
  "Stems"    = "#1b9e77",
  "Branches" = "#d95f02",
  "Residues" = "#7570b3"   
)

# ── Aggregate Xjk: src_product × del_product, summed over all periods/consumers

flow_df <- ext$Xjk %>%
  group_by(src_product, del_product) %>%
  summarise(volume = sum(value, na.rm = TRUE), .groups = "drop") %>%
  filter(volume > 0) %>%
  mutate(
    # axis1 = biomass fraction (source side)
    src = factor(SRC_LABEL[as.character(src_product)], levels = SRC_LABEL),
    # axis2 = consumer category (delivery/sink side)
    del = factor(DEL_LABEL[as.character(del_product)], levels = DEL_LABEL)
  )

# ── Plot ──────────────────────────────────────────────────────────────────────
ggplot(flow_df,
       aes(axis1 = src, axis2 = del, y = volume, fill = src)) +

  geom_alluvium(
    alpha      = 0.72,
    width      = 1 / 3,
    knot.pos   = 0.45,
    curve_type = "sigmoid"
  ) +

  geom_stratum(
    width     = 1 / 3,
    fill      = "grey94",
    colour    = "grey50",
    linewidth = 0.3
  ) +

  geom_text(
    stat       = "stratum",
    aes(label  = after_stat(stratum)),
    size       = 2.8,
    colour     = "grey15",
    lineheight = 0.88
  ) +

  # Axis header annotations — placed below the stratum columns
  # x = 1 (axis1 = Feedstock fraction), x = 2 (axis2 = End use)
  annotate("text",
    x        = c(1, 2),
    y        = -Inf,
    label    = c("Feedstock", "End use"),
    vjust    = 1.6,
    size     = 3.0,
    fontface = "bold",
    colour   = "grey30"
  ) +

  scale_fill_manual(
    name   = "Biomass fraction",
    values = P_COLS_SRC          # only 3 source entries; no ghost colours
  ) +

  # ggalluvial uses integer axis positions (1, 2, …); do NOT set
  # limits = c("src", "del") — that breaks stratum positioning
  scale_x_discrete(
    expand = expansion(add = c(0.25, 0.25))  # widen margins for stratum labels
  ) +

  scale_y_continuous(
    name   = "Volume (t dry matter)",
    labels = comma_format(big.mark = ".", decimal.mark = ",")
  ) +

  coord_cartesian(clip = "off") +   # needed for below-axis annotations
  theme_minimal(base_size = 9) +
  theme(
    axis.title.x       = element_blank(),
    axis.text.x        = element_blank(),
    axis.ticks.x       = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    legend.position    = "none",
    legend.key.size    = unit(0.42, "cm"),
    plot.margin        = margin(t = 6, r = 6, b = 28, l = 6)
  )
```


<!-- ```{r fig-consumer-demand, cache=F, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Total biomass demand by consumer type and grade (P1 chemical, P2 pulp \\& paper, P3 energy/biogas). Bars show aggregated annual demand in kt wood input per year across all individual plants within each consumer type.", fig.width = 5.5, fig.height = 2.8, out.width="\\textwidth"} -->

<!-- # Re-use consumers_typed from above -->
<!-- # create unique types of consumers -->

<!-- consumers_typed <- consumers %>% -->
<!--   mutate( -->
<!--     type = case_when( -->
<!--       demand_P1 >= demand_P2 & demand_P1 >= demand_P3 & demand_P1 > 0 ~ "Industrial", -->
<!--       demand_P2 >= demand_P1 & demand_P2 >= demand_P3 & demand_P2 > 0 ~ "Paper & pulp", -->
<!--       demand_P3 > 0 ~ "Biogas", -->
<!--       TRUE          ~ "Other" -->
<!--     ) -->
<!--   ) -->


<!-- # Aggregate and pivot to long -->
<!-- demand_agg <- consumers_typed %>% -->
<!--   group_by(type) %>% -->
<!--   summarise( -->
<!--     P1 = sum(demand_P1), -->
<!--     P2 = sum(demand_P2), -->
<!--     P3 = sum(demand_P3), -->
<!--     .groups = "drop" -->
<!--   ) %>% -->
<!--   pivot_longer(cols = c(P1, P2, P3), -->
<!--                names_to  = "grade", -->
<!--                values_to = "demand_kt") %>% -->
<!--   filter(demand_kt > 0) %>% -->
<!--   mutate( -->
<!--     grade = factor(grade, levels = c("P1", "P2", "P3"), -->
<!--                    labels = c("P1 – Chemical", "P2 – Pulp", "P3 – Energy")), -->
<!--     type  = factor(type, -->
<!--                    levels = c("Industrial", "Paper & pulp", "Biogas")) -->
<!--   ) -->

<!-- GRADE_COLS <- c("P1 – Chemical" = "#1b7837", -->
<!--                 "P2 – Pulp"     = "#4393c3", -->
<!--                 "P3 – Energy"   = "#d6604d") -->

<!-- ggplot(demand_agg, aes(x = type, y = demand_kt, fill = grade)) + -->
<!--   geom_col(position = "dodge", width = 0.65, colour = "white", linewidth = 0.2) + -->
<!--   geom_text(aes(label = round(demand_kt, 0)), -->
<!--             position = position_dodge(width = 0.65), -->
<!--             vjust = -0.35, size = 2.4, colour = "grey30") + -->
<!--   scale_fill_manual(values = GRADE_COLS, name = "Biomass grade") + -->
<!--   scale_y_continuous( -->
<!--     expand = expansion(mult = c(0, 0.15)), -->
<!--     name   = "Annual demand (kt wood input / yr)" -->
<!--   ) + -->
<!--   labs(x = NULL) + -->
<!--   theme_minimal(base_size = 9) + -->
<!--   theme( -->
<!--     panel.grid.major.x = element_blank(), -->
<!--     panel.grid.minor   = element_blank(), -->
<!--     legend.position    = "bottom", -->
<!--     legend.key.size    = unit(0.35, "cm"), -->
<!--     legend.text        = element_text(size = 7), -->
<!--     axis.text.x        = element_text(size = 8) -->
<!--   ) -->
<!-- ``` -->


<!-- ```{r fig-results-inventory, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="End-of-period inventory levels at storage/pre-treatment hubs, disaggregated by product type. Each panel corresponds to one active hub.", fig.width=7, fig.height=3.5, dev="pdf", out.width="\\textwidth", cache=F, cache.extra = c("ext","sites_clustered", "storages" , "customers", "milp_instance")} -->

<!-- if( nrow(ext$S) > 0){ -->
<!-- inv_ts <- ext$S %>% -->
<!--   mutate(p = factor(p, labels = c("Stem", "Branches", "Residues")[seq_len(max(p))])) %>% -->
<!--   left_join( -->
<!--     storages %>% mutate(hub_id = row_number()) %>% select(hub_id, hub_name = name), -->
<!--     by = "hub_id" -->
<!--   ) %>% -->
<!--   filter(!is.na(value), value > 0) -->

<!-- ggplot(inv_ts, aes(x = period, y = value, fill = p)) + -->
<!--   geom_area(position = "stack", alpha = 0.78, colour = "white", linewidth = 0.2) + -->
<!--   facet_wrap(~hub_name, scales = "free_y") + -->
<!--   scale_fill_manual( -->
<!--     name   = "Product", -->
<!--     values = c("Stem" = "#1b9e77", "Branches" = "#d95f02", "Residues" = "#7570b3") -->
<!--   ) + -->
<!--   scale_x_continuous(breaks = seq(1, milp_instance$n_periods, by = 4)) + -->
<!--   scale_y_continuous(name = "Inventory (t)", labels = scales::comma) + -->
<!--   labs(x = "Planning period (year)") + -->
<!--   theme_minimal(base_size = 8) + -->
<!--   theme(legend.position = "bottom", -->
<!--         panel.grid.minor = element_blank(), -->
<!--         strip.text = element_text(face = "bold", size = 7))   -->
<!-- }else{ -->
<!--   NULL -->
<!-- } -->


<!-- ``` -->



<!-- ```{r tab-scenario-kpis, echo=FALSE, message=FALSE, warning=FALSE} -->
<!-- # Helper to compute KPIs for one scenario result ------------------------------ -->
<!-- compute_kpis <- function(res, scenario_name = NA_character_) { -->

<!--   milp_instance <- res$milp_instance -->
<!--   ext           <- res$ext -->
<!--   #result        <- res$result -->
<!--   obj.val       <- result$objective_value -->

<!--   # [NEU] Szenario-Definitions-Parameter aus res$setting extrahieren ---------- -->
<!--   setting       <- res$setting -->
<!--   scen_opp      <- if (!any(is.na(setting)))  setting$opp      else 150   # €/ha Opportunitätskosten-Basiswert -->
<!--   scen_cost     <- if (!any(is.na(setting)))  setting$cost     else 1.0   # Kostenmultiplikator (1 = Basis) -->
<!--   scen_revenue  <- if (!any(is.na(setting)))  setting$revenue  else 1.0   # Erlösmultiplikator  (1 = Basis) -->
<!--   # --------------------------------------------------------------------------- -->

<!--   # 1) Total profit over horizon (revenue – all costs) -->
<!--   price_df <- as.data.frame(milp_instance$consumer_prices) %>% -->
<!--     setNames(c("consumer_id", "del_product", "price")) %>% -->
<!--     mutate( -->
<!--       consumer_id = as.integer(consumer_id), -->
<!--       del_product = as.integer(del_product) -->
<!--     ) -->

<!--   rev_ts <- ext$Xjk %>% -->
<!--     left_join(price_df, by = c("consumer_id", "del_product")) %>% -->
<!--     mutate(rev = value * coalesce(price, 0)) %>% -->
<!--     group_by(period) %>% -->
<!--     summarise(rev = sum(rev, na.rm = TRUE), .groups = "drop") -->

<!--   # Site meta & active sites -->
<!--   active_site_ids <- ext$z %>% -->
<!--     filter(arc_type == "harvest") %>% -->
<!--     pull(site_id) %>% unique() -->

<!--   site_meta <- milp_instance$sites %>% -->
<!--     mutate(site_id = row_number()) %>% -->
<!--     filter(site_id %in% active_site_ids) -->

<!--   # Establishment, harvest, maintenance, opportunity costs -->
<!--   cost_est <- ext$z %>% -->
<!--     filter(arc_type == "establishment") %>% -->
<!--     left_join(site_meta, by = "site_id") %>% -->
<!--     mutate(cost = C_est * value, period = t) %>% -->
<!--     group_by(period) %>% -->
<!--     summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") -->

<!--   cost_harv <- ext$z %>% -->
<!--     filter(arc_type == "harvest") %>% -->
<!--     mutate(age_len = t - s) %>% -->
<!--     left_join(site_meta, by = "site_id") %>% -->
<!--     mutate(cost = C_harv * value, period = t) %>% -->
<!--     group_by(period) %>% -->
<!--     summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") -->

<!--   cost_main <- ext$z %>% -->
<!--     filter(arc_type == "harvest") %>% -->
<!--     mutate(age_len = t - s) %>% -->
<!--     left_join(site_meta, by = "site_id") %>% -->
<!--     mutate(cost = C_main * value * age_len, period = t) %>% -->
<!--     group_by(period) %>% -->
<!--     summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") -->

<!--   cost_opp <- ext$z %>% -->
<!--     filter(arc_type == "harvest") %>% -->
<!--     mutate(age_len = t - s) %>% -->
<!--     left_join(site_meta, by = "site_id") %>% -->
<!--     mutate(cost = C_opp * value * age_len, period = t) %>% -->
<!--     group_by(period) %>% -->
<!--     summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") -->

<!--   # Transport costs -->
<!--   dist_ij_m <- as.matrix(milp_instance$d_ij) -->
<!--   dist_jk_m <- as.matrix(milp_instance$d_jk) -->

<!--   cost_tr_raw <- ext$Xij %>% -->
<!--     mutate( -->
<!--       dist_km = mapply(function(i, j) dist_ij_m[i, j], site_id, hub_id), -->
<!--       cost    = milp_instance$c_tr_raw * dist_km * value -->
<!--     ) %>% -->
<!--     group_by(period) %>% -->
<!--     summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") -->

<!--   cost_tr_pre <- ext$Xjk %>% -->
<!--     mutate( -->
<!--       dist_km = mapply(function(j, k) dist_jk_m[j, k], hub_id, consumer_id), -->
<!--       cost    = milp_instance$c_tr_pre * dist_km * value -->
<!--     ) %>% -->
<!--     group_by(period) %>% -->
<!--     summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") -->

<!--   # Storage cost -->
<!--   cost_stor <- ext$S %>% -->
<!--     left_join( -->
<!--       milp_instance$storages %>% -->
<!--         mutate(hub_id = row_number()) %>% -->
<!--         select(hub_id, c_stor), -->
<!--       by = "hub_id" -->
<!--     ) %>% -->
<!--     mutate(cost = c_stor * value) %>% -->
<!--     group_by(period) %>% -->
<!--     summarise(cost = sum(cost, na.rm = TRUE), .groups = "drop") -->

<!--   # Combine all cost components -->
<!--   cost_ts <- bind_rows( -->
<!--     cost_est    %>% mutate(component = "Establishment"), -->
<!--     cost_harv   %>% mutate(component = "Harvesting"), -->
<!--     cost_main   %>% mutate(component = "Maintenance"), -->
<!--     cost_opp    %>% mutate(component = "Opportunity"), -->
<!--     cost_tr_raw %>% mutate(component = "Transport (raw)"), -->
<!--     cost_tr_pre %>% mutate(component = "Transport (pre-proc.)"), -->
<!--     cost_stor   %>% mutate(component = "Storage") -->
<!--   ) -->

<!--   total_rev    <- sum(rev_ts$rev,   na.rm = TRUE) -->
<!--   total_cost   <- sum(cost_ts$cost, na.rm = TRUE) -->
<!--   total_profit <- total_rev - total_cost -->

<!--   # 2) Number of active sites -->
<!--   n_active_sites    <- length(active_site_ids) -->
<!--   area_active_sites <- sum(site_meta$area_ha) -->

<!--   # 3) Total quantity of wood produced -->
<!--   total_wood <- ext$Xij %>% -->
<!--     group_by(p) %>% -->
<!--     summarise(prod_total = sum(value, na.rm = TRUE), .groups = "drop") %>% -->
<!--     rename(product = p) -->

<!--   # 4) Share of total customer demand satisfied -->
<!--   demand_df <- milp_instance$consumers %>% -->
<!--     select(starts_with("demand")) %>% -->
<!--     colSums() %>% -->
<!--     as.data.frame() %>% -->
<!--     rownames_to_column("product") -->

<!--   colnames(demand_df)[2] <- "demand" -->

<!--   demand_df <- demand_df %>% -->
<!--     mutate( -->
<!--       product = as.integer(as.character(factor(product, labels = 1:3))), -->
<!--       demand  = demand * milp_instance$n_periods -->
<!--     ) -->

<!--   total_delivered <- ext$Xjk %>% -->
<!--     group_by(del_product) %>% -->
<!--     summarise(total_consume = sum(value, na.rm = TRUE), .groups = "drop") %>% -->
<!--     rename(product = del_product) -->

<!--   dem_prod_df <- demand_df %>% -->
<!--     left_join(total_wood,      by = "product") %>% -->
<!--     left_join(total_delivered, by = "product") %>% -->
<!--     mutate( -->
<!--       share_demand = total_consume / demand, -->
<!--       share_prod   = total_consume / prod_total -->
<!--     ) -->

<!--   Share_demand_satisfied <- sum(dem_prod_df$total_consume) / sum(dem_prod_df$demand) -->

<!--   tibble( -->
<!--     Scenario                   = scenario_name, -->
<!--     Opp_cost_EUR_ha            = scen_opp,       # Opportunitätskosten €/ha -->
<!--     Cost_multiplier            = scen_cost,       # Kostenmultiplikator  (Basis = 1) -->
<!--     Revenue_multiplier         = scen_revenue,    # Erlösmultiplikator   (Basis = 1) -->
<!--     Rep_Total_profit_MEUR      = obj.val / 1e6, -->
<!--     Total_profit_MEUR          = total_profit / 1e6, -->
<!--     rel_profit                 = total_profit / area_active_sites / milp_instance$n_periods, -->
<!--     Active_sites               = n_active_sites, -->
<!--     Total_wood_units_kt        = sum(total_wood$prod_total) / milp_instance$n_periods / 1000, -->
<!--     Share_demand_satisfied_P1  = dem_prod_df$share_demand[2], -->
<!--     Share_demand_satisfied_P2  = dem_prod_df$share_demand[1], -->
<!--     Share_demand_satisfied_P3  = dem_prod_df$share_demand[3], -->
<!--     Share_demand_satisfied     = Share_demand_satisfied -->
<!--   ) -->
<!-- } -->

<!-- # ------------------------------------------------------------------ -->
<!-- #  Compute KPIs for benchmark and all sensitivity scenarios -->
<!-- # ------------------------------------------------------------------ -->
<!-- res_bench <- list(setting = NA, milp_instance = milp_instance, ext = ext, result = result) -->
<!-- kpi_bench <- compute_kpis(res_bench, scenario_name = "Bench.") -->

<!-- names(res.sens.list) <- 1:length(res.sens.list) -->

<!-- kpi_sens <- purrr::imap_dfr( -->
<!--   res.sens.list, -->
<!--   ~ compute_kpis(.x, scenario_name = .y) -->
<!-- ) -->

<!-- kpi_all <- bind_rows(kpi_bench, kpi_sens) %>% -->
<!--   group_by(Opp_cost_EUR_ha, Cost_multiplier, Revenue_multiplier) %>%  -->
<!--   summarize( -->
<!--     av_rel_profit = mean(rel_profit) ,  -->
<!--     av_active_sites = mean(Active_sites),  -->
<!--     av_wood = mean(Total_wood_units_kt),  -->
<!--     av_dem_sat = mean(Share_demand_satisfied)) %>%  -->
<!--   mutate( -->
<!--     av_rel_profit      = round(av_rel_profit, 0), -->
<!--     av_active_sites      = round(av_active_sites, 0), -->
<!--     av_wood        = round(av_wood, 0), -->
<!--     Opp_cost_EUR_ha            = round(Opp_cost_EUR_ha, 0), -->
<!--     Cost_multiplier            = round(Cost_multiplier, 2), -->
<!--     Revenue_multiplier         = round(Revenue_multiplier, 2), -->
<!--     av_dem_sat     = scales::percent(av_dem_sat,    accuracy = 0.1) -->
<!--   ) -->

<!-- kpi_all %>% -->
<!--   knitr::kable( -->
<!--     format   = "latex", -->
<!--     booktabs = TRUE, -->
<!--     align    = c("l", "r", "r", "r", "r", "r", "r", "r"), -->
<!--     col.names = c( -->
<!--       #"Scen.", -->
<!--       "Opp. cost (€/ha)",   # [NEU] -->
<!--       "Cost mult.",         # [NEU] -->
<!--       "Rev. mult.",         # [NEU] -->
<!--       "Profit (€/ha·a)", -->
<!--       "Act. sites", -->
<!--       "Wood (kt/a)", -->
<!--       "Demand sat. (%)" -->
<!--     ), -->
<!--     caption = "Key performance indicators for benchmark and sensitivity scenarios. -->
<!--                Opp.~cost: opportunity cost per hectare; Cost/Rev.~mult.: multipliers -->
<!--                relative to the baseline scenario (1.0 = baseline). -->
<!--                Total profit is net revenue minus all costs over the full planning -->
<!--                horizon; demand satisfaction reports the share of total customer -->
<!--                demand served." -->
<!--   ) %>% -->
<!--   kable_styling(font_size = 8, latex_options = c("striped", "scale_down")) -->
<!-- ``` -->




<!-- ```{r lm-marginal-effects, fig.cap="Partielle Marginaleffekte der kontinuierlichen Prädiktoren auf den Standortprofit. Alle anderen Variablen werden auf ihrem Mittelwert gehalten. Grau: 95\\%-Konfidenzband.", fig.width=7, fig.height=5} -->

<!-- library(marginaleffects) -->

<!-- vars_continuous <- c("opp_cost", "cost_level", "revenue_level", -->
<!--                      "avg_dist_hub_km", "avg_dist_consumer_km", -->
<!--                      "avg_rotation_yr", "share_p1") -->

<!-- plots_list <- lapply(vars_continuous, function(v) { -->
<!--   plot_predictions(lm_full, condition = v) + -->
<!--     labs(x = v, y = "Profit (€ ha yr)") + -->
<!--     theme_minimal(base_size = 8) + -->
<!--     theme(panel.grid.minor = element_blank()) -->
<!-- }) -->

<!-- patchwork::wrap_plots(plots_list, ncol = 3) -->
<!-- ``` -->


<!-- ```{r lm-avplots, fig.cap="Added-Variable Plots (partielle Regressionen): isolierter Effekt jedes Prädiktors auf den Standortprofit, kontrolliert für alle anderen Variablen.", fig.width=7, fig.height=6} -->

<!-- library(car) -->
<!-- avPlots(lm_full, cex = 0.5, col = "grey40", -->
<!--         col.lines = "#1b9e77", lwd = 1.5) -->
<!-- ``` -->

